### 2022-08-17 latex/article.tex
first commit

\documentclass{article}
\usepackage{natbib}
\usepackage[labelfont=bf]{caption}
\usepackage{graphicx}
\bibliographystyle{apalike}
\usepackage{lineno}
\linenumbers

\usepackage{geometry}
\geometry{letterpaper}
\geometry{margin=1in}

\usepackage{setspace}
\doublespacing
\captionsetup[figure]{font={stretch=2}}

\usepackage{authblk}

\newcommand{\matr}[1]{\mathbf{#1}}

\title{Application of the XT3D Multi-Point Flux Approximation to Vertically Staggered Grids}
\author{Xt3d Enthusiast1  \\
	Some Company  \\
	\and 
	Xt3d Enthusiast2 \\
	Another Company \\
	}

\date{\today}
% Hint: \title{what ever}, \author{who care} and \date{when ever} could stand 
% before or after the \begin{document} command 
% BUT the \maketitle command MUST come AFTER the \begin{document} command! 
\begin{document}

\maketitle

\begin{abstract}
This is the best paper ever... 
\end{abstract}

\section{Introduction}

Some intro stuff here about MODFLOW \citep{modflow6framework, modflow6gwf, modflow6gwt} and XT3D \citep{modflow6xt3d}...

%\begin{figure}[ht]
%	\centering
%	\includegraphics[scale=0.6]{figures/fig1.pdf}
%	\caption{This is the first figure.}
%	\label{fig:firstfigure}
%\end{figure}

\section{Example}

Some examples here...

\subsection{Example 1}

Example 1...

\subsection{Example 2}

Example 2...

\section{Summary and Conclusions}

\section{Acknowledgments}
Thank all those reviewers.

\section{Software Availability}
MODFLOW 6 is open source; software is developed following modern software development principles.  FloPy \citep{bakker2016scripting} contains full support for all MODFLOW 6 models and packages.  We welcome input to the community through our public software repository.  MODFLOW 6 is developed in the open, designed to be teachable, runs on multilple

\bibliography{references.bib}

\end{document}

### 2022-08-17 python/dis2disu.py

@@ -1,0 +1,331 @@
+import numpy as np
+
+class ConnectionProperty():
+    def __init__(self, connected_node, this_node, ihc, cl1, cl2, hwva, angldegx):
+        self.connected_node = connected_node
+        self.this_node = this_node
+        self.ihc = ihc
+        self.cl1 = cl1
+        self.cl2 = cl2
+        self.hwva = hwva
+        self.angldegx = angldegx
+
+
+class ConnectionPropertyList():
+    def __init__(self):
+        self.number_of_connections = None
+        self.connection_dict = {}
+    
+    def new_connection(self, connected_node, this_node, ihc, cl1, cl2, hwva, angldegx):
+        connection = ConnectionProperty(connected_node, this_node, ihc, cl1, cl2, hwva, angldegx)
+        if self.number_of_connections is None:
+            connection_number = 0
+            self.number_of_connections = 0
+        else:
+            connection_number = self.number_of_connections
+        self.connection_dict[connection_number] = connection
+        self.number_of_connections += 1
+        return connection_number
+
+
+class Dis2Disu():
+
+    def __init__(self, delr, delc, top, botm, staggered):
+        
+        assert isinstance(delr, np.ndarray)
+        assert isinstance(delc, np.ndarray)
+        assert isinstance(top, np.ndarray)
+        assert isinstance(botm, np.ndarray)
+
+        assert delr.ndim == 1, f"delr must have ndim = 1 but found {delr.ndim}"
+        assert delc.ndim == 1, f"delc must have ndim = 1 but found {delc.ndim}"
+        assert top.ndim == 2, f"top must have ndim = 2 but found {top.ndim}"
+        assert botm.ndim == 3, f"botm must have ndim = 3 but found {botm.ndim}"
+
+        self.nlay, self.nrow, self.ncol = botm.shape
+        self._nodes = self.nlay * self.nrow * self.ncol
+
+        self.delr = delr
+        self.delc = delc
+        self.top = top
+        self.botm = botm
+        self.staggered = staggered
+
+        self.connection_list = ConnectionPropertyList()
+        self._initialize()
+        self._build_connectivity()
+
+        return
+
+    def _initialize(self):
+
+        # unknown at this point
+        self._nja = None
+        self._nvert = None
+        self._iac = np.zeros(self._nodes, dtype=int)
+        self._ja = None
+        self._ihc = None
+        self._cl12 = None
+        self._hwva = None
+        self._angldegx = None
[…]

### 2022-08-17 latex/article.tex
* Added Kerry and Jim's regional PRAMS model with cross-connections as a possible test model

@@
 
-\title{Application of the XT3D Multi-Point Flux Approximation to Vertically Staggered Grids}
+\title{Application of the XT3D Multi-Point Flux Approximation to Vertically Staggered Grids \\
+	\normalsize or maybe \\
+	\LARGE Application of the XT3D Multi-Point Flux Approximation and Enhanced Grid Connectivity to Improve Accuracy of Flows in MODFLOW 6 Models With Steeply Sloping Layers}
 \author{Xt3d Enthusiast1  \\
@@
 \maketitle
+
+\textbf{Conflict of interest:} None.
+
+\textbf{Key words:} Key words ...
+
+\textbf{Article impact statement:} Article impact statement ...
 
@@
 
-\section{Example}
+\section{Theoretical Background}
 
-Some examples here...
+Introduce vertically offset grids and explain the connection angle/length issue.
 
-\subsection{Example 1}
+Summarize XT3D and how it accounts for connection angle/length.
+Reference Kerry and Jim’s demonstration that, in spite of that, XT3D doesn’t really improve things for a steeply sloping grid.
+Must be something else going on.
 
-Example 1...
+Explain wormholes and how they induce horizontal flow in a sloping channel regardless of XT3D.
+\begin{itemize}
+	\item Grid with “connector cells”
+	\item Role of flows between layers
+	\item Shutting off flows between layers using extreme anisotropy in connector cells
+	\begin{itemize}
+		\item Sloping flow in connector cells
+		\item Horizontal flow in flat-top cells
+	\end{itemize}
+	\item Squashing of connector cells VO grid with horizontal flows and “wormholes”
+\end{itemize}
 
-\subsection{Example 2}
+Proposed solution is to introduce cross-connections between layers.
 
-Example 2...
+\section{Approach}
 
-\section{Summary and Conclusions}
+Summarize the overall approach here.
+
+Will use a DISV plan-view model with connector cells and XT3D to demonstrate the "wormhole" effect discussed in the Theoretical Background in the limit as connector cells are squashed out.
+(Also can look at the other limit, as flat-top cells are squashed and connector cells dominate, so grid follows the channel boundary.)
+
+Will use a DIS cross-sectional model to show results you get on a vertically staggered grid (without cross-connections), with and without XT3D.  (Basically what Kerry and Jim showed.)
+
+Will convert the DIS grid to a DISU grid with cross-connections and show improved results, with and without XT3D.
+
+Will rerun Kerry and Jim's PRAMS (regional) model with cross-connections to demonstrate their practical effect?
+
+\section{Description of Test Problems}
+
+Describe the test problem setups here.
+
+\subsection{Test problem 1 (DISV plan-view with connector cells)}
+
+Test problem 1...
+
+\subsection{Test problem 2 (DIS cross-sectional)}
+
+Test problem 2...
+
+\subsection{Test problem 3 (DISU cross-sectional with cross-connections)}
+
+Test problem 3...
+
+\subsection{Test problem 4 (PRAMS model with cross-connections)?}
+
+Test problem 4...
+
+\section{Results and Discussion}
+
+\section{Conclusions}
 
@@
 
+\section{Supporting Information}
+
+\section{Appendix}
+
 \bibliography{references.bib}

### 2022-08-19 latex/article.tex
Rethought the inclusion of the PRAMS model, since this is envisioned as a succinct Methods Note, so it's probably best to focus on demonstrating the concept using just a few simple test problems and leave more complex applications for another time.
* Added Bardot et al (2022) to references * 'Blocked in' more of the intro * Removed Software Availability section * Minor tweaks

@@
 	\LARGE Application of the XT3D Multi-Point Flux Approximation and Enhanced Grid Connectivity to Improve Accuracy of Flows in MODFLOW 6 Models With Steeply Sloping Layers}
-\author{Xt3d Enthusiast1  \\
-	Some Company  \\
+\author{
+	Xt3d Enthusiast1, Affiliation1  \\
 	\and 
-	Xt3d Enthusiast2 \\
-	Another Company \\
+	Xt3d Enthusiast2, Affiliation2 \\
+	\and 
+	Xt3d Enthusiast3, Affiliation3 \\
+	\and 
+	Xt3d Enthusiast4, Affiliation4 \\
 	}
@@
 
-Some intro stuff here about MODFLOW \citep{modflow6framework, modflow6gwf, modflow6gwt} and XT3D \citep{modflow6xt3d}...
+Some intro stuff here about MODFLOW \citep{modflow6framework, modflow6gwf, modflow6gwt} and XT3D \citep{modflow6xt3d}... Horizontal tops and bottoms of cells... Connection length issue...
+
+Reference and summarize VO results in \cite{bardot2022}.
+This is an extreme case (virtually impermeable domain surrounding the channel) that nicely accentuates the problem, but there should be some error in any application involving VO grids.
+Obviously expect it to be a bigger issue for greater offsets.
+XT3D didn't help, which was initially surprising since it takes into account connection angles and lengths.
+Not specifying all the angles in the NPF input would cause XT3D to default to "nozee" and therefore assume "horizontal" connections are strictly horizontal, but that was not the case in \cite{bardot2022}.
+
+We explain in this note that it's not just a connection angle and length issue, though that has an effect.
+The main problem with a VO grid is insufficient hydraulic communication between grid cells in adjacent layers.
+(It's not restricted to layered models, but it's convenient to describe it in terms of layers.)  We propose and test a solution to the problem that involves enhanced connectivity between cells.
+
+We present three test models that demonstrate 1) how a VO grid effectively creates wormholes for flow, 2) results on an unmodified VO DIS grid, 3) results on corresponding DISU grid with enhanced connectivity.
+Results suggest that enhanced connectivity resolves the issue.
 
@@
 
-Introduce vertically offset grids and explain the connection angle/length issue.
+Explain vertically offset (VO) grids and the connection angle/length issue in more detail than in the Intro. Discuss areas.
 
 Summarize XT3D and how it accounts for connection angle/length.
-Reference Kerry and Jim’s demonstration that, in spite of that, XT3D doesn’t really improve things for a steeply sloping grid.
-Must be something else going on.
+Reference demonstration by \cite{bardot2022} that, in spite of that, XT3D doesn’t really improve things for a steeply sloping grid.
+Must be something else going on in VO grids.
 
-Explain wormholes and how they induce horizontal flow in a sloping channel regardless of XT3D.
+Explain wormholes and how they induce horizontal flow in the sloping channel regardless of XT3D.
 \begin{itemize}
@@
 
-Will use a DIS cross-sectional model to show results you get on a vertically staggered grid (without cross-connections), with and without XT3D.  (Basically what Kerry and Jim showed.)
+Will use a DIS cross-sectional model to show results you get on a vertically staggered grid (without cross-connections), with and without XT3D.
+Basically show that we can reproduce what \cite{bardot2022} found.
 
 Will convert the DIS grid to a DISU grid with cross-connections and show improved results, with and without XT3D.
-
-Will rerun Kerry and Jim's PRAMS (regional) model with cross-connections to demonstrate their practical effect?
 
@@
 
-\subsection{Test problem 4 (PRAMS model with cross-connections)?}
-
-Test problem 4...
-
 \section{Results and Discussion}
@@
 Thank all those reviewers.
-
-\section{Software Availability}
-MODFLOW 6 is open source; software is developed following modern software development principles.
-FloPy \citep{bakker2016scripting} contains full support for all MODFLOW 6 models and packages.
-We welcome input to the community through our public software repository.
-MODFLOW 6 is developed in the open, designed to be teachable, runs on multilple
 

### 2022-08-26 python/dis2disu.py
* Changes to the disu_approach notebook:     -- Generalized to allow any channel slope (not just the slope that gives 1/2-cell overlap)     -- Reworked the specification and calculation of grid parameters to be more like in wormholes and result in square cells     -- Changed the boundary heads from uniform at each end to gradient-honoring as in Bardot et al (2022)     -- Added a routine that print a summary of specific discharge results     -- Added code to recalculate qx to account for overlap areas * In the dis-disu converter, removed what appeared to be adjustment for overlap area that was

@@ -302,19 +302,23 @@ class Dis2Disu():
                 if angldegx == 0: # right
                     cl1 = 0.5 * self.delr[j]
                     cl2 = 0.5 * self.delr[j + 1]
-                    hwva = 0.5 * self.delc[i]
+                    #hwva = 0.5 * self.delc[i]
+                    hwva = self.delc[i]
                 elif angldegx == 90.: # back
                     cl1 = 0.5 * self.delc[i]
                     cl2 = 0.5 * self.delc[i - 1]
-                    hwva = 0.5 * self.delr[j]
+                    #hwva = 0.5 * self.delr[j]
+                    hwva = self.delr[j]
                 elif angldegx == 180: # left
                     cl1 = 0.5 * self.delr[j]
                     cl2 = 0.5 * self.delr[j - 1]
-                    hwva = 0.5 * self.delc[i]
+                    #hwva = 0.5 * self.delc[i]
+                    hwva = self.delc[i]
                 elif angldegx == 270.: # front
                     cl1 = 0.5 * self.delc[i]
                     cl2 = 0.5 * self.delc[i + 1]
-                    hwva = 0.5 * self.delr[j]
+                    #hwva = 0.5 * self.delr[j]
+                    hwva = self.delr[j]
                 connection_number = self.connection_list.new_connection(nn, this_node, ihc, cl1, cl2, hwva, angldegx)
                 connected_cells.append((nn, connection_number))
         return connected_cells

### 2022-09-09 latex/article.tex
Clumsily inserted some figures into the manscript
* Modified the disu_approach notebook to automatically run 4 scenarios: [without cross-connections, with cross-connections] by [without xt3d, with xt3d]; each scenario has its own subfolder * Updated a figure filename in the manuscript
* Extensive additions and changes to the manuscript. Notable additions include fleshing out the Theoretical Background section and appending a temporary summary of preliminary results at the end. * Addition of crossflow scenarios to disu_approach notebook. Also changed/shortened the root filename for input and ou

@@
 \usepackage[labelfont=bf]{caption}
+\usepackage{subcaption}
 \usepackage{graphicx}
+\usepackage{float}
 \bibliographystyle{apalike}
@@
 
-\title{Application of the XT3D Multi-Point Flux Approximation to Vertically Staggered Grids \\
-	\normalsize or maybe \\
-	\LARGE Application of the XT3D Multi-Point Flux Approximation and Enhanced Grid Connectivity to Improve Accuracy of Flows in MODFLOW 6 Models With Steeply Sloping Layers}
+\title{Application of the XT3D Multi-Point Flux Approximation to Vertically Staggered Grids in MODFLOW 6 to Improve Accuracy of Simulated Flows in Steeply Dipping Layers}
 \author{
@@
 
-\textbf{Conflict of interest:} None.
+\textbf{Conflict of interest:} There is no conflict that is of interest to us.
 
-\textbf{Key words:} Key words ...
+\textbf{Key words:} keyboard, keynote, turnkey, turkey, monkey
 
-\textbf{Article impact statement:} Article impact statement ...
+\textbf{Article impact statement:} This will likely leave a sizable impact crater.
 
@@
 Some intro stuff here about MODFLOW \citep{modflow6framework, modflow6gwf, modflow6gwt} and XT3D \citep{modflow6xt3d}... Horizontal tops and bottoms of cells... Connection length issue...
+
+If cell-cell flows are  off, advective and dispersive transport will be off.
+
+Horizontal cell tops and bottoms - benefits.  Vertical offsets.  Drawbacks - staircase boundaries, CVFD violations.  XT3D can address connection angle and length.
+
+Having flat tops and bottoms but "distorting" the grid vertically (having vertical offsets) is discussed in \cite{modflow84} on p.
+140 and fig 27, though tops and bottoms are not drawn as horizontal.
+Benefits cited: "minimize the number of model layers required to simulate an aquifer system" (compared to "horizontal discretization" using a "rectilinear grid"); and "each distorted cell is simulated as if it were rectangular so that flow may be approximated by the standard finite-difference equation." Note that the latter benefit extends to the case of a partially saturated cell, in which the conductance is based on the saturated thickness measured between the horizontal water table and the horizontal cell bottom.
+Also simplifies particle tracking, though a relative z mapping is needed when passing particles between cells.
+
+Vertical distortion is also discussed in \cite{modflow2005} and pictured with flat tops and bottoms in fig 4-2 of that report.
+
+Thought: Vertical offset happens even on a completely rectilinear grid if the water table elevation varies in horizontally adjacent cells. Doesn't seem like the same kind of issue, but think on it.
+
+Thought: What happens with a head gradient perpendicular to the dip of the channel (across rather than along the channel)? Preliminary results suggest cross-connections and xt3d help.
+
+Thought: Difference between truly single-layer model of channel vs embedded single-layer channel -- no vertical connections vs vertical connections that are no-flux.
+Effect on specific discharge calculations?
 
@@
 
-We present three test models that demonstrate 1) how a VO grid effectively creates wormholes for flow, 2) results on an unmodified VO DIS grid, 3) results on corresponding DISU grid with enhanced connectivity.
+We present two test models that demonstrate 1) results on an unmodified VO DIS grid and 2) results on corresponding DISU grid with enhanced connectivity.
 Results suggest that enhanced connectivity resolves the issue.
@@
 
-Explain vertically offset (VO) grids and the connection angle/length issue in more detail than in the Intro. Discuss areas.
+\begin{figure}
+	\begin{center}
+%	\includegraphics[scale=0.8]{../figures/schem_conn_area_flux.png}
+	\includegraphics[scale=0.4]{../figures/schem_conn_area_flux_resaved.png}
+	\caption{Schematics showing (a - c) grid connectivity and vertical cell interface areas and (c - f) cell interface fluxes in two-layer models of a dipping channel using three different grid configurations: (a, d) dip-following, (b, e) vertically offset, and (c, f) cross-connected.
+Hatching denotes impermeable boundaries along the top and bottom of the channel.
+In (a) - (c), gray circles and gray lines represent cell centers and grid connections between cells, respectively, and two-headed arrows indicate cell thicknesses used to compute cell interface areas.
+In (d) - (f), gray arrows represent specific-discharge (groundwater flux) vectors for uniform flow through the channel, and black arrows represent the corresponding components of flux normal to cell interfaces.
+For visual clarity, (f) shows gray arrows only for the cell interfaces that do not exist in (e).}
+	\label{fig:schem-conn-area-flux}
+	\end{center}
+\end{figure}
 
-Summarize XT3D and how it accounts for connection angle/length.
-Reference demonstration by \cite{bardot2022} that, in spite of that, XT3D doesn’t really improve things for a steeply sloping grid.
-Must be something else going on in VO grids.
+Figure \ref{fig:schem-conn-area-flux}a shows a group of four cells in a hypothetical model in which the dipping channel is discretized vertically into two layers of cells.
+Although such a configuration of cells can be constructed in plan view using the DISV grid type of MODFLOW 6, it cannot be constructed in cross-section because MODFLOW requires cells to have horizontal tops and bottoms.
+The sloping tops and bottoms of these hypothetical cells allow the cell geometry to conform to the dip of the channel, and we will therefore refer to this grid configuration as "dip-following." The grid connectivity is analogous to the connectivity found in a MODFLOW 6 DIS (structured, or regular) grid: a cell is hydraulically connected to each cell with which it shares a face.
+For example, cell 1A is connected to cells 1B and 2A, but not to cell 2B.
+Because cells share entire faces, i.e., adjacent cell faces overlap completely, the entire cell thickness is relevant in determining the area for flow between cells in the same layer.
 
-Explain wormholes and how they induce horizontal flow in the sloping channel regardless of XT3D.
-\begin{itemize}
-	\item Grid with “connector cells”
-	\item Role of flows between layers
-	\item Shutting off flows between layers using extreme anisotropy in connector cells
-	\begin{itemize}
-		\item Sloping flow in connector cells
-		\item Horizontal flow in flat-top cells
-	\end{itemize}
-	\item Squashing of connector cells VO grid with horizontal flows and “wormholes”
-\end{itemize}
+One would intuitively expect a dip-following grid to be well suited to simulating uniform flow along the channel.
+In figure \ref{fig:schem-conn-area-flux}d, a specific discharge (groundwater flux) vector representative of uniform flow along the channel is superimposed on each cell interface, together with a vector representing the component of the flux normal to the interface.
+The normal flux vectors show the directions in which groundwater must be able to flow between cells to accurately represent the uniform flow field.
+In the case of a dip-following grid, flow must simply pass from left to right between adjacent cells in the same layer (e.g., from cell 1A to cell 1B, and from cell 2A to cell 2B).
+There is no flow of groundwater between cells in different layers (e.g., between cells 1A and 2A, or between cells 1B and 2B).
+The grid connectivity shown in figure \ref{fig:schem-conn-area-flux}a is sufficient to accommodate this straighforward pattern of flows between cells; in fact, vertical connections between cells in different layers are not even needed, and the channel could be represented just as well using a single layer.
 
-Proposed solution is to introduce cross-connections between layers.
+Figure \ref{fig:schem-conn-area-flux}b shows the four cells as they would be configured for a DIS or DISV grid in a two-layer, cross-sectional MODFLOW 6 model of the channel.
+The cells have horizontal tops and bottoms and can therefore follow the dip of the channel only on average, in what \cite{bardot2022} refer to as "staircase" fashion, and we will refer to this grid configuration as "vertically offset." The pattern of grid connectivity is analogous to that in figure \ref{fig:schem-conn-area-flux}a, but it is not based strictly on which cell face overlap.
+Rather, a cell is connected to each cell in the same layer with which it has an overlapping face and with each cell in an adjacent layer with which it shares a horizontal face.
+Significantly, cells that have overlapping vertical faces but are in different layers (e.g., cells 1A and 2B) are not connected.
+In the standard and Newton-Raphson "conductance-based" formulations for flow, the flow between two cells in the same layer of a DIS or DISV grid (XXXXX or DISU that's not vertically staggered, but do we want to mention that here?
+XXXXX) is proportional to the difference in heads computed at the two cell centers and a conductance based on an effective hydraulic conductivity for the connection between the cells.
+If the two cells are fully saturated and have the same hydraulic conductivity and thickness, as in this example, the cell interface area used to calculate the conductance is effectively based on the full cell thickness, as shown in figure \ref{fig:schem-conn-area-flux}b.
+Also, the length on which the conductance is based is the horizontal distance, not the straight-line distance, between the cell centers.
+
+Conductance-based formulations for flow can be expected to give accurate results if the model grid satisfies a set of geometric "control-volume finite-difference (CVFD) conditions" (XXXXX; XXXXX).
+One of those conditions is that the straight-line connection between cell centers must be perpendicular to the cell interface.
+In a vertically offset grid, connections between cells across vertical cell interfaces are generally not horizontal and, therefore, not perpendicular to the interface.
+(XXXXX Also, in the grids considered by XXXXX, there was no partial overlap between cells faces, but do we want to get into that here?
+XXXXX) For example, in \ref{fig:schem-conn-area-flux}b, the nominally "horizontal" connection between cells 1A and 1B is not perpendicular to the cell interface, which compormises the accuracy of conductance-based formulations.
+The XT3D capability is a gradient reconstruction (interpolation) scheme introduced in MODFLOW 6 that allows simulation of fully three-dimensional anisotropy and improves the accuracy of flows between cells on grids that do not satisfy the CVFD conditions.
+
+In their dipping-channel benchmark problem, \cite{bardot2022} found that, as expected, the standard, conductance-based formulation gave poor results for a steeply ($30^{\circ}$) dipping channel.
+The simulated flow was not along the dip of the channel, but nearly horizontal, and the magnitude was too large.
+Given the ability of XT3D to compensate for grid connections that are not perpendicular to cell interfaces, rerunning the simulation with XT3D activated could have been expected to improve the flow solution substantially.
+However, \cite{bardot2022} obtained similarly poor results with and without XT3D.
+It appeared that something other than a simple violation of the CVFD perpendicularity condition was responsible, and \cite{bardot2022} noted a hypothesis suggested by two authors of the present paper (Provost and Langevin).
+
+The hypothesis can be explained with the help of figures \ref{fig:schem-conn-area-flux}b and \ref{fig:schem-conn-area-flux}e.
+Figure \ref{fig:schem-conn-area-flux}e shows the specific discharge vector superimposed on each cell interface of the vertically offset grid, together with a vector representing the component of the flux normal to the interface.
+As in the case of a dip-following grid (figure \ref{fig:schem-conn-area-flux}d), flow passes from left to right between adjacent cells in the same layer.
+However, unlike in the dip-following grid, there must also be vertical flow from each cell in layer 2 and to the cell immediately above it in layer 1 (e.g., from cell 2A and to cell 1A).
+This pattern of flow between cells obviously cannot be a solution to uniform, steady-state flow through the channel, since the exclusively upward vertical flows would deplete layer 2 and accumulate water in layer 1.
+Short of resorting to alternating upward and downward vertical flows, which would compromise the uniformity of the flow, the only alternative for representing uniform, steady-state flow on the vertically offset grid is to have exclusively horizontal flow.
+Although this concept is illustrated for a two-layer model in figure \ref{fig:schem-conn-area-flux}e, the same reasoning applies to a model of the channel with any number of layers.
+This explains why \cite{bardot2022} observed nearly horizontal flow in the middle portion of the channel, away from the lateral boundaries, where nearly uniform flow was established.
+Thus, it appears the nearly horizontal flows are not primarily due to what \cite{bardot2022} call the "cosine error" associated with the orientation and length of the sloping grid connections, but to the inability of the vertically offset grid to route flow appropriately between cells.
+
+This suggests an approach for constructing the grid to improve the accuracy of the flow solution.
+Figure \ref{fig:schem-conn-area-flux}c shows a grid that includes additional connections between cells in different layers compared to the grid in \ref{fig:schem-conn-area-flux}b.
+For example, cell 1A is connected not only cells 2A and 1B, but also the cell 2B.
+The grid connectivity is determined by overlap between cell faces; if two cells have faces that overlap even partially, the cells are connected across those faces.
+The area of the interface is the area of overlap between the cell faces.
+This grid configuration, which was introduced MODFLOW-USG (XXXXXXXXX) and is available for the DISU grid type in MODFLOW 6, is called "vertically staggered" and allows a cell to have nominally "horizontal" connections with more than one cell across a given cell face.
+In this case, vertical staggering is used to introduce "cross-connections" between cells in different layers.
+The ramifications of cross-connections for flow between cells are illustrated in figure \ref{fig:schem-conn-area-flux}f.
+Inclusion of cross-connections (e.g., between cells 1A and 2B) allows upward vertical flows from layer 2 to layer 1 (e.g., from cell 2A to cell 1A) to be routed back down to layer 2 so water does not accumulate in layer 1.
+The remainder of this paper describes test problems that evaluate the effectiveness of cross-connections for improving the accuracy of simulated flow though the channel.
+
+XXXXX Note: This idea is not limited to layered grids, but it's easiest to discuss in terms of layers. XXXXX
 
@@
 
-Will use a DISV plan-view model with connector cells and XT3D to demonstrate the "wormhole" effect discussed in the Theoretical Background in the limit as connector cells are squashed out.
-(Also can look at the other limit, as flat-top cells are squashed and connector cells dominate, so grid follows the channel boundary.)
-
-Will use a DIS cross-sectional model to show results you get on a vertically staggered grid (without cross-connections), with and without XT3D.
+Will use a DIS grid (XXXXX converted to an equivalent DISU grid without cross=connections?
+XXXXX) cross-sectional model to show results you get on a vertically offset grid, with and without XT3D.
 Basically show that we can reproduce what \cite{bardot2022} found.
@@
 
-\subsection{Test problem 1 (DISV plan-view with connector cells)}
+\subsection{Test problem 1 (DIS cross-sectional)}
 
@@
 
-\subsection{Test problem 2 (DIS cross-sectional)}
+XXXXX Idea: See how heterogeneity affects the error? XXXXX
+
+\subsection{Test problem 2 (DISU cross-sectional with cross-connections)}
 
@@
 
-\subsection{Test problem 3 (DISU cross-sectional with cross-connections)}
-
-Test problem 3...
+\begin{figure}[H]
+	\begin{center}
+	\includegraphics[scale=0.5]{../figures/disu-x-cc-head.png}
+	\caption{Cross-connections are good. XT3D is good.}
+	\label{fig:mymodel-head}
+	\end{center}
+\end{figure}
 
 \section{Results and Discussion}
+
+XXXXX See the temporary "preliminary results" section at the end of the manuscript XXXXX
 
@@
 
+\section{Oblique reference}
+
+\begin{figure}[H]
+	\includegraphics[scale=0.5]{../figures/oblique_reference.png}
+\end{figure}
+
+\section{preliminary results}
+This temporary section is a summary of preliminary results that can help provide context and guide our thinking as we construct the story.  Specific results may or may not make it into the paper.
+
+The models referenced are:
+\begin{itemize}
+	\item wormholes -- This is a plan-view, DISV model that emulates a cross-sectional model and includes "connector" cells that can have sloped tops and bottoms.
+Cell dimensions and hydraulic conductivities can be adjusted to approach a "dip-following" mesh (in which cell tops and bottoms follow the dip of the channel) at one extreme and a vertically offset mesh (the kind MODFLOW actually uses in cross-section) at another extreme.
+A vertically offset mesh is approached by making the width of the connector cells very small in the x direction and, in the case of xt3d, making them anisotropic with very high K (1000 x the base K in these runs) along the slope of the cell and very low K (0.001 x the base K) perpendicular to the slope.
+The case of a vertically offset mesh using the standard formulation, which cannot implement such anisotropy, was run but is not generally of interest.
+	\item crossconnected -- This is a plan-view, DISV model that emulates a cross-sectional model with "cross-connections" between model layers.
+This model is not in the repo, and results are noted here primarily to confirm the expected agreement with the disu\_approach model.
+	\item disu\_approach -- This implements cross-connections "for real" in a cross-sectional, DISU model.
+Specific discharge is recalculated in the notebook to account for overlap areas; the recent "patch" to MF6 is not being used (yet).
+	\item transect\_benchmark -- This is the vertically offset model of \cite{bardot2022} updated to allow cross-connections. Results from this model are not summarized here yet.
+\end{itemize}
+
+In the model of \cite{bardot2022} and its transect\_benchmark update, the channel is embedded in a surrounding "domain."  In each of the other three models, there is no surrounding domain -- the channel is the entire model.
+The distinction may be significant because the surrounding matrix provides vertical connections to the channel.
+When the surrounding domain is (effectively) impermeable, these vertical connections contribute "zero vertical flux" information to the specific discharge calculation, implying that the specific discharge is literally horizontal.
+In the absence of a surrounding domain, there is no vertical flux component information.
+Need to check what difference this might make.
+
+In the plan-view models, rows function as "layers." Unless stated otherwise, all results are for a 30-degree channel dip, and the analytical solution is a specific discharge of 1.
+inclined at 30 degrees.
+Total flow is the sum of the flows reported for the CHD cells along the left end of the channel.
+Numerical results are rounded arbitrarily.
+The terms "specific discharge" and "flux" are using interchangeably.
+
+\subsection{Single-layer channel}
+The channel is represented using a single layer of cells.
+
+\subsubsection{wormholes, 1x11 cells:}
+Results were the same whether the mesh was vertically offset (xt3d and standard, though not sure if standard results are meaningful) or dip-following (xt3d and standard), which is not surprising since there are no connections across the tops and bottoms.
+The errors in computed heads were miniscule in all cases.
+
+\paragraph{with xt3d:} Specific discharge is 0.8660 and "seemingly" horizontal.
+This makes sense if you consider that the only connections in the model are "horizontal" ones along the layer.
+A horizontal flux component of 0.8660 corresponds (via the cosine factor) to a unit flux from left to right along the dip of the channel, so in that sense it's correct.
+Furthermore, that's the only flux component that exists in this model, and therefore the only one xt3d can calculate; the vertical flux component is reported as zero by default.
+So rather than say the flux is horizontal, perhaps it's fairer to say MODFLOW gets the horizontal component right and simply can't tell us what the vertical component is.
+Total flow is 0.8660, which is correct.
+
+\paragraph{without xt3d (standard formulation)} Specific discharge is 1.1547 and, again, "seemingly" horizontal.
+This increase in magnitude relative to xt3d makes sense because the standard formulation ignores both the angle of the connection relative to the vertical cell interface and the reduction in area for flow (area perpendicular to the long axis of the channel) when the channel dips.
+Therefore, the  magnitude differs by a factor of $1/{(cosine factor)}^2$.
+Total flow is 1.1547, which reflects the error in the flux.
+
+\subsubsection{crossconnected, 1x11 cells:} Specific discharge results are essentially identical to those in the vertically offset wormholes model, which is not surprising given that there can be no cross-connections in a single-layer model.
+The errors in computed heads were miniscule with and without xt3d.
+
+\paragraph{with xt3d:} Total flow is 0.3660, which reflects the smaller area available for flow between two cells (the overlap area).
+
+\paragraph{without xt3d (standard formulation)} Total flow is 0.4880, which (given the incorrect flux) also reflects the smaller area available for flow between two cells (the overlap area).
+
+\subsubsection{disu\_approach, 1x11 cells:} Specific discharge results are essentially identical to those in the vertically offset wormholes model and crossconnected model.
+Total flow matched either the vertically offset wormholes result or the crossconnected result depending on how the "staggered" flag was set in the notebook, which determined whether the specific discharge remained based on average areas or was recalculated using overlap areas.
+(There are no cross-connections in a single-layer model, so the cell-cell flows were unaffected, so it was just a matter of whether the fluxes were postprocessed to adjust the areas.) The errors in computed heads were miniscule with and without xt3d.
+
+\subsection{Two-layer channel}
+The channel is represented using two layers of cells. This introduces the possibility of vertical flows between layers.
+
+\subsubsection{wormholes, 2x11 cells}
+
+\paragraph{dip-following mesh with xt3d:} In the connector cells (which are of primary interest in this case), specific discharge is 0.9998 mid-channel and 1.00003 on average, with a max error of +0.0003.
+Angle of specific discharge is 29.9913 deg mid-channel and 29.9879 on average, with a max error of +0.03 deg.
+Errors in computed heads (including all cells) are on the order of 0.002\%.
+Total flow is 1.7321 (0.8661 per cell).
+So, good agreement with the analytical solution.
+
+\paragraph{dip-following mesh without xt3d (standard formulation)} In the connector cells (which are of primary interest in this case), specific discharge is 1.3336 mid-channel and 1.3687 on average, with a max error of +0.69.
+Angle of specific discharge is 30.05 deg mid-channel and 32.24 on average, with a max error of +9.8 deg.
+Errors in computed heads (including all cells) are on the order of 2\%.
+Total flow is 2.3094 (1.1547 per cell).
+
+\paragraph{vertically offset mesh with xt3d:} In the flat-top cells (which are of primary interest in this case), specific discharge is 1.1117 mid-channel and 1.1211 on average, with a max error of +0.35.
+Angle of specific discharge is -0.53 deg mid-channel and 6.0095 on average, with a max error of -30.54 deg; the flow generally skews toward horizontal.
+Errors in computed heads (including all cells) are on the order of 4\%.
+Total flow is 2.1993 (1.0997 per cell).
+
+\subsubsection{crossconnected, 2x11 cells:}
+
+\paragraph{with xt3d:} Specific discharge matches the analytical solution very closely, with a max error of -5e-9.
+Angle of specific discharge also matches very closely, with a max error of -9e-8 deg.
+Errors in computed heads are minscule.
+Total flow is 1.2321 (0.6161 per cell), which reflects the reduced area for flow; the non-overlapping area of the lowest left-hand boundary cell is effectively "lost." Otherwise, a great result.
+
+\paragraph{without xt3d (standard formulation)}  Specific discharge is 1.0922 mid-channel and 1.1074 on average, with a max error of +0.30.
+Angle of specific discharge is 22.69 deg mid-channel and 23.28 on average, with a max error of -7.5 deg.
+Errors in computed heads are on the order of 0.5\%.
+Total flow is 1.3942 (0.6971 per cell).
+Overall, not a great result, but not terrible; the cross-connections alone seem to have helped substantially.
+
+\subsubsection{disu\_approach, 2x11 cells:}
+
+\paragraph{without cross-connections, with xt3d:} Results are comparable to those in the wormhole model with a vertically offset mesh and xt3d.
+Specific discharge is 1.1045 mid-channel and 1.1262 on average, with a max error of +0.38.
+Angle of specific discharge is 0.0731 deg mid-channel and 6.2516 on average, with a max error of -29.93 deg; the flow generally skews toward horizontal.
+Errors in computed heads are on the order of 3\%.
+Total flow is 2.2091 (1.1045 per cell).
+
+\paragraph{without cross-connections, without xt3d (standard formulation):} Specific discharge is 1.1547 mid-channel and 1.1755 on average, with a max error of +0.43.
+Angle of specific discharge is 0.0685 deg mid-channel and 5.9963 on average, with a max error of -29.93 deg; the flow generally skews toward horizontal.
+Errors in computed heads (including all cells) are on the order of 2\%.
+Total flow is 2.3094 (1.1547 per cell).
+Overall, comparable to the results with xt3d.
+
+\paragraph{with cross-connections, with xt3d:} Results are comparable to those in the crossconnected model with xt3d.
+Specific discharge matches the analytical solution very closely, with a max error of +1e-8.
+Angle of specific discharge also matches very closely, with a max error of -3e-7 deg.
+Errors in computed heads are minscule.
+Total flow is 1.2321 (0.6161 per cell), which reflects the reduced area for flow; the non-overlapping area of the lowest left-hand boundary cell is effectively "lost." Otherwise, a great result.
+
+\paragraph{with cross-connections, without xt3d (standard formulation):} Results are comparable to those in the crossconnected model without xt3d.
+Specific discharge is 1.0948 mid-channel and 1.1101 on average, with a max error of +0.30.
+Angle of specific discharge is 22.63 deg mid-channel and 23.22 on average, with a max error of -7.5 deg.
+Errors in computed heads are on the order of 0.5\%.
+Total flow is 1.3942 (0.6971 per cell).
+Overall, not a great result, but not terrible; the cross-connections alone seem to have helped substantially.
+
+\subsection{Five-layer channel}
+The channel is represented using five layers of cells.
+
+\subsubsection{wormholes, 5x11 cells:}
+
+\paragraph{dip-following mesh with xt3d:} Results are basically similar to those in the 2-layer case, except the total flow is greater in proportion to the greater number of boundary cells.
+
+\paragraph{dip-following mesh without xt3d (standard formulation)} Results are roughly comparable to but a bit worse overall than those in the 2-layer case.
+In the connector cells (which are of primary interest in this case), specific discharge is 1.3579 mid-channel and 1.4277 on average, with a max error of +1.14.
+Angle of specific discharge is 31.7497 deg mid-channel and 35.6924 on average, with a max error of +19.57 deg.
+Errors in computed heads (including all cells) are on the order of 8\%.
+Total flow is 5.7735 (1.1547 per cell).
+
+\paragraph{vertically offset mesh with xt3d:} Results are roughly comparable to those in the 2-layer case.
+In the flat-top cells (which are of primary interest in this case), specific discharge is 0.8906 mid-channel and 1.0013 on average, with a max error of +0.54.
+Angle of specific discharge is 2.8558 deg mid-channel and 10.9171 on average, with a max error of -31.5 deg; the flow generally skews toward horizontal.
+Errors in computed heads (including all cells) are on the order of 10\%.
+Total flow is 4.8426 (0.9685 per cell).
+
+\subsubsection{crossconnected, 5x11 cells:}
+
+\paragraph{with xt3d:} Results are basically similar to those in the 2-layer case.
+Total flow is 3.8301 (0.7660 per cell), which is better than the 2-layer result due to the "lost" (non-overlapping) area being a smaller proportion of the total area for flow.
+
+\paragraph{without xt3d (standard formulation)}  Results are basically comparable to those in the 2-layer case, except the errors in computed heads are substantially higher at around 1.5\%, and the total flow of 4.0944 gives a somewhat higher flow per cell at 0.8189.
+Again, not great, but not terrible.
+
+\subsubsection{disu\_approach, 5x11 cells:}
+
+\paragraph{without cross-connections, with xt3d:} Results are roughly comparable to but a bit worse overall than those in the 2-layer case.
+Specific discharge is 1.0547 mid-channel and 1.0926 on average, with a max error of +0.62.
+Angle of specific discharge is 2.9627 deg mid-channel and 10.3771 on average, with a max error of -28.1 deg; the flow generally skews toward horizontal.
+Errors in computed heads are on the order of 8\%.
+Total flow is 5.3106 (1.0621 per cell).
+
+\paragraph{without cross-connections, without xt3d (standard formulation):} Results are roughly comparable to but a bit worse overall than those in the 2-layer case.
+Specific discharge is 1.1560 mid-channel and 1.1827 on average, with a max error of +0.70.
+Angle of specific discharge is 2.7571 deg mid-channel and 9.5527 on average, with a max error of -28.3 deg; the flow generally skews toward horizontal.
+Errors in computed heads (including all cells) are on the order of 8\%.
+Total flow is 5.7735 (1.1547 per cell).
+Overall, comparable to the results with xt3d.
+
+\paragraph{with cross-connections, with xt3d:} Results are basically similar to those in the 2-layer case.
+Total flow is 3.8301 (0.7660 per cell), which is better than the 2-layer result due to the "lost" (non-overlapping) area being a smaller proportion of the total area for flow.
+
+\paragraph{with cross-connections, without xt3d (standard formulation):} Results are comparable to those in the crossconnected model without xt3d. Again, not great, but not terrible.
+
+\subsection{Crossflow}
+What happens if the flow is across the channel rather than along it?
+Added scenarios to the disu\_approach notebook to simulate this.
+Haven't analyzed the results in detail, but using cross-connections together with xt3d appears to help in this case, as well.
+
 \end{document}

### 2022-09-14 latex/article.tex
Substantial additions and changes to the manuscript, most notably in the Introduction and Description of Test Problems

@@
 \date{\today}
-% Hint: \title{what ever}, \author{who care} and \date{when ever} could stand 
-% before or after the \begin{document} command 
-% BUT the \maketitle command MUST come AFTER the \begin{document} command! 
+
 \begin{document}
@@
 
-Some intro stuff here about MODFLOW \citep{modflow6framework, modflow6gwf, modflow6gwt} and XT3D \citep{modflow6xt3d}... Horizontal tops and bottoms of cells... Connection length issue...
+Some intro stuff here about MODFLOW \citep{modflow6framework, modflow6gwf, modflow6gwt}...
 
-If cell-cell flows are  off, advective and dispersive transport will be off.
+XXXXX If cell-cell flows are inaccurate, advective and dispersive transport will be inaccurate XXXXX
 
-Horizontal cell tops and bottoms - benefits.  Vertical offsets.  Drawbacks - staircase boundaries, CVFD violations.  XT3D can address connection angle and length.
+MODFLOW requires model cells to have horizontal tops and bottoms.
+In versions of MODFLOW up to and including MODFLOW-2005 \citep{modflow2005}, which offered only structured grids composed of rows, columns and layers of cells, a horizontal top and bottom allowed each cell to be ``simulated as if it were rectangular so that flow may be approximated by the standard finite-difference equation" \citep{modflow84}.
+In MODFLOW-USG \citep{modflowusg} and MODFLOW 6 \citep{modflow6gwf}, which offer unstructured grids and use the control-volume finite-difference method to formulate the discrete balance equations, horizontal cell tops and bottoms still simplify the definition and calculation of water-table elevations, cell saturations, and conductances between cells.
+In all versions of MODFLOW, cells in the same model layer can be vertically offset from each other to allow model layers to ``deform'" with the hydrostratigraphy and thereby ``minimize the number of model layers required to simulate an aquifer system" \citep{modflow84} relative to a rectilinear grid with strictly horizontal layers.
 
-Having flat tops and bottoms but "distorting" the grid vertically (having vertical offsets) is discussed in \cite{modflow84} on p.
-140 and fig 27, though tops and bottoms are not drawn as horizontal.
-Benefits cited: "minimize the number of model layers required to simulate an aquifer system" (compared to "horizontal discretization" using a "rectilinear grid"); and "each distorted cell is simulated as if it were rectangular so that flow may be approximated by the standard finite-difference equation." Note that the latter benefit extends to the case of a partially saturated cell, in which the conductance is based on the saturated thickness measured between the horizontal water table and the horizontal cell bottom.
-Also simplifies particle tracking, though a relative z mapping is needed when passing particles between cells.
+Prior to MODFLOW 6, all versions of MODFLOW used a ``conductance-based" formulation to approximate the flow between two adjacent cells.
+The conductance-based formulation is a ``two-point" approximation in which the flow is proportional to the difference in the heads computed at the two cell centers and a conductance that depends on the hydraulic conductivities of the cells along the direction of the cell connection, the area of the cell interface, and a measure of the distance between the cell centers.
+On structured MODFLOW grids with no vertical offset between cells, i.e., strictly rectilinear grids, with hydraulic conductivity that is isotropic or has anisotropy aligned with the three mutually perpendicular grid directions, the conductance-based formulation is second-order accurate \citep{dehotin2010modeling, modflow6gwf}.
+This is because structured MODFLOW grids with no vertical offset satisfy the ``control-volume finite-difference (CVFD) requirement" that the straight-line connection between two cell centers must intersect the midpoint of the cell interface at a right angle \citep{narasimhan1976integrated}.
+Structured MODFLOW grids with vertical offset between cells, which we will call "vertically offset" grids, violate the CFVD requirement because nominally "horizontal" connections between cell centers in the same layer are not strictly horizontal and therefore not perpendicular to the vertical cell interfaces they intersect.
+(XXXXX Also, in the grids considered by \cite{narasimhan1976integrated}, there was no partial overlap between cells faces, but is that relevant, and do we want to get into that?
+XXXXX) The accuracy of the conductance-based flow formulation decreases with increasing vertical offset.
+For this reason, the MODFLOW 6 documentation \citep{modflow6gwf} warns that "[s]teeply dipping layers generally should not be represented" using vertical offsets.
+\cite{anderson2015applied} recommend a $10^{\circ}$ dip as a practical upper limit for using vertical offsets.
+A traditional alternative to using a vertically offset grid is to map the hydrologic properties of the aquifer system onto a strictly rectilinear grid, which typically requires more model layers to adequately discretize dipping hydrostratigraphic units \citep{modflow84}.
 
-Vertical distortion is also discussed in \cite{modflow2005} and pictured with flat tops and bottoms in fig 4-2 of that report.
+The ability to use unstructured, or ``irregular," grids in MODFLOW-USG and MODFLOW 6 introduced the potential for grids to violate the CVFD requirement and thereby render the conductance-based formulation less accurate even in the absence of vertical offsets.
+To improve accuracy on irregular grids, and to allow accurate simulation of arbitrarily oriented two- or three-dimensional anisotropy, the optional XT3D flow formulation \citep{modflow6xt3d} was introduced in MODFLOW 6.
+XT3D formulates the flow between two cells by interpolating head values from the two cells and their neighboring cells to construct an estimate of the full, two- or three-dimensional head-gradient vector on each side of the cell interface; applying the cell conductivity tensors to obtain an estimate of the groundwater flux vector on each side of the interface; and reconciling the two flux estimates to ensure continuity of flow across the interface.
+By performing flow calculations using Darcy's Law in its tensorial form, based on a ``multi-point" approximation of the head gradient, XT3D accounts for both arbitrarily oriented anisotropy and geometric irregularity of the grid.
 
-Thought: Vertical offset happens even on a completely rectilinear grid if the water table elevation varies in horizontally adjacent cells. Doesn't seem like the same kind of issue, but think on it.
+\cite{bardot2022} evaluated the effects of grid design and XT3D on the ability of MODFLOW 6 to accurately simulate flow through sedimentary structures by modeling a two-dimensional, permeable channel embedded in a nearly impermeable surrounding medium, or ``domain." The channel was angled counterclockwise from the $x$ axis of the model coordinates (by $30^{\circ}$ in most cases), and constant-head boundary conditions at the ends of the channel were set to induce uniform flow along the channel in the analytical solution.
+In their plan-view benchmark, the channel and surrounding domain were discretized using either a ``Regular Cartesian (RC)" grid that is rectilinear and satisfies the CVFD requirement or a ``Flexible Triangular (FT)" grid that is unstructured with triangular cells and does not satisfy the CVFD requirement.
+They found that the standard, conductance-based flow formulation gave accurate results on the RC grid but was subject to significant error on the FC grid, whereas the XT3D flow formulation gave accurate results on both grids, as expected.
+However, in their transect (cross-sectional) benchmark, which used a vertically offset grid to discretize the domain and a channel inclined by $30^{\circ}$, similarly poor results were observed with and without XT3D.
+The simulated groundwater flux was nearly horizontal ($0.01 - 0.03^{\circ}$ incline), and its magnitude was overestimated by 12 - 15\%.
+Given the ability of XT3D to compensate for violations of the CVFD requirement, it appeared likely that another aspect of the problem formulation or the inner workings of MODFLOW 6 was responsible for the unexpected results.
+(XXXXX Specifically mention that the NPF angle-specification "nozee" issue was looked into a found not to be the problem?
+Kind of an awkward detail to shoehorn in, but it would advertise this potential pitfall, which probably isn't well-known XXXXX) \cite{bardot2022} noted a suggestion by two authors of the present paper (Provost and Langevin) that inadequate grid connectivity was at the root of the problem.
 
-Thought: What happens with a head gradient perpendicular to the dip of the channel (across rather than along the channel)? Preliminary results suggest cross-connections and xt3d help.
+"Vertically staggered" grid connections were introduced in MODFLOW-USG \citep{modflowusg} and included for the DISU grid type in MODFLOW 6 \citep{modflow6gwf} primarily as a means to vertically subdiscretize portions of a grid that is conceptualized as being layered.
+In a vertically staggered grid, i.e., a grid that includes vertically staggered connections, some cells have nominally ``horizontal" connections with more than one cell across a given cell face.
+Vertically staggered MODFLOW 6 DISU grids allow more flexibility in assigning grid connectivity than do the more commonly used DIS and DISV grid types \citep{modflow6gwf}.
+To our knowledge, the implications of limited grid connectivity for the accuracy of simulated flows in models with steeply dipping layers, and the potential for using vertically staggered grids to overcome this limitation, have not been previously appreciated or investigated.
 
-Thought: Difference between truly single-layer model of channel vs embedded single-layer channel -- no vertical connections vs vertical connections that are no-flux.
+In the remainder of this paper we first present a theoretical justification for the hypothesis that inadequate grid connectivity is primarily responsible for the inaccurate simulation of flows in the steeply dipping channel benchmark on a vertically offset grid.
+Based on that theory, we propose a method for improving the accuracy of the flow solution using XT3D together with a vertically staggered grid, and we evaluate the effectiveness of that approach in a set of benchmark problems similar to those of \cite{bardot2022}.
+Finally, we discuss the potential implications of our findings for practical groundwater models that include high permeability contrasts and steeply dipping layers.
+
+XXXXX Random thoughts:
+\begin{itemize}
+	\item Vertical offset happens even on a completely rectilinear grid if the water table elevation varies in horizontally adjacent cells. Doesn't seem like the same kind of issue, but think on it.
+	\item What happens with a head gradient perpendicular to the dip of the channel (across rather than along the channel)? Preliminary results suggest cross-connections and xt3d help.
+	\item Difference between truly single-layer model of channel vs embedded single-layer channel -- no vertical connections vs vertical connections that are no-flux.
 Effect on specific discharge calculations?
-
-Reference and summarize VO results in \cite{bardot2022}.
-This is an extreme case (virtually impermeable domain surrounding the channel) that nicely accentuates the problem, but there should be some error in any application involving VO grids.
-Obviously expect it to be a bigger issue for greater offsets.
-XT3D didn't help, which was initially surprising since it takes into account connection angles and lengths.
-Not specifying all the angles in the NPF input would cause XT3D to default to "nozee" and therefore assume "horizontal" connections are strictly horizontal, but that was not the case in \cite{bardot2022}.
-
-We explain in this note that it's not just a connection angle and length issue, though that has an effect.
-The main problem with a VO grid is insufficient hydraulic communication between grid cells in adjacent layers.
-(It's not restricted to layered models, but it's convenient to describe it in terms of layers.)  We propose and test a solution to the problem that involves enhanced connectivity between cells.
-
-We present two test models that demonstrate 1) results on an unmodified VO DIS grid and 2) results on corresponding DISU grid with enhanced connectivity.
-Results suggest that enhanced connectivity resolves the issue.
-
-%\begin{figure}[ht]
-%	\centering
-%	\includegraphics[scale=0.6]{figures/fig1.pdf}
-%	\caption{This is the first figure.}
-%	\label{fig:firstfigure}
-%\end{figure}
+\end{itemize}
 
@@
 Although such a configuration of cells can be constructed in plan view using the DISV grid type of MODFLOW 6, it cannot be constructed in cross-section because MODFLOW requires cells to have horizontal tops and bottoms.
-The sloping tops and bottoms of these hypothetical cells allow the cell geometry to conform to the dip of the channel, and we will therefore refer to this grid configuration as "dip-following." The grid connectivity is analogous to the connectivity found in a MODFLOW 6 DIS (structured, or regular) grid: a cell is hydraulically connected to each cell with which it shares a face.
+The sloping tops and bottoms of these hypothetical cells allow the cell geometry to conform to the dip of the channel, and we will therefore refer to this grid configuration as ``dip-following." The grid connectivity is analogous to the connectivity found in a MODFLOW 6 DIS (structured, or regular) grid: a cell is hydraulically connected to each cell with which it shares a face.
 For example, cell 1A is connected to cells 1B and 2A, but not to cell 2B.
@@
 
-Figure \ref{fig:schem-conn-area-flux}b shows the four cells as they would be configured for a DIS or DISV grid in a two-layer, cross-sectional MODFLOW 6 model of the channel.
-The cells have horizontal tops and bottoms and can therefore follow the dip of the channel only on average, in what \cite{bardot2022} refer to as "staircase" fashion, and we will refer to this grid configuration as "vertically offset." The pattern of grid connectivity is analogous to that in figure \ref{fig:schem-conn-area-flux}a, but it is not based strictly on which cell face overlap.
+Figure \ref{fig:schem-conn-area-flux}b shows the four cells as they would be configured for a vertically offset DIS or DISV grid in a two-layer, cross-sectional MODFLOW 6 model of the channel.
+The cells have horizontal tops and bottoms and can therefore follow the dip of the channel only on average, in what \cite{bardot2022} refer to as ``staircase" fashion.
+The pattern of grid connectivity is analogous to that in figure \ref{fig:schem-conn-area-flux}a, but it is not based strictly on which cell face overlap.
 Rather, a cell is connected to each cell in the same layer with which it has an overlapping face and with each cell in an adjacent layer with which it shares a horizontal face.
 Significantly, cells that have overlapping vertical faces but are in different layers (e.g., cells 1A and 2B) are not connected.
-In the standard and Newton-Raphson "conductance-based" formulations for flow, the flow between two cells in the same layer of a DIS or DISV grid (XXXXX or DISU that's not vertically staggered, but do we want to mention that here?
+In the conductance-based formulation for flow, the flow between two cells in the same layer of a DIS or DISV grid (XXXXX or DISU that's not vertically staggered, but do we want to mention that here?
 XXXXX) is proportional to the difference in heads computed at the two cell centers and a conductance based on an effective hydraulic conductivity for the connection between the cells.
@@
 
-Conductance-based formulations for flow can be expected to give accurate results if the model grid satisfies a set of geometric "control-volume finite-difference (CVFD) conditions" (XXXXX; XXXXX).
-One of those conditions is that the straight-line connection between cell centers must be perpendicular to the cell interface.
+As mentioned earlier, for the conductance-based flow formulation to provide maximum accuracy, the model grid must satisfy the CVFD requirement that the straight-line connection between two cell centers must intersect the midpoint of the cell interface at a right angle.
 In a vertically offset grid, connections between cells across vertical cell interfaces are generally not horizontal and, therefore, not perpendicular to the interface.
-(XXXXX Also, in the grids considered by XXXXX, there was no partial overlap between cells faces, but do we want to get into that here?
-XXXXX) For example, in \ref{fig:schem-conn-area-flux}b, the nominally "horizontal" connection between cells 1A and 1B is not perpendicular to the cell interface, which compormises the accuracy of conductance-based formulations.
-The XT3D capability is a gradient reconstruction (interpolation) scheme introduced in MODFLOW 6 that allows simulation of fully three-dimensional anisotropy and improves the accuracy of flows between cells on grids that do not satisfy the CVFD conditions.
+For example, in \ref{fig:schem-conn-area-flux}b, the nominally ``horizontal" connection between cells 1A and 1B is not perpendicular to the cell interface, which compromises the accuracy of conductance-based formulations.
+(XXXXX Also, in the grids considered by \cite{narasimhan1976integrated}, there was no partial overlap between cells faces, but do we want to get into that here?
+XXXXX)
 
-In their dipping-channel benchmark problem, \cite{bardot2022} found that, as expected, the standard, conductance-based formulation gave poor results for a steeply ($30^{\circ}$) dipping channel.
-The simulated flow was not along the dip of the channel, but nearly horizontal, and the magnitude was too large.
+XXXXX Mostly covered in intro, so condense here XXXXX In their dipping-channel benchmark problem, \cite{bardot2022} found that, as expected, the standard, conductance-based formulation gave poor results for a steeply dipping channel discretized using a vertically offset grid.
+The simulated groundwater flux in the middle of the channel was not along the $30^{\circ}$ incline of the channel, but nearly horizontal ($0.03^{\circ}$ incline), and its magnitude was overestimated by 15\%.
 Given the ability of XT3D to compensate for grid connections that are not perpendicular to cell interfaces, rerunning the simulation with XT3D activated could have been expected to improve the flow solution substantially.
-However, \cite{bardot2022} obtained similarly poor results with and without XT3D.
-It appeared that something other than a simple violation of the CVFD perpendicularity condition was responsible, and \cite{bardot2022} noted a hypothesis suggested by two authors of the present paper (Provost and Langevin).
+However, \cite{bardot2022} observed similarly poor results with XT3D: the flux was still nearly horizontal ($0.01^{\circ}$ incline), and its magnitude was overestimated by 12\%.
+Something other than a simple violation of the CVFD perpendicularity condition was evidently responsible, and \cite{bardot2022} noted a suggestion by two authors of the present paper (Provost and Langevin) that inadequate grid connectivity was at the root of the problem.
 
-The hypothesis can be explained with the help of figures \ref{fig:schem-conn-area-flux}b and \ref{fig:schem-conn-area-flux}e.
+The influence of grid connectivity on the results observed by \cite{bardot2022} can be explained with the help of figures \ref{fig:schem-conn-area-flux}b and \ref{fig:schem-conn-area-flux}e.
 Figure \ref{fig:schem-conn-area-flux}e shows the specific discharge vector superimposed on each cell interface of the vertically offset grid, together with a vector representing the component of the flux normal to the interface.
@@
 This explains why \cite{bardot2022} observed nearly horizontal flow in the middle portion of the channel, away from the lateral boundaries, where nearly uniform flow was established.
-Thus, it appears the nearly horizontal flows are not primarily due to what \cite{bardot2022} call the "cosine error" associated with the orientation and length of the sloping grid connections, but to the inability of the vertically offset grid to route flow appropriately between cells.
+Thus, it appears the nearly horizontal flows are not primarily due to what \cite{bardot2022} call the ``cosine error" associated with the orientation and length of the sloping grid connections, but to the inability of the vertically offset grid to route flow appropriately between cells.
 
@@
 The area of the interface is the area of overlap between the cell faces.
-This grid configuration, which was introduced MODFLOW-USG (XXXXXXXXX) and is available for the DISU grid type in MODFLOW 6, is called "vertically staggered" and allows a cell to have nominally "horizontal" connections with more than one cell across a given cell face.
-In this case, vertical staggering is used to introduce "cross-connections" between cells in different layers.
+This grid configuration, which was introduced MODFLOW-USG \citep{modflowusg} and is available for the DISU grid type in MODFLOW 6, is called ``vertically staggered" and allows a cell to have nominally ``horizontal" connections with more than one cell across a given cell face.
+In this case, vertical staggering is used to introduce ``cross-connections" between cells in different layers.
 The ramifications of cross-connections for flow between cells are illustrated in figure \ref{fig:schem-conn-area-flux}f.
@@
 
-XXXXX Note: This idea is not limited to layered grids, but it's easiest to discuss in terms of layers. XXXXX
-
-\section{Approach}
-
-Summarize the overall approach here.
-
-Will use a DIS grid (XXXXX converted to an equivalent DISU grid without cross=connections?
-XXXXX) cross-sectional model to show results you get on a vertically offset grid, with and without XT3D.
-Basically show that we can reproduce what \cite{bardot2022} found.
-
-Will convert the DIS grid to a DISU grid with cross-connections and show improved results, with and without XT3D.
-
 \section{Description of Test Problems}
 
-Describe the test problem setups here.
+The test problems are patterned after the cross-sectional (transect) benchmarks of \cite{bardot2022}.
+They attempt simulate uniform flow in a permeable channel embedded in a less permeable surrounding ``domain." The channel is of uniform width and hydraulic conductivity and is inclined relative to the horizontal.
 
-\subsection{Test problem 1 (DIS cross-sectional)}
+To simulate flow along a channel inclined at angle $\theta$, heads are specified at the centers of cells along the perimeter of the model using an analytical solution that corresponds to a unit head gradient inclined at angle $\theta$ \citep{bardot2022}:
 
-Test problem 1...
+\begin{equation}
+\label{eqn:head_analyt_along}
+h = - x \cos \theta - z \sin \theta.
+\end{equation}
 
-XXXXX Idea: See how heterogeneity affects the error? XXXXX
+\noindent where $h$ is head in meters and $x$ and $z$ are the horizontal and vertical model coordinates, respectively, in meters.
+Flow across the channel (XXXXX seems like this would be good to include XXXXX) is simulated using the modified head function
+\begin{equation}
+\label{eqn:head_analyt_across}
+h = x \sin \theta - z \cos \theta,
+\end{equation}
+which represents a unit head gradient perpendicular to the long axis of the channel.
+In both cases, the hydraulic conductivity of the channel is set to 1 m/d (meter per day) so that the analytical flow solution is a groundwater flux of 1 m/d either along or across the channel.
+The conductivity of the domain surrounding the channel varies according to the scenario being investigated.
+In the homogeneous limit, the conductivity of the domain is set equal to that of the channel.
+In the highly heterogeneous limit, the conductivity of the domain is set much lower than that of the channel to effectively isolate the channel hydraulically.
 
-\subsection{Test problem 2 (DISU cross-sectional with cross-connections)}
+Simulations are performed using variations on two test problems that differ in their method of discretization.
+The first test problem uses the traditional approach of representing the inclined channel using a vertically offset grid of the MODFLOW 6 DIS (structured) type.
+(XXXXX To be decided: present results for the DIS grid, or for the equivalent of the DIS grid after conversion to DISU?
+XXXXX) This problem confirms the findings of \cite{bardot2022} and illustrates the limitations of the vertically offset grid approach.
+The second test problem is similar to the first but uses a vertically staggered grid of MODFLOW 6 DISU (unstructured) type.
+The second problem is designed to evaluate the effectiveness of cross-connections between layers (illustrated in figure \ref{fig:schem-conn-area-flux}c) in improving the accuracy of the flow solution, both with and without XT3D.
+Variations on both test problems are described in the ``Results and Discussion" section.
 
-Test problem 2...
+XXXXX Describe conversion XXXXX
+
+\section{Results and Discussion}
+
+XXXXX See the temporary ``preliminary results" section at the end of the manuscript XXXXX
+
+\subsection{Test Problem 1: Vertically Offset Discretization}
+
+XXXXX Idea: See how heterogeneity affects the error -- series of runs with varying permeability contrast between channel and surrounding domain (from 1:1 to very large), producing a plot along the lines of figs.
+3d and 4d in \cite{bardot2022} XXXXX
+
+\subsection{Test Problem 2: Vertically Staggered Discretization}
 
@@
 
-\section{Results and Discussion}
-
-XXXXX See the temporary "preliminary results" section at the end of the manuscript XXXXX
+XXXXX The concepts in this paper have been easiest to discuss in terms of layers, but they're not limited to layered grids. XXXXX
 
@@
 \begin{itemize}
-	\item wormholes -- This is a plan-view, DISV model that emulates a cross-sectional model and includes "connector" cells that can have sloped tops and bottoms.
-Cell dimensions and hydraulic conductivities can be adjusted to approach a "dip-following" mesh (in which cell tops and bottoms follow the dip of the channel) at one extreme and a vertically offset mesh (the kind MODFLOW actually uses in cross-section) at another extreme.
+	\item wormholes -- This is a plan-view, DISV model that emulates a cross-sectional model and includes ``connector" cells that can have sloped tops and bottoms.
+Cell dimensions and hydraulic conductivities can be adjusted to approach a ``dip-following" mesh (in which cell tops and bottoms follow the dip of the channel) at one extreme and a vertically offset mesh (the kind MODFLOW actually uses in cross-section) at another extreme.
 A vertically offset mesh is approached by making the width of the connector cells very small in the x direction and, in the case of xt3d, making them anisotropic with very high K (1000 x the base K in these runs) along the slope of the cell and very low K (0.001 x the base K) perpendicular to the slope.
 The case of a vertically offset mesh using the standard formulation, which cannot implement such anisotropy, was run but is not generally of interest.
-	\item crossconnected -- This is a plan-view, DISV model that emulates a cross-sectional model with "cross-connections" between model layers.
+This model is in the repo, but results from this model probably won't be included in the paper.
+	\item crossconnected -- This is a plan-view, DISV model that emulates a cross-sectional model with ``cross-connections" between model layers.
 This model is not in the repo, and results are noted here primarily to confirm the expected agreement with the disu\_approach model.
-	\item disu\_approach -- This implements cross-connections "for real" in a cross-sectional, DISU model.
-Specific discharge is recalculated in the notebook to account for overlap areas; the recent "patch" to MF6 is not being used (yet).
+	\item disu\_approach -- This implements cross-connections ``for real" in a cross-sectional, DISU model.
+Specific discharge is recalculated in the notebook to account for overlap areas; the recent ``patch" to MF6 is not being used (yet).
 	\item transect\_benchmark -- This is the vertically offset model of \cite{bardot2022} updated to allow cross-connections. Results from this model are not summarized here yet.
@@
 
-In the model of \cite{bardot2022} and its transect\_benchmark update, the channel is embedded in a surrounding "domain."  In each of the other three models, there is no surrounding domain -- the channel is the entire model.
+In the model of \cite{bardot2022} and its transect\_benchmark update, the channel is embedded in a surrounding ``domain."  In each of the other three models, there is no surrounding domain -- the channel is the entire model.
 The distinction may be significant because the surrounding matrix provides vertical connections to the channel.
-When the surrounding domain is (effectively) impermeable, these vertical connections contribute "zero vertical flux" information to the specific discharge calculation, implying that the specific discharge is literally horizontal.
+When the surrounding domain is (effectively) impermeable, these vertical connections contribute ``zero vertical flux" information to the specific discharge calculation, implying that the specific discharge is literally horizontal.
 In the absence of a surrounding domain, there is no vertical flux component information.
@@
 
-In the plan-view models, rows function as "layers." Unless stated otherwise, all results are for a 30-degree channel dip, and the analytical solution is a specific discharge of 1.
+In the plan-view models, rows function as ``layers." Unless stated otherwise, all results are for a 30-degree channel dip, and the analytical solution is a specific discharge of 1.
 inclined at 30 degrees.
@@
 Numerical results are rounded arbitrarily.
-The terms "specific discharge" and "flux" are using interchangeably.
+The terms ``specific discharge" and ``flux" are using interchangeably.
 
@@
 
-\paragraph{with xt3d:} Specific discharge is 0.8660 and "seemingly" horizontal.
-This makes sense if you consider that the only connections in the model are "horizontal" ones along the layer.
+\paragraph{with xt3d:} Specific discharge is 0.8660 and ``seemingly" horizontal.
+This makes sense if you consider that the only connections in the model are ``horizontal" ones along the layer.
 A horizontal flux component of 0.8660 corresponds (via the cosine factor) to a unit flux from left to right along the dip of the channel, so in that sense it's correct.
@@
 
-\paragraph{without xt3d (standard formulation)} Specific discharge is 1.1547 and, again, "seemingly" horizontal.
+\paragraph{without xt3d (standard formulation)} Specific discharge is 1.1547 and, again, ``seemingly" horizontal.
 This increase in magnitude relative to xt3d makes sense because the standard formulation ignores both the angle of the connection relative to the vertical cell interface and the reduction in area for flow (area perpendicular to the long axis of the channel) when the channel dips.
@@
 \subsubsection{disu\_approach, 1x11 cells:} Specific discharge results are essentially identical to those in the vertically offset wormholes model and crossconnected model.
-Total flow matched either the vertically offset wormholes result or the crossconnected result depending on how the "staggered" flag was set in the notebook, which determined whether the specific discharge remained based on average areas or was recalculated using overlap areas.
+Total flow matched either the vertically offset wormholes result or the crossconnected result depending on how the ``staggered" flag was set in the notebook, which determined whether the specific discharge remained based on average areas or was recalculated using overlap areas.
 (There are no cross-connections in a single-layer model, so the cell-cell flows were unaffected, so it was just a matter of whether the fluxes were postprocessed to adjust the areas.) The errors in computed heads were miniscule with and without xt3d.
@@
 Errors in computed heads are minscule.
-Total flow is 1.2321 (0.6161 per cell), which reflects the reduced area for flow; the non-overlapping area of the lowest left-hand boundary cell is effectively "lost." Otherwise, a great result.
+Total flow is 1.2321 (0.6161 per cell), which reflects the reduced area for flow; the non-overlapping area of the lowest left-hand boundary cell is effectively ``lost." Otherwise, a great result.
 
@@
 \paragraph{with xt3d:} Results are basically similar to those in the 2-layer case.
-Total flow is 3.8301 (0.7660 per cell), which is better than the 2-layer result due to the "lost" (non-overlapping) area being a smaller proportion of the total area for flow.
+Total flow is 3.8301 (0.7660 per cell), which is better than the 2-layer result due to the ``lost" (non-overlapping) area being a smaller proportion of the total area for flow.
 
@@
 \paragraph{with cross-connections, with xt3d:} Results are basically similar to those in the 2-layer case.
-Total flow is 3.8301 (0.7660 per cell), which is better than the 2-layer result due to the "lost" (non-overlapping) area being a smaller proportion of the total area for flow.
+Total flow is 3.8301 (0.7660 per cell), which is better than the 2-layer result due to the ``lost" (non-overlapping) area being a smaller proportion of the total area for flow.
 

### 2022-09-14 latex/article.tex
Added a thought about the conductivity contrast - a possible case to try
Inserted some figures with preliminary results into the Results and Discussion

@@
 	\item What happens with a head gradient perpendicular to the dip of the channel (across rather than along the channel)? Preliminary results suggest cross-connections and xt3d help.
+	\item Could also try making the channel less permeable than the domain; semi-confining unit.
 	\item Difference between truly single-layer model of channel vs embedded single-layer channel -- no vertical connections vs vertical connections that are no-flux.
@@
 The cells have horizontal tops and bottoms and can therefore follow the dip of the channel only on average, in what \cite{bardot2022} refer to as ``staircase" fashion.
-The pattern of grid connectivity is analogous to that in figure \ref{fig:schem-conn-area-flux}a, but it is not based strictly on which cell face overlap.
+The pattern of grid connectivity is analogous to that in figure \ref{fig:schem-conn-area-flux}a, but it is not based strictly on which cell faces overlap.
 Rather, a cell is connected to each cell in the same layer with which it has an overlapping face and with each cell in an adjacent layer with which it shares a horizontal face.
@@
 In the highly heterogeneous limit, the conductivity of the domain is set much lower than that of the channel to effectively isolate the channel hydraulically.
+XXXXX Could also try making the channel less permeable than the domain; semi-confining unit XXXXX
 
@@
 
-XXXXX Describe conversion XXXXX
+Jupyter notebooks for both test problems are available in the Supporting Information that accompanies this paper.
+The notebooks leverage the capabilities of FloPy \citep{bakker2016scripting} to assist in setting up input for and processing results from the MODFLOW 6 simulations.
+The second test problem is formulated by starting with a DIS grid, converting it to an equivalent DISU grid, and adding cross-connections.
+A Python script for converting and modifying the grid is included.
 
@@
 
+\begin{figure}[H]
+\centering
+\begin{subfigure}{0.4\textwidth}
+	\includegraphics[width=\textwidth]{../figures/disu-s-nocc-head.png}
+	\caption{standard conductance-based formulation}
+	\label{fig:disu-s-nocc-head}
+\end{subfigure}
+\hfill
+\begin{subfigure}{0.4\textwidth}
+	\includegraphics[width=\textwidth]{../figures/disu-x-nocc-head.png}
+	\caption{XT3D}
+	\label{fig:disu-x-nocc-head}
+\end{subfigure}
+\caption{Specific discharge and head for a vertically offset grid.
+The news is all bad; XT3D doesn't really help.
+(disu\_approach model, 5x11 grid without cross-connections, $30^{\circ}$ incline, set up for flow along the channel)}
+\label{fig:figures}
+\end{figure}
+
 XXXXX Idea: See how heterogeneity affects the error -- series of runs with varying permeability contrast between channel and surrounding domain (from 1:1 to very large), producing a plot along the lines of figs.
-3d and 4d in \cite{bardot2022} XXXXX
+3d and 4d in \cite{bardot2022}.
+Will need to use the transect\_benchmark model for this, as we probably should for all of the final simulations.
+XXXXX
+
+\begin{figure}[H]
+\centering
+\begin{subfigure}{0.4\textwidth}
+	\includegraphics[width=\textwidth]{../figures/disu-s-nocc-cf-head.png}
+	\caption{standard conductance-based formulation}
+	\label{fig:disu-s-nocc-cf-head.}
+\end{subfigure}
+\hfill
+\begin{subfigure}{0.4\textwidth}
+	\includegraphics[width=\textwidth]{../figures/disu-x-nocc-cf-head.png}
+	\caption{XT3D}
+	\label{fig:disu-x-nocc-cf-head}
+\end{subfigure}
+\caption{Specific discharge and head for a vertically offset grid and boundary heads set up for crossflow.
+The standard formulation is a fail, but XT3D nails it, even without cross-connections.
+(disu\_approach model, 5x11 grid without cross-connections, $30^{\circ}$ incline, set up for flow across the channel)}
+\label{fig:figures}
+\end{figure}
 
@@
 \begin{figure}[H]
-	\begin{center}
-	\includegraphics[scale=0.5]{../figures/disu-x-cc-head.png}
-	\caption{Cross-connections are good. XT3D is good.}
-	\label{fig:mymodel-head}
-	\end{center}
+\centering
+\begin{subfigure}{0.4\textwidth}
+	\includegraphics[width=\textwidth]{../figures/disu-s-cc-head.png}
+	\caption{standard conductance-based formulation}
+	\label{fig:disu-s-cc-head}
+\end{subfigure}
+\hfill
+\begin{subfigure}{0.4\textwidth}
+	\includegraphics[width=\textwidth]{../figures/disu-x-cc-head.png}
+	\caption{XT3D}
+	\label{fig:disu-x-cc-head}
+\end{subfigure}
+\caption{Specific discharge and head for a vertically staggered grid with cross-connections.
+The news is much better; the standard formulation isn't horrible, and XT3D nails it.
+(disu\_approach model, 5x11 grid with cross-connections, $30^{\circ}$ incline, set up for flow along the channel)}
+\label{fig:figures}
+\end{figure}
+
+\begin{figure}[H]
+\centering
+\begin{subfigure}{0.4\textwidth}
+	\includegraphics[width=\textwidth]{../figures/disu-s-cc-cf-head.png}
+	\caption{standard conductance-based formulation}
+	\label{fig:disu-s-cc-cf-head}
+\end{subfigure}
+\hfill
+\begin{subfigure}{0.4\textwidth}
+	\includegraphics[width=\textwidth]{../figures/disu-x-cc-cf-head.png}
+	\caption{XT3D}
+	\label{fig:disu-x-cc-cf-head}
+\end{subfigure}
+\caption{Specific discharge and head for a vertically staggered grid with cross-connections and boundary heads set up for crossflow.
+The standard formulation does pretty well, and XT3D nails it again.
+(disu\_approach model, 5x11 grid with cross-connections, $30^{\circ}$ incline, set up for flow across the channel)}
+\label{fig:figures}
 \end{figure}

### 2022-09-20 latex/article.tex
* Substantial additions/changes to the disu_approach notebook, mostly to add the option to include the "domain" surrounding the channel, plus other refactoring and tweaks. * Minor updates to the manuscript.  Note that the naming convention for the figures from the scenario runs has changed.

@@
 
-To simulate flow along a channel inclined at angle $\theta$, heads are specified at the centers of cells along the perimeter of the model using an analytical solution that corresponds to a unit head gradient inclined at angle $\theta$ \citep{bardot2022}:
+To simulate flow along a channel inclined at angle $\theta$, heads are specified at the centers of cells along the perimeter of the model (XXXXX this assumes we're using the transect\_benchmark version of the model, which includes the "domain," for all of our final runs XXXXX) using an analytical solution that corresponds to a unit head gradient inclined at angle $\theta$ \citep{bardot2022}:
 
@@
 \begin{subfigure}{0.4\textwidth}
-	\includegraphics[width=\textwidth]{../figures/disu-s-nocc-head.png}
+	\includegraphics[width=\textwidth]{../figures/disu-af-vo-s-head.png}
 	\caption{standard conductance-based formulation}
@@
 \begin{subfigure}{0.4\textwidth}
-	\includegraphics[width=\textwidth]{../figures/disu-x-nocc-head.png}
+	\includegraphics[width=\textwidth]{../figures/disu-af-vo-x-head.png}
 	\caption{XT3D}
@@
 The news is all bad; XT3D doesn't really help.
-(disu\_approach model, 5x11 grid without cross-connections, $30^{\circ}$ incline, set up for flow along the channel)}
+(preliminary result from disu\_approach model, 5x11 grid without cross-connections, $30^{\circ}$ incline, set up for flow along the channel)}
 \label{fig:figures}
@@
 \begin{subfigure}{0.4\textwidth}
-	\includegraphics[width=\textwidth]{../figures/disu-s-nocc-cf-head.png}
+	\includegraphics[width=\textwidth]{../figures/disu-cf-vo-s-head.png}
 	\caption{standard conductance-based formulation}
@@
 \begin{subfigure}{0.4\textwidth}
-	\includegraphics[width=\textwidth]{../figures/disu-x-nocc-cf-head.png}
+	\includegraphics[width=\textwidth]{../figures/disu-cf-vo-x-head.png}
 	\caption{XT3D}
@@
 The standard formulation is a fail, but XT3D nails it, even without cross-connections.
-(disu\_approach model, 5x11 grid without cross-connections, $30^{\circ}$ incline, set up for flow across the channel)}
+(preliminary result from disu\_approach model, 5x11 grid without cross-connections, $30^{\circ}$ incline, set up for flow across the channel)}
 \label{fig:figures}
@@
 \begin{subfigure}{0.4\textwidth}
-	\includegraphics[width=\textwidth]{../figures/disu-s-cc-head.png}
+	\includegraphics[width=\textwidth]{../figures/disu-af-vs-s-head.png}
 	\caption{standard conductance-based formulation}
@@
 \begin{subfigure}{0.4\textwidth}
-	\includegraphics[width=\textwidth]{../figures/disu-x-cc-head.png}
+	\includegraphics[width=\textwidth]{../figures/disu-af-vs-x-head.png}
 	\caption{XT3D}
@@
 The news is much better; the standard formulation isn't horrible, and XT3D nails it.
-(disu\_approach model, 5x11 grid with cross-connections, $30^{\circ}$ incline, set up for flow along the channel)}
+(preliminary result from disu\_approach model, 5x11 grid with cross-connections, $30^{\circ}$ incline, set up for flow along the channel)}
 \label{fig:figures}
@@
 \begin{subfigure}{0.4\textwidth}
-	\includegraphics[width=\textwidth]{../figures/disu-s-cc-cf-head.png}
+	\includegraphics[width=\textwidth]{../figures/disu-cf-vs-s-head.png}
 	\caption{standard conductance-based formulation}
@@
 \begin{subfigure}{0.4\textwidth}
-	\includegraphics[width=\textwidth]{../figures/disu-x-cc-cf-head.png}
+	\includegraphics[width=\textwidth]{../figures/disu-cf-vs-x-head.png}
 	\caption{XT3D}
@@
 The standard formulation does pretty well, and XT3D nails it again.
-(disu\_approach model, 5x11 grid with cross-connections, $30^{\circ}$ incline, set up for flow across the channel)}
+(preliminary result from disu\_approach model, 5x11 grid with cross-connections, $30^{\circ}$ incline, set up for flow across the channel)}
 \label{fig:figures}
@@
 
+\section{Conclusions}
+
 XXXXX The concepts in this paper have been easiest to discuss in terms of layers, but they're not limited to layered grids. XXXXX
-
-\section{Conclusions}
 
@@
 
+\subsection{Domain}
+Results with the surrounding "domain" (from an updated disu\_approach notebook) are similar but not identical, apparently due to the "staircase" effect.
+Might be best to presnt most of our results without the domain, at least when illustrating the main problem and its solution (to keep the "theory" clear), then discuss what happens when the domain is present (which is more like in practice).
+
 \end{document}

### 2022-09-27 latex/article.tex
* Added option to the disu_approach notebook to have the "domain" present. Refactored and expanded the setup of scenarios. * Inserted figs and captions with preliminary results of simulations with the domain present.

@@
 
+XXXXX Results without the domain are presented first to "cleanly" illustrate the main problem and its solution.
+Results with the domain are then presented to give more of an indication of how things might work in practice.
+Unless stated otherwise, the domain is 6 orders of magnitude less conductive than the channel, as in \cite{bardot2022}.
+XXXXX
+
 \subsection{Test Problem 1: Vertically Offset Discretization}
@@
 \caption{Specific discharge and head for a vertically offset grid.
-The news is all bad; XT3D doesn't really help.
-(preliminary result from disu\_approach model, 5x11 grid without cross-connections, $30^{\circ}$ incline, set up for flow along the channel)}
+As Kerry and Jim found previously, the news is all bad; the flow tends to go horizontal, and XT3D doesn't really help.
+(preliminary result from disu\_approach model, 5x11 grid without cross-connections, $30^{\circ}$ incline, set up for flow along the channel; scenarios disu-af-vo-s and disu-af-vo-x)}
 \label{fig:figures}
 \end{figure}
-
-XXXXX Idea: See how heterogeneity affects the error -- series of runs with varying permeability contrast between channel and surrounding domain (from 1:1 to very large), producing a plot along the lines of figs.
-3d and 4d in \cite{bardot2022}.
-Will need to use the transect\_benchmark model for this, as we probably should for all of the final simulations.
-XXXXX
 
@@
 The standard formulation is a fail, but XT3D nails it, even without cross-connections.
-(preliminary result from disu\_approach model, 5x11 grid without cross-connections, $30^{\circ}$ incline, set up for flow across the channel)}
+(preliminary result from disu\_approach model, 5x11 grid without cross-connections, $30^{\circ}$ incline, set up for flow across the channel; scenarios disu-cf-vo-s and disu-cf-vo-x)}
 \label{fig:figures}
@@
 \caption{Specific discharge and head for a vertically staggered grid with cross-connections.
-The news is much better; the standard formulation isn't horrible, and XT3D nails it.
-(preliminary result from disu\_approach model, 5x11 grid with cross-connections, $30^{\circ}$ incline, set up for flow along the channel)}
+The news is much better; the standard formulation is much improved, and XT3D nails it.
+(preliminary result from disu\_approach model, 5x11 grid with cross-connections, $30^{\circ}$ incline, set up for flow along the channel; scenarios disu-af-vs-s and disu-af-vs-x)}
 \label{fig:figures}
@@
 \caption{Specific discharge and head for a vertically staggered grid with cross-connections and boundary heads set up for crossflow.
-The standard formulation does pretty well, and XT3D nails it again.
-(preliminary result from disu\_approach model, 5x11 grid with cross-connections, $30^{\circ}$ incline, set up for flow across the channel)}
+The standard formulation does pretty well, and XT3D nails it.
+(preliminary result from disu\_approach model, 5x11 grid with cross-connections, $30^{\circ}$ incline, set up for flow across the channel; scenarios disu-cf-vs-s and disu-cf-vs-x)}
+\label{fig:figures}
+\end{figure}
+
+\subsection{Test Problem 3: Vertically Offset Discretization with Domain}
+
+\begin{figure}[H]
+\centering
+\begin{subfigure}{0.4\textwidth}
+	\includegraphics[width=\textwidth]{../figures/disu-d-af-vo-s-head.png}
+	\caption{standard conductance-based formulation}
+	\label{fig:disu-s-nocc-head}
+\end{subfigure}
+\hfill
+\begin{subfigure}{0.4\textwidth}
+	\includegraphics[width=\textwidth]{../figures/disu-d-af-vo-x-head.png}
+	\caption{XT3D}
+	\label{fig:disu-x-nocc-head}
+\end{subfigure}
+\caption{Specific discharge and head for a vertically offset grid with the domain.
+Results are similar to those without the domain and consistent with what Kerry and Jim found previously; the flow tends to go horizontal, and XT3D doesn't really help.
+(preliminary result from disu\_approach model, 5x11 channel with 5x11 subdomains above and below, without cross-connections, $30^{\circ}$ incline, set up for flow along the channel; scenarios disu-d-af-vo-s and disu-d-af-vo-x)}
+\label{fig:figures}
+\end{figure}
+
+XXXXX Idea: See how heterogeneity affects the error -- series of runs with varying permeability contrast between channel and surrounding domain (from 1:1 to very large), producing a plot along the lines of figs.
+3d and 4d in \cite{bardot2022}.
+Will need to use the transect\_benchmark model for this, as we probably should for all of the final simulations.
+XXXXX
+
+\begin{figure}[H]
+\centering
+\begin{subfigure}{0.4\textwidth}
+	\includegraphics[width=\textwidth]{../figures/disu-d-cf-vo-s-head.png}
+	\caption{standard conductance-based formulation}
+	\label{fig:disu-s-nocc-cf-head.}
+\end{subfigure}
+\hfill
+\begin{subfigure}{0.4\textwidth}
+	\includegraphics[width=\textwidth]{../figures/disu-d-cf-vo-x-head.png}
+	\caption{XT3D}
+	\label{fig:disu-x-nocc-cf-head}
+\end{subfigure}
+\caption{Specific discharge and head for a vertically offset grid with the domain and boundary heads set up for crossflow.
+The standard formulation gives nearly vertical flow in the channel, and XT3D also does poorly, even mid-channel.
+The huge specific discharge vectors near the channel boundary are apparently caused by XT3D's head-gradient interpolation drawing heads from the domain, where the head gradient is huge, when computing nominally horizontal flows between cells within the channel.
+(preliminary result from disu\_approach model, 5x11 channel with 5x11 subdomains above and below, without cross-connections, $30^{\circ}$ incline, set up for flow across the channel; scenarios disu-d-cf-vo-s and disu-d-cf-vo-x)}
+\label{fig:figures}
+\end{figure}
+
+\subsection{Test Problem 4: Vertically Staggered Discretization with Domain}
+
+\begin{figure}[H]
+\centering
+\begin{subfigure}{0.4\textwidth}
+	\includegraphics[width=\textwidth]{../figures/disu-d-af-vs-s-head.png}
+	\caption{standard conductance-based formulation}
+	\label{fig:disu-s-cc-head}
+\end{subfigure}
+\hfill
+\begin{subfigure}{0.4\textwidth}
+	\includegraphics[width=\textwidth]{../figures/disu-d-af-vs-x-head.png}
+	\caption{XT3D}
+	\label{fig:disu-x-cc-head}
+\end{subfigure}
+\caption{Specific discharge and head for a vertically staggered grid with the domain and with cross-connections.
+The standard formulation isn't horrible, and XT3D does very well except at the channel boundary, where there's presumably a "staircase" effect.
+(preliminary result from disu\_approach model, 5x11 channel with 5x11 subdomains above and below, with cross-connections, $30^{\circ}$ incline, set up for flow along the channel; scenarios disu-d-af-vs-s and disu-d-af-vs-x)}
+\label{fig:figures}
+\end{figure}
+
+\begin{figure}[H]
+\centering
+\begin{subfigure}{0.4\textwidth}
+	\includegraphics[width=\textwidth]{../figures/disu-d-cf-vs-s-head.png}
+	\caption{standard conductance-based formulation}
+	\label{fig:disu-s-cc-cf-head}
+\end{subfigure}
+\hfill
+\begin{subfigure}{0.4\textwidth}
+	\includegraphics[width=\textwidth]{../figures/disu-d-cf-vs-x-head.png}
+	\caption{XT3D}
+	\label{fig:disu-x-cc-cf-head}
+\end{subfigure}
+\caption{Specific discharge and head for a vertically staggered grid with the domain and with cross-connections and boundary heads set up for crossflow.
+The standard formulation is much improved.
+You can't tell because of the scale of the vectors, but XT3D does well mid-channel.
+However, near the channel boundary it still has huge flows due to interpolating heads from the domain.
+(preliminary result from disu\_approach model, 5x11 channel with 5x11 subdomains above and below, $30^{\circ}$ incline, set up for flow across the channel; scenarios disu-d-cf-vs-s and disu-d-cf-vs-x)}
 \label{fig:figures}
@@
 Results with the surrounding "domain" (from an updated disu\_approach notebook) are similar but not identical, apparently due to the "staircase" effect.
-Might be best to presnt most of our results without the domain, at least when illustrating the main problem and its solution (to keep the "theory" clear), then discuss what happens when the domain is present (which is more like in practice).
+Might be best to present most of our results without the domain, at least when illustrating the main problem and its solution (to keep the "theory" clear), then discuss what happens when the domain is present (which is more like in practice).
 

### 2023-08-22 latex/article.tex
Reorganization and tidying
* Modified title, filled in contact info and keywords
* Fleshed out and tightened up Introduction and Theoretical Background
* Simplified fig 1
* Organized side notes and modified/added based on chat of 2023-08-22

@@
 \usepackage{authblk}
+\usepackage{xcolor}
 
@@
 
-\title{Application of the XT3D Multi-Point Flux Approximation to Vertically Staggered Grids in MODFLOW 6 to Improve Accuracy of Simulated Flows in Steeply Dipping Layers}
+\title{Improving Accuracy of Simulated Flows in Steeply Dipping Layers Using Vertically Staggered Grids with the XT3D Multi-Point Flux Approximation in MODFLOW 6}
 \author{
-	Xt3d Enthusiast1, Affiliation1  \\
+	Alden M. Provost, U.S. Geological Survey, Integrated Modeling and Prediction Division, U.S. Geological Survey, 12201 Sunrise Valley Dr, Reston, VA, USA  \\
 	\and 
-	Xt3d Enthusiast2, Affiliation2 \\
+	Kerry Bardot, School of Earth Sciences, University of Western Australia, Perth, Australia \\
 	\and 
-	Xt3d Enthusiast3, Affiliation3 \\
+	Christian D. Langevin, U.S. Geological Survey, Integrated Modeling and Prediction Division, 2280 Woodale Drive, Mounds View, MN, USA \\
 	\and 
-	Xt3d Enthusiast4, Affiliation4 \\
+	James L. McCallum, School of Earth Sciences, University of Western Australia, Perth, Australia \\
 	}
@@
 
-\textbf{Conflict of interest:} There is no conflict that is of interest to us.
+\textbf{Conflict of interest:} None.
 
-\textbf{Key words:} keyboard, keynote, turnkey, turkey, monkey
+\textbf{Key words:} groundwater flow simulation, dipping model layers, vertically staggered grid, multi-point flux approximation
 
-\textbf{Article impact statement:} This will likely leave a sizable impact crater.
+\textbf{Article impact statement:} {\color{red} AMP: Need impact statement}
 
 \begin{abstract}
-This is the best paper ever...
+\noindent {\color{red} AMP: compose Abstract near the end}
 \end{abstract}
@@
 
-Some intro stuff here about MODFLOW \citep{modflow6framework, modflow6gwf, modflow6gwt}...
+In the MODFLOW family of hydrologic simulation codes, a groundwater model domain is discretized into a grid of hydraulically connected model cells.
+Versions of MODFLOW up to and including MODFLOW-2005 \citep{modflow2005} offered only structured grids composed of rows, columns, and layers of cells.
+Cells were conceptualized as having horizontal top and bottom surfaces, which allowed each cell to be ``simulated as if it were rectangular so that flow may be approximated by the standard finite-difference equation" \citep{modflow84}.
+In MODFLOW-USG \citep{modflowusg} and MODFLOW 6 \citep{modflow6gwf}, which offer unstructured grids and use the control-volume finite-difference method to formulate the discrete balance equations, cells continue to have horizontal tops and bottoms, which simplifies the definition and calculation of water-table elevations, cell saturations, and hydraulic conductances between cells.
 
-XXXXX If cell-cell flows are inaccurate, advective and dispersive transport will be inaccurate XXXXX
+A traditional approach to representing an aquifer system with dipping hydrogeologic layers in MODFLOW is to map the hydrologic properties onto a rectilinear grid \citep{modflow84}.
+Another approach available in all versions of MODFLOW is to allow the top and bottom elevations of cells in the same layer of the grid to vary spatially.
+Following \cite{bardot2022}, in this work we call this type of grid ``vertically offset." The vertical offsets between ``horizontally" adjacent cells allows model layers to ``deform'" with the hydrostratigraphy to ``minimize the number of model layers required to simulate an aquifer system" \citep{modflow84} relative to a rectilinear grid with horizontal layers.
+In both approaches, the model grid follows the boundaries of dipping layers approximately, in what \cite{bardot2022} call a ``staircase" fashion.
 
-MODFLOW requires model cells to have horizontal tops and bottoms.
-In versions of MODFLOW up to and including MODFLOW-2005 \citep{modflow2005}, which offered only structured grids composed of rows, columns and layers of cells, a horizontal top and bottom allowed each cell to be ``simulated as if it were rectangular so that flow may be approximated by the standard finite-difference equation" \citep{modflow84}.
-In MODFLOW-USG \citep{modflowusg} and MODFLOW 6 \citep{modflow6gwf}, which offer unstructured grids and use the control-volume finite-difference method to formulate the discrete balance equations, horizontal cell tops and bottoms still simplify the definition and calculation of water-table elevations, cell saturations, and conductances between cells.
-In all versions of MODFLOW, cells in the same model layer can be vertically offset from each other to allow model layers to ``deform'" with the hydrostratigraphy and thereby ``minimize the number of model layers required to simulate an aquifer system" \citep{modflow84} relative to a rectilinear grid with strictly horizontal layers.
-
-Prior to MODFLOW 6, all versions of MODFLOW used a ``conductance-based" formulation to approximate the flow between two adjacent cells.
-The conductance-based formulation is a ``two-point" approximation in which the flow is proportional to the difference in the heads computed at the two cell centers and a conductance that depends on the hydraulic conductivities of the cells along the direction of the cell connection, the area of the cell interface, and a measure of the distance between the cell centers.
-On structured MODFLOW grids with no vertical offset between cells, i.e., strictly rectilinear grids, with hydraulic conductivity that is isotropic or has anisotropy aligned with the three mutually perpendicular grid directions, the conductance-based formulation is second-order accurate \citep{dehotin2010modeling, modflow6gwf}.
-This is because structured MODFLOW grids with no vertical offset satisfy the ``control-volume finite-difference (CVFD) requirement" that the straight-line connection between two cell centers must intersect the midpoint of the cell interface at a right angle \citep{narasimhan1976integrated}.
-Structured MODFLOW grids with vertical offset between cells, which we will call "vertically offset" grids, violate the CFVD requirement because nominally "horizontal" connections between cell centers in the same layer are not strictly horizontal and therefore not perpendicular to the vertical cell interfaces they intersect.
-(XXXXX Also, in the grids considered by \cite{narasimhan1976integrated}, there was no partial overlap between cells faces, but is that relevant, and do we want to get into that?
-XXXXX) The accuracy of the conductance-based flow formulation decreases with increasing vertical offset.
+All versions of MODFLOW prior to MODFLOW 6 used a two-point, ``conductance-based" flow formulation in which the flow between two adjacent cells is proportional to the difference between the heads calculated in those two cells.
+The conductance-based flow formulation is most accurate when the grid satisfies the ``control-volume finite-difference (CVFD) requirement" that the straight-line connection between two cell centers must intersect the midpoint of the cell interface at a right angle \citep{narasimhan1976integrated}.
+However, vertically offset grids violate the CVFD requirement because cell interfaces are vertical but cell connections are not strictly horizontal.
 For this reason, the MODFLOW 6 documentation \citep{modflow6gwf} warns that "[s]teeply dipping layers generally should not be represented" using vertical offsets.
 \cite{anderson2015applied} recommend a $10^{\circ}$ dip as a practical upper limit for using vertical offsets.
-A traditional alternative to using a vertically offset grid is to map the hydrologic properties of the aquifer system onto a strictly rectilinear grid, which typically requires more model layers to adequately discretize dipping hydrostratigraphic units \citep{modflow84}.
 
-The ability to use unstructured, or ``irregular," grids in MODFLOW-USG and MODFLOW 6 introduced the potential for grids to violate the CVFD requirement and thereby render the conductance-based formulation less accurate even in the absence of vertical offsets.
-To improve accuracy on irregular grids, and to allow accurate simulation of arbitrarily oriented two- or three-dimensional anisotropy, the optional XT3D flow formulation \citep{modflow6xt3d} was introduced in MODFLOW 6.
-XT3D formulates the flow between two cells by interpolating head values from the two cells and their neighboring cells to construct an estimate of the full, two- or three-dimensional head-gradient vector on each side of the cell interface; applying the cell conductivity tensors to obtain an estimate of the groundwater flux vector on each side of the interface; and reconciling the two flux estimates to ensure continuity of flow across the interface.
-By performing flow calculations using Darcy's Law in its tensorial form, based on a ``multi-point" approximation of the head gradient, XT3D accounts for both arbitrarily oriented anisotropy and geometric irregularity of the grid.
+To improve the accuracy of flow simulations on unstructured grids, and to allow accurate simulation of arbitrarily oriented two- or three-dimensional anisotropy, the optional XT3D flow formulation \citep{modflow6xt3d} was introduced in MODFLOW 6.
+By performing flow calculations using Darcy's Law in its tensorial form based on a multi-point approximation of the head gradient vector, XT3D accounts for both arbitrarily oriented anisotropy and geometric irregularity of the grid.
+In theory, XT3D should be able to compensate for the geometic irregularity introduced by vertical offsets between adjacent cells in the same model layer.
 
-\cite{bardot2022} evaluated the effects of grid design and XT3D on the ability of MODFLOW 6 to accurately simulate flow through sedimentary structures by modeling a two-dimensional, permeable channel embedded in a nearly impermeable surrounding medium, or ``domain." The channel was angled counterclockwise from the $x$ axis of the model coordinates (by $30^{\circ}$ in most cases), and constant-head boundary conditions at the ends of the channel were set to induce uniform flow along the channel in the analytical solution.
-In their plan-view benchmark, the channel and surrounding domain were discretized using either a ``Regular Cartesian (RC)" grid that is rectilinear and satisfies the CVFD requirement or a ``Flexible Triangular (FT)" grid that is unstructured with triangular cells and does not satisfy the CVFD requirement.
-They found that the standard, conductance-based flow formulation gave accurate results on the RC grid but was subject to significant error on the FC grid, whereas the XT3D flow formulation gave accurate results on both grids, as expected.
-However, in their transect (cross-sectional) benchmark, which used a vertically offset grid to discretize the domain and a channel inclined by $30^{\circ}$, similarly poor results were observed with and without XT3D.
-The simulated groundwater flux was nearly horizontal ($0.01 - 0.03^{\circ}$ incline), and its magnitude was overestimated by 12 - 15\%.
+\cite{bardot2022} evaluated the effects of grid design and XT3D on the ability of MODFLOW 6 to accurately simulate flow through sedimentary structures by modeling an idealized, two-dimensional permeable channel or hydrogeologic layer embedded in a nearly impermeable surrounding medium, or ``domain." In their plan-view benchmark, the channel lay within the horizontal plane as was angled counterclockwise relative to the southern boundary of the model domain (by $30^{\circ}$ in most cases).
+Constant-head boundary conditions at the ends of the channel were set to induce uniform flow along the channel in the corresponding analytical solution.
+The channel and surrounding domain were discretized using either a rectilinear grid, which satisfies the CVFD requirement, or a ``flexible triangular" grid that is unstructured with triangular cells and does not satisfy the CVFD requirement.
+They found that the standard, conductance-based flow formulation gave accurate results on the rectilinear grid but was subject to significant error on the unstructured grid, whereas the XT3D flow formulation gave accurate results on both grids, as expected.
+However, in the transect (cross-sectional) version of their benchmark, which used a vertically offset grid to discretize the nearly impermeable domain and a permeable channel (hydrogeologic layer) inclined by $30^{\circ}$ from the horizontal, similarly poor results were observed with and without XT3D.
 Given the ability of XT3D to compensate for violations of the CVFD requirement, it appeared likely that another aspect of the problem formulation or the inner workings of MODFLOW 6 was responsible for the unexpected results.
-(XXXXX Specifically mention that the NPF angle-specification "nozee" issue was looked into a found not to be the problem?
-Kind of an awkward detail to shoehorn in, but it would advertise this potential pitfall, which probably isn't well-known XXXXX) \cite{bardot2022} noted a suggestion by two authors of the present paper (Provost and Langevin) that inadequate grid connectivity was at the root of the problem.
+As noted by \cite{bardot2022}, a preliminary analysis by two authors of the present work (Provost and Langevin) suggested that the cell connectivity offered by a vertically offset grid was inadequate for allowing an accurate flow solution.
 
-"Vertically staggered" grid connections were introduced in MODFLOW-USG \citep{modflowusg} and included for the DISU grid type in MODFLOW 6 \citep{modflow6gwf} primarily as a means to vertically subdiscretize portions of a grid that is conceptualized as being layered.
-In a vertically staggered grid, i.e., a grid that includes vertically staggered connections, some cells have nominally ``horizontal" connections with more than one cell across a given cell face.
-Vertically staggered MODFLOW 6 DISU grids allow more flexibility in assigning grid connectivity than do the more commonly used DIS and DISV grid types \citep{modflow6gwf}.
+"Vertically staggered" grid connections were introduced in MODFLOW-USG \citep{modflowusg} and included for the fully unstructured (DISU) grid type in MODFLOW 6 \citep{modflow6gwf} primarily as a means to vertically subdiscretize portions of a grid that are conceptualized as being layered.
+A grid that includes vertically staggered connections, which in this work we call a ``vertically staggered grid," some cells have nominally ``horizontal" connections with more than one cell across a given cell face.
+Vertically staggered DISU grids allow more flexibility in assigning grid connectivity than do the more commonly used DIS and DISV grid types \citep{modflow6gwf} in MODFLOW 6.
 To our knowledge, the implications of limited grid connectivity for the accuracy of simulated flows in models with steeply dipping layers, and the potential for using vertically staggered grids to overcome this limitation, have not been previously appreciated or investigated.
 
-In the remainder of this paper we first present a theoretical justification for the hypothesis that inadequate grid connectivity is primarily responsible for the inaccurate simulation of flows in the steeply dipping channel benchmark on a vertically offset grid.
+In the remainder of this paper we first present a theoretical justification for the hypothesis that inadequate grid connectivity is primarily responsible for the inaccurate simulation of flows in the steeply dipping layer benchmark on a vertically offset grid.
 Based on that theory, we propose a method for improving the accuracy of the flow solution using XT3D together with a vertically staggered grid, and we evaluate the effectiveness of that approach in a set of benchmark problems similar to those of \cite{bardot2022}.
@@
 
-XXXXX Random thoughts:
+{\color{red} AMP: terminology -- ``channel" vs ``hydrogeologic layer"? ``hydrogelogic layer" vs ``model layer"?}
+
+{\color{red} AMP: Random thoughts
 \begin{itemize}
-	\item Vertical offset happens even on a completely rectilinear grid if the water table elevation varies in horizontally adjacent cells. Doesn't seem like the same kind of issue, but think on it.
-	\item What happens with a head gradient perpendicular to the dip of the channel (across rather than along the channel)? Preliminary results suggest cross-connections and xt3d help.
-	\item Could also try making the channel less permeable than the domain; semi-confining unit.
+	\item Vertical offset happens even on a completely rectilinear grid if the water table elevation varies in horizontally adjacent cells.
+Doesn't seem like the same kind of issue, but think on it.
+But not now, I think.
+	\item Could mention the crossflow case briefly in this paper, like \cite{bardot2022} mentioned this work as something we were looking into at the time.
+	\item Could try making the channel less permeable than the domain; semi-confining unit. Thinking not now.
 	\item Difference between truly single-layer model of channel vs embedded single-layer channel -- no vertical connections vs vertical connections that are no-flux.
 Effect on specific discharge calculations?
+Relevance of the specific-discharge "patch"?
+But beyond that, Kerry suggested that it's important to investigate and describe the embedded single-layer case, as it's a common use case, and she did some runs -- email of 2022-09-29.
+With or without xt3d, the results were not good, and I'm not sure it's possible to get good results, since there's no way to introduce cross-connections within a single-layer channel.
+Would be good to explain this in the paper.
+(She also tried a vertical gradient for the embedded single-layer case, also with poor results, though this might also have something to do with the cross-flow, "heterogeneity correction" issue.)
 \end{itemize}
+}
 
@@
 %	\includegraphics[scale=0.8]{../figures/schem_conn_area_flux.png}
-	\includegraphics[scale=0.4]{../figures/schem_conn_area_flux_resaved.png}
-	\caption{Schematics showing (a - c) grid connectivity and vertical cell interface areas and (c - f) cell interface fluxes in two-layer models of a dipping channel using three different grid configurations: (a, d) dip-following, (b, e) vertically offset, and (c, f) cross-connected.
+	\includegraphics[scale=0.6]{../figures/schem_conn_area_flux.png}
+	\caption{Schematics showing (a) grid connectivity and (b) cell interface fluxes in a two-layer model of a dipping channel.
 Hatching denotes impermeable boundaries along the top and bottom of the channel.
-In (a) - (c), gray circles and gray lines represent cell centers and grid connections between cells, respectively, and two-headed arrows indicate cell thicknesses used to compute cell interface areas.
-In (d) - (f), gray arrows represent specific-discharge (groundwater flux) vectors for uniform flow through the channel, and black arrows represent the corresponding components of flux normal to cell interfaces.
-For visual clarity, (f) shows gray arrows only for the cell interfaces that do not exist in (e).}
+In (a), black circles represent cell centers.
+Blue lines represent hydraulic connections between cells in a vertically offset grid.
+Red lines represent additional connections that result in a vertically staggered grid.
+In (b), gray arrows represent the uniform groundwater flux the model is attempting to simulate.
+Blue arrows represent components of the uniform groundwater flux normal to the horizontal and vertical cell-cell interfaces in the vertically offset grid.
+Red arrows represent components of the uniform groundwater flux normal to the additional vertical cell-cell interfaces introduced in the vertically staggered grid.}
 	\label{fig:schem-conn-area-flux}
@@
 
-Figure \ref{fig:schem-conn-area-flux}a shows a group of four cells in a hypothetical model in which the dipping channel is discretized vertically into two layers of cells.
-Although such a configuration of cells can be constructed in plan view using the DISV grid type of MODFLOW 6, it cannot be constructed in cross-section because MODFLOW requires cells to have horizontal tops and bottoms.
-The sloping tops and bottoms of these hypothetical cells allow the cell geometry to conform to the dip of the channel, and we will therefore refer to this grid configuration as ``dip-following." The grid connectivity is analogous to the connectivity found in a MODFLOW 6 DIS (structured, or regular) grid: a cell is hydraulically connected to each cell with which it shares a face.
-For example, cell 1A is connected to cells 1B and 2A, but not to cell 2B.
-Because cells share entire faces, i.e., adjacent cell faces overlap completely, the entire cell thickness is relevant in determining the area for flow between cells in the same layer.
+Figure \ref{fig:schem-conn-area-flux}a shows a group of four cells in a two-layer, cross-sectional MODFLOW 6 model of fully saturated flow through a permeable channel (aquifer) with impermeable top and bottom boundaries.
+The cells have horizontal tops and bottoms and can therefore follow the dip of the channel only on average, in what \cite{bardot2022} refer to as ``staircase" fashion.
+If the grid is represented using a layered grid type (DIS or DISV) in MODFLOW 6, a cell is hydraulically connected to each cell in the same layer with which it has an overlapping face and with each cell in an adjacent layer with which it shares a horizontal face, as indicated by the blue lines that connect cell centers in figure \ref{fig:schem-conn-area-flux}a.
+Following \cite{bardot2022}, in this work we call this type of grid ``vertically offset" because the tops and bottoms of adjacent cells in the same layer are not necessarily aligned.
+Note that cells that have overlapping vertical faces but are in different layers (e.g., cells 1A and 2B) do not have a direct hydraulic connection in a vertically offset grid.
+The addition of such connections between cells in different layers, as indicated by the red lines that connect cell centers in figure \ref{fig:schem-conn-area-flux}a, results in a ``vertically staggered" grid.
+The ability to handle vertically staggered grids was first introduced to MODFLOW by \cite{modflowusg} in MODFLOW-USG, and the term ``vertically staggered" was first used by \cite{modflow6gwf} in the context of MODFLOW 6.
+The term describes a grid in which which a cell can have nominally ``horizontal" connections with with multiple cells across the same cell face.
+In this work, ``vertically offset" and ``vertically staggered" are used to distinguish grids that lack nominally ``horizontal" connections between cells in adjacent layers (red lines in figure \ref{fig:schem-conn-area-flux}a) from those that include such connections, respectively.
+In the vertically offset grid in \ref{fig:schem-conn-area-flux}a (blue connections only), the interfacial area across which flow occurs between adjacent cells in the same layer is based on the vertical cell thickness (the entire thickness of the channel).
+In the vertically staggered grid (blue and red connections), the interfacial area for nominally ``horizontal" flow between adjacent cells is based on the area over which the two cell faces overlap.
+Thus, while the total area for nominally ``horizontal" flow across a cell face is the same as for a vertically offset grid, that total area is divided between an adjacent cell in the same layer (blue connection) and an adjacent cell in the adjacent layer (red connection).
+In MODFLOW 6, vertically offset grids can be represented using the structured (DIS), vertex-based (DISV), or fully unstructured (DISU) grid type, whereas vertically staggered grids can be represented only using the DISU grid type.
 
-One would intuitively expect a dip-following grid to be well suited to simulating uniform flow along the channel.
-In figure \ref{fig:schem-conn-area-flux}d, a specific discharge (groundwater flux) vector representative of uniform flow along the channel is superimposed on each cell interface, together with a vector representing the component of the flux normal to the interface.
-The normal flux vectors show the directions in which groundwater must be able to flow between cells to accurately represent the uniform flow field.
-In the case of a dip-following grid, flow must simply pass from left to right between adjacent cells in the same layer (e.g., from cell 1A to cell 1B, and from cell 2A to cell 2B).
-There is no flow of groundwater between cells in different layers (e.g., between cells 1A and 2A, or between cells 1B and 2B).
-The grid connectivity shown in figure \ref{fig:schem-conn-area-flux}a is sufficient to accommodate this straighforward pattern of flows between cells; in fact, vertical connections between cells in different layers are not even needed, and the channel could be represented just as well using a single layer.
+In MODFLOW 6, the default method for representing the flow between two hydraulically connected model cells is the ``conductance-based flow formulation" \citep{modflow6gwf}.
+In this formulation, which is based on the commonly used two-point flux approximation {\color{red} *** ref?
+***}, the flow is proportional to the difference in the hydraulic heads computed at the two cell centers and a hydraulic conductance based on an effective hydraulic conductivity for the connection between the cells.
+On strictly rectilinear grids, such as structured MODFLOW grids with no vertical offset between cells in the same layer, with hydraulic conductivity that is isotropic or has anisotropy aligned with the three mutually perpendicular grid directions, the conductance-based formulation is second-order accurate \citep{dehotin2010modeling, modflow6gwf}.
+This is because strictly rectilinear grids satisfy the ``control-volume finite-difference (CVFD) requirement" that the straight-line connection between two cell centers must intersect the midpoint of the cell interface at a right angle \citep{narasimhan1976integrated}.
+Vertically offset grids, however, violate the CFVD requirement because nominally ``horizontal" connections between cell centers in the same layer are not strictly horizontal and therefore not perpendicular to the vertical cell interfaces they intersect.
+For example, in figure \ref{fig:schem-conn-area-flux}a, the connection between cells 1A and 1B is not perpendicular to the cell interface, which compromises the accuracy of the conductance-based formulation.
 
-Figure \ref{fig:schem-conn-area-flux}b shows the four cells as they would be configured for a vertically offset DIS or DISV grid in a two-layer, cross-sectional MODFLOW 6 model of the channel.
-The cells have horizontal tops and bottoms and can therefore follow the dip of the channel only on average, in what \cite{bardot2022} refer to as ``staircase" fashion.
-The pattern of grid connectivity is analogous to that in figure \ref{fig:schem-conn-area-flux}a, but it is not based strictly on which cell faces overlap.
-Rather, a cell is connected to each cell in the same layer with which it has an overlapping face and with each cell in an adjacent layer with which it shares a horizontal face.
-Significantly, cells that have overlapping vertical faces but are in different layers (e.g., cells 1A and 2B) are not connected.
-In the conductance-based formulation for flow, the flow between two cells in the same layer of a DIS or DISV grid (XXXXX or DISU that's not vertically staggered, but do we want to mention that here?
-XXXXX) is proportional to the difference in heads computed at the two cell centers and a conductance based on an effective hydraulic conductivity for the connection between the cells.
-If the two cells are fully saturated and have the same hydraulic conductivity and thickness, as in this example, the cell interface area used to calculate the conductance is effectively based on the full cell thickness, as shown in figure \ref{fig:schem-conn-area-flux}b.
-Also, the length on which the conductance is based is the horizontal distance, not the straight-line distance, between the cell centers.
+The conductance-based formulation was the only flow formulation available in versions of MODFLOW prior to MODFLOW 6.
+As mentioned earlier, the limitations of the conductance-based formulation on vertically offset grids were well recognized.
+However, the ability to use unstructured grids in MODFLOW-USG and MODFLOW 6 introduced the potential to violate the CVFD requirement, and thereby render the conductance-based formulation less accurate, even in the absence of vertical offsets.
+To improve accuracy on unstructured grids, and to allow accurate simulation of arbitrarily oriented two- or three-dimensional anisotropy, the optional XT3D flow formulation \citep{modflow6xt3d} was introduced in MODFLOW 6.
+XT3D formulates the flow between two cells by interpolating head values from the two cells and their neighboring cells to construct an estimate of the full, two- or three-dimensional head-gradient vector on each side of the cell interface; applying the cell conductivity tensors to obtain an estimate of the groundwater flux vector on each side of the interface; and reconciling the two flux estimates to ensure continuity of flow across the interface.
+By performing flow calculations using Darcy's Law in its tensorial form, based on a ``multi-point" approximation of the head gradient, XT3D accounts for both arbitrarily oriented anisotropy and geometric irregularity of the grid.
 
-As mentioned earlier, for the conductance-based flow formulation to provide maximum accuracy, the model grid must satisfy the CVFD requirement that the straight-line connection between two cell centers must intersect the midpoint of the cell interface at a right angle.
-In a vertically offset grid, connections between cells across vertical cell interfaces are generally not horizontal and, therefore, not perpendicular to the interface.
-For example, in \ref{fig:schem-conn-area-flux}b, the nominally ``horizontal" connection between cells 1A and 1B is not perpendicular to the cell interface, which compromises the accuracy of conductance-based formulations.
-(XXXXX Also, in the grids considered by \cite{narasimhan1976integrated}, there was no partial overlap between cells faces, but do we want to get into that here?
-XXXXX)
+The dipping-channel benchmark problem of \cite{bardot2022} used a vertically offset grid, which is the type of grid used most often in MODFLOW models.
+Test simulations using the conductance-based flow formulation for a steeply dipping channel (hydrogeologic layer) embedded in a domain that is six orders of magnitude less hydraulically conductive yielded significant errors in the simulated flows.
+The simulated groundwater flux in the middle of the channel was not along the $30^{\circ}$ incline of the channel, but nearly horizontal ($0.03^{\circ}$ incline), and its magnitude was overestimated by 15\%.
+Significant errors in the simulated flows were expected in this case, given that the conductance-based flow formulation does not rigorously account for violations of the CVFD requirement that the cell-cell connection be perpedicular to the cell-cell interface, and that it takes the connection length to be the horizontal distance between cell centers.
+Given the ability of XT3D to account rigorously for grid connections that are not perpendicular to cell interfaces, and for the increase in connection length due to the slope of the connection, rerunning the simulation with XT3D activated could have been expected to improve the flow solution substantially.
+However, \cite{bardot2022} observed similarly poor results with XT3D: the flux was still nearly horizontal ($0.01^{\circ}$ incline), and its magnitude was overestimated by 12\%.
+Based on a preliminary analysis by two authors of the present work (Provost and Langevin), \cite{bardot2022} hypothesized that the unexpectedly large error in the flow solution with XT3D was related to indequate grid connectivity was at the root of the problem.
 
-XXXXX Mostly covered in intro, so condense here XXXXX In their dipping-channel benchmark problem, \cite{bardot2022} found that, as expected, the standard, conductance-based formulation gave poor results for a steeply dipping channel discretized using a vertically offset grid.
-The simulated groundwater flux in the middle of the channel was not along the $30^{\circ}$ incline of the channel, but nearly horizontal ($0.03^{\circ}$ incline), and its magnitude was overestimated by 15\%.
-Given the ability of XT3D to compensate for grid connections that are not perpendicular to cell interfaces, rerunning the simulation with XT3D activated could have been expected to improve the flow solution substantially.
-However, \cite{bardot2022} observed similarly poor results with XT3D: the flux was still nearly horizontal ($0.01^{\circ}$ incline), and its magnitude was overestimated by 12\%.
-Something other than a simple violation of the CVFD perpendicularity condition was evidently responsible, and \cite{bardot2022} noted a suggestion by two authors of the present paper (Provost and Langevin) that inadequate grid connectivity was at the root of the problem.
+The role of grid connectivity in enabling accurate simulation of flow along a dipping channel can be understood by considering figure \ref{fig:schem-conn-area-flux}b, in which components of the groundwater flux (specific discharge) vector representative of steady, uniform flow along the channel are superimposed on each cell interface.
+Gray vectors represent the uniform flux oriented along the channel, which the model is attempting to simulate.
+Blue vectors represent the flux components normal to cell interfaces that correspond to blue connections in figure \ref{fig:schem-conn-area-flux}a, which are common to both the vertically offset and vertically staggered grids.
+Red vectors represent the flux components normal to cell interfaces that correspond to red connections in figure \ref{fig:schem-conn-area-flux}a, which are lacking in the vertically offset grid.
 
-The influence of grid connectivity on the results observed by \cite{bardot2022} can be explained with the help of figures \ref{fig:schem-conn-area-flux}b and \ref{fig:schem-conn-area-flux}e.
-Figure \ref{fig:schem-conn-area-flux}e shows the specific discharge vector superimposed on each cell interface of the vertically offset grid, together with a vector representing the component of the flux normal to the interface.
-As in the case of a dip-following grid (figure \ref{fig:schem-conn-area-flux}d), flow passes from left to right between adjacent cells in the same layer.
-However, unlike in the dip-following grid, there must also be vertical flow from each cell in layer 2 and to the cell immediately above it in layer 1 (e.g., from cell 2A and to cell 1A).
-This pattern of flow between cells obviously cannot be a solution to uniform, steady-state flow through the channel, since the exclusively upward vertical flows would deplete layer 2 and accumulate water in layer 1.
-Short of resorting to alternating upward and downward vertical flows, which would compromise the uniformity of the flow, the only alternative for representing uniform, steady-state flow on the vertically offset grid is to have exclusively horizontal flow.
-Although this concept is illustrated for a two-layer model in figure \ref{fig:schem-conn-area-flux}e, the same reasoning applies to a model of the channel with any number of layers.
-This explains why \cite{bardot2022} observed nearly horizontal flow in the middle portion of the channel, away from the lateral boundaries, where nearly uniform flow was established.
-Thus, it appears the nearly horizontal flows are not primarily due to what \cite{bardot2022} call the ``cosine error" associated with the orientation and length of the sloping grid connections, but to the inability of the vertically offset grid to route flow appropriately between cells.
+The blue vectors in figure \ref{fig:schem-conn-area-flux}b show that to simulate steady, uniform flow along the dipping channel, cells must exchange water not only ``horizontally" with neighboring cells in the same layer, but also vertically with cells in the adjacent layer.
+Specifically, there must be flow from cells in the bottom layer to cells in the top layer, which corresponds to the vertical component of the groundwater flux.
+Note, however, that the blue vectors alone are incompatible with steady, uniform flow along the channel because the exchange of water between the two model layers is in one direction only: upward.
+Such a steady-state flow configuration would imply continuous depletion of flow within the bottom layer and accumulation of flow within the top layer as one moves along the channel in the direction of flow.
+Thus, steady, uniform flow along the channel cannot be simulated accurately without some mechanism for returning flow from the top layer to the bottom layer.
+The vertically staggered grid provides such a mechanism by offering additional connections between cells in adjacent layers.
+Vertical flows from the bottom layer to the top layer, represented by vertical blue vectors in figure \ref{fig:schem-conn-area-flux}b, can be returned to the bottom layer by nominally ``horizontal" flows, represented by red vectors in figure \ref{fig:schem-conn-area-flux}b.
+Although this concept has been illustrated using a two-layer model for simplicity, the same reasoning applies given any number of layers.
 
-This suggests an approach for constructing the grid to improve the accuracy of the flow solution.
-Figure \ref{fig:schem-conn-area-flux}c shows a grid that includes additional connections between cells in different layers compared to the grid in \ref{fig:schem-conn-area-flux}b.
-For example, cell 1A is connected not only cells 2A and 1B, but also the cell 2B.
-The grid connectivity is determined by overlap between cell faces; if two cells have faces that overlap even partially, the cells are connected across those faces.
-The area of the interface is the area of overlap between the cell faces.
-This grid configuration, which was introduced MODFLOW-USG \citep{modflowusg} and is available for the DISU grid type in MODFLOW 6, is called ``vertically staggered" and allows a cell to have nominally ``horizontal" connections with more than one cell across a given cell face.
-In this case, vertical staggering is used to introduce ``cross-connections" between cells in different layers.
-The ramifications of cross-connections for flow between cells are illustrated in figure \ref{fig:schem-conn-area-flux}f.
-Inclusion of cross-connections (e.g., between cells 1A and 2B) allows upward vertical flows from layer 2 to layer 1 (e.g., from cell 2A to cell 1A) to be routed back down to layer 2 so water does not accumulate in layer 1.
-The remainder of this paper describes test problems that evaluate the effectiveness of cross-connections for improving the accuracy of simulated flow though the channel.
+Once the appropriate connectivity is established using a vertically staggered grid, it is the role of the flow formulation to account properly for the geometry of the grid, a task for which XT3D was specifically designed.
+However, XT3D alone cannot compensate for grid connectivity that does not provide adequate pathways for flow.
+This explains why using XT3D did not improve the simulation results substantially on a vertically offset grid in the benchmark tests of \cite{bardot2022}.
+In the middle section of the channel, away from the end boundary conditions, the flow solution approached steady, uniform flow the only way that it could given the limited connectivity of the vertically offset grid: by suppressing vertical flows between layers, thereby rendering the overall flow approximately horizontal.
+
+The arguments presented in this section suggest that, in addition to the use of XT3D, the enhanced connectivity provided by a vertically staggered grid can be important for obtaining an accurate flow solution in MODFLOW 6 simulations that involve steeply dipping layers.
+The next section presents results of simulations designed to test this hypothesis.
 
@@
 
-To simulate flow along a channel inclined at angle $\theta$, heads are specified at the centers of cells along the perimeter of the model (XXXXX this assumes we're using the transect\_benchmark version of the model, which includes the "domain," for all of our final runs XXXXX) using an analytical solution that corresponds to a unit head gradient inclined at angle $\theta$ \citep{bardot2022}:
+To simulate flow along a channel inclined at angle $\theta$, heads are specified at the centers of cells along the perimeter of the model {\color{red} (AMP: this assumes we're using the transect\_benchmark version of the model, which includes the "domain," for all of our final runs)} using an analytical solution that corresponds to a unit head gradient inclined at angle $\theta$ \citep{bardot2022}:
 
@@
 \noindent where $h$ is head in meters and $x$ and $z$ are the horizontal and vertical model coordinates, respectively, in meters.
-Flow across the channel (XXXXX seems like this would be good to include XXXXX) is simulated using the modified head function
-\begin{equation}
-\label{eqn:head_analyt_across}
-h = x \sin \theta - z \cos \theta,
-\end{equation}
-which represents a unit head gradient perpendicular to the long axis of the channel.
-In both cases, the hydraulic conductivity of the channel is set to 1 m/d (meter per day) so that the analytical flow solution is a groundwater flux of 1 m/d either along or across the channel.
+The hydraulic conductivity of the channel is set to 1 m/d (meter per day) so that the analytical flow solution is a groundwater flux of 1 m/d either along or across the channel.
 The conductivity of the domain surrounding the channel varies according to the scenario being investigated.
@@
 In the highly heterogeneous limit, the conductivity of the domain is set much lower than that of the channel to effectively isolate the channel hydraulically.
-XXXXX Could also try making the channel less permeable than the domain; semi-confining unit XXXXX
 
@@
 The first test problem uses the traditional approach of representing the inclined channel using a vertically offset grid of the MODFLOW 6 DIS (structured) type.
-(XXXXX To be decided: present results for the DIS grid, or for the equivalent of the DIS grid after conversion to DISU?
-XXXXX) This problem confirms the findings of \cite{bardot2022} and illustrates the limitations of the vertically offset grid approach.
+{\color{red} (AMP: present results for the DIS grid, or for the equivalent of the DIS grid after conversion to DISU?)} This problem confirms the findings of \cite{bardot2022} and illustrates the limitations of the vertically offset grid approach.
 The second test problem is similar to the first but uses a vertically staggered grid of MODFLOW 6 DISU (unstructured) type.
@@
 
-XXXXX See the temporary ``preliminary results" section at the end of the manuscript XXXXX
+{\color{red} AMP: Note that there's a temporary ``preliminary results" section at the end of the manuscript.}
 
-XXXXX Results without the domain are presented first to "cleanly" illustrate the main problem and its solution.
+{\color{red} AMP: Obviously need to substitute in our "real" results with "real" discussion and captions here.}
+
+{\color{red} AMP: Based on the release notes, looks like the specific-discharge ``patch" went into v6.3.0. Mention it? Note that here is still the issue of coordinate dependence.}
+
+{\color{red} AMP: Results without the domain are presented first to "cleanly" illustrate the main problem and its solution.
 Results with the domain are then presented to give more of an indication of how things might work in practice.
-Unless stated otherwise, the domain is 6 orders of magnitude less conductive than the channel, as in \cite{bardot2022}.
-XXXXX
+Kerry has suggested that perhaps the cases with the domain are sufficient -- let's consider.}
+
+{\color{red} AMP: Unless stated otherwise, the domain is 6 orders of magnitude less conductive than the channel, as in \cite{bardot2022}.}
+
+{\color{red} AMP: Add and discuss results for how the flow error varies with discretization.  See plots in Kerry's emails of 2022-10-06 and 2022-12-14.}
+
+{\color{red} AMP: Add and discuss results for how the flow error varies with dip angle. See plots in Kerry's email of 2022-12-14. "Spike" at 45 degrees due to "disconnection?"}
+
+{\color{red} AMP: Add and discuss results for how the flow error varies with K contrast between channel and domain (heterogeneity).  See plots in Kerry's email of 2022-12-14.}
+
+{\color{red} AMP: For consistency, should choose one notebook or the other for all runs.}
 
@@
 	\includegraphics[width=\textwidth]{../figures/disu-af-vo-s-head.png}
-	\caption{standard conductance-based formulation}
+	\caption{vertically offset grid, standard conductance-based formulation}
 	\label{fig:disu-s-nocc-head}
@@
 	\includegraphics[width=\textwidth]{../figures/disu-af-vo-x-head.png}
-	\caption{XT3D}
+	\caption{vertically offset grid, XT3D}
 	\label{fig:disu-x-nocc-head}
@@
 (preliminary result from disu\_approach model, 5x11 grid without cross-connections, $30^{\circ}$ incline, set up for flow along the channel; scenarios disu-af-vo-s and disu-af-vo-x)}
-\label{fig:figures}
-\end{figure}
-
-\begin{figure}[H]
-\centering
-\begin{subfigure}{0.4\textwidth}
-	\includegraphics[width=\textwidth]{../figures/disu-cf-vo-s-head.png}
-	\caption{standard conductance-based formulation}
-	\label{fig:disu-s-nocc-cf-head.}
-\end{subfigure}
-\hfill
-\begin{subfigure}{0.4\textwidth}
-	\includegraphics[width=\textwidth]{../figures/disu-cf-vo-x-head.png}
-	\caption{XT3D}
-	\label{fig:disu-x-nocc-cf-head}
-\end{subfigure}
-\caption{Specific discharge and head for a vertically offset grid and boundary heads set up for crossflow.
-The standard formulation is a fail, but XT3D nails it, even without cross-connections.
-(preliminary result from disu\_approach model, 5x11 grid without cross-connections, $30^{\circ}$ incline, set up for flow across the channel; scenarios disu-cf-vo-s and disu-cf-vo-x)}
 \label{fig:figures}
@@
 	\includegraphics[width=\textwidth]{../figures/disu-af-vs-s-head.png}
-	\caption{standard conductance-based formulation}
+	\caption{vertically staggered grid, standard conductance-based formulation}
 	\label{fig:disu-s-cc-head}
@@
 	\includegraphics[width=\textwidth]{../figures/disu-af-vs-x-head.png}
-	\caption{XT3D}
+	\caption{vertically staggered grid, XT3D}
 	\label{fig:disu-x-cc-head}
@@
 The news is much better; the standard formulation is much improved, and XT3D nails it.
-(preliminary result from disu\_approach model, 5x11 grid with cross-connections, $30^{\circ}$ incline, set up for flow along the channel; scenarios disu-af-vs-s and disu-af-vs-x)}
-\label{fig:figures}
-\end{figure}
-
-\begin{figure}[H]
-\centering
-\begin{subfigure}{0.4\textwidth}
-	\includegraphics[width=\textwidth]{../figures/disu-cf-vs-s-head.png}
-	\caption{standard conductance-based formulation}
-	\label{fig:disu-s-cc-cf-head}
-\end{subfigure}
-\hfill
-\begin{subfigure}{0.4\textwidth}
-	\includegraphics[width=\textwidth]{../figures/disu-cf-vs-x-head.png}
-	\caption{XT3D}
-	\label{fig:disu-x-cc-cf-head}
-\end{subfigure}
-\caption{Specific discharge and head for a vertically staggered grid with cross-connections and boundary heads set up for crossflow.
-The standard formulation does pretty well, and XT3D nails it.
-(preliminary result from disu\_approach model, 5x11 grid with cross-connections, $30^{\circ}$ incline, set up for flow across the channel; scenarios disu-cf-vs-s and disu-cf-vs-x)}
+(preliminary result from disu\_approach model, 5x11 grid with cross-connections, $30^{\circ}$ incline, set up for flow along the channel; scenarios disu-af-vs-s and disu-af-vs-x)  {\color{red} AMP: If keeping, combine with fig 3.}}
 \label{fig:figures}
@@
 	\includegraphics[width=\textwidth]{../figures/disu-d-af-vo-s-head.png}
-	\caption{standard conductance-based formulation}
+	\caption{vertically offset grid, standard conductance-based formulation}
 	\label{fig:disu-s-nocc-head}
@@
 	\includegraphics[width=\textwidth]{../figures/disu-d-af-vo-x-head.png}
-	\caption{XT3D}
+	\caption{vertically offset grid, XT3D}
 	\label{fig:disu-x-nocc-head}
@@
 (preliminary result from disu\_approach model, 5x11 channel with 5x11 subdomains above and below, without cross-connections, $30^{\circ}$ incline, set up for flow along the channel; scenarios disu-d-af-vo-s and disu-d-af-vo-x)}
-\label{fig:figures}
-\end{figure}
-
-XXXXX Idea: See how heterogeneity affects the error -- series of runs with varying permeability contrast between channel and surrounding domain (from 1:1 to very large), producing a plot along the lines of figs.
-3d and 4d in \cite{bardot2022}.
-Will need to use the transect\_benchmark model for this, as we probably should for all of the final simulations.
-XXXXX
-
-\begin{figure}[H]
-\centering
-\begin{subfigure}{0.4\textwidth}
-	\includegraphics[width=\textwidth]{../figures/disu-d-cf-vo-s-head.png}
-	\caption{standard conductance-based formulation}
-	\label{fig:disu-s-nocc-cf-head.}
-\end{subfigure}
-\hfill
-\begin{subfigure}{0.4\textwidth}
-	\includegraphics[width=\textwidth]{../figures/disu-d-cf-vo-x-head.png}
-	\caption{XT3D}
-	\label{fig:disu-x-nocc-cf-head}
-\end{subfigure}
-\caption{Specific discharge and head for a vertically offset grid with the domain and boundary heads set up for crossflow.
-The standard formulation gives nearly vertical flow in the channel, and XT3D also does poorly, even mid-channel.
-The huge specific discharge vectors near the channel boundary are apparently caused by XT3D's head-gradient interpolation drawing heads from the domain, where the head gradient is huge, when computing nominally horizontal flows between cells within the channel.
-(preliminary result from disu\_approach model, 5x11 channel with 5x11 subdomains above and below, without cross-connections, $30^{\circ}$ incline, set up for flow across the channel; scenarios disu-d-cf-vo-s and disu-d-cf-vo-x)}
 \label{fig:figures}
@@
 	\includegraphics[width=\textwidth]{../figures/disu-d-af-vs-s-head.png}
-	\caption{standard conductance-based formulation}
+	\caption{vertically staggered grid, standard conductance-based formulation}
 	\label{fig:disu-s-cc-head}
@@
 	\includegraphics[width=\textwidth]{../figures/disu-d-af-vs-x-head.png}
-	\caption{XT3D}
+	\caption{vertically staggered grid, XT3D}
 	\label{fig:disu-x-cc-head}
@@
 The standard formulation isn't horrible, and XT3D does very well except at the channel boundary, where there's presumably a "staircase" effect.
-(preliminary result from disu\_approach model, 5x11 channel with 5x11 subdomains above and below, with cross-connections, $30^{\circ}$ incline, set up for flow along the channel; scenarios disu-d-af-vs-s and disu-d-af-vs-x)}
-\label{fig:figures}
-\end{figure}
-
-\begin{figure}[H]
-\centering
-\begin{subfigure}{0.4\textwidth}
-	\includegraphics[width=\textwidth]{../figures/disu-d-cf-vs-s-head.png}
-	\caption{standard conductance-based formulation}
-	\label{fig:disu-s-cc-cf-head}
-\end{subfigure}
-\hfill
-\begin{subfigure}{0.4\textwidth}
-	\includegraphics[width=\textwidth]{../figures/disu-d-cf-vs-x-head.png}
-	\caption{XT3D}
-	\label{fig:disu-x-cc-cf-head}
-\end{subfigure}
-\caption{Specific discharge and head for a vertically staggered grid with the domain and with cross-connections and boundary heads set up for crossflow.
-The standard formulation is much improved.
-You can't tell because of the scale of the vectors, but XT3D does well mid-channel.
-However, near the channel boundary it still has huge flows due to interpolating heads from the domain.
-(preliminary result from disu\_approach model, 5x11 channel with 5x11 subdomains above and below, $30^{\circ}$ incline, set up for flow across the channel; scenarios disu-d-cf-vs-s and disu-d-cf-vs-x)}
+(preliminary result from disu\_approach model, 5x11 channel with 5x11 subdomains above and below, with cross-connections, $30^{\circ}$ incline, set up for flow along the channel; scenarios disu-d-af-vs-s and disu-d-af-vs-x) {\color{red} AMP: Combine with fig 4.} {\color{red} AMP: Explain boundary anomalies.}}
 \label{fig:figures}
@@
 
-XXXXX The concepts in this paper have been easiest to discuss in terms of layers, but they're not limited to layered grids. XXXXX
+{\color{red} AMP: Need conclusions.}
+
+{\color{red} AMP: The concepts in this paper have been easiest to discuss in terms of grid layers, but they're not limited to layered grids.}
 
 \section{Acknowledgments}
-Thank all those reviewers.
+{\color{red} AMP: Need acknowledgments.}
 
@@
 \section{preliminary results}
-This temporary section is a summary of preliminary results that can help provide context and guide our thinking as we construct the story.  Specific results may or may not make it into the paper.
+{\color{red} AMP: This temporary section is a summary of preliminary results that can help provide context and guide our thinking as we construct the story.
+Specific results may or may not make it into the paper.}
 
@@
 What happens if the flow is across the channel rather than along it?
-Added scenarios to the disu\_approach notebook to simulate this.
-Haven't analyzed the results in detail, but using cross-connections together with xt3d appears to help in this case, as well.
+Using cross-connections together with xt3d appears to help when only the channel is simulated (no domain).
+With the domain present, the results are thrown off considerably by the lack of a ``heterogeneity correction" in the xt3d interpolation, though the errors may be concentrated near the channel boundary.
+Will not get too much into the crossflow case in this paper, but could mention the heterogeneity correction as ongoing work.
+
+\begin{figure}[H]
+\centering
+\begin{subfigure}{0.4\textwidth}
+	\includegraphics[width=\textwidth]{../figures/disu-cf-vo-s-head.png}
+	\caption{standard conductance-based formulation}
+	\label{fig:disu-s-nocc-cf-head.}
+\end{subfigure}
+\hfill
+\begin{subfigure}{0.4\textwidth}
+	\includegraphics[width=\textwidth]{../figures/disu-cf-vo-x-head.png}
+	\caption{XT3D}
+	\label{fig:disu-x-nocc-cf-head}
+\end{subfigure}
+\caption{Specific discharge and head for a vertically offset grid and boundary heads set up for crossflow.
+The standard formulation is a fail, but XT3D nails it, even without cross-connections.
+(preliminary result from disu\_approach model, 5x11 grid without cross-connections, $30^{\circ}$ incline, set up for flow across the channel; scenarios disu-cf-vo-s and disu-cf-vo-x)}
+\label{fig:figures}
+\end{figure}
+
+\begin{figure}[H]
+\centering
+\begin{subfigure}{0.4\textwidth}
+	\includegraphics[width=\textwidth]{../figures/disu-cf-vs-s-head.png}
+	\caption{standard conductance-based formulation}
+	\label{fig:disu-s-cc-cf-head}
+\end{subfigure}
+\hfill
+\begin{subfigure}{0.4\textwidth}
+	\includegraphics[width=\textwidth]{../figures/disu-cf-vs-x-head.png}
+	\caption{XT3D}
+	\label{fig:disu-x-cc-cf-head}
+\end{subfigure}
+\caption{Specific discharge and head for a vertically staggered grid with cross-connections and boundary heads set up for crossflow.
+The standard formulation does pretty well, and XT3D nails it.
+(preliminary result from disu\_approach model, 5x11 grid with cross-connections, $30^{\circ}$ incline, set up for flow across the channel; scenarios disu-cf-vs-s and disu-cf-vs-x)}
+\label{fig:figures}
+\end{figure}
+
+\begin{figure}[H]
+\centering
+\begin{subfigure}{0.4\textwidth}
+	\includegraphics[width=\textwidth]{../figures/disu-d-cf-vo-s-head.png}
+	\caption{standard conductance-based formulation}
+	\label{fig:disu-s-nocc-cf-head.}
+\end{subfigure}
+\hfill
+\begin{subfigure}{0.4\textwidth}
+	\includegraphics[width=\textwidth]{../figures/disu-d-cf-vo-x-head.png}
+	\caption{XT3D}
+	\label{fig:disu-x-nocc-cf-head}
+\end{subfigure}
+\caption{Specific discharge and head for a vertically offset grid with the domain and boundary heads set up for crossflow.
+The standard formulation gives nearly vertical flow in the channel, and XT3D also does poorly, even mid-channel.
+The huge specific discharge vectors near the channel boundary are apparently caused by XT3D's head-gradient interpolation drawing heads from the domain, where the head gradient is huge, when computing nominally horizontal flows between cells within the channel.
+(preliminary result from disu\_approach model, 5x11 channel with 5x11 subdomains above and below, without cross-connections, $30^{\circ}$ incline, set up for flow across the channel; scenarios disu-d-cf-vo-s and disu-d-cf-vo-x)}
+\label{fig:figures}
+\end{figure}
+
+\begin{figure}[H]
+\centering
+\begin{subfigure}{0.4\textwidth}
+	\includegraphics[width=\textwidth]{../figures/disu-d-cf-vs-s-head.png}
+	\caption{standard conductance-based formulation}
+	\label{fig:disu-s-cc-cf-head}
+\end{subfigure}
+\hfill
+\begin{subfigure}{0.4\textwidth}
+	\includegraphics[width=\textwidth]{../figures/disu-d-cf-vs-x-head.png}
+	\caption{XT3D}
+	\label{fig:disu-x-cc-cf-head}
+\end{subfigure}
+\caption{Specific discharge and head for a vertically staggered grid with the domain and with cross-connections and boundary heads set up for crossflow.
+The standard formulation is much improved.
+You can't tell because of the scale of the vectors, but XT3D does well mid-channel.
+However, near the channel boundary it still has huge flows due to interpolating heads from the domain.
+(preliminary result from disu\_approach model, 5x11 channel with 5x11 subdomains above and below, $30^{\circ}$ incline, set up for flow across the channel; scenarios disu-d-cf-vs-s and disu-d-cf-vs-x)}
+\label{fig:figures}
+\end{figure}
 
@@
 Results with the surrounding "domain" (from an updated disu\_approach notebook) are similar but not identical, apparently due to the "staircase" effect.
-Might be best to present most of our results without the domain, at least when illustrating the main problem and its solution (to keep the "theory" clear), then discuss what happens when the domain is present (which is more like in practice).
+Original thought was to present most of our results without the domain, at least when illustrating the main problem and its solution (to keep the "theory" clear), then discuss what happens when the domain is present (which is more like in practice).
+Kerry has suggested the results with the domain are sufficient.
 

### 2023-10-17 latex/article.tex
Kerry & Jim's updates
Incorporate Kerry & Jim's updates to paper and scripts

@@
 Constant-head boundary conditions at the ends of the channel were set to induce uniform flow along the channel in the corresponding analytical solution.
-The channel and surrounding domain were discretized using either a rectilinear grid, which satisfies the CVFD requirement, or a ``flexible triangular" grid that is unstructured with triangular cells and does not satisfy the CVFD requirement.
-They found that the standard, conductance-based flow formulation gave accurate results on the rectilinear grid but was subject to significant error on the unstructured grid, whereas the XT3D flow formulation gave accurate results on both grids, as expected.
+The channel and surrounding domain were discretized using either a rectilinear grid, which satisfies the CVFD requirement, or a ``flexible triangular" grid that is unstructured with triangular cells that conform to the channel boundary but does not satisfy the CVFD requirement.
+They found that the standard, conductance-based flow formulation gave accurate results on the rectilinear grid using isotropic hydraulic conductivity tensors only {\color{red} AMP: intended emphasis of ``only"?} but was subject to significant error on the unstructured grid, whereas the XT3D flow formulation gave accurate results on both grids for both isotropic and anisotropic hydraulic conductivity tensors, as expected.
 However, in the transect (cross-sectional) version of their benchmark, which used a vertically offset grid to discretize the nearly impermeable domain and a permeable channel (hydrogeologic layer) inclined by $30^{\circ}$ from the horizontal, similarly poor results were observed with and without XT3D.
@@
 Finally, we discuss the potential implications of our findings for practical groundwater models that include high permeability contrasts and steeply dipping layers.
-
-{\color{red} AMP: terminology -- ``channel" vs ``hydrogeologic layer"? ``hydrogelogic layer" vs ``model layer"?}
-
-{\color{red} AMP: Random thoughts
-\begin{itemize}
-	\item Vertical offset happens even on a completely rectilinear grid if the water table elevation varies in horizontally adjacent cells.
-Doesn't seem like the same kind of issue, but think on it.
-But not now, I think.
-	\item Could mention the crossflow case briefly in this paper, like \cite{bardot2022} mentioned this work as something we were looking into at the time.
-	\item Could try making the channel less permeable than the domain; semi-confining unit. Thinking not now.
-	\item Difference between truly single-layer model of channel vs embedded single-layer channel -- no vertical connections vs vertical connections that are no-flux.
-Effect on specific discharge calculations?
-Relevance of the specific-discharge "patch"?
-But beyond that, Kerry suggested that it's important to investigate and describe the embedded single-layer case, as it's a common use case, and she did some runs -- email of 2022-09-29.
-With or without xt3d, the results were not good, and I'm not sure it's possible to get good results, since there's no way to introduce cross-connections within a single-layer channel.
-Would be good to explain this in the paper.
-(She also tried a vertical gradient for the embedded single-layer case, also with poor results, though this might also have something to do with the cross-flow, "heterogeneity correction" issue.)
-\end{itemize}
-}
 
@@
 The cells have horizontal tops and bottoms and can therefore follow the dip of the channel only on average, in what \cite{bardot2022} refer to as ``staircase" fashion.
-If the grid is represented using a layered grid type (DIS or DISV) in MODFLOW 6, a cell is hydraulically connected to each cell in the same layer with which it has an overlapping face and with each cell in an adjacent layer with which it shares a horizontal face, as indicated by the blue lines that connect cell centers in figure \ref{fig:schem-conn-area-flux}a.
+If the grid is represented using a layered grid type (DIS or DISV) in MODFLOW 6, a cell is hydraulically connected to each cell in the same layer with which it has an overlapping face and with each cell in an adjacent layer with which it shares a horizontal face {\color{red} (JM: Maybe state that this is not implicit in the formulation - that faces have to overlap.
+From my experience with people modelling faults...)}, as indicated by the blue lines that connect cell centers in figure \ref{fig:schem-conn-area-flux}a.
 Following \cite{bardot2022}, in this work we call this type of grid ``vertically offset" because the tops and bottoms of adjacent cells in the same layer are not necessarily aligned.
@@
 In this work, ``vertically offset" and ``vertically staggered" are used to distinguish grids that lack nominally ``horizontal" connections between cells in adjacent layers (red lines in figure \ref{fig:schem-conn-area-flux}a) from those that include such connections, respectively.
-In the vertically offset grid in \ref{fig:schem-conn-area-flux}a (blue connections only), the interfacial area across which flow occurs between adjacent cells in the same layer is based on the vertical cell thickness (the entire thickness of the channel).
+In the vertically offset grid in figure \ref{fig:schem-conn-area-flux}a (blue connections only), the interfacial area across which flow occurs between adjacent cells in the same layer is based on the vertical cell thickness (the entire thickness of the channel).
 In the vertically staggered grid (blue and red connections), the interfacial area for nominally ``horizontal" flow between adjacent cells is based on the area over which the two cell faces overlap.
@@
 
-In MODFLOW 6, the default method for representing the flow between two hydraulically connected model cells is the ``conductance-based flow formulation" \citep{modflow6gwf}.
+The default method for representing the flow between two hydraulically connected model cells in MODFLOW 6 is the ``conductance-based flow formulation" \citep{modflow6gwf}.
 In this formulation, which is based on the commonly used two-point flux approximation {\color{red} *** ref?
@@
 This is because strictly rectilinear grids satisfy the ``control-volume finite-difference (CVFD) requirement" that the straight-line connection between two cell centers must intersect the midpoint of the cell interface at a right angle \citep{narasimhan1976integrated}.
-Vertically offset grids, however, violate the CFVD requirement because nominally ``horizontal" connections between cell centers in the same layer are not strictly horizontal and therefore not perpendicular to the vertical cell interfaces they intersect.
-For example, in figure \ref{fig:schem-conn-area-flux}a, the connection between cells 1A and 1B is not perpendicular to the cell interface, which compromises the accuracy of the conductance-based formulation.
+However, vertically offset grids violate the CFVD requirement because nominally ``horizontal" connections between cell centers in the same layer are not strictly horizontal and therefore not perpendicular to the vertical cell interfaces they intersect.
+For example, in {\color{red} AMP: capitalize ``Figure" throughout} Figure \ref{fig:schem-conn-area-flux}a, the connection between cells 1A and 1B is not perpendicular to the cell interface, which compromises the accuracy of the conductance-based formulation.
 
@@
 However, the ability to use unstructured grids in MODFLOW-USG and MODFLOW 6 introduced the potential to violate the CVFD requirement, and thereby render the conductance-based formulation less accurate, even in the absence of vertical offsets.
-To improve accuracy on unstructured grids, and to allow accurate simulation of arbitrarily oriented two- or three-dimensional anisotropy, the optional XT3D flow formulation \citep{modflow6xt3d} was introduced in MODFLOW 6.
-XT3D formulates the flow between two cells by interpolating head values from the two cells and their neighboring cells to construct an estimate of the full, two- or three-dimensional head-gradient vector on each side of the cell interface; applying the cell conductivity tensors to obtain an estimate of the groundwater flux vector on each side of the interface; and reconciling the two flux estimates to ensure continuity of flow across the interface.
+{\color{red} (JM: Elude to ghost nodes?)} To improve accuracy on unstructured grids, and to allow accurate simulation of arbitrarily oriented two- or three-dimensional anisotropy, the optional XT3D flow formulation \citep{modflow6xt3d} was introduced in MODFLOW 6.
+XT3D formulates the flow between two cells by interpolating head values from the two cells and their neighboring cells to construct an estimate of the full, two- or three-dimensional head-gradient vector on each side of the cell interface; applying the cell conductivity tensors to obtain an estimate of the groundwater flux  normal to the interface on each side of the interface; and reconciling the two flux estimates to ensure continuity of flow across the interface.
 By performing flow calculations using Darcy's Law in its tensorial form, based on a ``multi-point" approximation of the head gradient, XT3D accounts for both arbitrarily oriented anisotropy and geometric irregularity of the grid.
@@
 The vertically staggered grid provides such a mechanism by offering additional connections between cells in adjacent layers.
-Vertical flows from the bottom layer to the top layer, represented by vertical blue vectors in figure \ref{fig:schem-conn-area-flux}b, can be returned to the bottom layer by nominally ``horizontal" flows, represented by red vectors in figure \ref{fig:schem-conn-area-flux}b.
+Vertical flows from the bottom layer to the top layer, represented by vertical blue vectors in Figure \ref{fig:schem-conn-area-flux}b, can be returned to the bottom layer by nominally ``horizontal" flows, represented by red vectors in figure \ref{fig:schem-conn-area-flux}b.
 Although this concept has been illustrated using a two-layer model for simplicity, the same reasoning applies given any number of layers.
@@
 
-\section{Description of Test Problems}
+\section{Description of Test Problem}
 
-The test problems are patterned after the cross-sectional (transect) benchmarks of \cite{bardot2022}.
-They attempt simulate uniform flow in a permeable channel embedded in a less permeable surrounding ``domain." The channel is of uniform width and hydraulic conductivity and is inclined relative to the horizontal.
+The test problem is patterned after the cross-sectional (transect) benchmark of \cite{bardot2022}.
+It attempts to simulate uniform flow in an inclined permeable channel embedded in a less permeable surrounding ``domain." The channel is of uniform width and hydraulic conductivity and is inclined relative to the horizontal.
 
-To simulate flow along a channel inclined at angle $\theta$, heads are specified at the centers of cells along the perimeter of the model {\color{red} (AMP: this assumes we're using the transect\_benchmark version of the model, which includes the "domain," for all of our final runs)} using an analytical solution that corresponds to a unit head gradient inclined at angle $\theta$ \citep{bardot2022}:
-
+To simulate flow along a channel inclined at angle $\theta$ ($30^{\circ}$ for the base case), heads are specified at the centers of cells along the perimeter of the model using an analytical solution that corresponds to a unit head gradient:
 \begin{equation}
@@
 \noindent where $h$ is head in meters and $x$ and $z$ are the horizontal and vertical model coordinates, respectively, in meters.
-The hydraulic conductivity of the channel is set to 1 m/d (meter per day) so that the analytical flow solution is a groundwater flux of 1 m/d either along or across the channel.
-The conductivity of the domain surrounding the channel varies according to the scenario being investigated.
-In the homogeneous limit, the conductivity of the domain is set equal to that of the channel.
-In the highly heterogeneous limit, the conductivity of the domain is set much lower than that of the channel to effectively isolate the channel hydraulically.
+The hydraulic conductivity of the channel is set to 1 m/d (meter per day) so that the analytical flow solution is a groundwater flux of 1 m/d along the channel.
+In the base case, the conductivity of the domain is $10^{-6}$ m/d to effectively isolate the channel hydraulically.
 
-Simulations are performed using variations on two test problems that differ in their method of discretization.
-The first test problem uses the traditional approach of representing the inclined channel using a vertically offset grid of the MODFLOW 6 DIS (structured) type.
-{\color{red} (AMP: present results for the DIS grid, or for the equivalent of the DIS grid after conversion to DISU?)} This problem confirms the findings of \cite{bardot2022} and illustrates the limitations of the vertically offset grid approach.
-The second test problem is similar to the first but uses a vertically staggered grid of MODFLOW 6 DISU (unstructured) type.
-The second problem is designed to evaluate the effectiveness of cross-connections between layers (illustrated in figure \ref{fig:schem-conn-area-flux}c) in improving the accuracy of the flow solution, both with and without XT3D.
-Variations on both test problems are described in the ``Results and Discussion" section.
+Two methods of discretization are used to simulate the test problem: (1) vertically offset grid of the MODFLOW 6 DIS (structured) type, and (2) vertically staggered grid of MODFLOW 6 DISU (unstructured) type.
+The cell geometry in these two discretization methods is identical, but the cell connectivity is different.
+The vertically staggered grid included additional ``cross-connections" between cells in adjacent model layers, as described later in the paper.
+The input files for a vertically staggered scenario are formulated by starting with a DIS grid, converting it to an equivalent DISU grid, and adding cross-connections.
+The base case uses grids consisting of 11 columns and 9 layers of cells, with 3 layers within the channel, 3 layers above the channel, and 3 layers below the channel.
 
-Jupyter notebooks for both test problems are available in the Supporting Information that accompanies this paper.
+The above description represents the base case test problem.
+However, the results also include simulations designed to investigate the effects of varying the gridding resolution, the dip angle of the channel, and the conductivity contrast between the channel and the domain.
+
+Jupyter notebooks created for the test problem are available in the Supporting Information that accompanies this paper.
 The notebooks leverage the capabilities of FloPy \citep{bakker2016scripting} to assist in setting up input for and processing results from the MODFLOW 6 simulations.
-The second test problem is formulated by starting with a DIS grid, converting it to an equivalent DISU grid, and adding cross-connections.
-A Python script for converting and modifying the grid is included.
+A Python script for converting structured, DIS grids to unstructured, DISU grids is included.
 
@@
 
-{\color{red} AMP: Note that there's a temporary ``preliminary results" section at the end of the manuscript.}
+The test problem allows the numerical solution for head and fluxes to be compared with an analytical solution.
+This section presents results for the base case scenario when comparing gridding strategy (i.e.
+vertically offset versus vertically staggered) and flow solution formulation (i.e.
+standard conductance based formulation versus XT3D option) to test the hypothesis that a vertically staggered grid produces a more accurate solution for dipping layers than a vertically offset grid.
+The base case scenario is then modified to further investigate other aspects including grid resolution, dip angle, conductivity contrast and anisotropy.
 
-{\color{red} AMP: Obviously need to substitute in our "real" results with "real" discussion and captions here.}
+Results for flux magnitude, flux direction and volumetric flowthrough are used to assess the accuracy of the numerical representations of the test problem.
+Note that the terms ``flux"  and ``specific discharge" are used interchangely.
+Reported fluxes for each model are for the central model cell and therefore uninfluenced by the model boundary or the interface between channel and domain.
+Volumetric flow through the channel is therefore also examined as it encompasses fluxes through the entire channel.
+This paper uses the sum of the flows reported for the CHD cells along the left end of the channel as the volumetric flow.
 
-{\color{red} AMP: Based on the release notes, looks like the specific-discharge ``patch" went into v6.3.0. Mention it? Note that here is still the issue of coordinate dependence.}
+\subsection{Gridding Strategy and Flow Formulation Comparison}
 
-{\color{red} AMP: Results without the domain are presented first to "cleanly" illustrate the main problem and its solution.
-Results with the domain are then presented to give more of an indication of how things might work in practice.
-Kerry has suggested that perhaps the cases with the domain are sufficient -- let's consider.}
+The numerical solution for the vertically offset grid confirms the findings of \cite{bardot2022} in that fluxes are predominantly horizontal (top row, Figure \ref{fig:fig2}).
+We see that despite an enforced hydraulic gradient of $30^{\circ}$, the vertical component of the flux cannot propogate through the channel given that cells along the channel top are hydraulically disconnected from the low permeability domain, hence forcing flow horizontally (red arrows on right panel).
+The numerical solution for this grid type is a problem of cell connectivity, and thus the XT3D formulation does not resolve this issue (second row, Figure \ref{fig:fig2}).
+Results for the vertically staggered grid with the standard conductance based formulation shows an improvement in the flow solution (third row, Figure \ref{fig:fig2}), with the XT3D formulation successfully reproducing the analytical solution (bottom row, Figure \ref{fig:fig2}).
+We see that the staggered grid cell connections allows incorporation of the vertical flow component, thus permitting cross-connection between model layers.
 
-{\color{red} AMP: Unless stated otherwise, the domain is 6 orders of magnitude less conductive than the channel, as in \cite{bardot2022}.}
+Through working on this test problem, we have encountered some nuances in working with specific discharges calculated using MODFLOW.
+There are two main MODFLOW output files pertaining to fluxes.
+The first is an array of face flows which specifies the normal flux across each cell face (``FLOW-JA").
+The second is the cell-averaged specific discharge for each cell in its directional components ($q_{x}$, $q_{y}$, $q_{z}$) (``DATA-SPDIS"), and is calculated using distance-weighted averages of the cell face flows.
+Therefore, the specific discharge magnitude and direction normally reported by MODFLOW for cells within and along the channel edge are ``dampened" due to averaging of the face flows from the cell top and bottom.
+The result is a ``staircase" effect whereby these edge cells have a reduced flux magnitude and ``flattened" flux directions, as shown in Figure 3c of \cite{bardot2022}.
+The plots in Figure 2 display corrected cell-averaged specific discharges which exclude including the near-zero face flows along the channel boundary, and therefore represents the actual simulated flux magnitude and direction within the cell.
 
-{\color{red} AMP: Add and discuss results for how the flow error varies with discretization.  See plots in Kerry's emails of 2022-10-06 and 2022-12-14.}
+%Although correcting the cell-averaged specific discharge makes sense for this theoretical test problem, we must then consider the potential impact, if any, of the standard post-processing procedure used by MODFLOW to derive cell-averaged specific discharge.
+Applications, such as MODPATH (ref), which use cell face flows to interpolate velocities are exempt from this issue.
+However, applications such as MT3D (ref) which bases a dispersion tensor on the cell specific discharge, may incur some error in transport simulations, although it is anticpated this would only occur in extreme scenarios where solute is moving against a highly impermeable dipping boundary.
+The only other application of the cell averaged specific discharge is for plotting groundwater flow which is of minor importance.
+{\color{red} Probably a bit peripheral for this}.
 
-{\color{red} AMP: Add and discuss results for how the flow error varies with dip angle. See plots in Kerry's email of 2022-12-14. "Spike" at 45 degrees due to "disconnection?"}
+\begin{figure}[p!]
+	\begin{center}
+	\includegraphics[scale=0.8]{../figures/fig2_paper.png}
+	\caption{Numerical results for the test problem using base case settings for vertically offset and vertically staggered grids, with the standard conductance based formulation as well as the XT3D formulation.
+The left panel of each scenario shows the calculated specific discharge at each cell centre (black arrows) with head contours (orange).
+The right panel shows the face flows (red arrows).}
+	\label{fig:fig2}
+	\end{center}
+\end{figure}
 
-{\color{red} AMP: Add and discuss results for how the flow error varies with K contrast between channel and domain (heterogeneity).  See plots in Kerry's email of 2022-12-14.}
+\subsection{Grid Resolution}
 
-{\color{red} AMP: For consistency, should choose one notebook or the other for all runs.}
+{\color{red} Should we also consider the horizontal discretisation too?}
 
-\subsection{Test Problem 1: Vertically Offset Discretization}
+\begin{figure}
+	\begin{center}
+	\includegraphics[scale=0.9]{../figures/fig3paper.png}
+	\caption{Graphs indicating how increasing the number of model layers within a single hydrogeological layer (in this case, the channel) affects (a) flux magnitude of middle cell, (b) flux direction of middle cell and (c) volumetric flow through the channel.
+We immediately notice the errors vertically offset grids incur (blue line), and how using a vertically staggered grid reproduces the analytical solution from two model layers used to discretise the channel (orange).}
+	\label{fig:fig3}
+	\end{center}
+\end{figure}
 
-\begin{figure}[H]
+The base scenario arbitrarily uses 3 model layers to represent the channel, but it is important to address the question: \emph{How many model layers are required to adequately represent a hydrogeological layer?}.
+Vertically offset grids, which have been used by groundwater modellers for decades (ref?) {\color{red} Jim to check, I think there is some refinement guidabce for aquitards}, are almost always configured as one model layer per hydrogeological unit.
+However, although computationally efficient, we investigate the effect of altering resolution within the channel by using the base case and varying number of cells per channel width from one to nine (Figure \ref{fig:fig3}).
+We compare modelled flux magnitude and direction in the centre cell, as well as the volumetric flow against the analytical solution.
+
+Flux magnitude error for vertically offset grids peaked at around 10\% but improved and tended to zero with refinement.
+Flux direction for vertically offset grids is clearly a major issue, being reporteded as horizontal for one model layer per channel width, and only improved to 12$^{\circ}$ for 9 layers per channel width.
+Flow through the channel peaked at 13\% error but decreased to only a 4\% error for 9 layers per channel width.
+
+The results for the vertically staggered grid clearly shows that a minimum of two layers per channel width is required to match the analytical result for flux magnitude, direction and volumetric flow.
+This is an important finding as it demonstrates that computational efficiency can be made using a staggered grid over a rectilinear grid where parameters are overlayed, but it is important to use a minimum of two model layers per hydrogeological unit to allow vertical fluxes to be incorporated into the flow solution.
+
+\subsection{Dip angle and conductivity contrast}
+
+The base case scenario uses a dip angle of 30$^{\circ}$ and an extreme conductivity contrast between channel and domain of $1:10^{-6}$.
+Here, we investigate the behaviour of the flow solution for a wide range of channel inclines by systematically changing the dip angle by 2.5$^{\circ}$ from 0 to 80 $^{\circ}$.
+Similarly, we examine the solution for multiple conductivity contrasts by changing the domain conductivity to 2, 5, 10 and 100 times less than the channel.
+Simulated flux magnitude and direction of the centre cell for all combinations are plotted against the analytical solution in Figure \ref{fig:fig4}.
+
+Traditional vertically offset grids without XT3D shows rapid diversion of flux magnitude and direction from the analytical solution as dip increases (Figure \ref{fig:fig4a}).
+Use of the XT3D option keeps the flux magnitude somewhat on-track until about 30 $^{\circ}$ when the solution starts peeling away from the analytical.
+Flux direction is improved for the homogeneous scenario, but not for the heterogeneous scenarios.
+The vertically staggered grid on the other hand, even without XT3D, roughly follows the analytical solution for flux magnitude and direction until around 65 $^{\circ}$ when it starts to ``fall-apart." There are also spikes at certain dip angles which are related to transitions between cell connections due to changing cell face overlaps.
+The vertically staggered grid with XT3D clearly boasts the best results with the flux magnitude and direction matching the analytical until 65 $^{\circ}$, with the exception of a slight decrease in flux direction after 45 $^{\circ}$.
+
+These results confirm that vertically staggered grids with XT3D not only outperform the alternatives when modelling dipping hydrogeological unit, but come extremely close to the analytical solution.
+
+\begin{figure}[p!]
 \centering
-\begin{subfigure}{0.4\textwidth}
-	\includegraphics[width=\textwidth]{../figures/disu-af-vo-s-head.png}
+\begin{subfigure}{0.9\textwidth}
+	\includegraphics[width=\textwidth]{../figures/fig4_0_paper.png}
 	\caption{vertically offset grid, standard conductance-based formulation}
-	\label{fig:disu-s-nocc-head}
+	\label{fig:fig4a}
 \end{subfigure}
-\hfill
-\begin{subfigure}{0.4\textwidth}
-	\includegraphics[width=\textwidth]{../figures/disu-af-vo-x-head.png}
+\begin{subfigure}{0.9\textwidth}
+	\includegraphics[width=\textwidth]{../figures/fig4_1_paper.png}
 	\caption{vertically offset grid, XT3D}
-	\label{fig:disu-x-nocc-head}
+	\label{fig:fig4b}
 \end{subfigure}
-\caption{Specific discharge and head for a vertically offset grid.
-As Kerry and Jim found previously, the news is all bad; the flow tends to go horizontal, and XT3D doesn't really help.
-(preliminary result from disu\_approach model, 5x11 grid without cross-connections, $30^{\circ}$ incline, set up for flow along the channel; scenarios disu-af-vo-s and disu-af-vo-x)}
+\begin{subfigure}{0.9\textwidth}
+	\includegraphics[width=\textwidth]{../figures/fig4_2_paper.png}
+	\caption{vertically staggered grid, standard conductance-based formulation}
+	\label{fig:fig4c}
+\end{subfigure}
+\begin{subfigure}{0.9\textwidth}
+	\includegraphics[width=\textwidth]{../figures/fig4_3_paper.png}
+	\caption{vertically staggered grid, XT3D}
+	\label{fig:fig4d}
+\end{subfigure}
+
+\caption{Flux magnitude and direction for centre cells for varying contrasts in log conductivity (blue, orange, green, red lines) for varying dip angles (x-axis).
+The analytical value is shown in black dashed line.}
 \label{fig:figures}
@@
 
-\subsection{Test Problem 2: Vertically Staggered Discretization}
+\subsection{Anisotropy}
 
-\begin{figure}[H]
-\centering
-\begin{subfigure}{0.4\textwidth}
-	\includegraphics[width=\textwidth]{../figures/disu-af-vs-s-head.png}
-	\caption{vertically staggered grid, standard conductance-based formulation}
-	\label{fig:disu-s-cc-head}
-\end{subfigure}
-\hfill
-\begin{subfigure}{0.4\textwidth}
-	\includegraphics[width=\textwidth]{../figures/disu-af-vs-x-head.png}
-	\caption{vertically staggered grid, XT3D}
-	\label{fig:disu-x-cc-head}
-\end{subfigure}
-\caption{Specific discharge and head for a vertically staggered grid with cross-connections.
-The news is much better; the standard formulation is much improved, and XT3D nails it.
-(preliminary result from disu\_approach model, 5x11 grid with cross-connections, $30^{\circ}$ incline, set up for flow along the channel; scenarios disu-af-vs-s and disu-af-vs-x)  {\color{red} AMP: If keeping, combine with fig 3.}}
-\label{fig:figures}
-\end{figure}
+The base case and subsequent scenarios have assumed isotropic conductivity.
+However, sedimentary layers often exhibit anisotropic conductivity, and therefore we examine the gridding strategies using anisotropic and dipping conductivity tensors, with reduced conductivity perpendicular to the channel by 10, 100, 1000 and 10,000.
+We also include the isotropic scenario (ratio of 1).
+Results are presented in Figure \ref{fig:fig5}.
+Vertically offset grids produce flux magnitude and flow results that significantly deviate from the analytical solution (blue line), whilst vertically staggered grids reproduce the analytical solution (orange line).
+These results verify, once again, our proposed method of using vertically staggered grids to model dipping anisotropic layers.
 
-\subsection{Test Problem 3: Vertically Offset Discretization with Domain}
-
-\begin{figure}[H]
-\centering
-\begin{subfigure}{0.4\textwidth}
-	\includegraphics[width=\textwidth]{../figures/disu-d-af-vo-s-head.png}
-	\caption{vertically offset grid, standard conductance-based formulation}
-	\label{fig:disu-s-nocc-head}
-\end{subfigure}
-\hfill
-\begin{subfigure}{0.4\textwidth}
-	\includegraphics[width=\textwidth]{../figures/disu-d-af-vo-x-head.png}
-	\caption{vertically offset grid, XT3D}
-	\label{fig:disu-x-nocc-head}
-\end{subfigure}
-\caption{Specific discharge and head for a vertically offset grid with the domain.
-Results are similar to those without the domain and consistent with what Kerry and Jim found previously; the flow tends to go horizontal, and XT3D doesn't really help.
-(preliminary result from disu\_approach model, 5x11 channel with 5x11 subdomains above and below, without cross-connections, $30^{\circ}$ incline, set up for flow along the channel; scenarios disu-d-af-vo-s and disu-d-af-vo-x)}
-\label{fig:figures}
-\end{figure}
-
-\subsection{Test Problem 4: Vertically Staggered Discretization with Domain}
-
-\begin{figure}[H]
-\centering
-\begin{subfigure}{0.4\textwidth}
-	\includegraphics[width=\textwidth]{../figures/disu-d-af-vs-s-head.png}
-	\caption{vertically staggered grid, standard conductance-based formulation}
-	\label{fig:disu-s-cc-head}
-\end{subfigure}
-\hfill
-\begin{subfigure}{0.4\textwidth}
-	\includegraphics[width=\textwidth]{../figures/disu-d-af-vs-x-head.png}
-	\caption{vertically staggered grid, XT3D}
-	\label{fig:disu-x-cc-head}
-\end{subfigure}
-\caption{Specific discharge and head for a vertically staggered grid with the domain and with cross-connections.
-The standard formulation isn't horrible, and XT3D does very well except at the channel boundary, where there's presumably a "staircase" effect.
-(preliminary result from disu\_approach model, 5x11 channel with 5x11 subdomains above and below, with cross-connections, $30^{\circ}$ incline, set up for flow along the channel; scenarios disu-d-af-vs-s and disu-d-af-vs-x) {\color{red} AMP: Combine with fig 4.} {\color{red} AMP: Explain boundary anomalies.}}
-\label{fig:figures}
+\begin{figure}
+	\begin{center}
+	\includegraphics[scale=0.9]{../figures/fig5paper.png}
+	\caption{Varying anisotropy ratios of conductivity withing the channel and the resulting (a) flux magnitude of middle cell, (b) flux direction of middle cell and (c) volumetric flow through the channel.
+}
+	\label{fig:fig5}
+	\end{center}
 \end{figure}
@@
 
-{\color{red} AMP: Need conclusions.}
+This paper presents a solution for accurately representing steeply dipping flow structures groundwater utilising MODFLOW 6.
+\citep{bardot2022} demonstrated that the flux magnitude in highly dipping scenarios was erroneous when utilising vertically offset grids.
+More significantly, the direction of fluxes were inaccurate, neglecting the vertical component of the flux.
+The issues observed in the existing study of \citep{bardot2022} partially related to the implementation of only within-layer horizontally connections, limiting the model capacity to simulate the vertical component of the flow.
+The present study developed an approach which utilised vertically staggered grids \cite{modflow6gwf} to overcome the perceived problems.
+Through comparison of the previous vertically offset and proposed vertically staggered approached to the analytical solution presented in \citep{bardot2022}, we demonstrate that the use of vertically staggered grids (in conjunction with the XT3D formulation) is an accurate method to simulate flow through dipping hydrogeological layers.
+However, accurate representation of flows requires a minimum of two model layers per hydro geological unit to allow for vertical connections to induce flows.
 
-{\color{red} AMP: The concepts in this paper have been easiest to discuss in terms of grid layers, but they're not limited to layered grids.}
+Points:
+
+1) Furthermore, we point out the nuances in specific discharge calculations and the ambiguity introduced when assessing model behaviour.  MODPATH and MT3D
+
+2) The concepts in this paper have been easiest to discuss in terms of grid layers, but they're not limited to layered grids.
+
+3) (Random thought) The use of vertically staggered grids improved the accuracy of the transect solutions; however, the results of \citep{bardot2022} using unstructured meshes that conformed to the channel geometry preformed better in isotropic grids.
+If you force the normal flux to be the same at the interface of the domain and channel, the analytical solution would suggest a uniform gradient either side, do we introduce an error if the faces don't align with the channel correctly?
+Would figure 2 show a correction for this if the anisotropic scenario was tested?
 
 \section{Acknowledgments}
-{\color{red} AMP: Need acknowledgments.}
+We would like to acknowledge....
 
@@
 
-\section{Oblique reference}
-
-\begin{figure}[H]
-	\includegraphics[scale=0.5]{../figures/oblique_reference.png}
-\end{figure}
-
-\section{preliminary results}
-{\color{red} AMP: This temporary section is a summary of preliminary results that can help provide context and guide our thinking as we construct the story.
-Specific results may or may not make it into the paper.}
-
-The models referenced are:
-\begin{itemize}
-	\item wormholes -- This is a plan-view, DISV model that emulates a cross-sectional model and includes ``connector" cells that can have sloped tops and bottoms.
-Cell dimensions and hydraulic conductivities can be adjusted to approach a ``dip-following" mesh (in which cell tops and bottoms follow the dip of the channel) at one extreme and a vertically offset mesh (the kind MODFLOW actually uses in cross-section) at another extreme.
-A vertically offset mesh is approached by making the width of the connector cells very small in the x direction and, in the case of xt3d, making them anisotropic with very high K (1000 x the base K in these runs) along the slope of the cell and very low K (0.001 x the base K) perpendicular to the slope.
-The case of a vertically offset mesh using the standard formulation, which cannot implement such anisotropy, was run but is not generally of interest.
-This model is in the repo, but results from this model probably won't be included in the paper.
-	\item crossconnected -- This is a plan-view, DISV model that emulates a cross-sectional model with ``cross-connections" between model layers.
-This model is not in the repo, and results are noted here primarily to confirm the expected agreement with the disu\_approach model.
-	\item disu\_approach -- This implements cross-connections ``for real" in a cross-sectional, DISU model.
-Specific discharge is recalculated in the notebook to account for overlap areas; the recent ``patch" to MF6 is not being used (yet).
-	\item transect\_benchmark -- This is the vertically offset model of \cite{bardot2022} updated to allow cross-connections. Results from this model are not summarized here yet.
-\end{itemize}
-
-In the model of \cite{bardot2022} and its transect\_benchmark update, the channel is embedded in a surrounding ``domain."  In each of the other three models, there is no surrounding domain -- the channel is the entire model.
-The distinction may be significant because the surrounding matrix provides vertical connections to the channel.
-When the surrounding domain is (effectively) impermeable, these vertical connections contribute ``zero vertical flux" information to the specific discharge calculation, implying that the specific discharge is literally horizontal.
-In the absence of a surrounding domain, there is no vertical flux component information.
-Need to check what difference this might make.
-
-In the plan-view models, rows function as ``layers." Unless stated otherwise, all results are for a 30-degree channel dip, and the analytical solution is a specific discharge of 1.
-inclined at 30 degrees.
-Total flow is the sum of the flows reported for the CHD cells along the left end of the channel.
-Numerical results are rounded arbitrarily.
-The terms ``specific discharge" and ``flux" are using interchangeably.
-
-\subsection{Single-layer channel}
-The channel is represented using a single layer of cells.
-
-\subsubsection{wormholes, 1x11 cells:}
-Results were the same whether the mesh was vertically offset (xt3d and standard, though not sure if standard results are meaningful) or dip-following (xt3d and standard), which is not surprising since there are no connections across the tops and bottoms.
-The errors in computed heads were miniscule in all cases.
-
-\paragraph{with xt3d:} Specific discharge is 0.8660 and ``seemingly" horizontal.
-This makes sense if you consider that the only connections in the model are ``horizontal" ones along the layer.
-A horizontal flux component of 0.8660 corresponds (via the cosine factor) to a unit flux from left to right along the dip of the channel, so in that sense it's correct.
-Furthermore, that's the only flux component that exists in this model, and therefore the only one xt3d can calculate; the vertical flux component is reported as zero by default.
-So rather than say the flux is horizontal, perhaps it's fairer to say MODFLOW gets the horizontal component right and simply can't tell us what the vertical component is.
-Total flow is 0.8660, which is correct.
-
-\paragraph{without xt3d (standard formulation)} Specific discharge is 1.1547 and, again, ``seemingly" horizontal.
-This increase in magnitude relative to xt3d makes sense because the standard formulation ignores both the angle of the connection relative to the vertical cell interface and the reduction in area for flow (area perpendicular to the long axis of the channel) when the channel dips.
-Therefore, the  magnitude differs by a factor of $1/{(cosine factor)}^2$.
-Total flow is 1.1547, which reflects the error in the flux.
-
-\subsubsection{crossconnected, 1x11 cells:} Specific discharge results are essentially identical to those in the vertically offset wormholes model, which is not surprising given that there can be no cross-connections in a single-layer model.
-The errors in computed heads were miniscule with and without xt3d.
-
-\paragraph{with xt3d:} Total flow is 0.3660, which reflects the smaller area available for flow between two cells (the overlap area).
-
-\paragraph{without xt3d (standard formulation)} Total flow is 0.4880, which (given the incorrect flux) also reflects the smaller area available for flow between two cells (the overlap area).
-
-\subsubsection{disu\_approach, 1x11 cells:} Specific discharge results are essentially identical to those in the vertically offset wormholes model and crossconnected model.
-Total flow matched either the vertically offset wormholes result or the crossconnected result depending on how the ``staggered" flag was set in the notebook, which determined whether the specific discharge remained based on average areas or was recalculated using overlap areas.
-(There are no cross-connections in a single-layer model, so the cell-cell flows were unaffected, so it was just a matter of whether the fluxes were postprocessed to adjust the areas.) The errors in computed heads were miniscule with and without xt3d.
-
-\subsection{Two-layer channel}
-The channel is represented using two layers of cells. This introduces the possibility of vertical flows between layers.
-
-\subsubsection{wormholes, 2x11 cells}
-
-\paragraph{dip-following mesh with xt3d:} In the connector cells (which are of primary interest in this case), specific discharge is 0.9998 mid-channel and 1.00003 on average, with a max error of +0.0003.
-Angle of specific discharge is 29.9913 deg mid-channel and 29.9879 on average, with a max error of +0.03 deg.
-Errors in computed heads (including all cells) are on the order of 0.002\%.
-Total flow is 1.7321 (0.8661 per cell).
-So, good agreement with the analytical solution.
-
-\paragraph{dip-following mesh without xt3d (standard formulation)} In the connector cells (which are of primary interest in this case), specific discharge is 1.3336 mid-channel and 1.3687 on average, with a max error of +0.69.
-Angle of specific discharge is 30.05 deg mid-channel and 32.24 on average, with a max error of +9.8 deg.
-Errors in computed heads (including all cells) are on the order of 2\%.
-Total flow is 2.3094 (1.1547 per cell).
-
-\paragraph{vertically offset mesh with xt3d:} In the flat-top cells (which are of primary interest in this case), specific discharge is 1.1117 mid-channel and 1.1211 on average, with a max error of +0.35.
-Angle of specific discharge is -0.53 deg mid-channel and 6.0095 on average, with a max error of -30.54 deg; the flow generally skews toward horizontal.
-Errors in computed heads (including all cells) are on the order of 4\%.
-Total flow is 2.1993 (1.0997 per cell).
-
-\subsubsection{crossconnected, 2x11 cells:}
-
-\paragraph{with xt3d:} Specific discharge matches the analytical solution very closely, with a max error of -5e-9.
-Angle of specific discharge also matches very closely, with a max error of -9e-8 deg.
-Errors in computed heads are minscule.
-Total flow is 1.2321 (0.6161 per cell), which reflects the reduced area for flow; the non-overlapping area of the lowest left-hand boundary cell is effectively ``lost." Otherwise, a great result.
-
-\paragraph{without xt3d (standard formulation)}  Specific discharge is 1.0922 mid-channel and 1.1074 on average, with a max error of +0.30.
-Angle of specific discharge is 22.69 deg mid-channel and 23.28 on average, with a max error of -7.5 deg.
-Errors in computed heads are on the order of 0.5\%.
-Total flow is 1.3942 (0.6971 per cell).
-Overall, not a great result, but not terrible; the cross-connections alone seem to have helped substantially.
-
-\subsubsection{disu\_approach, 2x11 cells:}
-
-\paragraph{without cross-connections, with xt3d:} Results are comparable to those in the wormhole model with a vertically offset mesh and xt3d.
-Specific discharge is 1.1045 mid-channel and 1.1262 on average, with a max error of +0.38.
-Angle of specific discharge is 0.0731 deg mid-channel and 6.2516 on average, with a max error of -29.93 deg; the flow generally skews toward horizontal.
-Errors in computed heads are on the order of 3\%.
-Total flow is 2.2091 (1.1045 per cell).
-
-\paragraph{without cross-connections, without xt3d (standard formulation):} Specific discharge is 1.1547 mid-channel and 1.1755 on average, with a max error of +0.43.
-Angle of specific discharge is 0.0685 deg mid-channel and 5.9963 on average, with a max error of -29.93 deg; the flow generally skews toward horizontal.
-Errors in computed heads (including all cells) are on the order of 2\%.
-Total flow is 2.3094 (1.1547 per cell).
-Overall, comparable to the results with xt3d.
-
-\paragraph{with cross-connections, with xt3d:} Results are comparable to those in the crossconnected model with xt3d.
-Specific discharge matches the analytical solution very closely, with a max error of +1e-8.
-Angle of specific discharge also matches very closely, with a max error of -3e-7 deg.
-Errors in computed heads are minscule.
-Total flow is 1.2321 (0.6161 per cell), which reflects the reduced area for flow; the non-overlapping area of the lowest left-hand boundary cell is effectively ``lost." Otherwise, a great result.
-
-\paragraph{with cross-connections, without xt3d (standard formulation):} Results are comparable to those in the crossconnected model without xt3d.
-Specific discharge is 1.0948 mid-channel and 1.1101 on average, with a max error of +0.30.
-Angle of specific discharge is 22.63 deg mid-channel and 23.22 on average, with a max error of -7.5 deg.
-Errors in computed heads are on the order of 0.5\%.
-Total flow is 1.3942 (0.6971 per cell).
-Overall, not a great result, but not terrible; the cross-connections alone seem to have helped substantially.
-
-\subsection{Five-layer channel}
-The channel is represented using five layers of cells.
-
-\subsubsection{wormholes, 5x11 cells:}
-
-\paragraph{dip-following mesh with xt3d:} Results are basically similar to those in the 2-layer case, except the total flow is greater in proportion to the greater number of boundary cells.
-
-\paragraph{dip-following mesh without xt3d (standard formulation)} Results are roughly comparable to but a bit worse overall than those in the 2-layer case.
-In the connector cells (which are of primary interest in this case), specific discharge is 1.3579 mid-channel and 1.4277 on average, with a max error of +1.14.
-Angle of specific discharge is 31.7497 deg mid-channel and 35.6924 on average, with a max error of +19.57 deg.
-Errors in computed heads (including all cells) are on the order of 8\%.
-Total flow is 5.7735 (1.1547 per cell).
-
-\paragraph{vertically offset mesh with xt3d:} Results are roughly comparable to those in the 2-layer case.
-In the flat-top cells (which are of primary interest in this case), specific discharge is 0.8906 mid-channel and 1.0013 on average, with a max error of +0.54.
-Angle of specific discharge is 2.8558 deg mid-channel and 10.9171 on average, with a max error of -31.5 deg; the flow generally skews toward horizontal.
-Errors in computed heads (including all cells) are on the order of 10\%.
-Total flow is 4.8426 (0.9685 per cell).
-
-\subsubsection{crossconnected, 5x11 cells:}
-
-\paragraph{with xt3d:} Results are basically similar to those in the 2-layer case.
-Total flow is 3.8301 (0.7660 per cell), which is better than the 2-layer result due to the ``lost" (non-overlapping) area being a smaller proportion of the total area for flow.
-
-\paragraph{without xt3d (standard formulation)}  Results are basically comparable to those in the 2-layer case, except the errors in computed heads are substantially higher at around 1.5\%, and the total flow of 4.0944 gives a somewhat higher flow per cell at 0.8189.
-Again, not great, but not terrible.
-
-\subsubsection{disu\_approach, 5x11 cells:}
-
-\paragraph{without cross-connections, with xt3d:} Results are roughly comparable to but a bit worse overall than those in the 2-layer case.
-Specific discharge is 1.0547 mid-channel and 1.0926 on average, with a max error of +0.62.
-Angle of specific discharge is 2.9627 deg mid-channel and 10.3771 on average, with a max error of -28.1 deg; the flow generally skews toward horizontal.
-Errors in computed heads are on the order of 8\%.
-Total flow is 5.3106 (1.0621 per cell).
-
-\paragraph{without cross-connections, without xt3d (standard formulation):} Results are roughly comparable to but a bit worse overall than those in the 2-layer case.
-Specific discharge is 1.1560 mid-channel and 1.1827 on average, with a max error of +0.70.
-Angle of specific discharge is 2.7571 deg mid-channel and 9.5527 on average, with a max error of -28.3 deg; the flow generally skews toward horizontal.
-Errors in computed heads (including all cells) are on the order of 8\%.
-Total flow is 5.7735 (1.1547 per cell).
-Overall, comparable to the results with xt3d.
-
-\paragraph{with cross-connections, with xt3d:} Results are basically similar to those in the 2-layer case.
-Total flow is 3.8301 (0.7660 per cell), which is better than the 2-layer result due to the ``lost" (non-overlapping) area being a smaller proportion of the total area for flow.
-
-\paragraph{with cross-connections, without xt3d (standard formulation):} Results are comparable to those in the crossconnected model without xt3d. Again, not great, but not terrible.
-
-\subsection{Crossflow}
-What happens if the flow is across the channel rather than along it?
-Using cross-connections together with xt3d appears to help when only the channel is simulated (no domain).
-With the domain present, the results are thrown off considerably by the lack of a ``heterogeneity correction" in the xt3d interpolation, though the errors may be concentrated near the channel boundary.
-Will not get too much into the crossflow case in this paper, but could mention the heterogeneity correction as ongoing work.
-
-\begin{figure}[H]
-\centering
-\begin{subfigure}{0.4\textwidth}
-	\includegraphics[width=\textwidth]{../figures/disu-cf-vo-s-head.png}
-	\caption{standard conductance-based formulation}
-	\label{fig:disu-s-nocc-cf-head.}
-\end{subfigure}
-\hfill
-\begin{subfigure}{0.4\textwidth}
-	\includegraphics[width=\textwidth]{../figures/disu-cf-vo-x-head.png}
-	\caption{XT3D}
-	\label{fig:disu-x-nocc-cf-head}
-\end{subfigure}
-\caption{Specific discharge and head for a vertically offset grid and boundary heads set up for crossflow.
-The standard formulation is a fail, but XT3D nails it, even without cross-connections.
-(preliminary result from disu\_approach model, 5x11 grid without cross-connections, $30^{\circ}$ incline, set up for flow across the channel; scenarios disu-cf-vo-s and disu-cf-vo-x)}
-\label{fig:figures}
-\end{figure}
-
-\begin{figure}[H]
-\centering
-\begin{subfigure}{0.4\textwidth}
-	\includegraphics[width=\textwidth]{../figures/disu-cf-vs-s-head.png}
-	\caption{standard conductance-based formulation}
-	\label{fig:disu-s-cc-cf-head}
-\end{subfigure}
-\hfill
-\begin{subfigure}{0.4\textwidth}
-	\includegraphics[width=\textwidth]{../figures/disu-cf-vs-x-head.png}
-	\caption{XT3D}
-	\label{fig:disu-x-cc-cf-head}
-\end{subfigure}
-\caption{Specific discharge and head for a vertically staggered grid with cross-connections and boundary heads set up for crossflow.
-The standard formulation does pretty well, and XT3D nails it.
-(preliminary result from disu\_approach model, 5x11 grid with cross-connections, $30^{\circ}$ incline, set up for flow across the channel; scenarios disu-cf-vs-s and disu-cf-vs-x)}
-\label{fig:figures}
-\end{figure}
-
-\begin{figure}[H]
-\centering
-\begin{subfigure}{0.4\textwidth}
-	\includegraphics[width=\textwidth]{../figures/disu-d-cf-vo-s-head.png}
-	\caption{standard conductance-based formulation}
-	\label{fig:disu-s-nocc-cf-head.}
-\end{subfigure}
-\hfill
-\begin{subfigure}{0.4\textwidth}
-	\includegraphics[width=\textwidth]{../figures/disu-d-cf-vo-x-head.png}
-	\caption{XT3D}
-	\label{fig:disu-x-nocc-cf-head}
-\end{subfigure}
-\caption{Specific discharge and head for a vertically offset grid with the domain and boundary heads set up for crossflow.
-The standard formulation gives nearly vertical flow in the channel, and XT3D also does poorly, even mid-channel.
-The huge specific discharge vectors near the channel boundary are apparently caused by XT3D's head-gradient interpolation drawing heads from the domain, where the head gradient is huge, when computing nominally horizontal flows between cells within the channel.
-(preliminary result from disu\_approach model, 5x11 channel with 5x11 subdomains above and below, without cross-connections, $30^{\circ}$ incline, set up for flow across the channel; scenarios disu-d-cf-vo-s and disu-d-cf-vo-x)}
-\label{fig:figures}
-\end{figure}
-
-\begin{figure}[H]
-\centering
-\begin{subfigure}{0.4\textwidth}
-	\includegraphics[width=\textwidth]{../figures/disu-d-cf-vs-s-head.png}
-	\caption{standard conductance-based formulation}
-	\label{fig:disu-s-cc-cf-head}
-\end{subfigure}
-\hfill
-\begin{subfigure}{0.4\textwidth}
-	\includegraphics[width=\textwidth]{../figures/disu-d-cf-vs-x-head.png}
-	\caption{XT3D}
-	\label{fig:disu-x-cc-cf-head}
-\end{subfigure}
-\caption{Specific discharge and head for a vertically staggered grid with the domain and with cross-connections and boundary heads set up for crossflow.
-The standard formulation is much improved.
-You can't tell because of the scale of the vectors, but XT3D does well mid-channel.
-However, near the channel boundary it still has huge flows due to interpolating heads from the domain.
-(preliminary result from disu\_approach model, 5x11 channel with 5x11 subdomains above and below, $30^{\circ}$ incline, set up for flow across the channel; scenarios disu-d-cf-vs-s and disu-d-cf-vs-x)}
-\label{fig:figures}
-\end{figure}
-
-\subsection{Domain}
-Results with the surrounding "domain" (from an updated disu\_approach notebook) are similar but not identical, apparently due to the "staircase" effect.
-Original thought was to present most of our results without the domain, at least when illustrating the main problem and its solution (to keep the "theory" clear), then discuss what happens when the domain is present (which is more like in practice).
-Kerry has suggested the results with the domain are sufficient.
-
 \end{document}

### 2023-10-18 latex/article.tex
langevin edits

@@
 
-\title{Improving Accuracy of Simulated Flows in Steeply Dipping Layers Using Vertically Staggered Grids with the XT3D Multi-Point Flux Approximation in MODFLOW 6}
+%\title{Improving Accuracy of Simulated Flows in Steeply Dipping Layers Using full connectivity Grids with the XT3D Multi-Point Flux Approximation in MODFLOW 6}
+
+% suggested title revision
+\title{Efficient and Accurate Modeling of Flow through Sedimentary Structures with MODFLOW 6}
+
 \author{
@@
 
-\textbf{Key words:} groundwater flow simulation, dipping model layers, vertically staggered grid, multi-point flux approximation
+\textbf{Key words:} groundwater flow simulation, dipping model layers, full connectivity grid, multi-point flux approximation
 
@@
 Versions of MODFLOW up to and including MODFLOW-2005 \citep{modflow2005} offered only structured grids composed of rows, columns, and layers of cells.
-Cells were conceptualized as having horizontal top and bottom surfaces, which allowed each cell to be ``simulated as if it were rectangular so that flow may be approximated by the standard finite-difference equation" \citep{modflow84}.
-In MODFLOW-USG \citep{modflowusg} and MODFLOW 6 \citep{modflow6gwf}, which offer unstructured grids and use the control-volume finite-difference method to formulate the discrete balance equations, cells continue to have horizontal tops and bottoms, which simplifies the definition and calculation of water-table elevations, cell saturations, and hydraulic conductances between cells.
+Cells were conceptualized as having horizontal top and bottom surfaces and vertical lateral faces, which allowed each cell to be ``simulated as if it were rectangular so that flow may be approximated by the standard finite-difference equation'' \citep{modflow84}.
+In MODFLOW-USG \citep{modflowusg} and MODFLOW 6 \citep{modflow6gwf}, which offer unstructured grids and use the control-volume finite-difference method to formulate the discrete balance equations, cells continue to have horizontal tops and bottoms and vertical lateral faces.
+Defining cells with horizontal tops and bottoms and vertical lateral faces simplifies the definition and calculation of water-table elevations, cell saturations, hydraulic conductances between cells, and transitions between confined and unconfined conditions.
 
 A traditional approach to representing an aquifer system with dipping hydrogeologic layers in MODFLOW is to map the hydrologic properties onto a rectilinear grid \citep{modflow84}.
+\cite{hoaglund2003} refer to this as the ``grid overlay'' method.
+In the grid overlay method, the boundaries of dipping layers are represented approximately, in what \cite{bardot2022} call a ``staircase'' fashion.
 Another approach available in all versions of MODFLOW is to allow the top and bottom elevations of cells in the same layer of the grid to vary spatially.
-Following \cite{bardot2022}, in this work we call this type of grid ``vertically offset." The vertical offsets between ``horizontally" adjacent cells allows model layers to ``deform'" with the hydrostratigraphy to ``minimize the number of model layers required to simulate an aquifer system" \citep{modflow84} relative to a rectilinear grid with horizontal layers.
-In both approaches, the model grid follows the boundaries of dipping layers approximately, in what \cite{bardot2022} call a ``staircase" fashion.
+Following \cite{bardot2022}, in this work we call this type of grid ``vertically offset,'' and we show how to improve the accuracy of flow simulations that use vertically offset grids.
+The vertical offsets in top and bottom elevations between ``horizontally'' adjacent cells allows model layers to ``deform'' with the hydrostratigraphy to ``minimize the number of model layers required to simulate an aquifer system'' \citep{modflow84} relative to the grid overlay method with a rectilinear grid consisting of horizontal layers.
 
-All versions of MODFLOW prior to MODFLOW 6 used a two-point, ``conductance-based" flow formulation in which the flow between two adjacent cells is proportional to the difference between the heads calculated in those two cells.
-The conductance-based flow formulation is most accurate when the grid satisfies the ``control-volume finite-difference (CVFD) requirement" that the straight-line connection between two cell centers must intersect the midpoint of the cell interface at a right angle \citep{narasimhan1976integrated}.
-However, vertically offset grids violate the CVFD requirement because cell interfaces are vertical but cell connections are not strictly horizontal.
-For this reason, the MODFLOW 6 documentation \citep{modflow6gwf} warns that "[s]teeply dipping layers generally should not be represented" using vertical offsets.
+All versions of MODFLOW prior to MODFLOW 6 used a two-point, ``conductance-based'' flow formulation in which the flow between two adjacent cells is proportional to the difference between the heads calculated in those two cells.
+The conductance-based flow formulation is most accurate when the grid satisfies the ``control-volume finite-difference (CVFD) requirement'' that the straight-line connection between two cell centers must intersect the midpoint of the cell interface at a right angle \citep{narasimhan1976integrated}.
+However, vertically offset grids violate the CVFD requirement because lateral cell faces are vertical but cell connections are not strictly horizontal.
+For this reason, the MODFLOW 6 documentation \citep{modflow6gwf} warns that ``[s]teeply dipping layers generally should not be represented'' using vertical offsets.
 \cite{anderson2015applied} recommend a $10^{\circ}$ dip as a practical upper limit for using vertical offsets.
@@
 By performing flow calculations using Darcy's Law in its tensorial form based on a multi-point approximation of the head gradient vector, XT3D accounts for both arbitrarily oriented anisotropy and geometric irregularity of the grid.
-In theory, XT3D should be able to compensate for the geometic irregularity introduced by vertical offsets between adjacent cells in the same model layer.
+In theory, XT3D should be able to compensate for the geometric irregularity introduced by vertical offsets between adjacent cells in the same model layer.
 
-\cite{bardot2022} evaluated the effects of grid design and XT3D on the ability of MODFLOW 6 to accurately simulate flow through sedimentary structures by modeling an idealized, two-dimensional permeable channel or hydrogeologic layer embedded in a nearly impermeable surrounding medium, or ``domain." In their plan-view benchmark, the channel lay within the horizontal plane as was angled counterclockwise relative to the southern boundary of the model domain (by $30^{\circ}$ in most cases).
+\cite{bardot2022} evaluated the effects of grid design and XT3D on the ability of MODFLOW 6 to accurately simulate flow through sedimentary structures by modeling an idealized, two-dimensional permeable channel or hydrogeologic layer {\color{red} (CDL: use ``permeable feature'' instead of channel?)} embedded in a nearly impermeable surrounding medium, or ``domain.'' In their plan-view benchmark, the channel lay within the horizontal plane and was angled counterclockwise relative to the southern boundary of the model domain (by $30^{\circ}$ in most cases).
 Constant-head boundary conditions at the ends of the channel were set to induce uniform flow along the channel in the corresponding analytical solution.
-The channel and surrounding domain were discretized using either a rectilinear grid, which satisfies the CVFD requirement, or a ``flexible triangular" grid that is unstructured with triangular cells that conform to the channel boundary but does not satisfy the CVFD requirement.
-They found that the standard, conductance-based flow formulation gave accurate results on the rectilinear grid using isotropic hydraulic conductivity tensors only {\color{red} AMP: intended emphasis of ``only"?} but was subject to significant error on the unstructured grid, whereas the XT3D flow formulation gave accurate results on both grids for both isotropic and anisotropic hydraulic conductivity tensors, as expected.
+The channel and surrounding domain were discretized using either a rectilinear grid, which satisfies the CVFD requirement, or a ``flexible triangular'' grid that is unstructured with triangular cells that conform to the channel boundary but does not satisfy the CVFD requirement.
+They found that the standard, conductance-based flow formulation gave accurate results on the rectilinear grid using isotropic hydraulic conductivity tensors only (accuracy decreased substantially in the presence of anisotropy) {\color{red} AMP: intended emphasis of ``only''?
+(CDL: I believe ``only'' was intentional.
+I added the part about anisotropy effects in parentheses)} but was subject to significant error on the unstructured grid, whereas the XT3D flow formulation gave accurate results on both grids for both isotropic and anisotropic hydraulic conductivity tensors, as expected.
 However, in the transect (cross-sectional) version of their benchmark, which used a vertically offset grid to discretize the nearly impermeable domain and a permeable channel (hydrogeologic layer) inclined by $30^{\circ}$ from the horizontal, similarly poor results were observed with and without XT3D.
 Given the ability of XT3D to compensate for violations of the CVFD requirement, it appeared likely that another aspect of the problem formulation or the inner workings of MODFLOW 6 was responsible for the unexpected results.
-As noted by \cite{bardot2022}, a preliminary analysis by two authors of the present work (Provost and Langevin) suggested that the cell connectivity offered by a vertically offset grid was inadequate for allowing an accurate flow solution.
+As noted by \cite{bardot2022}, a preliminary analysis by two authors of the present work (Provost and Langevin) suggested that the cell connectivity offered by the vertically offset grid used for the problem was inadequate for allowing an accurate flow solution.
 
-"Vertically staggered" grid connections were introduced in MODFLOW-USG \citep{modflowusg} and included for the fully unstructured (DISU) grid type in MODFLOW 6 \citep{modflow6gwf} primarily as a means to vertically subdiscretize portions of a grid that are conceptualized as being layered.
-A grid that includes vertically staggered connections, which in this work we call a ``vertically staggered grid," some cells have nominally ``horizontal" connections with more than one cell across a given cell face.
+MODFLOW 6 supports three different types of grids: regular MODFLOW grids consisting of layers, rows, and columns (DIS),  grids discretized by vertices (DISV), and fully unstructured (DISU) grids, which are patterned after the DISU grid type in MODFLOW-USG  \citep{modflowusg}.
+The DIS and DISV grids in MODFLOW 6 are ``layered'', which means that the same two-dimensional grid (in plan view) applies to all model layers.
+The ``layered'' concept of DIS and DISV grids is used within MODFLOW to determine how cells are hydraulically connected.
+Cells in DIS and DISV grids are automatically assigned vertical connections to overlying and underlying cells and horizontal connections to adjacent cells in the same model layer.
+DISU grids are more flexible than DIS and DISV grids in that the user has complete freedom to specify the cell connectivity and the geometric properties for each connection.
+DISU grids support two types of horizontal connections: a ``layered'' horizontal connection and a ``vertically staggered'' horizontal connection \citep{modflow6gwf}.
+With a layered horizontal connection, the hydraulic conductance is calculated using the full saturated thickness of the two connected model cells.
+Alternatively, for a staggered horizontal connection, the hydraulic conductance is calculated using only the overlap between the two cell faces.
+Vertically staggered horizontal connections were introduced in MODFLOW-USG \citep{modflowusg} and included in MODFLOW 6 \citep{modflow6gwf} primarily as a means to vertically subdiscretize parts of a model domain.
+The grid used by \cite{bardot2022} for the steeply dipping benchmark example used layered horizontal connections.
+To our knowledge, the implications of limited horizontal connectivity in layered grids for the accuracy of simulated flows in models with steeply dipping layers, and the potential for using vertically staggered grids with full cell connectivity to overcome this limitation, have not been previously appreciated or investigated.
+
+% text removed
+% For a grid that includes vertically staggered connections, which in this work we call a ``vertically staggered grid,'' some cells have nominally ``horizontal'' connections with more than one cell in some direction.
 Vertically staggered DISU grids allow more flexibility in assigning grid connectivity than do the more commonly used DIS and DISV grid types \citep{modflow6gwf} in MODFLOW 6.
-To our knowledge, the implications of limited grid connectivity for the accuracy of simulated flows in models with steeply dipping layers, and the potential for using vertically staggered grids to overcome this limitation, have not been previously appreciated or investigated.
 
 In the remainder of this paper we first present a theoretical justification for the hypothesis that inadequate grid connectivity is primarily responsible for the inaccurate simulation of flows in the steeply dipping layer benchmark on a vertically offset grid.
-Based on that theory, we propose a method for improving the accuracy of the flow solution using XT3D together with a vertically staggered grid, and we evaluate the effectiveness of that approach in a set of benchmark problems similar to those of \cite{bardot2022}.
+Based on that theory, we propose a method for improving the accuracy of the flow solution using XT3D together with vertically offset grid with full connectivity, and we evaluate the effectiveness of that approach in a set of benchmark problems similar to those of \cite{bardot2022}.
 Finally, we discuss the potential implications of our findings for practical groundwater models that include high permeability contrasts and steeply dipping layers.
@@
 	\begin{center}
-%	\includegraphics[scale=0.8]{../figures/schem_conn_area_flux.png}
 	\includegraphics[scale=0.6]{../figures/schem_conn_area_flux.png}
@@
 In (a), black circles represent cell centers.
-Blue lines represent hydraulic connections between cells in a vertically offset grid.
-Red lines represent additional connections that result in a vertically staggered grid.
+Blue lines represent hydraulic connections between cells in a layered connectivity grid.
+Red lines represent additional connections that must be made in order to achieve full connectivity.
 In (b), gray arrows represent the uniform groundwater flux the model is attempting to simulate.
-Blue arrows represent components of the uniform groundwater flux normal to the horizontal and vertical cell-cell interfaces in the vertically offset grid.
-Red arrows represent components of the uniform groundwater flux normal to the additional vertical cell-cell interfaces introduced in the vertically staggered grid.}
+Blue arrows represent components of the uniform groundwater flux normal to the horizontal and vertical cell-cell interfaces in the layered connectivity grid.
+Red arrows represent components of the uniform groundwater flux normal to the additional vertical cell-cell interfaces introduced to represent the full connectivity.}
 	\label{fig:schem-conn-area-flux}
@@
 Figure \ref{fig:schem-conn-area-flux}a shows a group of four cells in a two-layer, cross-sectional MODFLOW 6 model of fully saturated flow through a permeable channel (aquifer) with impermeable top and bottom boundaries.
-The cells have horizontal tops and bottoms and can therefore follow the dip of the channel only on average, in what \cite{bardot2022} refer to as ``staircase" fashion.
-If the grid is represented using a layered grid type (DIS or DISV) in MODFLOW 6, a cell is hydraulically connected to each cell in the same layer with which it has an overlapping face and with each cell in an adjacent layer with which it shares a horizontal face {\color{red} (JM: Maybe state that this is not implicit in the formulation - that faces have to overlap.
-From my experience with people modelling faults...)}, as indicated by the blue lines that connect cell centers in figure \ref{fig:schem-conn-area-flux}a.
-Following \cite{bardot2022}, in this work we call this type of grid ``vertically offset" because the tops and bottoms of adjacent cells in the same layer are not necessarily aligned.
-Note that cells that have overlapping vertical faces but are in different layers (e.g., cells 1A and 2B) do not have a direct hydraulic connection in a vertically offset grid.
-The addition of such connections between cells in different layers, as indicated by the red lines that connect cell centers in figure \ref{fig:schem-conn-area-flux}a, results in a ``vertically staggered" grid.
-The ability to handle vertically staggered grids was first introduced to MODFLOW by \cite{modflowusg} in MODFLOW-USG, and the term ``vertically staggered" was first used by \cite{modflow6gwf} in the context of MODFLOW 6.
-The term describes a grid in which which a cell can have nominally ``horizontal" connections with with multiple cells across the same cell face.
-In this work, ``vertically offset" and ``vertically staggered" are used to distinguish grids that lack nominally ``horizontal" connections between cells in adjacent layers (red lines in figure \ref{fig:schem-conn-area-flux}a) from those that include such connections, respectively.
+The cells have horizontal tops and bottoms and can therefore follow the dip of the channel only on average.
+If the grid is represented using a layered grid type (DIS or DISV) in MODFLOW 6, a cell is hydraulically connected to each adjacent cell in the same layer, regardless of whether or not the cells overlap, and with each overlying and underlying cell {\color{red} (CDL: I clarified this to indicate that cells are connected horizontally even if they do not overlap) (JM: Maybe state that this is not implicit in the formulation - that faces have to overlap.
+From my experience with people modelling faults...)}, as indicated by the blue lines that connect cell centers in Figure \ref{fig:schem-conn-area-flux}a.
+In this work we call this a vertically offset grid with ``layered connectivity''.
+Note that cells that have overlapping vertical faces but are in different layers (e.g., cells 1A and 2B) do not have a direct hydraulic connection in a grid with layered connectivity.
+The addition of such connections between cells in different layers, as indicated by the red lines that connect cell centers in figure \ref{fig:schem-conn-area-flux}a, results in a grid with ``full connectivity''.
+In MODFLOW 6, layered connectivity is represented using the structured (DIS), vertex-based (DISV), or fully unstructured (DISU) grid types, whereas full connectivity can be represented only using the DISU grid type.
+
+% CDL: it seems like we could remove this following text.  
+%The ability to handle full connectivity was first introduced to MODFLOW by \cite{modflowusg} in MODFLOW-USG, and the term ``vertically staggered'' was first used by \cite{modflow6gwf} in the context of MODFLOW 6.
+The term describes a grid in which which a cell can have nominally ``horizontal'' connections with multiple cells across the same cell face.
+In this work, ``vertically offset'' and ``vertically staggered'' are used to distinguish grids that lack nominally ``horizontal'' connections between cells in adjacent layers (red lines in Figure \ref{fig:schem-conn-area-flux}a) from those that include such connections, respectively.
 In the vertically offset grid in figure \ref{fig:schem-conn-area-flux}a (blue connections only), the interfacial area across which flow occurs between adjacent cells in the same layer is based on the vertical cell thickness (the entire thickness of the channel).
-In the vertically staggered grid (blue and red connections), the interfacial area for nominally ``horizontal" flow between adjacent cells is based on the area over which the two cell faces overlap.
-Thus, while the total area for nominally ``horizontal" flow across a cell face is the same as for a vertically offset grid, that total area is divided between an adjacent cell in the same layer (blue connection) and an adjacent cell in the adjacent layer (red connection).
+In the vertically staggered grid (blue and red connections), the interfacial area for nominally ``horizontal'' flow between adjacent cells is based on the area over which the two cell faces overlap.
+Thus, while the total area for nominally ``horizontal'' flow across a cell face is the same as for a vertically offset grid, that total area is divided between an adjacent cell in the same layer (blue connection) and an adjacent cell in the adjacent layer (red connection).
 In MODFLOW 6, vertically offset grids can be represented using the structured (DIS), vertex-based (DISV), or fully unstructured (DISU) grid type, whereas vertically staggered grids can be represented only using the DISU grid type.
 
-The default method for representing the flow between two hydraulically connected model cells in MODFLOW 6 is the ``conductance-based flow formulation" \citep{modflow6gwf}.
+The default method for representing the flow between two hydraulically connected model cells in MODFLOW 6 is the ``conductance-based flow formulation'' \citep{modflow6gwf}.
 In this formulation, which is based on the commonly used two-point flux approximation {\color{red} *** ref?
@@
 On strictly rectilinear grids, such as structured MODFLOW grids with no vertical offset between cells in the same layer, with hydraulic conductivity that is isotropic or has anisotropy aligned with the three mutually perpendicular grid directions, the conductance-based formulation is second-order accurate \citep{dehotin2010modeling, modflow6gwf}.
-This is because strictly rectilinear grids satisfy the ``control-volume finite-difference (CVFD) requirement" that the straight-line connection between two cell centers must intersect the midpoint of the cell interface at a right angle \citep{narasimhan1976integrated}.
-However, vertically offset grids violate the CFVD requirement because nominally ``horizontal" connections between cell centers in the same layer are not strictly horizontal and therefore not perpendicular to the vertical cell interfaces they intersect.
-For example, in {\color{red} AMP: capitalize ``Figure" throughout} Figure \ref{fig:schem-conn-area-flux}a, the connection between cells 1A and 1B is not perpendicular to the cell interface, which compromises the accuracy of the conductance-based formulation.
+This is because strictly rectilinear grids satisfy the ``control-volume finite-difference (CVFD) requirement'' that the straight-line connection between two cell centers must intersect the midpoint of the cell interface at a right angle \citep{narasimhan1976integrated}.
+However, vertically offset grids violate the CFVD requirement because nominally ``horizontal'' connections between cell centers in the same layer are not strictly horizontal and therefore not perpendicular to the vertical cell interfaces they intersect.
+For example, in {\color{red} AMP: capitalize ``Figure'' throughout} Figure \ref{fig:schem-conn-area-flux}a, the connection between cells 1A and 1B is not perpendicular to the cell interface, which compromises the accuracy of the conductance-based formulation.
 
@@
 As mentioned earlier, the limitations of the conductance-based formulation on vertically offset grids were well recognized.
-However, the ability to use unstructured grids in MODFLOW-USG and MODFLOW 6 introduced the potential to violate the CVFD requirement, and thereby render the conductance-based formulation less accurate, even in the absence of vertical offsets.
-{\color{red} (JM: Elude to ghost nodes?)} To improve accuracy on unstructured grids, and to allow accurate simulation of arbitrarily oriented two- or three-dimensional anisotropy, the optional XT3D flow formulation \citep{modflow6xt3d} was introduced in MODFLOW 6.
+However, the ability to use unstructured grids in MODFLOW-USG and MODFLOW 6 introduced new ways to violate the CVFD requirement, and thereby render the conductance-based formulation less accurate, even in the absence of vertical offsets.
+The concept of a ``ghost node'' was included in MODFLOW-USG \citep{modflowusg}, and subsequently in MODFLOW 6, as a simple and optional way to improve the accuracy for grids that violated the CVFD requirement.
+As an alternative to the ghost-node method for improving accuracy, and as a way to allow accurate simulation of arbitrarily oriented two- or three-dimensional anisotropy, the optional XT3D flow formulation \citep{modflow6xt3d} was introduced in MODFLOW 6.
 XT3D formulates the flow between two cells by interpolating head values from the two cells and their neighboring cells to construct an estimate of the full, two- or three-dimensional head-gradient vector on each side of the cell interface; applying the cell conductivity tensors to obtain an estimate of the groundwater flux  normal to the interface on each side of the interface; and reconciling the two flux estimates to ensure continuity of flow across the interface.
-By performing flow calculations using Darcy's Law in its tensorial form, based on a ``multi-point" approximation of the head gradient, XT3D accounts for both arbitrarily oriented anisotropy and geometric irregularity of the grid.
+By performing flow calculations using Darcy's Law in its tensorial form, based on a ``multi-point'' approximation of the head gradient, XT3D accounts for both arbitrarily oriented anisotropy and geometric irregularity of the grid.
 
-The dipping-channel benchmark problem of \cite{bardot2022} used a vertically offset grid, which is the type of grid used most often in MODFLOW models.
+The dipping-channel benchmark problem of \cite{bardot2022} used a vertically offset layered grid, which is the type of grid used most often in MODFLOW models.
 Test simulations using the conductance-based flow formulation for a steeply dipping channel (hydrogeologic layer) embedded in a domain that is six orders of magnitude less hydraulically conductive yielded significant errors in the simulated flows.
 The simulated groundwater flux in the middle of the channel was not along the $30^{\circ}$ incline of the channel, but nearly horizontal ($0.03^{\circ}$ incline), and its magnitude was overestimated by 15\%.
-Significant errors in the simulated flows were expected in this case, given that the conductance-based flow formulation does not rigorously account for violations of the CVFD requirement that the cell-cell connection be perpedicular to the cell-cell interface, and that it takes the connection length to be the horizontal distance between cell centers.
+Significant errors in the simulated flows were expected in this case, given that the conductance-based flow formulation does not rigorously account for violations of the CVFD requirement that the cell-cell connection be perpendicular to the cell-cell interface, and that it takes the connection length to be the horizontal distance between cell centers.
 Given the ability of XT3D to account rigorously for grid connections that are not perpendicular to cell interfaces, and for the increase in connection length due to the slope of the connection, rerunning the simulation with XT3D activated could have been expected to improve the flow solution substantially.
 However, \cite{bardot2022} observed similarly poor results with XT3D: the flux was still nearly horizontal ($0.01^{\circ}$ incline), and its magnitude was overestimated by 12\%.
-Based on a preliminary analysis by two authors of the present work (Provost and Langevin), \cite{bardot2022} hypothesized that the unexpectedly large error in the flow solution with XT3D was related to indequate grid connectivity was at the root of the problem.
+Based on a preliminary analysis by two authors of the present work (Provost and Langevin), \cite{bardot2022} hypothesized that the unexpectedly large error in the flow solution with XT3D was related to inadequate grid connectivity.
 
-The role of grid connectivity in enabling accurate simulation of flow along a dipping channel can be understood by considering figure \ref{fig:schem-conn-area-flux}b, in which components of the groundwater flux (specific discharge) vector representative of steady, uniform flow along the channel are superimposed on each cell interface.
+The role of grid connectivity in enabling accurate simulation of flow along a dipping channel can be understood by considering Figure \ref{fig:schem-conn-area-flux}b, in which components of the groundwater flux (specific discharge) vector representative of steady, uniform flow along the channel are superimposed on each cell interface.
 Gray vectors represent the uniform flux oriented along the channel, which the model is attempting to simulate.
-Blue vectors represent the flux components normal to cell interfaces that correspond to blue connections in figure \ref{fig:schem-conn-area-flux}a, which are common to both the vertically offset and vertically staggered grids.
-Red vectors represent the flux components normal to cell interfaces that correspond to red connections in figure \ref{fig:schem-conn-area-flux}a, which are lacking in the vertically offset grid.
+Blue vectors represent the flux components normal to cell interfaces that correspond to blue connections in Figure \ref{fig:schem-conn-area-flux}a, which are common to both the layered and full connectivity grids.
+Red vectors represent the flux components normal to cell interfaces that correspond to red connections in Figure \ref{fig:schem-conn-area-flux}a, which are lacking in the layered connectivity grid.
 
-The blue vectors in figure \ref{fig:schem-conn-area-flux}b show that to simulate steady, uniform flow along the dipping channel, cells must exchange water not only ``horizontally" with neighboring cells in the same layer, but also vertically with cells in the adjacent layer.
+The blue vectors in Figure \ref{fig:schem-conn-area-flux}b show that to simulate steady, uniform flow along the dipping channel, cells must exchange water not only ``horizontally'' with neighboring cells in the same layer, but also vertically with cells in the adjacent layer.
 Specifically, there must be flow from cells in the bottom layer to cells in the top layer, which corresponds to the vertical component of the groundwater flux.
@@
 Thus, steady, uniform flow along the channel cannot be simulated accurately without some mechanism for returning flow from the top layer to the bottom layer.
-The vertically staggered grid provides such a mechanism by offering additional connections between cells in adjacent layers.
-Vertical flows from the bottom layer to the top layer, represented by vertical blue vectors in Figure \ref{fig:schem-conn-area-flux}b, can be returned to the bottom layer by nominally ``horizontal" flows, represented by red vectors in figure \ref{fig:schem-conn-area-flux}b.
+The full connectivity grid provides such a mechanism by offering additional connections between cells in adjacent layers.
+Vertical flows from the bottom layer to the top layer, represented by vertical blue vectors in Figure \ref{fig:schem-conn-area-flux}b, can be returned to the bottom layer by nominally ``horizontal'' flows, represented by red vectors in Figure \ref{fig:schem-conn-area-flux}b.
 Although this concept has been illustrated using a two-layer model for simplicity, the same reasoning applies given any number of layers.
 
-Once the appropriate connectivity is established using a vertically staggered grid, it is the role of the flow formulation to account properly for the geometry of the grid, a task for which XT3D was specifically designed.
+Once the full connectivity is established, it is the role of the flow formulation to account properly for the geometry of the grid, a task for which XT3D was specifically designed.
 However, XT3D alone cannot compensate for grid connectivity that does not provide adequate pathways for flow.
-This explains why using XT3D did not improve the simulation results substantially on a vertically offset grid in the benchmark tests of \cite{bardot2022}.
-In the middle section of the channel, away from the end boundary conditions, the flow solution approached steady, uniform flow the only way that it could given the limited connectivity of the vertically offset grid: by suppressing vertical flows between layers, thereby rendering the overall flow approximately horizontal.
+This explains why using XT3D did not improve the simulation results substantially on the layered connectivity grid in the benchmark tests of \cite{bardot2022}.
+In the middle section of the channel, away from the end boundary conditions, the flow solution approached steady, uniform flow the only way that it could given the limited connectivity: by suppressing vertical flows between layers, thereby rendering the overall flow approximately horizontal.
 
-The arguments presented in this section suggest that, in addition to the use of XT3D, the enhanced connectivity provided by a vertically staggered grid can be important for obtaining an accurate flow solution in MODFLOW 6 simulations that involve steeply dipping layers.
+The arguments presented in this section suggest that, in addition to the use of XT3D, full connectivity can be important for obtaining an accurate flow solution in MODFLOW 6 simulations that involve steeply dipping layers.
 The next section presents results of simulations designed to test this hypothesis.
@@
 The test problem is patterned after the cross-sectional (transect) benchmark of \cite{bardot2022}.
-It attempts to simulate uniform flow in an inclined permeable channel embedded in a less permeable surrounding ``domain." The channel is of uniform width and hydraulic conductivity and is inclined relative to the horizontal.
+It attempts to simulate uniform flow in an inclined permeable channel embedded in a less permeable surrounding ``domain.'' The channel is of uniform width and hydraulic conductivity and is inclined relative to the horizontal.
 
@@
 
-Two methods of discretization are used to simulate the test problem: (1) vertically offset grid of the MODFLOW 6 DIS (structured) type, and (2) vertically staggered grid of MODFLOW 6 DISU (unstructured) type.
+Two methods of discretization are used to simulate the test problem: (1) vertically offset grid of the MODFLOW 6 DIS (structured) type, and (2) vertically offset grid of MODFLOW 6 DISU (unstructured) type with full connectivity.
 The cell geometry in these two discretization methods is identical, but the cell connectivity is different.
-The vertically staggered grid included additional ``cross-connections" between cells in adjacent model layers, as described later in the paper.
-The input files for a vertically staggered scenario are formulated by starting with a DIS grid, converting it to an equivalent DISU grid, and adding cross-connections.
+The DISU grid included additional ``cross-connections'' between cells in adjacent model layers, as described later in the paper.
+The input files for the DISU grid are formulated by starting with a DIS grid, converting it to an equivalent DISU grid, and adding cross-connections.
 The base case uses grids consisting of 11 columns and 9 layers of cells, with 3 layers within the channel, 3 layers above the channel, and 3 layers below the channel.
@@
 Jupyter notebooks created for the test problem are available in the Supporting Information that accompanies this paper.
-The notebooks leverage the capabilities of FloPy \citep{bakker2016scripting} to assist in setting up input for and processing results from the MODFLOW 6 simulations.
+The notebooks leverage the capabilities of FloPy \citep{bakker2016scripting, hughes2023flopy} to assist in setting up input for and processing results from the MODFLOW 6 simulations.
 A Python script for converting structured, DIS grids to unstructured, DISU grids is included.
@@
 This section presents results for the base case scenario when comparing gridding strategy (i.e.
-vertically offset versus vertically staggered) and flow solution formulation (i.e.
-standard conductance based formulation versus XT3D option) to test the hypothesis that a vertically staggered grid produces a more accurate solution for dipping layers than a vertically offset grid.
+layered connectivity versus full connectivity) and flow solution formulation (i.e.
+standard conductance based formulation versus XT3D option) to test the hypothesis that a grid with full connectivity produces a more accurate solution for dipping layers than a grid with only layered connectivity.
 The base case scenario is then modified to further investigate other aspects including grid resolution, dip angle, conductivity contrast and anisotropy.
 
-Results for flux magnitude, flux direction and volumetric flowthrough are used to assess the accuracy of the numerical representations of the test problem.
-Note that the terms ``flux"  and ``specific discharge" are used interchangely.
-Reported fluxes for each model are for the central model cell and therefore uninfluenced by the model boundary or the interface between channel and domain.
+Results for flux magnitude, flux direction and volumetric flow through the channel are used to assess the accuracy of the numerical representations of the test problem.
+Note that the terms ``flux''  and ``specific discharge'' are used interchangeably.
+Reported fluxes for each model are for the central model cell and are therefore uninfluenced by the model boundary or the interface between the channel and the domain.
 Volumetric flow through the channel is therefore also examined as it encompasses fluxes through the entire channel.
@@
 
-The numerical solution for the vertically offset grid confirms the findings of \cite{bardot2022} in that fluxes are predominantly horizontal (top row, Figure \ref{fig:fig2}).
-We see that despite an enforced hydraulic gradient of $30^{\circ}$, the vertical component of the flux cannot propogate through the channel given that cells along the channel top are hydraulically disconnected from the low permeability domain, hence forcing flow horizontally (red arrows on right panel).
+The numerical solution for the grid with layered connectivity confirms the findings of \cite{bardot2022} in that fluxes are predominantly horizontal (top row, Figure \ref{fig:fig2}).
+We see that despite an enforced hydraulic gradient of $30^{\circ}$, the vertical component of the flux cannot propagate through the channel given that cells along the channel top are hydraulically disconnected from the low permeability domain, hence forcing flow horizontally (red arrows on right panel).
 The numerical solution for this grid type is a problem of cell connectivity, and thus the XT3D formulation does not resolve this issue (second row, Figure \ref{fig:fig2}).
-Results for the vertically staggered grid with the standard conductance based formulation shows an improvement in the flow solution (third row, Figure \ref{fig:fig2}), with the XT3D formulation successfully reproducing the analytical solution (bottom row, Figure \ref{fig:fig2}).
-We see that the staggered grid cell connections allows incorporation of the vertical flow component, thus permitting cross-connection between model layers.
+Results for the full connectivity grid with the standard conductance-based formulation shows an improvement in the flow solution (third row, Figure \ref{fig:fig2}), with the XT3D formulation successfully reproducing the analytical solution (bottom row, Figure \ref{fig:fig2}).
+We see that the grid with full connectivity allows incorporation of the vertical flow component, thus permitting cross-connection between model layers.
 
@@
 There are two main MODFLOW output files pertaining to fluxes.
-The first is an array of face flows which specifies the normal flux across each cell face (``FLOW-JA").
-The second is the cell-averaged specific discharge for each cell in its directional components ($q_{x}$, $q_{y}$, $q_{z}$) (``DATA-SPDIS"), and is calculated using distance-weighted averages of the cell face flows.
-Therefore, the specific discharge magnitude and direction normally reported by MODFLOW for cells within and along the channel edge are ``dampened" due to averaging of the face flows from the cell top and bottom.
-The result is a ``staircase" effect whereby these edge cells have a reduced flux magnitude and ``flattened" flux directions, as shown in Figure 3c of \cite{bardot2022}.
-The plots in Figure 2 display corrected cell-averaged specific discharges which exclude including the near-zero face flows along the channel boundary, and therefore represents the actual simulated flux magnitude and direction within the cell.
+The first is an array of face flows which specifies the total flow across each cell face (``FLOW-JA-FACE'').
+The second is the cell-averaged specific discharge for each cell in its directional components ($q_{x}$, $q_{y}$, $q_{z}$) (``DATA-SPDIS''), and is calculated using distance-weighted averages of the cell face flows.
+Therefore, the specific discharge magnitude and direction normally reported by MODFLOW for cells within and along the channel edge are ``dampened'' due to averaging of the face flows from the cell top and bottom.
+The result is a ``staircase'' effect whereby these edge cells have a reduced flux magnitude and ``flattened'' flux directions, as shown in Figure 3c of \cite{bardot2022}.
+The plots in Figure 2 display corrected cell-averaged specific discharges which exclude the near-zero face flows along the channel boundary, and therefore represents the actual simulated flux magnitude and direction within the cell.
 
@@
 	\includegraphics[scale=0.8]{../figures/fig2_paper.png}
-	\caption{Numerical results for the test problem using base case settings for vertically offset and vertically staggered grids, with the standard conductance based formulation as well as the XT3D formulation.
+	\caption{{\color{red} (CDL: replace ``vertically offset'' with ``layered connectivity'' and ``vertically staggered'' with ``full connectivity''.)} Numerical results for the test problem using base case settings for layered connectivity and full connectivity grids, with the standard conductance based formulation as well as the XT3D formulation.
 The left panel of each scenario shows the calculated specific discharge at each cell centre (black arrows) with head contours (orange).
@@
 
-{\color{red} Should we also consider the horizontal discretisation too?}
+{\color{red} Should we also consider the horizontal discretisation too? (CDL: I would say no.  Paper is already getting a bit long.)}
 
@@
 	\includegraphics[scale=0.9]{../figures/fig3paper.png}
-	\caption{Graphs indicating how increasing the number of model layers within a single hydrogeological layer (in this case, the channel) affects (a) flux magnitude of middle cell, (b) flux direction of middle cell and (c) volumetric flow through the channel.
-We immediately notice the errors vertically offset grids incur (blue line), and how using a vertically staggered grid reproduces the analytical solution from two model layers used to discretise the channel (orange).}
+	\caption{{\color{red} (CDL: replace ``vertically offset'' with ``layered connectivity'' and ``vertically staggered'' with ``full connectivity.'')} Graphs indicating how increasing the number of model layers within a single hydrogeological layer (in this case, the channel) affects (a) flux magnitude of middle cell, (b) flux direction of middle cell and (c) volumetric flow through the channel.
+We immediately notice the errors layered connectivity grids incur (blue line), and how using a full connectivity grid reproduces the analytical solution from two model layers used to discretise the channel (orange).}
 	\label{fig:fig3}
@@
 The base scenario arbitrarily uses 3 model layers to represent the channel, but it is important to address the question: \emph{How many model layers are required to adequately represent a hydrogeological layer?}.
-Vertically offset grids, which have been used by groundwater modellers for decades (ref?) {\color{red} Jim to check, I think there is some refinement guidabce for aquitards}, are almost always configured as one model layer per hydrogeological unit.
+Layered connectivity grids, which have been used by groundwater modellers for decades (ref?) {\color{red} Jim to check, I think there is some refinement guidance for aquitards}, are almost always configured as one model layer per hydrogeological unit.
 However, although computationally efficient, we investigate the effect of altering resolution within the channel by using the base case and varying number of cells per channel width from one to nine (Figure \ref{fig:fig3}).
-We compare modelled flux magnitude and direction in the centre cell, as well as the volumetric flow against the analytical solution.
+We compare modeled flux magnitude and direction in the center cell, as well as the volumetric flow against the analytical solution.
 
-Flux magnitude error for vertically offset grids peaked at around 10\% but improved and tended to zero with refinement.
-Flux direction for vertically offset grids is clearly a major issue, being reporteded as horizontal for one model layer per channel width, and only improved to 12$^{\circ}$ for 9 layers per channel width.
+Flux magnitude error for layered connectivity grids peaked at around 10\% but improved and tended to zero with refinement.
+Flux direction for full connectivity grids is clearly a major issue, being reported as horizontal for one model layer per channel width, and only improved to 12$^{\circ}$ for 9 layers per channel width.
 Flow through the channel peaked at 13\% error but decreased to only a 4\% error for 9 layers per channel width.
 
-The results for the vertically staggered grid clearly shows that a minimum of two layers per channel width is required to match the analytical result for flux magnitude, direction and volumetric flow.
-This is an important finding as it demonstrates that computational efficiency can be made using a staggered grid over a rectilinear grid where parameters are overlayed, but it is important to use a minimum of two model layers per hydrogeological unit to allow vertical fluxes to be incorporated into the flow solution.
+The results for the full connectivity grid clearly shows that a minimum of two layers per channel width is required to match the analytical result for flux magnitude, direction and volumetric flow.
+This is an important finding as it demonstrates that computational efficiency can be made using a full connectivity grid over a rectilinear grid overlay approach, but it is important to use a minimum of two model layers per hydrogeological unit to allow vertical fluxes to be incorporated into the flow solution.
 
@@
 The base case scenario uses a dip angle of 30$^{\circ}$ and an extreme conductivity contrast between channel and domain of $1:10^{-6}$.
-Here, we investigate the behaviour of the flow solution for a wide range of channel inclines by systematically changing the dip angle by 2.5$^{\circ}$ from 0 to 80 $^{\circ}$.
+Here, we investigate the behavior of the flow solution for a wide range of channel inclines by systematically changing the dip angle by 2.5$^{\circ}$ from 0 to 80 $^{\circ}$.
 Similarly, we examine the solution for multiple conductivity contrasts by changing the domain conductivity to 2, 5, 10 and 100 times less than the channel.
@@
 
-Traditional vertically offset grids without XT3D shows rapid diversion of flux magnitude and direction from the analytical solution as dip increases (Figure \ref{fig:fig4a}).
+Traditional layered connectivity grids without XT3D shows rapid diversion of flux magnitude and direction from the analytical solution as dip increases (Figure \ref{fig:fig4a}).
 Use of the XT3D option keeps the flux magnitude somewhat on-track until about 30 $^{\circ}$ when the solution starts peeling away from the analytical.
 Flux direction is improved for the homogeneous scenario, but not for the heterogeneous scenarios.
-The vertically staggered grid on the other hand, even without XT3D, roughly follows the analytical solution for flux magnitude and direction until around 65 $^{\circ}$ when it starts to ``fall-apart." There are also spikes at certain dip angles which are related to transitions between cell connections due to changing cell face overlaps.
-The vertically staggered grid with XT3D clearly boasts the best results with the flux magnitude and direction matching the analytical until 65 $^{\circ}$, with the exception of a slight decrease in flux direction after 45 $^{\circ}$.
+The full connectivity grid on the other hand, even without XT3D, roughly follows the analytical solution for flux magnitude and direction until around 65 $^{\circ}$ when it starts to ``fall-apart.'' There are also spikes at certain dip angles which are related to transitions between cell connections due to changing cell face overlaps.
+The full connectivity grid with XT3D clearly boasts the best results with the flux magnitude and direction matching the analytical until 65 $^{\circ}$, with the exception of a slight decrease in flux direction after 45 $^{\circ}$.
 
-These results confirm that vertically staggered grids with XT3D not only outperform the alternatives when modelling dipping hydrogeological unit, but come extremely close to the analytical solution.
+These results confirm that full connectivity grids with XT3D not only outperform the alternatives when modelling dipping hydrogeological unit, but come extremely close to the analytical solution.
 
@@
 	\includegraphics[width=\textwidth]{../figures/fig4_0_paper.png}
-	\caption{vertically offset grid, standard conductance-based formulation}
+	\caption{layered connectivity grid, standard conductance-based formulation}
 	\label{fig:fig4a}
@@
 	\includegraphics[width=\textwidth]{../figures/fig4_1_paper.png}
-	\caption{vertically offset grid, XT3D}
+	\caption{layered connectivity grid, XT3D}
 	\label{fig:fig4b}
@@
 	\includegraphics[width=\textwidth]{../figures/fig4_2_paper.png}
-	\caption{vertically staggered grid, standard conductance-based formulation}
+	\caption{full connectivity grid, standard conductance-based formulation}
 	\label{fig:fig4c}
@@
 	\includegraphics[width=\textwidth]{../figures/fig4_3_paper.png}
-	\caption{vertically staggered grid, XT3D}
+	\caption{full connectivity grid, XT3D}
 	\label{fig:fig4d}
@@
 
-\caption{Flux magnitude and direction for centre cells for varying contrasts in log conductivity (blue, orange, green, red lines) for varying dip angles (x-axis).
+\caption{{\color{red} (CDL: replace ``vertically offset'' with ``layered connectivity'' and ``vertically staggered'' with ``full connectivity''.)}Flux magnitude and direction for centre cells for varying contrasts in log conductivity (blue, orange, green, red lines) for varying dip angles (x-axis).
 The analytical value is shown in black dashed line.}
@@
 Results are presented in Figure \ref{fig:fig5}.
-Vertically offset grids produce flux magnitude and flow results that significantly deviate from the analytical solution (blue line), whilst vertically staggered grids reproduce the analytical solution (orange line).
-These results verify, once again, our proposed method of using vertically staggered grids to model dipping anisotropic layers.
+Layered connectivity grids produce flux magnitude and flow results that significantly deviate from the analytical solution (blue line), whilst full connectivity grids reproduce the analytical solution (orange line).
+These results verify, once again, our proposed method of using full connectivity grids to model dipping anisotropic layers.
 
@@
 
+ {\color{red} CDL: I added the following two paragraphs in an attempt to summarize my understanding of the main conclusion.  Feel free to edit or scrap as you see fit.}
+
+\cite{bardot2022} revisited the capability of MODFLOW to accurately simulate groundwater flow through sedimentary structures.
+They concluded that the XT3D multi-point flux approximation in MODFLOW 6 significantly improves the accuracy of simulated flows for rectilinear grid overlay approaches in which locations of sedimentary structures are mapped onto a relatively fine model grid.
+They also concluded that the XT3D multi-point flux approximation did not perform as well as anticipated for vertically offset grids, which are advantageous for significantly reducing the number of layers required for many problems.
+They hypothesized that the inherent connectivity in DIS and DISV layered grids in MODFLOW 6 was inadequate for allowing the XT3D multi-point flux approximation to work as intended.
+
+In this paper, we confirm that the layered connectivity implemented in DIS and DISV model grids is inadequate for simulating groundwater flow through steeply dipping sedimentary structures, even when the XT3D multi-point flux approximation is used.
+However, when additional cross connections are added to comprise a grid with full connectivity, there is a substantial increase in the accuracy of simulated flows.
+Importantly, this accuracy improvement requires that at least two model layers are used to represent the steeply dipping sedimentary structure.
+The primary conclusion of this paper, therefore, is that that the XT3D multi-point flux approximation implemented in MODFLOW 6 can be combined with vertically offset grids to efficiently and accurately model flows through sedimentary structures.
+Prior to this work, the capability to efficiently and accurately model flow through sedimentary structures was generally restricted to finite element simulators.
+
 This paper presents a solution for accurately representing steeply dipping flow structures groundwater utilising MODFLOW 6.
-\citep{bardot2022} demonstrated that the flux magnitude in highly dipping scenarios was erroneous when utilising vertically offset grids.
-More significantly, the direction of fluxes were inaccurate, neglecting the vertical component of the flux.
+\citep{bardot2022} demonstrated that the flux magnitude in highly dipping scenarios was erroneous when utilising layered connectivity grids.
+{\color{red} CDL: Also applicable for MODFLOW-USG DISU grids.} More significantly, the direction of fluxes were inaccurate, neglecting the vertical component of the flux.
 The issues observed in the existing study of \citep{bardot2022} partially related to the implementation of only within-layer horizontally connections, limiting the model capacity to simulate the vertical component of the flow.
-The present study developed an approach which utilised vertically staggered grids \cite{modflow6gwf} to overcome the perceived problems.
-Through comparison of the previous vertically offset and proposed vertically staggered approached to the analytical solution presented in \citep{bardot2022}, we demonstrate that the use of vertically staggered grids (in conjunction with the XT3D formulation) is an accurate method to simulate flow through dipping hydrogeological layers.
+The present study developed an approach which utilised full connectivity grids \cite{modflow6gwf} to overcome the perceived problems.
+Through comparison of the previous layered connectivity and proposed full connectivity approached to the analytical solution presented in \citep{bardot2022}, we demonstrate that the use of full connectivity grids (in conjunction with the XT3D formulation) is an accurate method to simulate flow through dipping hydrogeological layers.
 However, accurate representation of flows requires a minimum of two model layers per hydro geological unit to allow for vertical connections to induce flows.
@@
 
-3) (Random thought) The use of vertically staggered grids improved the accuracy of the transect solutions; however, the results of \citep{bardot2022} using unstructured meshes that conformed to the channel geometry preformed better in isotropic grids.
+3) (Random thought) The use of full connectivity grids improved the accuracy of the transect solutions; however, the results of \citep{bardot2022} using unstructured meshes that conformed to the channel geometry preformed better in isotropic grids.
 If you force the normal flux to be the same at the interface of the domain and channel, the analytical solution would suggest a uniform gradient either side, do we introduce an error if the faces don't align with the channel correctly?
 Would figure 2 show a correction for this if the anisotropic scenario was tested?
+
+ {\color{red} CDL: Maybe worthwhile to implement full connectivity approaches for DIS and DISV grids in mf6.}
 

### 2023-10-19 latex/article.tex
provost edits
Various edits, including many aimed at consistent terminology

@@
 
-%\title{Improving Accuracy of Simulated Flows in Steeply Dipping Layers Using full connectivity Grids with the XT3D Multi-Point Flux Approximation in MODFLOW 6}
+%\title{Improving Accuracy of Simulated Flows in Steeply Dipping Layers Using Full-Connectivity Grids with the XT3D Multi-Point Flux Approximation in MODFLOW 6}
 
@@
 
-\textbf{Key words:} groundwater flow simulation, dipping model layers, full connectivity grid, multi-point flux approximation
+\textbf{Key words:} groundwater flow simulation, sedimentary structures, dipping layers, full-connectivity grid {\color{red} AMP: maybe ``cell connectivity''' or ``grid connectivity'' would make a better search term?}, multi-point flux approximation
 
@@
 \cite{hoaglund2003} refer to this as the ``grid overlay'' method.
-In the grid overlay method, the boundaries of dipping layers are represented approximately, in what \cite{bardot2022} call a ``staircase'' fashion.
+In the grid overlay method, the boundaries of dipping hydrogeologic layers are represented approximately, in what \cite{bardot2022} call a ``staircase'' fashion.
 Another approach available in all versions of MODFLOW is to allow the top and bottom elevations of cells in the same layer of the grid to vary spatially.
-Following \cite{bardot2022}, in this work we call this type of grid ``vertically offset,'' and we show how to improve the accuracy of flow simulations that use vertically offset grids.
-The vertical offsets in top and bottom elevations between ``horizontally'' adjacent cells allows model layers to ``deform'' with the hydrostratigraphy to ``minimize the number of model layers required to simulate an aquifer system'' \citep{modflow84} relative to the grid overlay method with a rectilinear grid consisting of horizontal layers.
+Following \cite{bardot2022}, we call this type of grid ``vertically offset.'' In this work, we show how to improve the accuracy of flow simulations that use vertically offset grids.
+The vertical offsets in top and bottom elevations between adjacent cells within a grid layer allows each grid layer to ``deform'' with the hydrostratigraphy.
+This ``minimize[s] the number of model [grid] layers required to simulate an aquifer system'' \citep{modflow84} relative to the grid overlay method.
 
 All versions of MODFLOW prior to MODFLOW 6 used a two-point, ``conductance-based'' flow formulation in which the flow between two adjacent cells is proportional to the difference between the heads calculated in those two cells.
-The conductance-based flow formulation is most accurate when the grid satisfies the ``control-volume finite-difference (CVFD) requirement'' that the straight-line connection between two cell centers must intersect the midpoint of the cell interface at a right angle \citep{narasimhan1976integrated}.
+The conductance-based flow formulation is most accurate when the grid satisfies the ``control-volume finite-difference (CVFD) requirement'' that the straight-line connection between two cell centers must intersect the midpoint of the cell-cell interface at a right angle \citep{narasimhan1976integrated}.
 However, vertically offset grids violate the CVFD requirement because lateral cell faces are vertical but cell connections are not strictly horizontal.
@@
 By performing flow calculations using Darcy's Law in its tensorial form based on a multi-point approximation of the head gradient vector, XT3D accounts for both arbitrarily oriented anisotropy and geometric irregularity of the grid.
-In theory, XT3D should be able to compensate for the geometric irregularity introduced by vertical offsets between adjacent cells in the same model layer.
+In theory, XT3D should be able to compensate for the geometric irregularity introduced by vertical offsets between adjacent cells in the same grid layer.
 
-\cite{bardot2022} evaluated the effects of grid design and XT3D on the ability of MODFLOW 6 to accurately simulate flow through sedimentary structures by modeling an idealized, two-dimensional permeable channel or hydrogeologic layer {\color{red} (CDL: use ``permeable feature'' instead of channel?)} embedded in a nearly impermeable surrounding medium, or ``domain.'' In their plan-view benchmark, the channel lay within the horizontal plane and was angled counterclockwise relative to the southern boundary of the model domain (by $30^{\circ}$ in most cases).
+\cite{bardot2022} evaluated the effects of grid design and XT3D on the ability of MODFLOW 6 to accurately simulate flow through sedimentary structures by modeling an idealized, two-dimensional permeable feature (``channel'') in plan view and a dipping hydrogeologic layer (which \cite{bardot2022} also generically called a ``channel'') in cross-section.
+{\color{red} (CDL: use ``permeable feature'' instead of channel?) (AMP: Think it's ok to use channel here in reference to Kerry and Jim's plan-view model, but not after this paragraph; made changes to that effect throughout.)} In each case, the channel or hydrogeologic layer was embedded in a nearly impermeable surrounding medium, or ``domain.'' In their plan-view benchmark, the channel lay within the horizontal plane and was angled counterclockwise relative to the southern boundary of the model (by $30^{\circ}$ in most cases).
 Constant-head boundary conditions at the ends of the channel were set to induce uniform flow along the channel in the corresponding analytical solution.
 The channel and surrounding domain were discretized using either a rectilinear grid, which satisfies the CVFD requirement, or a ``flexible triangular'' grid that is unstructured with triangular cells that conform to the channel boundary but does not satisfy the CVFD requirement.
-They found that the standard, conductance-based flow formulation gave accurate results on the rectilinear grid using isotropic hydraulic conductivity tensors only (accuracy decreased substantially in the presence of anisotropy) {\color{red} AMP: intended emphasis of ``only''?
-(CDL: I believe ``only'' was intentional.
-I added the part about anisotropy effects in parentheses)} but was subject to significant error on the unstructured grid, whereas the XT3D flow formulation gave accurate results on both grids for both isotropic and anisotropic hydraulic conductivity tensors, as expected.
-However, in the transect (cross-sectional) version of their benchmark, which used a vertically offset grid to discretize the nearly impermeable domain and a permeable channel (hydrogeologic layer) inclined by $30^{\circ}$ from the horizontal, similarly poor results were observed with and without XT3D.
+They found that the standard, conductance-based flow formulation gave accurate results on the rectilinear grid using isotropic hydraulic conductivity tensors (although accuracy decreased substantially in the presence of anisotropy) but was subject to significant error on the unstructured grid.
+By contrast, the XT3D flow formulation gave accurate results on both grids for both isotropic and anisotropic hydraulic conductivity tensors, as expected.
+However, in the cross-sectional (``transect'') version of their benchmark, which used a vertically offset grid to discretize the nearly impermeable domain and a highly permeable hydrogeologic layer inclined by $30^{\circ}$ from the horizontal, similarly poor results were observed with and without XT3D.
 Given the ability of XT3D to compensate for violations of the CVFD requirement, it appeared likely that another aspect of the problem formulation or the inner workings of MODFLOW 6 was responsible for the unexpected results.
-As noted by \cite{bardot2022}, a preliminary analysis by two authors of the present work (Provost and Langevin) suggested that the cell connectivity offered by the vertically offset grid used for the problem was inadequate for allowing an accurate flow solution.
+As noted by \cite{bardot2022}, a preliminary analysis by two authors of the present work (Provost and Langevin) suggested that the limited cell connectivity offered by the vertically offset grid was inadequate for allowing an accurate flow solution.
 
-MODFLOW 6 supports three different types of grids: regular MODFLOW grids consisting of layers, rows, and columns (DIS),  grids discretized by vertices (DISV), and fully unstructured (DISU) grids, which are patterned after the DISU grid type in MODFLOW-USG  \citep{modflowusg}.
-The DIS and DISV grids in MODFLOW 6 are ``layered'', which means that the same two-dimensional grid (in plan view) applies to all model layers.
+MODFLOW 6 supports three different types of grids: regular MODFLOW grids consisting of layers, rows, and columns (DIS);  grids discretized by vertices (DISV); and fully unstructured (DISU) grids, which are patterned after the DISU grid type in MODFLOW-USG  \citep{modflowusg}.
+The DIS and DISV grids in MODFLOW 6 are ``layered'', which means that the same two-dimensional grid structure (in plan view) applies to all grid layers.
 The ``layered'' concept of DIS and DISV grids is used within MODFLOW to determine how cells are hydraulically connected.
-Cells in DIS and DISV grids are automatically assigned vertical connections to overlying and underlying cells and horizontal connections to adjacent cells in the same model layer.
+Cells in DIS and DISV grids are automatically assigned vertical connections to overlying and underlying cells and horizontal connections to adjacent cells in the same grid layer.
 DISU grids are more flexible than DIS and DISV grids in that the user has complete freedom to specify the cell connectivity and the geometric properties for each connection.
 DISU grids support two types of horizontal connections: a ``layered'' horizontal connection and a ``vertically staggered'' horizontal connection \citep{modflow6gwf}.
-With a layered horizontal connection, the hydraulic conductance is calculated using the full saturated thickness of the two connected model cells.
-Alternatively, for a staggered horizontal connection, the hydraulic conductance is calculated using only the overlap between the two cell faces.
+With a layered horizontal connection, the hydraulic conductance is calculated using the full saturated thicknesses of the two connected model cells.
+By contrast, for a staggered horizontal connection, the hydraulic conductance is calculated using only the area of overlap between the two cell faces.
 Vertically staggered horizontal connections were introduced in MODFLOW-USG \citep{modflowusg} and included in MODFLOW 6 \citep{modflow6gwf} primarily as a means to vertically subdiscretize parts of a model domain.
+Vertically staggered connections offer the possibility of implementing ``full'' connectivity in the sense that every cell is hydraulically connected to every other cell with which its faces overlap.
+
 The grid used by \cite{bardot2022} for the steeply dipping benchmark example used layered horizontal connections.
-To our knowledge, the implications of limited horizontal connectivity in layered grids for the accuracy of simulated flows in models with steeply dipping layers, and the potential for using vertically staggered grids with full cell connectivity to overcome this limitation, have not been previously appreciated or investigated.
+To our knowledge, the implications of limited horizontal connectivity in layered grids for the accuracy of simulated flows in models with steeply dipping hydrogeologic layers, and the potential for using full cell connectivity via vertically staggered connections to overcome this limitation, have not been previously appreciated or investigated.
 
@@
 % For a grid that includes vertically staggered connections, which in this work we call a ``vertically staggered grid,'' some cells have nominally ``horizontal'' connections with more than one cell in some direction.
-Vertically staggered DISU grids allow more flexibility in assigning grid connectivity than do the more commonly used DIS and DISV grid types \citep{modflow6gwf} in MODFLOW 6.
+Vertically staggered DISU grids allow more flexibility in assigning cell connectivity than do the more commonly used DIS and DISV grid types \citep{modflow6gwf} in MODFLOW 6.
 
-In the remainder of this paper we first present a theoretical justification for the hypothesis that inadequate grid connectivity is primarily responsible for the inaccurate simulation of flows in the steeply dipping layer benchmark on a vertically offset grid.
+In the remainder of this paper we first present a theoretical justification for the hypothesis that inadequate cell connectivity is primarily responsible for the inaccurate simulation of flows in the steeply dipping hydrogeologic layer benchmark on a vertically offset grid.
 Based on that theory, we propose a method for improving the accuracy of the flow solution using XT3D together with vertically offset grid with full connectivity, and we evaluate the effectiveness of that approach in a set of benchmark problems similar to those of \cite{bardot2022}.
-Finally, we discuss the potential implications of our findings for practical groundwater models that include high permeability contrasts and steeply dipping layers.
+Finally, we discuss the potential implications of our findings for practical groundwater models that include high permeability contrasts and steeply dipping hydrogeologic layers.
 
@@
 	\includegraphics[scale=0.6]{../figures/schem_conn_area_flux.png}
-	\caption{Schematics showing (a) grid connectivity and (b) cell interface fluxes in a two-layer model of a dipping channel.
-Hatching denotes impermeable boundaries along the top and bottom of the channel.
+	\caption{Schematics showing (a) cell connectivity and (b) cell-cell interface fluxes in a two-grid-layer model of a dipping hydrogeologic layer.
+Hatching denotes impermeable boundaries along the top and bottom of the hydrogeologic layer.
 In (a), black circles represent cell centers.
-Blue lines represent hydraulic connections between cells in a layered connectivity grid.
-Red lines represent additional connections that must be made in order to achieve full connectivity.
+Blue lines represent hydraulic connections between cells in a layered-connectivity grid.
+Red lines represent additional connections that must be made to achieve full connectivity.
 In (b), gray arrows represent the uniform groundwater flux the model is attempting to simulate.
-Blue arrows represent components of the uniform groundwater flux normal to the horizontal and vertical cell-cell interfaces in the layered connectivity grid.
+Blue arrows represent components of the uniform groundwater flux normal to the horizontal and vertical cell-cell interfaces in the layered-connectivity grid.
 Red arrows represent components of the uniform groundwater flux normal to the additional vertical cell-cell interfaces introduced to represent the full connectivity.}
@@
 
-Figure \ref{fig:schem-conn-area-flux}a shows a group of four cells in a two-layer, cross-sectional MODFLOW 6 model of fully saturated flow through a permeable channel (aquifer) with impermeable top and bottom boundaries.
-The cells have horizontal tops and bottoms and can therefore follow the dip of the channel only on average.
-If the grid is represented using a layered grid type (DIS or DISV) in MODFLOW 6, a cell is hydraulically connected to each adjacent cell in the same layer, regardless of whether or not the cells overlap, and with each overlying and underlying cell {\color{red} (CDL: I clarified this to indicate that cells are connected horizontally even if they do not overlap) (JM: Maybe state that this is not implicit in the formulation - that faces have to overlap.
-From my experience with people modelling faults...)}, as indicated by the blue lines that connect cell centers in Figure \ref{fig:schem-conn-area-flux}a.
-In this work we call this a vertically offset grid with ``layered connectivity''.
-Note that cells that have overlapping vertical faces but are in different layers (e.g., cells 1A and 2B) do not have a direct hydraulic connection in a grid with layered connectivity.
-The addition of such connections between cells in different layers, as indicated by the red lines that connect cell centers in figure \ref{fig:schem-conn-area-flux}a, results in a grid with ``full connectivity''.
-In MODFLOW 6, layered connectivity is represented using the structured (DIS), vertex-based (DISV), or fully unstructured (DISU) grid types, whereas full connectivity can be represented only using the DISU grid type.
-
-% CDL: it seems like we could remove this following text.  
-%The ability to handle full connectivity was first introduced to MODFLOW by \cite{modflowusg} in MODFLOW-USG, and the term ``vertically staggered'' was first used by \cite{modflow6gwf} in the context of MODFLOW 6.
-The term describes a grid in which which a cell can have nominally ``horizontal'' connections with multiple cells across the same cell face.
-In this work, ``vertically offset'' and ``vertically staggered'' are used to distinguish grids that lack nominally ``horizontal'' connections between cells in adjacent layers (red lines in Figure \ref{fig:schem-conn-area-flux}a) from those that include such connections, respectively.
-In the vertically offset grid in figure \ref{fig:schem-conn-area-flux}a (blue connections only), the interfacial area across which flow occurs between adjacent cells in the same layer is based on the vertical cell thickness (the entire thickness of the channel).
-In the vertically staggered grid (blue and red connections), the interfacial area for nominally ``horizontal'' flow between adjacent cells is based on the area over which the two cell faces overlap.
-Thus, while the total area for nominally ``horizontal'' flow across a cell face is the same as for a vertically offset grid, that total area is divided between an adjacent cell in the same layer (blue connection) and an adjacent cell in the adjacent layer (red connection).
-In MODFLOW 6, vertically offset grids can be represented using the structured (DIS), vertex-based (DISV), or fully unstructured (DISU) grid type, whereas vertically staggered grids can be represented only using the DISU grid type.
+Figure \ref{fig:schem-conn-area-flux}a shows a group of four cells in a two-grid-layer, cross-sectional MODFLOW 6 model of fully saturated flow through a permeable hydrogeologic layer (aquifer) with impermeable top and bottom boundaries.
+The cells have horizontal tops and bottoms and can therefore follow the dip of the hydrogeologic layer only on average.
+If the grid is represented using a layered grid type (DIS or DISV) in MODFLOW 6, a cell is hydraulically connected to each adjacent cell in the same grid layer, regardless of whether or not the cells overlap, and with each overlying and underlying cell, as indicated by the blue lines that connect cell centers in Figure \ref{fig:schem-conn-area-flux}a.
+In this work we call this a vertically offset grid with ``layered connectivity.''  Note that cells that have overlapping vertical faces but are in different grid layers (e.g., cells 1A and 2B) do not have a direct hydraulic connection in a grid with layered connectivity.
+The addition of such connections between cells in different grid layers, as indicated by the red lines that connect cell centers in Figure \ref{fig:schem-conn-area-flux}a, results in a grid with ``full connectivity.''  In MODFLOW 6, only layered connectivity can be represented in structured (DIS) and vertex-based (DISV) grids.
+Fully unstructured (DISU) grids allow representation of any kind of cell connectivity, including layered or full connectivity.
 
@@
 ***}, the flow is proportional to the difference in the hydraulic heads computed at the two cell centers and a hydraulic conductance based on an effective hydraulic conductivity for the connection between the cells.
-On strictly rectilinear grids, such as structured MODFLOW grids with no vertical offset between cells in the same layer, with hydraulic conductivity that is isotropic or has anisotropy aligned with the three mutually perpendicular grid directions, the conductance-based formulation is second-order accurate \citep{dehotin2010modeling, modflow6gwf}.
-This is because strictly rectilinear grids satisfy the ``control-volume finite-difference (CVFD) requirement'' that the straight-line connection between two cell centers must intersect the midpoint of the cell interface at a right angle \citep{narasimhan1976integrated}.
-However, vertically offset grids violate the CFVD requirement because nominally ``horizontal'' connections between cell centers in the same layer are not strictly horizontal and therefore not perpendicular to the vertical cell interfaces they intersect.
-For example, in {\color{red} AMP: capitalize ``Figure'' throughout} Figure \ref{fig:schem-conn-area-flux}a, the connection between cells 1A and 1B is not perpendicular to the cell interface, which compromises the accuracy of the conductance-based formulation.
+On strictly rectilinear grids, such as structured MODFLOW grids with no vertical offset between cells in the same grid layer, with hydraulic conductivity that is isotropic or has anisotropy aligned with the three mutually perpendicular grid directions, the conductance-based formulation is second-order accurate \citep{dehotin2010modeling, modflow6gwf}.
+This is because strictly rectilinear grids satisfy the ``control-volume finite-difference (CVFD) requirement'' that the straight-line connection between two cell centers must intersect the midpoint of the cell-cell interface at a right angle \citep{narasimhan1976integrated}.
+However, vertically offset grids violate the CFVD requirement because nominally ``horizontal'' connections between cell centers in the same grid layer are not strictly horizontal and therefore not perpendicular to the vertical cell-cell interfaces they intersect.
+For example, in Figure \ref{fig:schem-conn-area-flux}a, the connection between cells 1A and 1B is not perpendicular to the cell-cell interface, which compromises the accuracy of the conductance-based formulation.
 
@@
 The concept of a ``ghost node'' was included in MODFLOW-USG \citep{modflowusg}, and subsequently in MODFLOW 6, as a simple and optional way to improve the accuracy for grids that violated the CVFD requirement.
-As an alternative to the ghost-node method for improving accuracy, and as a way to allow accurate simulation of arbitrarily oriented two- or three-dimensional anisotropy, the optional XT3D flow formulation \citep{modflow6xt3d} was introduced in MODFLOW 6.
-XT3D formulates the flow between two cells by interpolating head values from the two cells and their neighboring cells to construct an estimate of the full, two- or three-dimensional head-gradient vector on each side of the cell interface; applying the cell conductivity tensors to obtain an estimate of the groundwater flux  normal to the interface on each side of the interface; and reconciling the two flux estimates to ensure continuity of flow across the interface.
+However, implementation of the ghost-node correction is highly problem dependent, and it is not obvious how to specify optimal ghost-node locations and weights on complex, unstructured grids.
+As a more ``automatic'' alternative to the ghost-node method for improving accuracy, and as a way to allow accurate simulation of arbitrarily oriented two- or three-dimensional anisotropy, the optional XT3D flow formulation \citep{modflow6xt3d} was introduced in MODFLOW 6.
+XT3D formulates the flow between two cells by interpolating head values from the two cells and their neighboring cells to construct an estimate of the full, two- or three-dimensional head-gradient vector on each side of the cell-cell interface; applying the cell conductivity tensors to obtain an estimate of the groundwater flux  normal to the interface on each side of the interface; and reconciling the two flux estimates to ensure continuity of flow across the interface.
 By performing flow calculations using Darcy's Law in its tensorial form, based on a ``multi-point'' approximation of the head gradient, XT3D accounts for both arbitrarily oriented anisotropy and geometric irregularity of the grid.
 
-The dipping-channel benchmark problem of \cite{bardot2022} used a vertically offset layered grid, which is the type of grid used most often in MODFLOW models.
-Test simulations using the conductance-based flow formulation for a steeply dipping channel (hydrogeologic layer) embedded in a domain that is six orders of magnitude less hydraulically conductive yielded significant errors in the simulated flows.
-The simulated groundwater flux in the middle of the channel was not along the $30^{\circ}$ incline of the channel, but nearly horizontal ($0.03^{\circ}$ incline), and its magnitude was overestimated by 15\%.
+The dipping-hydrogeologic-layer benchmark problem of \cite{bardot2022} used a vertically offset layered grid, which is the type of grid used most often in MODFLOW models.
+Test simulations using the conductance-based flow formulation for a steeply dipping hydrogeologic layer (which we will call the ``aquifer'') embedded in a domain that is six orders of magnitude less hydraulically conductive yielded significant errors in the simulated flows.
+The simulated groundwater flux in the middle of the aquifer was not along the $30^{\circ}$ incline of the aquifer as expected, but nearly horizontal ($0.03^{\circ}$ incline), and its magnitude was overestimated by 15\%.
 Significant errors in the simulated flows were expected in this case, given that the conductance-based flow formulation does not rigorously account for violations of the CVFD requirement that the cell-cell connection be perpendicular to the cell-cell interface, and that it takes the connection length to be the horizontal distance between cell centers.
-Given the ability of XT3D to account rigorously for grid connections that are not perpendicular to cell interfaces, and for the increase in connection length due to the slope of the connection, rerunning the simulation with XT3D activated could have been expected to improve the flow solution substantially.
+Given the ability of XT3D to account rigorously for cell connections that are not perpendicular to cell-cell interfaces, and for the increase in connection length due to the slope of the connection, rerunning the simulation with XT3D activated could have been expected to improve the flow solution substantially.
 However, \cite{bardot2022} observed similarly poor results with XT3D: the flux was still nearly horizontal ($0.01^{\circ}$ incline), and its magnitude was overestimated by 12\%.
-Based on a preliminary analysis by two authors of the present work (Provost and Langevin), \cite{bardot2022} hypothesized that the unexpectedly large error in the flow solution with XT3D was related to inadequate grid connectivity.
+Based on a preliminary analysis by two authors of the present work (Provost and Langevin), \cite{bardot2022} hypothesized that the unexpectedly large error in the flow solution with XT3D was related to inadequate cell connectivity.
 
-The role of grid connectivity in enabling accurate simulation of flow along a dipping channel can be understood by considering Figure \ref{fig:schem-conn-area-flux}b, in which components of the groundwater flux (specific discharge) vector representative of steady, uniform flow along the channel are superimposed on each cell interface.
-Gray vectors represent the uniform flux oriented along the channel, which the model is attempting to simulate.
-Blue vectors represent the flux components normal to cell interfaces that correspond to blue connections in Figure \ref{fig:schem-conn-area-flux}a, which are common to both the layered and full connectivity grids.
-Red vectors represent the flux components normal to cell interfaces that correspond to red connections in Figure \ref{fig:schem-conn-area-flux}a, which are lacking in the layered connectivity grid.
+The role of cell connectivity in enabling accurate simulation of flow along a dipping hydrogeologic layer can be understood by considering Figure \ref{fig:schem-conn-area-flux}b, in which components of the groundwater flux (specific discharge) vector representative of steady, uniform flow along the hydrogeologic layer are superimposed on each cell-cell interface.
+Gray vectors represent the uniform flux oriented along the hydrogeologic layer, which the model is attempting to simulate.
+Blue vectors represent the flux components normal to cell-cell interfaces that correspond to blue connections in Figure \ref{fig:schem-conn-area-flux}a, which are common to both the layered-connectivity and full-connectivity grids.
+Red vectors represent the flux components normal to cell-cell interfaces that correspond to red connections in Figure \ref{fig:schem-conn-area-flux}a, which are lacking in the layered-connectivity grid.
 
-The blue vectors in Figure \ref{fig:schem-conn-area-flux}b show that to simulate steady, uniform flow along the dipping channel, cells must exchange water not only ``horizontally'' with neighboring cells in the same layer, but also vertically with cells in the adjacent layer.
-Specifically, there must be flow from cells in the bottom layer to cells in the top layer, which corresponds to the vertical component of the groundwater flux.
-Note, however, that the blue vectors alone are incompatible with steady, uniform flow along the channel because the exchange of water between the two model layers is in one direction only: upward.
-Such a steady-state flow configuration would imply continuous depletion of flow within the bottom layer and accumulation of flow within the top layer as one moves along the channel in the direction of flow.
-Thus, steady, uniform flow along the channel cannot be simulated accurately without some mechanism for returning flow from the top layer to the bottom layer.
-The full connectivity grid provides such a mechanism by offering additional connections between cells in adjacent layers.
-Vertical flows from the bottom layer to the top layer, represented by vertical blue vectors in Figure \ref{fig:schem-conn-area-flux}b, can be returned to the bottom layer by nominally ``horizontal'' flows, represented by red vectors in Figure \ref{fig:schem-conn-area-flux}b.
-Although this concept has been illustrated using a two-layer model for simplicity, the same reasoning applies given any number of layers.
+The blue vectors in Figure \ref{fig:schem-conn-area-flux}b show that to simulate steady, uniform flow along the dipping hydrogeologic layer, cells must exchange water not only ``horizontally'' with neighboring cells in the same grid layer, but also vertically with cells in the adjacent grid layer.
+Specifically, there must be flow from cells in the bottom grid layer to cells in the top grid layer, which corresponds to the vertical component of the groundwater flux.
+Note, however, that the blue vectors alone are incompatible with steady, uniform flow along the hydrogeologic layer because the exchange of water between the two grid layers is in one direction only: upward.
+Such a steady-state flow configuration would imply continuous depletion of flow within the bottom grid layer and accumulation of flow within the top grid layer as one moves along the hydrogeologic layer in the direction of flow.
+Thus, steady, uniform flow along the hydrogeologic layer cannot be simulated accurately without some mechanism for returning flow from the top grid layer to the bottom grid layer.
+The full-connectivity grid provides such a mechanism by offering additional connections between cells in adjacent grid layers.
+Vertical flows from the bottom grid layer to the top grid layer, represented by vertical blue vectors in Figure \ref{fig:schem-conn-area-flux}b, can be returned to the bottom grid layer by nominally ``horizontal'' flows, represented by red vectors in Figure \ref{fig:schem-conn-area-flux}b.
+Although this concept has been illustrated using a two-grid-layer model for simplicity, the same reasoning applies given any number of grid layers.
 
 Once the full connectivity is established, it is the role of the flow formulation to account properly for the geometry of the grid, a task for which XT3D was specifically designed.
-However, XT3D alone cannot compensate for grid connectivity that does not provide adequate pathways for flow.
-This explains why using XT3D did not improve the simulation results substantially on the layered connectivity grid in the benchmark tests of \cite{bardot2022}.
-In the middle section of the channel, away from the end boundary conditions, the flow solution approached steady, uniform flow the only way that it could given the limited connectivity: by suppressing vertical flows between layers, thereby rendering the overall flow approximately horizontal.
+However, XT3D alone cannot compensate for cell connectivity that does not provide adequate pathways for flow.
+This explains why using XT3D did not improve the simulation results substantially on the layered-connectivity grid in the benchmark tests of \cite{bardot2022}.
+In the middle section of the hydrogeologic layer, away from the end boundary conditions, the flow solution approached steady, uniform flow the only way that it could given the limited connectivity: by suppressing vertical flows between grid layers, thereby rendering the overall flow approximately horizontal.
 
-The arguments presented in this section suggest that, in addition to the use of XT3D, full connectivity can be important for obtaining an accurate flow solution in MODFLOW 6 simulations that involve steeply dipping layers.
+The arguments presented in this section suggest that, in addition to the use of XT3D, full connectivity can be important for obtaining an accurate flow solution in MODFLOW 6 simulations that involve steeply dipping hydrogeologic layers.
 The next section presents results of simulations designed to test this hypothesis.
@@
 The test problem is patterned after the cross-sectional (transect) benchmark of \cite{bardot2022}.
-It attempts to simulate uniform flow in an inclined permeable channel embedded in a less permeable surrounding ``domain.'' The channel is of uniform width and hydraulic conductivity and is inclined relative to the horizontal.
+It attempts to simulate uniform flow in an inclined permeable hydrogeologic layer (``aquifer'') embedded in less permeable material (the ``domain'').
+The aquifer is of uniform width and hydraulic conductivity and is inclined relative to the horizontal.
 
-To simulate flow along a channel inclined at angle $\theta$ ($30^{\circ}$ for the base case), heads are specified at the centers of cells along the perimeter of the model using an analytical solution that corresponds to a unit head gradient:
+To simulate flow along an aquifer inclined at angle $\theta$ ($30^{\circ}$ for the base case), heads are specified at the centers of cells along the perimeter of the model using a formula that corresponds to a uniform, unit head gradient:
 \begin{equation}
@@
 \noindent where $h$ is head in meters and $x$ and $z$ are the horizontal and vertical model coordinates, respectively, in meters.
-The hydraulic conductivity of the channel is set to 1 m/d (meter per day) so that the analytical flow solution is a groundwater flux of 1 m/d along the channel.
-In the base case, the conductivity of the domain is $10^{-6}$ m/d to effectively isolate the channel hydraulically.
+The hydraulic conductivity of the aquifer is set to 1 m/d (meter per day) so that the analytical flow solution is a groundwater flux of 1 m/d along the aquifer.
+In the base case, the conductivity of the domain is $10^{-6}$ m/d to effectively isolate the aquifer hydraulically.
 
-Two methods of discretization are used to simulate the test problem: (1) vertically offset grid of the MODFLOW 6 DIS (structured) type, and (2) vertically offset grid of MODFLOW 6 DISU (unstructured) type with full connectivity.
+Two methods of discretization are used to simulate the test problem: (1) vertically offset grid of the MODFLOW 6 DIS (structured) type, and (2) vertically offset grid of MODFLOW 6 DISU (unstructured) type.
 The cell geometry in these two discretization methods is identical, but the cell connectivity is different.
-The DISU grid included additional ``cross-connections'' between cells in adjacent model layers, as described later in the paper.
-The input files for the DISU grid are formulated by starting with a DIS grid, converting it to an equivalent DISU grid, and adding cross-connections.
-The base case uses grids consisting of 11 columns and 9 layers of cells, with 3 layers within the channel, 3 layers above the channel, and 3 layers below the channel.
+The DIS grid uses layered connectivity, and the DISU grid includes the additional connections between cells in adjacent grid layers needed to realize full connectivity.
+The DISU grid is constructed by starting with a DIS grid, converting it to an equivalent DISU grid, and then implementing full connectivity.
+The base case uses grids consisting of 11 columns and 9 layers of cells, with 3 layers within the aquifer, 3 layers in the domain above the aquifer, and 3 layers in the domain below the aquifer.
 
 The above description represents the base case test problem.
-However, the results also include simulations designed to investigate the effects of varying the gridding resolution, the dip angle of the channel, and the conductivity contrast between the channel and the domain.
+However, the results also include simulations designed to investigate the effects of varying the gridding resolution, the dip angle of the aquifer, the conductivity contrast between the aquifer and the domain, and the anisotropy of the conductivity tensor.
 
@@
 The notebooks leverage the capabilities of FloPy \citep{bakker2016scripting, hughes2023flopy} to assist in setting up input for and processing results from the MODFLOW 6 simulations.
-A Python script for converting structured, DIS grids to unstructured, DISU grids is included.
+A Python script for converting structured, DIS grids to unstructured, DISU grids with full connectivity is included.
 
@@
 The test problem allows the numerical solution for head and fluxes to be compared with an analytical solution.
-This section presents results for the base case scenario when comparing gridding strategy (i.e.
-layered connectivity versus full connectivity) and flow solution formulation (i.e.
-standard conductance based formulation versus XT3D option) to test the hypothesis that a grid with full connectivity produces a more accurate solution for dipping layers than a grid with only layered connectivity.
+This section presents results for the base case scenario when comparing gridding strategy (i.e., layered connectivity versus full connectivity) and flow solution formulation (i.e., standard conductance based formulation versus XT3D option) to test the hypothesis that a grid with full connectivity produces a more accurate solution for dipping hydrogeologic layers than a grid with only layered connectivity.
 The base case scenario is then modified to further investigate other aspects including grid resolution, dip angle, conductivity contrast and anisotropy.
 
-Results for flux magnitude, flux direction and volumetric flow through the channel are used to assess the accuracy of the numerical representations of the test problem.
+Results for flux magnitude, flux direction and volumetric flow through the aquifer are used to assess the accuracy of the numerical representations of the test problem.
 Note that the terms ``flux''  and ``specific discharge'' are used interchangeably.
-Reported fluxes for each model are for the central model cell and are therefore uninfluenced by the model boundary or the interface between the channel and the domain.
-Volumetric flow through the channel is therefore also examined as it encompasses fluxes through the entire channel.
-This paper uses the sum of the flows reported for the CHD cells along the left end of the channel as the volumetric flow.
+Reported fluxes for each model are for the central model cell and are therefore uninfluenced by the model boundary or the interface between the aquifer and the domain.
+Volumetric flow through the aquifer is therefore also examined as it encompasses fluxes through the entire aquifer.
+This paper uses the sum of the flows reported for the CHD cells along the left end of the aquifer as the volumetric flow.
 
@@
 The numerical solution for the grid with layered connectivity confirms the findings of \cite{bardot2022} in that fluxes are predominantly horizontal (top row, Figure \ref{fig:fig2}).
-We see that despite an enforced hydraulic gradient of $30^{\circ}$, the vertical component of the flux cannot propagate through the channel given that cells along the channel top are hydraulically disconnected from the low permeability domain, hence forcing flow horizontally (red arrows on right panel).
+We see that despite an enforced hydraulic gradient of $30^{\circ}$, the vertical component of the flux cannot propagate through the aquifer given that cells along the aquifer top are hydraulically disconnected from the low permeability domain, hence forcing flow horizontally (red arrows on right panel).
 The numerical solution for this grid type is a problem of cell connectivity, and thus the XT3D formulation does not resolve this issue (second row, Figure \ref{fig:fig2}).
-Results for the full connectivity grid with the standard conductance-based formulation shows an improvement in the flow solution (third row, Figure \ref{fig:fig2}), with the XT3D formulation successfully reproducing the analytical solution (bottom row, Figure \ref{fig:fig2}).
-We see that the grid with full connectivity allows incorporation of the vertical flow component, thus permitting cross-connection between model layers.
+Results for the full-connectivity grid with the standard conductance-based formulation shows an improvement in the flow solution (third row, Figure \ref{fig:fig2}), with the XT3D formulation successfully reproducing the analytical solution (bottom row, Figure \ref{fig:fig2}).
+We see that the grid with full connectivity allows incorporation of the vertical flow component, thus permitting cross-connection between grid layers.
 
@@
 The second is the cell-averaged specific discharge for each cell in its directional components ($q_{x}$, $q_{y}$, $q_{z}$) (``DATA-SPDIS''), and is calculated using distance-weighted averages of the cell face flows.
-Therefore, the specific discharge magnitude and direction normally reported by MODFLOW for cells within and along the channel edge are ``dampened'' due to averaging of the face flows from the cell top and bottom.
+Therefore, the specific discharge magnitude and direction normally reported by MODFLOW for cells within and along the aquifer edge are ``dampened'' due to averaging of the face flows from the cell top and bottom.
 The result is a ``staircase'' effect whereby these edge cells have a reduced flux magnitude and ``flattened'' flux directions, as shown in Figure 3c of \cite{bardot2022}.
-The plots in Figure 2 display corrected cell-averaged specific discharges which exclude the near-zero face flows along the channel boundary, and therefore represents the actual simulated flux magnitude and direction within the cell.
+The plots in Figure \ref{fig:fig2} display corrected cell-averaged specific discharges which exclude the near-zero face flows along the aquifer boundary, and therefore represents the actual simulated flux magnitude and direction within the cell.
 
@@
 	\includegraphics[scale=0.8]{../figures/fig2_paper.png}
-	\caption{{\color{red} (CDL: replace ``vertically offset'' with ``layered connectivity'' and ``vertically staggered'' with ``full connectivity''.)} Numerical results for the test problem using base case settings for layered connectivity and full connectivity grids, with the standard conductance based formulation as well as the XT3D formulation.
+	\caption{Numerical results for the test problem using base case settings for layered-connectivity and full-connectivity grids, with the standard conductance based formulation as well as the XT3D formulation.
 The left panel of each scenario shows the calculated specific discharge at each cell centre (black arrows) with head contours (orange).
@@
 
-{\color{red} Should we also consider the horizontal discretisation too? (CDL: I would say no.  Paper is already getting a bit long.)}
+{\color{red} Should we also consider the horizontal discretisation too?
+(CDL: I would say no.
+Paper is already getting a bit long.) (AMP: I agree with CDL.
+But let's be sure to mention that we've only looked at square cells in the aquifer.)}
 
@@
 	\includegraphics[scale=0.9]{../figures/fig3paper.png}
-	\caption{{\color{red} (CDL: replace ``vertically offset'' with ``layered connectivity'' and ``vertically staggered'' with ``full connectivity.'')} Graphs indicating how increasing the number of model layers within a single hydrogeological layer (in this case, the channel) affects (a) flux magnitude of middle cell, (b) flux direction of middle cell and (c) volumetric flow through the channel.
-We immediately notice the errors layered connectivity grids incur (blue line), and how using a full connectivity grid reproduces the analytical solution from two model layers used to discretise the channel (orange).}
+	\caption{Graphs indicating how increasing the number of grid layers within a single hydrogeologic layer (in this case, the aquifer) affects (a) flux magnitude of middle cell, (b) flux direction of middle cell and (c) volumetric flow through the hydrogeologic layer.
+We immediately notice the errors layered-connectivity grids incur (blue line), and how using a full-connectivity grid reproduces the analytical solution from two grid layers used to discretise the hydrogeologic layer (orange).}
 	\label{fig:fig3}
@@
 
-The base scenario arbitrarily uses 3 model layers to represent the channel, but it is important to address the question: \emph{How many model layers are required to adequately represent a hydrogeological layer?}.
-Layered connectivity grids, which have been used by groundwater modellers for decades (ref?) {\color{red} Jim to check, I think there is some refinement guidance for aquitards}, are almost always configured as one model layer per hydrogeological unit.
-However, although computationally efficient, we investigate the effect of altering resolution within the channel by using the base case and varying number of cells per channel width from one to nine (Figure \ref{fig:fig3}).
+The base scenario arbitrarily uses 3 grid layers to represent the aquifer, but it is important to address the question: \emph{How many grid layers are required to adequately represent a hydrogeological layer?}.
+Layered-connectivity grids, which have been used by groundwater modellers for decades (ref?) {\color{red} Jim to check, I think there is some refinement guidance for aquitards}, are almost always configured as one grid layer per hydrogeological unit.
+However, although computationally efficient, we investigate the effect of altering resolution within the aquifer by using the base case and varying number of cells per aquifer width from one to nine (Figure \ref{fig:fig3}).
 We compare modeled flux magnitude and direction in the center cell, as well as the volumetric flow against the analytical solution.
 
-Flux magnitude error for layered connectivity grids peaked at around 10\% but improved and tended to zero with refinement.
-Flux direction for full connectivity grids is clearly a major issue, being reported as horizontal for one model layer per channel width, and only improved to 12$^{\circ}$ for 9 layers per channel width.
-Flow through the channel peaked at 13\% error but decreased to only a 4\% error for 9 layers per channel width.
+Flux magnitude error for layered-connectivity grids peaked at around 10\% but improved and tended to zero with refinement.
+Flux direction for full-connectivity grids is clearly a major issue, being reported as horizontal for one grid layer per aquifer width, and only improved to 12$^{\circ}$ for 9 grid layers per aquifer width.
+Flow through the aquifer peaked at 13\% error but decreased to only a 4\% error for 9 grid layers per aquifer width.
 
-The results for the full connectivity grid clearly shows that a minimum of two layers per channel width is required to match the analytical result for flux magnitude, direction and volumetric flow.
-This is an important finding as it demonstrates that computational efficiency can be made using a full connectivity grid over a rectilinear grid overlay approach, but it is important to use a minimum of two model layers per hydrogeological unit to allow vertical fluxes to be incorporated into the flow solution.
+The results for the full-connectivity grid clearly show that a minimum of two grid layers per aquifer width is required to match the analytical result for flux magnitude, direction and volumetric flow.
+This is an important finding as it demonstrates that computational efficiency can be made using a full-connectivity grid over a rectilinear grid overlay approach, but it is important to use a minimum of two grid layers per hydrogeological unit to allow vertical fluxes to be incorporated into the flow solution.
 
@@
 
-The base case scenario uses a dip angle of 30$^{\circ}$ and an extreme conductivity contrast between channel and domain of $1:10^{-6}$.
-Here, we investigate the behavior of the flow solution for a wide range of channel inclines by systematically changing the dip angle by 2.5$^{\circ}$ from 0 to 80 $^{\circ}$.
-Similarly, we examine the solution for multiple conductivity contrasts by changing the domain conductivity to 2, 5, 10 and 100 times less than the channel.
+The base case scenario uses a dip angle of 30$^{\circ}$ and an extreme conductivity contrast between aquifer and domain of $1:10^{-6}$.
+Here, we investigate the behavior of the flow solution for a wide range of aquifer inclines by systematically changing the dip angle by 2.5$^{\circ}$ from 0 to 80 $^{\circ}$.
+Similarly, we examine the solution for multiple conductivity contrasts by changing the domain conductivity to 2, 5, 10 and 100 times less than the aquifer.
 Simulated flux magnitude and direction of the centre cell for all combinations are plotted against the analytical solution in Figure \ref{fig:fig4}.
 
-Traditional layered connectivity grids without XT3D shows rapid diversion of flux magnitude and direction from the analytical solution as dip increases (Figure \ref{fig:fig4a}).
+Traditional layered-connectivity grids without XT3D shows rapid diversion of flux magnitude and direction from the analytical solution as dip increases (Figure \ref{fig:fig4a}).
 Use of the XT3D option keeps the flux magnitude somewhat on-track until about 30 $^{\circ}$ when the solution starts peeling away from the analytical.
 Flux direction is improved for the homogeneous scenario, but not for the heterogeneous scenarios.
-The full connectivity grid on the other hand, even without XT3D, roughly follows the analytical solution for flux magnitude and direction until around 65 $^{\circ}$ when it starts to ``fall-apart.'' There are also spikes at certain dip angles which are related to transitions between cell connections due to changing cell face overlaps.
-The full connectivity grid with XT3D clearly boasts the best results with the flux magnitude and direction matching the analytical until 65 $^{\circ}$, with the exception of a slight decrease in flux direction after 45 $^{\circ}$.
+The full-connectivity grid on the other hand, even without XT3D, roughly follows the analytical solution for flux magnitude and direction until around 65 $^{\circ}$ when it starts to ``fall-apart.'' There are also spikes at certain dip angles which are related to transitions between cell connections due to changing cell face overlaps.
+The full-connectivity grid with XT3D clearly boasts the best results with the flux magnitude and direction matching the analytical until 65 $^{\circ}$, with the exception of a slight decrease in flux direction after 45 $^{\circ}$.
 
-These results confirm that full connectivity grids with XT3D not only outperform the alternatives when modelling dipping hydrogeological unit, but come extremely close to the analytical solution.
+These results confirm that full-connectivity grids with XT3D not only outperform the alternatives when modelling dipping hydrogeological unit, but come extremely close to the analytical solution.
 
@@
 	\includegraphics[width=\textwidth]{../figures/fig4_0_paper.png}
-	\caption{layered connectivity grid, standard conductance-based formulation}
+	\caption{layered-connectivity grid, standard conductance-based formulation}
 	\label{fig:fig4a}
@@
 	\includegraphics[width=\textwidth]{../figures/fig4_1_paper.png}
-	\caption{layered connectivity grid, XT3D}
+	\caption{layered-connectivity grid, XT3D}
 	\label{fig:fig4b}
@@
 	\includegraphics[width=\textwidth]{../figures/fig4_2_paper.png}
-	\caption{full connectivity grid, standard conductance-based formulation}
+	\caption{full-connectivity grid, standard conductance-based formulation}
 	\label{fig:fig4c}
@@
 	\includegraphics[width=\textwidth]{../figures/fig4_3_paper.png}
-	\caption{full connectivity grid, XT3D}
+	\caption{full-connectivity grid, XT3D}
 	\label{fig:fig4d}
@@
 
-\caption{{\color{red} (CDL: replace ``vertically offset'' with ``layered connectivity'' and ``vertically staggered'' with ``full connectivity''.)}Flux magnitude and direction for centre cells for varying contrasts in log conductivity (blue, orange, green, red lines) for varying dip angles (x-axis).
+\caption{Flux magnitude and direction for centre cells for varying contrasts in log conductivity (blue, orange, green, red lines) for varying dip angles (x-axis).
 The analytical value is shown in black dashed line.}
@@
 The base case and subsequent scenarios have assumed isotropic conductivity.
-However, sedimentary layers often exhibit anisotropic conductivity, and therefore we examine the gridding strategies using anisotropic and dipping conductivity tensors, with reduced conductivity perpendicular to the channel by 10, 100, 1000 and 10,000.
+However, sedimentary layers often exhibit anisotropic conductivity, and therefore we examine the gridding strategies using anisotropic and dipping conductivity tensors {\color{red} (AMP: throughout, or only in the aquifer?)}, with reduced conductivity perpendicular to the aquifer by 10, 100, 1000 and 10,000.
 We also include the isotropic scenario (ratio of 1).
 Results are presented in Figure \ref{fig:fig5}.
-Layered connectivity grids produce flux magnitude and flow results that significantly deviate from the analytical solution (blue line), whilst full connectivity grids reproduce the analytical solution (orange line).
-These results verify, once again, our proposed method of using full connectivity grids to model dipping anisotropic layers.
+Layered-connectivity grids produce flux magnitude and flow results that significantly deviate from the analytical solution (blue line), whilst full-connectivity grids reproduce the analytical solution (orange line).
+These results verify, once again, our proposed method of using full-connectivity grids to model dipping anisotropic hydrogeologic layers.
 
@@
 	\includegraphics[scale=0.9]{../figures/fig5paper.png}
-	\caption{Varying anisotropy ratios of conductivity withing the channel and the resulting (a) flux magnitude of middle cell, (b) flux direction of middle cell and (c) volumetric flow through the channel.
+	\caption{Varying anisotropy ratios of conductivity withing the aquifer and the resulting (a) flux magnitude of middle cell, (b) flux direction of middle cell and (c) volumetric flow through the aquifer.
 }
@@
 They concluded that the XT3D multi-point flux approximation in MODFLOW 6 significantly improves the accuracy of simulated flows for rectilinear grid overlay approaches in which locations of sedimentary structures are mapped onto a relatively fine model grid.
-They also concluded that the XT3D multi-point flux approximation did not perform as well as anticipated for vertically offset grids, which are advantageous for significantly reducing the number of layers required for many problems.
+They also concluded that the XT3D multi-point flux approximation did not perform as well as anticipated for vertically offset grids, which are advantageous for significantly reducing the number of grid layers required for many problems.
 They hypothesized that the inherent connectivity in DIS and DISV layered grids in MODFLOW 6 was inadequate for allowing the XT3D multi-point flux approximation to work as intended.
@@
 However, when additional cross connections are added to comprise a grid with full connectivity, there is a substantial increase in the accuracy of simulated flows.
-Importantly, this accuracy improvement requires that at least two model layers are used to represent the steeply dipping sedimentary structure.
+Importantly, this accuracy improvement requires that at least two grid layers are used to represent the steeply dipping sedimentary structure.
 The primary conclusion of this paper, therefore, is that that the XT3D multi-point flux approximation implemented in MODFLOW 6 can be combined with vertically offset grids to efficiently and accurately model flows through sedimentary structures.
@@
 This paper presents a solution for accurately representing steeply dipping flow structures groundwater utilising MODFLOW 6.
-\citep{bardot2022} demonstrated that the flux magnitude in highly dipping scenarios was erroneous when utilising layered connectivity grids.
+\citep{bardot2022} demonstrated that the flux magnitude in highly dipping scenarios was erroneous when utilising layered-connectivity grids.
 {\color{red} CDL: Also applicable for MODFLOW-USG DISU grids.} More significantly, the direction of fluxes were inaccurate, neglecting the vertical component of the flux.
-The issues observed in the existing study of \citep{bardot2022} partially related to the implementation of only within-layer horizontally connections, limiting the model capacity to simulate the vertical component of the flow.
-The present study developed an approach which utilised full connectivity grids \cite{modflow6gwf} to overcome the perceived problems.
-Through comparison of the previous layered connectivity and proposed full connectivity approached to the analytical solution presented in \citep{bardot2022}, we demonstrate that the use of full connectivity grids (in conjunction with the XT3D formulation) is an accurate method to simulate flow through dipping hydrogeological layers.
-However, accurate representation of flows requires a minimum of two model layers per hydro geological unit to allow for vertical connections to induce flows.
+The issues observed in the existing study of \citep{bardot2022} partially related to the implementation of only within-grid-layer horizontal connections, limiting the model capacity to simulate the vertical component of the flow.
+The present study developed an approach which utilised full-connectivity grids \cite{modflow6gwf} to overcome the perceived problems.
+Through comparison of the previous layered-connectivity and proposed full-connectivity approach to the analytical solution presented in \citep{bardot2022}, we demonstrate that the use of full-connectivity grids (in conjunction with the XT3D formulation) is an accurate method to simulate flow through dipping hydrogeologic layers.
+However, accurate representation of flows requires a minimum of two grid layers per hydrogeologic unit to allow for vertical connections to induce flows.
 
@@
 
-3) (Random thought) The use of full connectivity grids improved the accuracy of the transect solutions; however, the results of \citep{bardot2022} using unstructured meshes that conformed to the channel geometry preformed better in isotropic grids.
-If you force the normal flux to be the same at the interface of the domain and channel, the analytical solution would suggest a uniform gradient either side, do we introduce an error if the faces don't align with the channel correctly?
-Would figure 2 show a correction for this if the anisotropic scenario was tested?
+3) (Random thought) The use of full-connectivity grids improved the accuracy of the transect solutions; however, the results of \citep{bardot2022} using unstructured meshes that conformed to the aquifer geometry preformed better in isotropic grids.
+If you force the normal flux to be the same at the interface of the domain and aquifer, the analytical solution would suggest a uniform gradient either side, do we introduce an error if the faces don't align with the aquifer correctly?
+Would figure \ref{fig:fig2} show a correction for this if the anisotropic scenario was tested?
 

### 2023-10-19 latex/article.tex
abstract and article impact statement

@@
 
-\textbf{Article impact statement:} {\color{red} AMP: Need impact statement}
+\textbf{Article impact statement: Groundwater flow through dipping sedimentary structures can be accurately simulated using the XT3D multi-point flow formulation in MODFLOW 6.}
 
 \begin{abstract}
-\noindent {\color{red} AMP: compose Abstract near the end}
+Groundwater flow through dipping hydrogeologic layers can be modeled with MODFLOW by specifying top and bottom elevations of model layers to coincide with top and bottom elevations of hydrogeologic layers.
+With this approach, adjacent cells within a model layer are vertically offset from one another.
+While the standard formulation in MODFLOW does not strictly handle these offsets in the underlying flow calculations, results are generally considered to be accurate provided dip angles do not exceed $10^{\circ}$.
+To improve the accuracy of flow simulations on unstructured grids, and to allow accurate simulation of arbitrarily oriented two- or three-dimensional anisotropy, the optional XT3D multi-point flow formulation was introduced in MODFLOW 6.
+The XT3D flow formulation was designed to handle grid irregularities, such as vertical offsets between laterally adjacent cells, and to provide accurate results for both isotropic and anisotropic groundwater flow.
+\cite{bardot2022} showed, however, that XT3D simulations did not produce the correct flow direction and magnitude for simple tests involving flow through a steeply dipping permeable channel.
+In this paper, we identify the cause of the poor XT3D performance reported by \cite{bardot2022} for the steeply dipping test problem: for XT3D to perform as intended, the full cell connectivity is required for vertically offset grids.
+The common discretization packages in MODFLOW 6 are based on layered connectivity in which cells are laterally connected only with cells in the same model layer.
+When the fully unstructured grid type in MODFLOW 6 is used for this problem, and when additional cross connections are specified to represent lateral connections between cells in different model layers, XT3D is shown to give the correct solution both in terms of flow magnitude and direction.
+An important requirement for this improved accuracy, which is available for both isotropic and anisotropic conditions, is that permeable hydrogeologic features must be represented with at least two model layers.
+The combined use of vertically offset grids and the XT3D formulation in MODFLOW 6 provides a computationally efficient and accurate method for simulating isotropic and anisotropic flow through sedimentary structures, a capability that was previously reserved to finite-element simulators.
 \end{abstract}

### 2023-10-23 latex/article.tex
provost edits
various edits aimed at fleshing things out, flow, and consistent style

--- 2023-10-19 latex/article.tex
@@
 
-\textbf{Article impact statement:} {\color{red} AMP: Need impact statement}
+\textbf{Article impact statement: Groundwater flow through dipping sedimentary structures can be accurately simulated using the XT3D multi-point flow formulation in MODFLOW 6.}
 
 \begin{abstract}
-\noindent {\color{red} AMP: compose Abstract near the end}
+Groundwater flow through dipping hydrogeologic layers is often modeled with MODFLOW by specifying top and bottom elevations of model layers to coincide with top and bottom elevations of hydrogeologic layers.
+With this approach, adjacent cells within a model layer are vertically offset from one another.
+While the standard formulation in MODFLOW does not rigorously account for these offsets in the underlying flow calculations, results are generally expected to be accurate provided dip angles do not exceed $10^{\circ}$.
+To improve the accuracy of flow simulations on unstructured grids, and to allow accurate simulation of arbitrarily oriented two- or three-dimensional anisotropy, the optional XT3D multi-point flow formulation was introduced in MODFLOW 6.
+The XT3D flow formulation is designed to account for geometric irregularities in the grid, including vertical offsets between laterally adjacent cells, and to provide accurate results for both isotropic and anisotropic groundwater flow.
+\cite{bardot2022} showed, however, that neither the standard formulation nor the XT3D formulation produced the correct flow direction and magnitude in simple tests involving flow through a steeply dipping, permeable hydrogeologic layer.
+In this paper, we explain that the inability of XT3D to give accurate results in the steeply dipping test problem of \cite{bardot2022} was caused by the limited cell connectivity that is inherent in the most commonly used discretization packages in MODFLOW 6.
+The DIS and DISV discretization packages are based on layered connectivity in which cells have lateral connections only with cells in the same grid layer.
+We demonstrate that layered lateral connectivity is insufficient to allow accurate simulation of flow through a steeply dipping hydrogeologic layer on a vertically offset grid.
+When lateral connections between cells in adjacent grid layers are included using the fully unstructured (DISU) grid type, XT3D is able to calculate the flow magnitude and direction accurately for both isotropic flow and for anisotropy oriented along the permeable hydrogeologic layer.
+To realize this improved accuracy, the permeable hydrogeologic layer must be discretized vertically using at least two model layers.
+{\color{red} (AMP: Given the idealized nature of the test problem, how general can we be with our conclusions?
+I pulled it back here relative to what you had originally, CDL, but left a slightly stronger statement in the Conclusions, at least for now.})  Our findings suggest that, given appropriate discretization and cell connectivity, the combination of vertically offset grids with the XT3D flow formulation in MODFLOW 6 can provide a computationally efficient and accurate method for simulating isotropic and anisotropic flow through steeply dipping sedimentary structures, a capability that was previously reserved to finite-element-based simulators.
+{\color{red} (AMP: Any Pal and Edwards-type MPFA stuff that's relevant to that final phrase?
+See also Conclusions)}
 \end{abstract}
@@
 It attempts to simulate uniform flow in an inclined permeable hydrogeologic layer (``aquifer'') embedded in less permeable material (the ``domain'').
-The aquifer is of uniform width and hydraulic conductivity and is inclined relative to the horizontal.
 
-To simulate flow along an aquifer inclined at angle $\theta$ ($30^{\circ}$ for the base case), heads are specified at the centers of cells along the perimeter of the model using a formula that corresponds to a uniform, unit head gradient:
+To simulate flow along an aquifer of uniform thickness inclined at angle $\theta$ relative to the horizontal ($30^{\circ}$ for the base case), heads are specified at the centers of cells along the perimeter of the model using a formula that corresponds to a uniform, unit head gradient:
 \begin{equation}
@@
 The hydraulic conductivity of the aquifer is set to 1 m/d (meter per day) so that the analytical flow solution is a groundwater flux of 1 m/d along the aquifer.
-In the base case, the conductivity of the domain is $10^{-6}$ m/d to effectively isolate the aquifer hydraulically.
+In the base case, the conductivity of the domain surrounding the aquifer is set to $10^{-6}$ m/d to effectively isolate the aquifer hydraulically.
 
@@
 The base case uses grids consisting of 11 columns and 9 layers of cells, with 3 layers within the aquifer, 3 layers in the domain above the aquifer, and 3 layers in the domain below the aquifer.
+{\color{red} (AMP: Need dimensions of aquifer and domain.)}
 
-The above description represents the base case test problem.
+The above description represents the base-case test problem.
 However, the results also include simulations designed to investigate the effects of varying the gridding resolution, the dip angle of the aquifer, the conductivity contrast between the aquifer and the domain, and the anisotropy of the conductivity tensor.
@@
 
-The test problem allows the numerical solution for head and fluxes to be compared with an analytical solution.
-This section presents results for the base case scenario when comparing gridding strategy (i.e., layered connectivity versus full connectivity) and flow solution formulation (i.e., standard conductance based formulation versus XT3D option) to test the hypothesis that a grid with full connectivity produces a more accurate solution for dipping hydrogeologic layers than a grid with only layered connectivity.
-The base case scenario is then modified to further investigate other aspects including grid resolution, dip angle, conductivity contrast and anisotropy.
+{\color{red} (AMP: Did some editing of this section for flow and consistent style, but I'll have another go once we're satisfied that all the pieces are basically in place.
+We're really close, if not there already.)}
 
-Results for flux magnitude, flux direction and volumetric flow through the aquifer are used to assess the accuracy of the numerical representations of the test problem.
-Note that the terms ``flux''  and ``specific discharge'' are used interchangeably.
-Reported fluxes for each model are for the central model cell and are therefore uninfluenced by the model boundary or the interface between the aquifer and the domain.
-Volumetric flow through the aquifer is therefore also examined as it encompasses fluxes through the entire aquifer.
-This paper uses the sum of the flows reported for the CHD cells along the left end of the aquifer as the volumetric flow.
+To demonstrate the effect of cell connectivity on the accuracy of simulated flow through a steeply dipping hydrogeologic layer (aquifer), results for the base case described in the previous section are compared for simulations that use either layered or full connectivity, and either the standard or the XT3D flow formulation.
+The base case is then modified to further investigate the effects of grid resolution, aquifer dip angle, conductivity contrast between the aquifer and the surrounding domain, and anisotropic hydraulic conductivity.
 
-\subsection{Gridding Strategy and Flow Formulation Comparison}
+Plots of specific discharge (groundwater flux) at cell centers and flows across cell-cell interfaces (face flows) calculated by MODFLOW 6 are used to illustrate the accruacy of the simulated flows.
+The magnitude and direction of specific discharge at the cell in the center of the aquifer, at which boundary effects are presumably minimized, and the total volumetric flow rate entering the aquifer through the left-hand boundary, which consists of constant-head (CHD) cells, are also reported as ``representative'' values of flow.
+{\color{red} (AMP: Heads are also plotted.
+Do we want to also report, say, maximum error in head?)}
+
+Specific discharge at cell centers is calculated in a postprocessing step using distance-weighted interpolation of face flows calculated by MODFLOW 6 and reported in the binary budget output file.
+For the purpose of illustrating and evaluating the results of this idealized benchmark, specific discharge in cells within the aquifer is calculated using only face flows between cells within the aquifer.
+Likewise, specific discharge in cells within the surrounding domain is calculated using only face flows between cells within the domain.
+Specific discharge values reported in the binary budget output file by MODFLOW 6 are different than the values reported here for cells that are adjacent to the boundary between the aquifer and the domain.
+This is because MODFLOW interpolates specific discharge in a cell using all of the face flows for that cell, including flows across cell faces on the aquifer-domain boundary.
+Those flows represent an ``average'' of the different flows on either side of the aquifer-domain boundary.
+Therefore, specific discharge calculated in a cell adjacent to the aquifer-domain boundary is affected by the different flow rate on the other side of the boundary.
+While this is an unavoidable consequence of vertically offset discretization in practical applications, and not necessarily a problem, users should be aware of these effects when interpreting the results of MODFLOW 6 dispersive-transport and MODPATH particle-tracking simulations.
+
+\subsection{Effects of Cell Connectivity and Flow Formulation}
 
@@
 
-Through working on this test problem, we have encountered some nuances in working with specific discharges calculated using MODFLOW.
+The results shown in the top four panels of Figure \ref{fig:fig2} {\color{red} (AMP: Label the figure panels a through h so we can refer to them by letter?)}, which are based on a vertically offset grid with layered connectivity, are consistent with the findings of \cite{bardot2022}.
+Despite boundary heads designed to induce specific discharge of 1 m/d along the $30^{\circ}$ incline of the aquifer, the simulated flow within the aquifer is predominantly horizontal, and the magnitude of the specific discharge at the center of the aquifer and the total flow entering the aquifer are both overestimated.
+The XT3D flow formulation is unable to compensate for the inadequate cell connectivity, which does not allow flow between laterally adjacent cells in different grid layers.
+
+The results shown in the bottom four panels of Figure \ref{fig:fig2}, which are based on a vertically offset grid with full connectivity, show significantly improved results, even for the standard flow formulation.
+Using the XT3D flow formulation provides the most accurate results, as expected.
+
+%Through working on this test problem, we have encountered some nuances in working with specific discharges calculated using MODFLOW.
 There are two main MODFLOW output files pertaining to fluxes.
@@
 The plots in Figure \ref{fig:fig2} display corrected cell-averaged specific discharges which exclude the near-zero face flows along the aquifer boundary, and therefore represents the actual simulated flux magnitude and direction within the cell.
+{\color{red} (AMP: An edited version of this now appears near the beginning of the Results and Discussion section)}
 
@@
 	\includegraphics[scale=0.8]{../figures/fig2_paper.png}
-	\caption{Numerical results for the test problem using base case settings for layered-connectivity and full-connectivity grids, with the standard conductance based formulation as well as the XT3D formulation.
+	\caption{Numerical results for the test problem using base-case settings for layered-connectivity and full-connectivity grids, with the standard conductance based formulation as well as the XT3D formulation.
 The left panel of each scenario shows the calculated specific discharge at each cell centre (black arrows) with head contours (orange).
-The right panel shows the face flows (red arrows).}
+The right panel shows the face flows (red arrows).
+{\color{red} (AMP: Change ``Vertically Offset'' to ``Layered Connectivity'' and ``Vertically Staggered'' to ``Full Connectivity'' in the script that generates the figure)} {\color{red} (AMP: Magnify tiny vectors in the domain [by the conductivity contrast, say] so they can be seen?
+Face flows along the aquifer/domain boundary are tricky because they're ``in between'' in terms of magnitude.)}}
 	\label{fig:fig2}
@@
 
-\subsection{Grid Resolution}
+\subsection{Effects of Grid Resolution}
 
@@
 	\caption{Graphs indicating how increasing the number of grid layers within a single hydrogeologic layer (in this case, the aquifer) affects (a) flux magnitude of middle cell, (b) flux direction of middle cell and (c) volumetric flow through the hydrogeologic layer.
-We immediately notice the errors layered-connectivity grids incur (blue line), and how using a full-connectivity grid reproduces the analytical solution from two grid layers used to discretise the hydrogeologic layer (orange).}
+We immediately notice the errors layered-connectivity grids incur (blue line), and how using a full-connectivity grid reproduces the analytical solution from two grid layers used to discretise the hydrogeologic layer (orange).
+{\color{red} (AMP: Change ``Vertically Offset'' to ``Layered Connectivity'', ``Vertically Staggered'' to ``Full Connectivity'', and ``Model layers'' to ``Grid layers'' in the script that generates the figure)}}
 	\label{fig:fig3}
@@
 
-The base scenario arbitrarily uses 3 grid layers to represent the aquifer, but it is important to address the question: \emph{How many grid layers are required to adequately represent a hydrogeological layer?}.
+The base case arbitrarily uses 3 grid layers to represent the aquifer, but it is important to address the question: \emph{How many grid layers are required to adequately represent a hydrogeological layer?}.
 Layered-connectivity grids, which have been used by groundwater modellers for decades (ref?) {\color{red} Jim to check, I think there is some refinement guidance for aquitards}, are almost always configured as one grid layer per hydrogeological unit.
-However, although computationally efficient, we investigate the effect of altering resolution within the aquifer by using the base case and varying number of cells per aquifer width from one to nine (Figure \ref{fig:fig3}).
-We compare modeled flux magnitude and direction in the center cell, as well as the volumetric flow against the analytical solution.
+{\color{red} (AMP: Could maybe mention that it's generally well known you want multiple layers for accurate transport?)} However, although computationally efficient, we investigate the effect of altering resolution within the aquifer by using the base case and varying number of cells per aquifer width from one to nine {\color{red} (AMP: Explain how aquifer and domain dimensions vary with the discretization.
+And is that how we want them to vary?)} Figure \ref{fig:fig3}).
+{\color{red} (AMP: Are all results in this figure with XT3D?
+If so, would there be any benefit to showing the standard formulation results, as well?)} We compare modeled flux magnitude and direction in the center cell, as well as the volumetric flow against the analytical solution.
+{\color{red} (AMP: Should we also say something about the error in the head solution?
+Report the maximum head error in each case?)}
 
 Flux magnitude error for layered-connectivity grids peaked at around 10\% but improved and tended to zero with refinement.
-Flux direction for full-connectivity grids is clearly a major issue, being reported as horizontal for one grid layer per aquifer width, and only improved to 12$^{\circ}$ for 9 grid layers per aquifer width.
+Flux direction for layered-connectivity grids is clearly a major issue, being reported as horizontal for one grid layer per aquifer width, and only improved to 12$^{\circ}$ for 9 grid layers per aquifer width.
 Flow through the aquifer peaked at 13\% error but decreased to only a 4\% error for 9 grid layers per aquifer width.
@@
 
-\subsection{Dip angle and conductivity contrast}
+\subsection{Effects of Dip Angle and Conductivity Contrast}
 
-The base case scenario uses a dip angle of 30$^{\circ}$ and an extreme conductivity contrast between aquifer and domain of $1:10^{-6}$.
-Here, we investigate the behavior of the flow solution for a wide range of aquifer inclines by systematically changing the dip angle by 2.5$^{\circ}$ from 0 to 80 $^{\circ}$.
+The base case uses a dip angle of 30$^{\circ}$ and an extreme conductivity contrast between aquifer and domain of 1:$10^{-6}$.
+Here, we investigate the behavior of the flow solution for a wide range of aquifer inclines by systematically changing the dip angle by 2.5$^{\circ}$ from 0 to 80$^{\circ}$.
 Similarly, we examine the solution for multiple conductivity contrasts by changing the domain conductivity to 2, 5, 10 and 100 times less than the aquifer.
-Simulated flux magnitude and direction of the centre cell for all combinations are plotted against the analytical solution in Figure \ref{fig:fig4}.
+{\color{red} (AMP: Dare we try making the domain more conductive than the ``aquifer''?)}  Simulated flux magnitude and direction of the centre cell for all combinations are plotted against the analytical solution in Figure \ref{fig:fig4}.
 
 Traditional layered-connectivity grids without XT3D shows rapid diversion of flux magnitude and direction from the analytical solution as dip increases (Figure \ref{fig:fig4a}).
-Use of the XT3D option keeps the flux magnitude somewhat on-track until about 30 $^{\circ}$ when the solution starts peeling away from the analytical.
+Use of the XT3D option keeps the flux magnitude somewhat on-track until about 30$^{\circ}$ when the solution starts peeling away from the analytical.
 Flux direction is improved for the homogeneous scenario, but not for the heterogeneous scenarios.
-The full-connectivity grid on the other hand, even without XT3D, roughly follows the analytical solution for flux magnitude and direction until around 65 $^{\circ}$ when it starts to ``fall-apart.'' There are also spikes at certain dip angles which are related to transitions between cell connections due to changing cell face overlaps.
-The full-connectivity grid with XT3D clearly boasts the best results with the flux magnitude and direction matching the analytical until 65 $^{\circ}$, with the exception of a slight decrease in flux direction after 45 $^{\circ}$.
+The full-connectivity grid on the other hand, even without XT3D, roughly follows the analytical solution for flux magnitude and direction until around 65$^{\circ}$ when it starts to ``fall-apart.'' There are also spikes at certain dip angles which are related to transitions between cell connections due to changing cell face overlaps.
+The full-connectivity grid with XT3D clearly boasts the best results with the flux magnitude and direction matching the analytical until 65$^{\circ}$, with the exception of a slight decrease in flux direction after 45$^{\circ}$.
+{\color{red} (AMP: Can explain the wheels coming off around 65 deg by the fact that with only 3 cells vertically in the aquifer, the aquifer then looks like a long, zigzag Tetris piece that's no more than one cell wide, either vertically or horizontally.
+Finer vertical discretization improves things.)}
 
@@
 
-\subsection{Anisotropy}
+\subsection{Effect of Anisotropy}
 
 The base case and subsequent scenarios have assumed isotropic conductivity.
-However, sedimentary layers often exhibit anisotropic conductivity, and therefore we examine the gridding strategies using anisotropic and dipping conductivity tensors {\color{red} (AMP: throughout, or only in the aquifer?)}, with reduced conductivity perpendicular to the aquifer by 10, 100, 1000 and 10,000.
+However, sedimentary layers often exhibit anisotropic conductivity, and therefore we examine the gridding strategies using anisotropic and dipping conductivity tensors {\color{red} (AMP: in the domain and aquifer, or only in the aquifer?)}, with reduced conductivity perpendicular to the aquifer by 10, 100, 1000 and 10,000.
 We also include the isotropic scenario (ratio of 1).
-Results are presented in Figure \ref{fig:fig5}.
-Layered-connectivity grids produce flux magnitude and flow results that significantly deviate from the analytical solution (blue line), whilst full-connectivity grids reproduce the analytical solution (orange line).
+{\color{red} (AMP: Dare we try increased perpendicular conductivity?)} Results are presented in Figure \ref{fig:fig5}.
+{\color{red} (AMP: Are all results in this figure with XT3D?
+If so, would there be any benefit to showing the standard formulation results, as well?)} Layered-connectivity grids produce flux magnitude and flow results that significantly deviate from the analytical solution (blue line), whilst full-connectivity grids reproduce the analytical solution (orange line).
 These results verify, once again, our proposed method of using full-connectivity grids to model dipping anisotropic hydrogeologic layers.
@@
 \cite{bardot2022} revisited the capability of MODFLOW to accurately simulate groundwater flow through sedimentary structures.
-They concluded that the XT3D multi-point flux approximation in MODFLOW 6 significantly improves the accuracy of simulated flows for rectilinear grid overlay approaches in which locations of sedimentary structures are mapped onto a relatively fine model grid.
-They also concluded that the XT3D multi-point flux approximation did not perform as well as anticipated for vertically offset grids, which are advantageous for significantly reducing the number of grid layers required for many problems.
-They hypothesized that the inherent connectivity in DIS and DISV layered grids in MODFLOW 6 was inadequate for allowing the XT3D multi-point flux approximation to work as intended.
+They concluded that the XT3D multi-point flux approximation in MODFLOW 6 significantly improves the accuracy of simulated flows when used with a grid-overlay approach in which locations of sedimentary structures, which can be steeply dipping, are mapped onto a relatively fine, rectilinear model grid.
+They also concluded that the XT3D multi-point flux approximation does not perform as well as anticipated when simulating flow through a steeply dipping hydrogeologic layer using vertically offset grids, which are advantageous for significantly reducing the number of grid layers required for many problems.
+They hypothesized that the inherently limited lateral connectivity between cells in DIS and DISV layered grids in MODFLOW 6 was responsible for the inability to obtain accurate flow results in their steeply dipping test problems using vertically offsets grids, with or without XT3D.
 
-In this paper, we confirm that the layered connectivity implemented in DIS and DISV model grids is inadequate for simulating groundwater flow through steeply dipping sedimentary structures, even when the XT3D multi-point flux approximation is used.
-However, when additional cross connections are added to comprise a grid with full connectivity, there is a substantial increase in the accuracy of simulated flows.
-Importantly, this accuracy improvement requires that at least two grid layers are used to represent the steeply dipping sedimentary structure.
-The primary conclusion of this paper, therefore, is that that the XT3D multi-point flux approximation implemented in MODFLOW 6 can be combined with vertically offset grids to efficiently and accurately model flows through sedimentary structures.
-Prior to this work, the capability to efficiently and accurately model flow through sedimentary structures was generally restricted to finite element simulators.
+In this paper, we have confirmed that the layered lateral connectivity implemented in DIS and DISV model grids is inadequate for simulating groundwater flow through steeply dipping sedimentary structures, even when the XT3D multi-point flux approximation is used.
+However, when connections between laterally adjacent cells in different grid layers are added to comprise a grid with full connectivity, there is a substantial increase in the accuracy of simulated flows.
+The additional connectivity is implemented using the fully unstructured (DISU) grid type.
+Importantly, this accuracy improvement requires that at least two grid layers be used to represent the steeply dipping sedimentary structure.
+The primary conclusion of this paper is that, given appropriate discretization and cell connectivity, the XT3D multi-point flux approximation implemented in MODFLOW 6 can be combined with vertically offset grids to efficiently and accurately model flows through {\color{red} steeply dipping??} sedimentary structures.
+Prior to this work, the capability to efficiently and accurately model flow through {\color{red} steeply dipping??} sedimentary structures was generally restricted to finite element simulators.
 
-This paper presents a solution for accurately representing steeply dipping flow structures groundwater utilising MODFLOW 6.
-\citep{bardot2022} demonstrated that the flux magnitude in highly dipping scenarios was erroneous when utilising layered-connectivity grids.
-{\color{red} CDL: Also applicable for MODFLOW-USG DISU grids.} More significantly, the direction of fluxes were inaccurate, neglecting the vertical component of the flux.
-The issues observed in the existing study of \citep{bardot2022} partially related to the implementation of only within-grid-layer horizontal connections, limiting the model capacity to simulate the vertical component of the flow.
-The present study developed an approach which utilised full-connectivity grids \cite{modflow6gwf} to overcome the perceived problems.
-Through comparison of the previous layered-connectivity and proposed full-connectivity approach to the analytical solution presented in \citep{bardot2022}, we demonstrate that the use of full-connectivity grids (in conjunction with the XT3D formulation) is an accurate method to simulate flow through dipping hydrogeologic layers.
-However, accurate representation of flows requires a minimum of two grid layers per hydrogeologic unit to allow for vertical connections to induce flows.
+The importance of adequate connectivity has been explained and evaluated in this work specifically in terms of hydrogeologic layers that correspond to sedimentary structures.
+However, the impact of cell connectivity on the accuracy of simulated flows should be considered in any MODFLOW application that models flow through a heterogeneous hydraulic conductivity field using a grid with large vertical offsets between laterally adjacent cells, whether or not the grid has a layered structure.
+The full connectivity needed to improve accuracy can be implemented using the DISU grid types in both MODFLOW 6 and MODFLOW-USG.
 
@@
 
-2) The concepts in this paper have been easiest to discuss in terms of grid layers, but they're not limited to layered grids.
-
 3) (Random thought) The use of full-connectivity grids improved the accuracy of the transect solutions; however, the results of \citep{bardot2022} using unstructured meshes that conformed to the aquifer geometry preformed better in isotropic grids.
@@
 Would figure \ref{fig:fig2} show a correction for this if the anisotropic scenario was tested?
+{\color{red} (AMP: Is the underlying question whether using an unstructured mesh that conforms to the aquifer geometry [which MF6 can't do in transect] would improve the accuracy more than a full-connectivity vertically offset grid, assuming xt3d is used in both cases?
+Let's discuss on our next call.)}
 
@@
 \section{Acknowledgments}
-We would like to acknowledge....
+We would like to acknowledge.... {\color{red} that this simple paper has been in the works far too long. Kindly address complaints to the first author.}
 

### 2023-12-08 latex/article.tex
Comment
Add comment
Edits by JM and KB

@@
 
-%\title{Improving Accuracy of Simulated Flows in Steeply Dipping Layers Using Full-Connectivity Grids with the XT3D Multi-Point Flux Approximation in MODFLOW 6}
-
-% suggested title revision
-\title{Efficient and Accurate Modeling of Flow through Sedimentary Structures with MODFLOW 6}
+\title{Efficient and Accurate Modeling of Flow through Dipping Aquifers with MODFLOW 6}
 
@@
 
-\textbf{Key words:} groundwater flow simulation, sedimentary structures, dipping layers, full-connectivity grid {\color{red} AMP: maybe ``cell connectivity''' or ``grid connectivity'' would make a better search term?}, multi-point flux approximation
+\textbf{Key words:} groundwater flow simulation, discretization, dipping layers, cell connectivity, multi-point flux approximation, sedimentary structures
 
-\textbf{Article impact statement: Groundwater flow through dipping sedimentary structures can be accurately simulated using the XT3D multi-point flow formulation in MODFLOW 6.}
+\textbf{Article impact statement: Groundwater flow through dipping aquifers can be accurately simulated using a vertically offset grid combined with an unstructured grid approach and the XT3D multi-point flow formulation in MODFLOW 6.}
 
@@
 With this approach, adjacent cells within a model layer are vertically offset from one another.
-While the standard formulation in MODFLOW does not rigorously account for these offsets in the underlying flow calculations, results are generally expected to be accurate provided dip angles do not exceed $10^{\circ}$.
+While the standard formulation in MODFLOW does not rigorously account for these offsets in the underlying flow calculations, results are generally expected to be accurate provided dip angles do not exceed $10^{\circ}$  {\color{red} (KB: I would remove this sentence as even 10 degrees if anisotropic say 10:1 is still  a 50\% error)}.
 To improve the accuracy of flow simulations on unstructured grids, and to allow accurate simulation of arbitrarily oriented two- or three-dimensional anisotropy, the optional XT3D multi-point flow formulation was introduced in MODFLOW 6.
@@
 We demonstrate that layered lateral connectivity is insufficient to allow accurate simulation of flow through a steeply dipping hydrogeologic layer on a vertically offset grid.
-When lateral connections between cells in adjacent grid layers are included using the fully unstructured (DISU) grid type, XT3D is able to calculate the flow magnitude and direction accurately for both isotropic flow and for anisotropy oriented along the permeable hydrogeologic layer.
+When lateral connections between cells in adjacent grid layers are included using the fully unstructured (DISU) grid type, XT3D is able to calculate the flow magnitude and direction accurately for both isotropic and anisotropic conductivity oriented along the permeable hydrogeologic layer.
 To realize this improved accuracy, the permeable hydrogeologic layer must be discretized vertically using at least two model layers.
 {\color{red} (AMP: Given the idealized nature of the test problem, how general can we be with our conclusions?
-I pulled it back here relative to what you had originally, CDL, but left a slightly stronger statement in the Conclusions, at least for now.})  Our findings suggest that, given appropriate discretization and cell connectivity, the combination of vertically offset grids with the XT3D flow formulation in MODFLOW 6 can provide a computationally efficient and accurate method for simulating isotropic and anisotropic flow through steeply dipping sedimentary structures, a capability that was previously reserved to finite-element-based simulators.
+I pulled it back here relative to what you had originally, CDL, but left a slightly stronger statement in the Conclusions, at least for now.})  Our findings suggest that, given appropriate discretization and cell connectivity, the combination of vertically offset grids with the XT3D flow formulation in MODFLOW 6 can provide a computationally efficient and accurate method for simulating flow through isotropic and anisotropic steeply dipping layers, a capability that was previously reserved to finite-element-based simulators.
 {\color{red} (AMP: Any Pal and Edwards-type MPFA stuff that's relevant to that final phrase?
 See also Conclusions)}
+
+{\color{red} Here's another option for a simpler abstract! }
+
+	Dipping aquifers are often modeled using a deformed grid whereby each hydrogeological layer is represented using a single grid layer.
+However, until the release of MODFLOW 6, accurate modeling of groundwater flow through dipping anisotropic layers has not been possible due to limitations of its finite difference conductance-based solution.
+The introduction of the XT3D multi-point flow formulation has allowed accurate simulation of the full conductivity tensor, as well as handling the flow solution for irregular cell geometries for unstructured grids.
+Through the use of a benchmark, \cite{bardot2022} showcased MODFLOW 6's capability to model flow through complex sedimentary structures, previously restricted to finite element simulators.
+However, the benchmark study also brought to light the inadequacy of the deformed vertically offset grid for accurate modeling of dipping layers, with or without XT3D.
+This paper explains that the limitation of the deformed grid is a result of poor cell connectivity between vertically adjacent model layers.
+Nonetheless, we also demonstrate that dipping layers may be efficiently and accurately modeled using a deformed vertically offset grid with full-connectivity, which can be achieved by combining MODFLOW 6's DISU discretization package with the XT3D flow formulation.
+A repeat benchmark study shows that this method produces accurate head and flux results in isotropic and anisotropic dipping layers, providing a minimum of two grid layers per hydrogeological layer are deployed which allows for sufficient vertical connectivity.
+
 \end{abstract}
@@
 
-In the MODFLOW family of hydrologic simulation codes, a groundwater model domain is discretized into a grid of hydraulically connected model cells.
-Versions of MODFLOW up to and including MODFLOW-2005 \citep{modflow2005} offered only structured grids composed of rows, columns, and layers of cells.
-Cells were conceptualized as having horizontal top and bottom surfaces and vertical lateral faces, which allowed each cell to be ``simulated as if it were rectangular so that flow may be approximated by the standard finite-difference equation'' \citep{modflow84}.
-In MODFLOW-USG \citep{modflowusg} and MODFLOW 6 \citep{modflow6gwf}, which offer unstructured grids and use the control-volume finite-difference method to formulate the discrete balance equations, cells continue to have horizontal tops and bottoms and vertical lateral faces.
-Defining cells with horizontal tops and bottoms and vertical lateral faces simplifies the definition and calculation of water-table elevations, cell saturations, hydraulic conductances between cells, and transitions between confined and unconfined conditions.
+MODFLOW is commonly used to simulate groundwater flow in a wide variety of geological settings including sedimentary, alluvial and fractured rock aquifers \cite{modflow2005,Kumar2019}.
+Accurate simulation of the hydrostratigraphic sequences encountered in these geological settings requires that the model grid discretisation and solution technique can represent the scale and effects of important geologial features \cite{Reilly2004,bardot2022}, where permeability tensors are aligned along key structures \cite{hoaglund2003}.
+MODFLOW versions up to and including MODFLOW-2005 employ a structured grid based on a fixed number of layers, rows and columns \citep{modflow2005,modflow84,modflow88}.
+Further, through assuming that the hydraulic conductivity tensor aligns with the grid coordinates, the standard finite difference approximation can be applied accurately and efficiently to simulate groundwater flow \citep{modflow84}.
+The main trade-off in this approximation is that important hydrological and geological features are required to be mapped to the structured grids and orinetations need to be aligned to Cartesian coordinates.
+These requirements can result in pixelated features in plan view and, more significantly, create problems where geological layers are discontinuous at unconformities and faults, and where they are dipping and misaligned with the model grid \citep{bardot2022}.
 
-A traditional approach to representing an aquifer system with dipping hydrogeologic layers in MODFLOW is to map the hydrologic properties onto a rectilinear grid \citep{modflow84}.
-\cite{hoaglund2003} refer to this as the ``grid overlay'' method.
-In the grid overlay method, the boundaries of dipping hydrogeologic layers are represented approximately, in what \cite{bardot2022} call a ``staircase'' fashion.
-Another approach available in all versions of MODFLOW is to allow the top and bottom elevations of cells in the same layer of the grid to vary spatially.
-Following \cite{bardot2022}, we call this type of grid ``vertically offset.'' In this work, we show how to improve the accuracy of flow simulations that use vertically offset grids.
-The vertical offsets in top and bottom elevations between adjacent cells within a grid layer allows each grid layer to ``deform'' with the hydrostratigraphy.
-This ``minimize[s] the number of model [grid] layers required to simulate an aquifer system'' \citep{modflow84} relative to the grid overlay method.
+In meeting the requirements for the standard finite difference simulation, two methods have traditionally been used to represent dipping geological features in MODFLOW in versions up to and including MODFLOW-2005.
+One approach is to map the geological features to a rectilinear grid \citep{modflow84} in a method referred to as ‘grid overlay’  \citep{hoaglund2003}.
+Grid overlay preserves the control volume finite difference approximation but may require a high level of vertical discretization to accurately represent variable layering \citep{Zyvoloski2006}.
+An alternative approach is to set each hydrostratigraphic unit as its own model layer, resulting in 'deformed layers' \cite{anderson2015applied,Reilly2004} in a 'vertically offset' or 'conforming finite difference grid' \citep{Zyvoloski2006}.
+When the layers in the offset grid are not strictly horizontal and of a constant thickness, the standard finite difference approximation is violated \citep{Zheng1994,anderson2015applied,Reilly2004}.
+To maintain a reasonable level of accuracy, \cite{anderson2015applied} recommend a $10^{\circ}$ dip as a practical upper limit for using vertical offsets.
 
-All versions of MODFLOW prior to MODFLOW 6 used a two-point, ``conductance-based'' flow formulation in which the flow between two adjacent cells is proportional to the difference between the heads calculated in those two cells.
-The conductance-based flow formulation is most accurate when the grid satisfies the ``control-volume finite-difference (CVFD) requirement'' that the straight-line connection between two cell centers must intersect the midpoint of the cell-cell interface at a right angle \citep{narasimhan1976integrated}.
-However, vertically offset grids violate the CVFD requirement because lateral cell faces are vertical but cell connections are not strictly horizontal.
-For this reason, the MODFLOW 6 documentation \citep{modflow6gwf} warns that ``[s]teeply dipping layers generally should not be represented'' using vertical offsets.
-\cite{anderson2015applied} recommend a $10^{\circ}$ dip as a practical upper limit for using vertical offsets.
+The capability to represent complex ‘un-layered’ geology using unstructured grids was developed in versions of MODFLOW after MODFLOW 2005, namely MODFLOW-USG \citep{modflowusg} and MODFLOW 6 \citep{modflow6software}.
+Unstructured versions offered greater flexibility in representing the shape of geological features in plan view.
+To overcome the violations of the control volume finite difference approximations bought about by unsturctured grids, Ghost Node corrections were implemented in MODFLOW-USG \citep{modflowusg}, with MODFLOW 6 incorporating the XT3D option, which solves the full tensorial version of Darcy’s law \citep{modflow6xt3d}.
+Two of the discretisation options in MODFLOW 6 (known as the DIS and DISV options) preserve the layered connectivity approach from earlier MODFLOW versions \citep{modflow6framework}; however, both MODFLOW-USG, and the DISU option within MODLFOW 6 offer the capability to hydraulically connect multiple cells horizontally to the same face using the vertically staggered option with full connectivity, hence revolutionising the traditional layered-approach.
+We note that this type of connection is not implicit within the unstructured versions of the code, but needs to be defined by the user.
 
-To improve the accuracy of flow simulations on unstructured grids, and to allow accurate simulation of arbitrarily oriented two- or three-dimensional anisotropy, the optional XT3D flow formulation \citep{modflow6xt3d} was introduced in MODFLOW 6.
-By performing flow calculations using Darcy's Law in its tensorial form based on a multi-point approximation of the head gradient vector, XT3D accounts for both arbitrarily oriented anisotropy and geometric irregularity of the grid.
-In theory, XT3D should be able to compensate for the geometric irregularity introduced by vertical offsets between adjacent cells in the same grid layer.
+Recently Bardot et al.
+(2022) revisited the capacity of MODFLOW 6 to represent non-orthogonal geologic structures.
+A simple test problem was created using a highly conductive channel misaligned with a Cartesian grid in both plan and transect to test the code's capability to solve flow using flexible gridding in plan, and dipping anisotropic conductivity in transect.
+Through comparison with an analytical solution, the authors demonstrated that the application of XT3D produced excellent results in plan view with flexible gridding, highlighting the improved capabilities of MODFLOW 6.
+However, when attempting to represent an analytically identical problem in transect view, the authors experienced large errors in flow magnitudes and directions when using a vertically offset grid both with and without the use of XT3D.
+They recommended the use of the grid overlay method in conjunction with XT3D for simulating steeply dipping layers.
+A preliminary analysis by two authors of the present work (Provost and Langevin) hypothesized that the limited cell connectivity offered by the vertically offset grid was inadequate for allowing an accurate flow solution as it does not consider full connectivity between the offset layers.
 
-\cite{bardot2022} evaluated the effects of grid design and XT3D on the ability of MODFLOW 6 to accurately simulate flow through sedimentary structures by modeling an idealized, two-dimensional permeable feature (``channel'') in plan view and a dipping hydrogeologic layer (which \cite{bardot2022} also generically called a ``channel'') in cross-section.
-{\color{red} (CDL: use ``permeable feature'' instead of channel?) (AMP: Think it's ok to use channel here in reference to Kerry and Jim's plan-view model, but not after this paragraph; made changes to that effect throughout.)} In each case, the channel or hydrogeologic layer was embedded in a nearly impermeable surrounding medium, or ``domain.'' In their plan-view benchmark, the channel lay within the horizontal plane and was angled counterclockwise relative to the southern boundary of the model (by $30^{\circ}$ in most cases).
-Constant-head boundary conditions at the ends of the channel were set to induce uniform flow along the channel in the corresponding analytical solution.
-The channel and surrounding domain were discretized using either a rectilinear grid, which satisfies the CVFD requirement, or a ``flexible triangular'' grid that is unstructured with triangular cells that conform to the channel boundary but does not satisfy the CVFD requirement.
-They found that the standard, conductance-based flow formulation gave accurate results on the rectilinear grid using isotropic hydraulic conductivity tensors (although accuracy decreased substantially in the presence of anisotropy) but was subject to significant error on the unstructured grid.
-By contrast, the XT3D flow formulation gave accurate results on both grids for both isotropic and anisotropic hydraulic conductivity tensors, as expected.
-However, in the cross-sectional (``transect'') version of their benchmark, which used a vertically offset grid to discretize the nearly impermeable domain and a highly permeable hydrogeologic layer inclined by $30^{\circ}$ from the horizontal, similarly poor results were observed with and without XT3D.
-Given the ability of XT3D to compensate for violations of the CVFD requirement, it appeared likely that another aspect of the problem formulation or the inner workings of MODFLOW 6 was responsible for the unexpected results.
-As noted by \cite{bardot2022}, a preliminary analysis by two authors of the present work (Provost and Langevin) suggested that the limited cell connectivity offered by the vertically offset grid was inadequate for allowing an accurate flow solution.
-
-MODFLOW 6 supports three different types of grids: regular MODFLOW grids consisting of layers, rows, and columns (DIS);  grids discretized by vertices (DISV); and fully unstructured (DISU) grids, which are patterned after the DISU grid type in MODFLOW-USG  \citep{modflowusg}.
-The DIS and DISV grids in MODFLOW 6 are ``layered'', which means that the same two-dimensional grid structure (in plan view) applies to all grid layers.
-The ``layered'' concept of DIS and DISV grids is used within MODFLOW to determine how cells are hydraulically connected.
-Cells in DIS and DISV grids are automatically assigned vertical connections to overlying and underlying cells and horizontal connections to adjacent cells in the same grid layer.
-DISU grids are more flexible than DIS and DISV grids in that the user has complete freedom to specify the cell connectivity and the geometric properties for each connection.
-DISU grids support two types of horizontal connections: a ``layered'' horizontal connection and a ``vertically staggered'' horizontal connection \citep{modflow6gwf}.
-With a layered horizontal connection, the hydraulic conductance is calculated using the full saturated thicknesses of the two connected model cells.
-By contrast, for a staggered horizontal connection, the hydraulic conductance is calculated using only the area of overlap between the two cell faces.
-Vertically staggered horizontal connections were introduced in MODFLOW-USG \citep{modflowusg} and included in MODFLOW 6 \citep{modflow6gwf} primarily as a means to vertically subdiscretize parts of a model domain.
-Vertically staggered connections offer the possibility of implementing ``full'' connectivity in the sense that every cell is hydraulically connected to every other cell with which its faces overlap.
-
-The grid used by \cite{bardot2022} for the steeply dipping benchmark example used layered horizontal connections.
-To our knowledge, the implications of limited horizontal connectivity in layered grids for the accuracy of simulated flows in models with steeply dipping hydrogeologic layers, and the potential for using full cell connectivity via vertically staggered connections to overcome this limitation, have not been previously appreciated or investigated.
-
-% text removed
-% For a grid that includes vertically staggered connections, which in this work we call a ``vertically staggered grid,'' some cells have nominally ``horizontal'' connections with more than one cell in some direction.
-Vertically staggered DISU grids allow more flexibility in assigning cell connectivity than do the more commonly used DIS and DISV grid types \citep{modflow6gwf} in MODFLOW 6.
-
-In the remainder of this paper we first present a theoretical justification for the hypothesis that inadequate cell connectivity is primarily responsible for the inaccurate simulation of flows in the steeply dipping hydrogeologic layer benchmark on a vertically offset grid.
-Based on that theory, we propose a method for improving the accuracy of the flow solution using XT3D together with vertically offset grid with full connectivity, and we evaluate the effectiveness of that approach in a set of benchmark problems similar to those of \cite{bardot2022}.
-Finally, we discuss the potential implications of our findings for practical groundwater models that include high permeability contrasts and steeply dipping hydrogeologic layers.
+Although the grid overlay method accurately simulates groundwater flow in dipping layers, the high vertical resolution required to represent geological structure becomes increasingly significant as dip increases, making the solution computationally prohibitive in some cases due to the large number of cells.
+This paper investigates the limitations of vertically offset grids and develops a solution for accurate and efficient grid representation in the vertical direction.
+We present a theoretical background which explains in detail the connectivity issues associated with vertically distorted grids, and propose a method to upgrade its accuracy using full connectivity grids together with XT3D option.
+The effectiveness of the proposed approach is validated through the benchmark problem of \cite{bardot2022}.
+Finally, we discuss the implications of our findings for practical groundwater problems involving steeply dipping hydrogeologic layers.
 
@@
 The default method for representing the flow between two hydraulically connected model cells in MODFLOW 6 is the ``conductance-based flow formulation'' \citep{modflow6gwf}.
-In this formulation, which is based on the commonly used two-point flux approximation {\color{red} *** ref?
-***}, the flow is proportional to the difference in the hydraulic heads computed at the two cell centers and a hydraulic conductance based on an effective hydraulic conductivity for the connection between the cells.
-On strictly rectilinear grids, such as structured MODFLOW grids with no vertical offset between cells in the same grid layer, with hydraulic conductivity that is isotropic or has anisotropy aligned with the three mutually perpendicular grid directions, the conductance-based formulation is second-order accurate \citep{dehotin2010modeling, modflow6gwf}.
+In this formulation, which is based on the commonly used two-point flux approximation which arises from the starndard finite difference formulation \citep{modflow84}, the flow is proportional to the difference in the hydraulic heads computed at the two cell centers and a hydraulic conductance based on an effective hydraulic conductivity for the connection between the cells.
+On strictly rectilinear grids, such as structured MODFLOW grids with no vertical offset between cells in the same grid layer, and hydraulic conductivity that is isotropic or has anisotropy aligned with the three mutually perpendicular grid directions, the conductance-based formulation is second-order accurate \citep{dehotin2010modeling, modflow6gwf}.
 This is because strictly rectilinear grids satisfy the ``control-volume finite-difference (CVFD) requirement'' that the straight-line connection between two cell centers must intersect the midpoint of the cell-cell interface at a right angle \citep{narasimhan1976integrated}.
@@
 The dipping-hydrogeologic-layer benchmark problem of \cite{bardot2022} used a vertically offset layered grid, which is the type of grid used most often in MODFLOW models.
-Test simulations using the conductance-based flow formulation for a steeply dipping hydrogeologic layer (which we will call the ``aquifer'') embedded in a domain that is six orders of magnitude less hydraulically conductive yielded significant errors in the simulated flows.
+Test simulations using the conductance-based flow formulation for a steeply dipping hydrogeologic layer (which we will call the ``aquifer'') embedded in a near-impermable domain yielded significant errors in the simulated flows.
 The simulated groundwater flux in the middle of the aquifer was not along the $30^{\circ}$ incline of the aquifer as expected, but nearly horizontal ($0.03^{\circ}$ incline), and its magnitude was overestimated by 15\%.
@@
 
-The role of cell connectivity in enabling accurate simulation of flow along a dipping hydrogeologic layer can be understood by considering Figure \ref{fig:schem-conn-area-flux}b, in which components of the groundwater flux (specific discharge) vector representative of steady, uniform flow along the hydrogeologic layer are superimposed on each cell-cell interface.
+The role of cell connectivity in enabling accurate simulation of flow along a dipping hydrogeologic layer can be understood by considering Figure \ref{fig:schem-conn-area-flux}b.
+Components of the groundwater flux (specific discharge) vector are displayed on each cell-cell interface and are representative of steady, uniform flow along the hydrogeologic layer.
 Gray vectors represent the uniform flux oriented along the hydrogeologic layer, which the model is attempting to simulate.
@@
 The base case uses grids consisting of 11 columns and 9 layers of cells, with 3 layers within the aquifer, 3 layers in the domain above the aquifer, and 3 layers in the domain below the aquifer.
-{\color{red} (AMP: Need dimensions of aquifer and domain.)}
+The model width was set to 11 m such that each column was 1m wide.
+Cells within the aquifer were chosen to stay square (i.e.
+1 x 1 m) for the purpose of the test problem to eliminate distortion effects.
+Corner cells in the lower left and upper right corner were also set to be the same dimension as the aquifer cells.
+The resulting base model vertical height is 14.77 m.
 
@@
 
-{\color{red} (AMP: Did some editing of this section for flow and consistent style, but I'll have another go once we're satisfied that all the pieces are basically in place.
-We're really close, if not there already.)}
-
 To demonstrate the effect of cell connectivity on the accuracy of simulated flow through a steeply dipping hydrogeologic layer (aquifer), results for the base case described in the previous section are compared for simulations that use either layered or full connectivity, and either the standard or the XT3D flow formulation.
@@
 
-Plots of specific discharge (groundwater flux) at cell centers and flows across cell-cell interfaces (face flows) calculated by MODFLOW 6 are used to illustrate the accruacy of the simulated flows.
+Plots of specific discharge (groundwater flux) at cell centers and flows across cell-cell interfaces (face flows) calculated by MODFLOW 6 are used to illustrate the accuracy of the simulated flows.
 The magnitude and direction of specific discharge at the cell in the center of the aquifer, at which boundary effects are presumably minimized, and the total volumetric flow rate entering the aquifer through the left-hand boundary, which consists of constant-head (CHD) cells, are also reported as ``representative'' values of flow.
-{\color{red} (AMP: Heads are also plotted.
-Do we want to also report, say, maximum error in head?)}
+Maximum head errors are also reported for the base case.
 
@@
 
-The numerical solution for the grid with layered connectivity confirms the findings of \cite{bardot2022} in that fluxes are predominantly horizontal (top row, Figure \ref{fig:fig2}).
+The numerical solution for the layered connectivity grid confirms the findings of \cite{bardot2022} in that fluxes are predominantly horizontal (top row, Figure \ref{fig:fig2}).
 We see that despite an enforced hydraulic gradient of $30^{\circ}$, the vertical component of the flux cannot propagate through the aquifer given that cells along the aquifer top are hydraulically disconnected from the low permeability domain, hence forcing flow horizontally (red arrows on right panel).
@@
 
-The results shown in the top four panels of Figure \ref{fig:fig2} {\color{red} (AMP: Label the figure panels a through h so we can refer to them by letter?)}, which are based on a vertically offset grid with layered connectivity, are consistent with the findings of \cite{bardot2022}.
+The results shown in Figure \ref{fig:fig2}a and b , which are based on a vertically offset grid with layered connectivity, are consistent with the findings of \cite{bardot2022}.
 Despite boundary heads designed to induce specific discharge of 1 m/d along the $30^{\circ}$ incline of the aquifer, the simulated flow within the aquifer is predominantly horizontal, and the magnitude of the specific discharge at the center of the aquifer and the total flow entering the aquifer are both overestimated.
-The XT3D flow formulation is unable to compensate for the inadequate cell connectivity, which does not allow flow between laterally adjacent cells in different grid layers.
+The XT3D flow formulation is unable to compensate for the inadequate cell connectivity (Figure \ref{fig:fig2}b), which does not allow flow between laterally adjacent cells in different grid layers.
 
-The results shown in the bottom four panels of Figure \ref{fig:fig2}, which are based on a vertically offset grid with full connectivity, show significantly improved results, even for the standard flow formulation.
-Using the XT3D flow formulation provides the most accurate results, as expected.
-
-%Through working on this test problem, we have encountered some nuances in working with specific discharges calculated using MODFLOW.
-There are two main MODFLOW output files pertaining to fluxes.
-The first is an array of face flows which specifies the total flow across each cell face (``FLOW-JA-FACE'').
-The second is the cell-averaged specific discharge for each cell in its directional components ($q_{x}$, $q_{y}$, $q_{z}$) (``DATA-SPDIS''), and is calculated using distance-weighted averages of the cell face flows.
-Therefore, the specific discharge magnitude and direction normally reported by MODFLOW for cells within and along the aquifer edge are ``dampened'' due to averaging of the face flows from the cell top and bottom.
-The result is a ``staircase'' effect whereby these edge cells have a reduced flux magnitude and ``flattened'' flux directions, as shown in Figure 3c of \cite{bardot2022}.
-The plots in Figure \ref{fig:fig2} display corrected cell-averaged specific discharges which exclude the near-zero face flows along the aquifer boundary, and therefore represents the actual simulated flux magnitude and direction within the cell.
-{\color{red} (AMP: An edited version of this now appears near the beginning of the Results and Discussion section)}
-
-%Although correcting the cell-averaged specific discharge makes sense for this theoretical test problem, we must then consider the potential impact, if any, of the standard post-processing procedure used by MODFLOW to derive cell-averaged specific discharge.
-Applications, such as MODPATH (ref), which use cell face flows to interpolate velocities are exempt from this issue.
-However, applications such as MT3D (ref) which bases a dispersion tensor on the cell specific discharge, may incur some error in transport simulations, although it is anticpated this would only occur in extreme scenarios where solute is moving against a highly impermeable dipping boundary.
-The only other application of the cell averaged specific discharge is for plotting groundwater flow which is of minor importance.
-{\color{red} Probably a bit peripheral for this}.
+The vertically offset grid with full connectivity (Figure \ref{fig:fig2}c and d) show significantly improved results, even for the standard flow formulation (Figure \ref{fig:fig2}c).
+Using the XT3D flow formulation provides the most accurate results, as expected (Figure \ref{fig:fig2}d).
 
@@
 	\caption{Numerical results for the test problem using base-case settings for layered-connectivity and full-connectivity grids, with the standard conductance based formulation as well as the XT3D formulation.
-The left panel of each scenario shows the calculated specific discharge at each cell centre (black arrows) with head contours (orange).
-The right panel shows the face flows (red arrows).
-{\color{red} (AMP: Change ``Vertically Offset'' to ``Layered Connectivity'' and ``Vertically Staggered'' to ``Full Connectivity'' in the script that generates the figure)} {\color{red} (AMP: Magnify tiny vectors in the domain [by the conductivity contrast, say] so they can be seen?
-Face flows along the aquifer/domain boundary are tricky because they're ``in between'' in terms of magnitude.)}}
+The left panel of each scenario shows the calculated specific discharge at each cell centre within the aquifer (black arrows) and domain (grey arrows).
+The middle panel shows the face flows (red arrows).
+The right panel shows head contours for the analytical solution (black) as well as the numerical solution (blue).}
 	\label{fig:fig2}
@@
 
-{\color{red} Should we also consider the horizontal discretisation too?
-(CDL: I would say no.
-Paper is already getting a bit long.) (AMP: I agree with CDL.
-But let's be sure to mention that we've only looked at square cells in the aquifer.)}
-
 \begin{figure}
@@
 	\includegraphics[scale=0.9]{../figures/fig3paper.png}
-	\caption{Graphs indicating how increasing the number of grid layers within a single hydrogeologic layer (in this case, the aquifer) affects (a) flux magnitude of middle cell, (b) flux direction of middle cell and (c) volumetric flow through the hydrogeologic layer.
-We immediately notice the errors layered-connectivity grids incur (blue line), and how using a full-connectivity grid reproduces the analytical solution from two grid layers used to discretise the hydrogeologic layer (orange).
-{\color{red} (AMP: Change ``Vertically Offset'' to ``Layered Connectivity'', ``Vertically Staggered'' to ``Full Connectivity'', and ``Model layers'' to ``Grid layers'' in the script that generates the figure)}}
+	\caption{Graphs indicating how increasing the number of grid layers within a single hydrogeologic layer (in this case, the aquifer) affects flux magnitude of middle cell (top), flux direction of middle cell (middle) and volumetric flow through the hydrogeologic layer (bottom).
+We immediately notice the errors layered-connectivity grids incur (left panel), and how using a full-connectivity grid (right panel), particularly in conjunction with XT3D (orange line) reproduces the analytical solution from two grid layers used to discretize the hydrogeologic layer.}
 	\label{fig:fig3}
@@
 The base case arbitrarily uses 3 grid layers to represent the aquifer, but it is important to address the question: \emph{How many grid layers are required to adequately represent a hydrogeological layer?}.
-Layered-connectivity grids, which have been used by groundwater modellers for decades (ref?) {\color{red} Jim to check, I think there is some refinement guidance for aquitards}, are almost always configured as one grid layer per hydrogeological unit.
-{\color{red} (AMP: Could maybe mention that it's generally well known you want multiple layers for accurate transport?)} However, although computationally efficient, we investigate the effect of altering resolution within the aquifer by using the base case and varying number of cells per aquifer width from one to nine {\color{red} (AMP: Explain how aquifer and domain dimensions vary with the discretization.
-And is that how we want them to vary?)} Figure \ref{fig:fig3}).
-{\color{red} (AMP: Are all results in this figure with XT3D?
-If so, would there be any benefit to showing the standard formulation results, as well?)} We compare modeled flux magnitude and direction in the center cell, as well as the volumetric flow against the analytical solution.
-{\color{red} (AMP: Should we also say something about the error in the head solution?
-Report the maximum head error in each case?)}
+Layered-connectivity grids, which are commonly used by groundwater modelers \citep{Reilly2004}, are almost always configured as one grid layer per hydrogeological unit when simulating groundwater flow.
+{\color{red} (AMP: Could maybe mention that it's generally well known you want multiple layers for accurate transport?
+JM: Added "when simulating groundwater flow" to just specify flow explicitly and keep things simple)} However, although computationally efficient, we investigate the effect of altering resolution within the aquifer by using the base case and varying number of cells per aquifer width from one to nine (Figure \ref{fig:fig3}).
+When increasing resolution, cells inside the aquifer were maintained as a square shape to allow fair comparison between models of varying refinement, uncomplicated by cell distortion.
+Therefore, the height of the model extended as each additional grid layer was added to represent the aquifer.
+We compare modeled flux magnitude and direction in the center cell, as well as volumetric flow against the analytical solution.
 
@@
 Similarly, we examine the solution for multiple conductivity contrasts by changing the domain conductivity to 2, 5, 10 and 100 times less than the aquifer.
-{\color{red} (AMP: Dare we try making the domain more conductive than the ``aquifer''?)}  Simulated flux magnitude and direction of the centre cell for all combinations are plotted against the analytical solution in Figure \ref{fig:fig4}.
+Simulated flux magnitude and direction of the center cell for all combinations are plotted against the analytical solution in Figure 4.
 
-Traditional layered-connectivity grids without XT3D shows rapid diversion of flux magnitude and direction from the analytical solution as dip increases (Figure \ref{fig:fig4a}).
-Use of the XT3D option keeps the flux magnitude somewhat on-track until about 30$^{\circ}$ when the solution starts peeling away from the analytical.
+Traditional layered-connectivity grids without XT3D shows rapid diversion of flux magnitude and direction from the analytical solution as dip increases (Figure 4a).
+Use of the XT3D option keeps the flux magnitude somewhat on-track until about 30$^{\circ}$ when the solution starts peeling away from the analytical (Figure 4b).
 Flux direction is improved for the homogeneous scenario, but not for the heterogeneous scenarios.
-The full-connectivity grid on the other hand, even without XT3D, roughly follows the analytical solution for flux magnitude and direction until around 65$^{\circ}$ when it starts to ``fall-apart.'' There are also spikes at certain dip angles which are related to transitions between cell connections due to changing cell face overlaps.
-The full-connectivity grid with XT3D clearly boasts the best results with the flux magnitude and direction matching the analytical until 65$^{\circ}$, with the exception of a slight decrease in flux direction after 45$^{\circ}$.
-{\color{red} (AMP: Can explain the wheels coming off around 65 deg by the fact that with only 3 cells vertically in the aquifer, the aquifer then looks like a long, zigzag Tetris piece that's no more than one cell wide, either vertically or horizontally.
-Finer vertical discretization improves things.)}
+The full-connectivity grid on the other hand, even without XT3D, roughly follows the analytical solution for flux magnitude and direction until around 65$^{\circ}$ when it starts to ``fall-apart'' (Figure 4c).
+There are also spikes at certain dip angles which are related to transitions between cell connections due to changing cell face overlaps.
+The full-connectivity grid with XT3D clearly boasts the best results (Figure 4d) with the flux magnitude and direction matching the analytical until 65$^{\circ}$, with the exception of a slight decrease in flux direction after 45$^{\circ}$.
+The errors reflect the discretization choice of the model.
+For the 1 m by 1 m cells, the aquifer center cell is fully connected to other aquifer cells up to an angle of 45$^{\circ}$.
+After 45$^{\circ}$, some aquifer cell faces share greater overlap with domain cell faces than the adjacent aquifer cell face.
+After 65$^{\circ}$, the aquifer center cell loses all lateral connections to adjacent aquifer cells such that only vertical aquifer connections remain, forcing flow vertically (90$^{\circ}$).
+To rectify this issue when encountering dips greater than 45$^{\circ}$, and especially after 65$^{\circ}$, finer discretization in the vertical and horizontal directions should be made to maximize inter-aquifer cell connections.
 
-These results confirm that full-connectivity grids with XT3D not only outperform the alternatives when modelling dipping hydrogeological unit, but come extremely close to the analytical solution.
+These results confirm that full-connectivity grids with XT3D not only outperform the alternatives when modelling dipping hydrogeological units, but come extremely close to the analytical solution.
+We note that these results relate to the chosen discretization.
+The key element is to ensure that the discretization maintains connectivity at the most extreme angles.
+{\color{red} (JLM: is the advice here that the ratio of vertical to horizontal dicretisation needs to be approximately $\arctan(\theta)$ for three sub-layers.
+Could be a general formula for $\theta$ and number of sub cells)}
 
@@
 	\includegraphics[width=\textwidth]{../figures/fig4_0_paper.png}
-	\caption{layered-connectivity grid, standard conductance-based formulation}
-	\label{fig:fig4a}
 \end{subfigure}
@@
 	\includegraphics[width=\textwidth]{../figures/fig4_1_paper.png}
-	\caption{layered-connectivity grid, XT3D}
-	\label{fig:fig4b}
 \end{subfigure}
@@
 	\includegraphics[width=\textwidth]{../figures/fig4_2_paper.png}
-	\caption{full-connectivity grid, standard conductance-based formulation}
-	\label{fig:fig4c}
 \end{subfigure}
@@
 	\includegraphics[width=\textwidth]{../figures/fig4_3_paper.png}
-	\caption{full-connectivity grid, XT3D}
-	\label{fig:fig4d}
 \end{subfigure}
@@
 The base case and subsequent scenarios have assumed isotropic conductivity.
-However, sedimentary layers often exhibit anisotropic conductivity, and therefore we examine the gridding strategies using anisotropic and dipping conductivity tensors {\color{red} (AMP: in the domain and aquifer, or only in the aquifer?)}, with reduced conductivity perpendicular to the aquifer by 10, 100, 1000 and 10,000.
+However, sedimentary layers often exhibit anisotropic conductivity, and therefore we examine the gridding strategies using anisotropic and dipping conductivity tensors within the aquifer, with reduced conductivity perpendicular to the aquifer by 10, 100, 1000 and 10,000.
 We also include the isotropic scenario (ratio of 1).
-{\color{red} (AMP: Dare we try increased perpendicular conductivity?)} Results are presented in Figure \ref{fig:fig5}.
-{\color{red} (AMP: Are all results in this figure with XT3D?
-If so, would there be any benefit to showing the standard formulation results, as well?)} Layered-connectivity grids produce flux magnitude and flow results that significantly deviate from the analytical solution (blue line), whilst full-connectivity grids reproduce the analytical solution (orange line).
+Results are presented in Figure \ref{fig:fig5}.
+Layered-connectivity grids produce flux magnitude and flow results that significantly deviate from the analytical solution both with and without the XT3D formulation (Figure \ref{fig:fig5}a), whilst full-connectivity grids with XT3D reproduce the analytical solution (Figure \ref{fig:fig5}b, orange line).
 These results verify, once again, our proposed method of using full-connectivity grids to model dipping anisotropic hydrogeologic layers.
@@
 	\includegraphics[scale=0.9]{../figures/fig5paper.png}
-	\caption{Varying anisotropy ratios of conductivity withing the aquifer and the resulting (a) flux magnitude of middle cell, (b) flux direction of middle cell and (c) volumetric flow through the aquifer.
-}
+	\caption{Varying anisotropy ratios of conductivity withing the aquifer and the resulting flux magnitude of middle cell (top panel), flux direction of middle cell (middle panel) and volumetric flow through the aquifer (bottom panel).
+Plots include results for the standard conductance-based formulation (blue) as well as the XT3D formulation (orange).}
 	\label{fig:fig5}
@@
 
- {\color{red} CDL: I added the following two paragraphs in an attempt to summarize my understanding of the main conclusion.  Feel free to edit or scrap as you see fit.}
-
  \cite{bardot2022} revisited the capability of MODFLOW to accurately simulate groundwater flow through sedimentary structures.
 They concluded that the XT3D multi-point flux approximation in MODFLOW 6 significantly improves the accuracy of simulated flows when used with a grid-overlay approach in which locations of sedimentary structures, which can be steeply dipping, are mapped onto a relatively fine, rectilinear model grid.
-They also concluded that the XT3D multi-point flux approximation does not perform as well as anticipated when simulating flow through a steeply dipping hydrogeologic layer using vertically offset grids, which are advantageous for significantly reducing the number of grid layers required for many problems.
+They also concluded that the XT3D multi-point flux approximation does not perform as well as anticipated when simulating flow through a steeply dipping hydrogeologic layer using deformed vertically offset grids, which are advantageous for significantly reducing the number of grid layers required for many problems.
 They hypothesized that the inherently limited lateral connectivity between cells in DIS and DISV layered grids in MODFLOW 6 was responsible for the inability to obtain accurate flow results in their steeply dipping test problems using vertically offsets grids, with or without XT3D.
@@
 The additional connectivity is implemented using the fully unstructured (DISU) grid type.
-Importantly, this accuracy improvement requires that at least two grid layers be used to represent the steeply dipping sedimentary structure.
-The primary conclusion of this paper is that, given appropriate discretization and cell connectivity, the XT3D multi-point flux approximation implemented in MODFLOW 6 can be combined with vertically offset grids to efficiently and accurately model flows through {\color{red} steeply dipping??} sedimentary structures.
-Prior to this work, the capability to efficiently and accurately model flow through {\color{red} steeply dipping??} sedimentary structures was generally restricted to finite element simulators.
+Importantly, this accuracy improvement requires that at least two grid layers be used to represent the dipping aquifers, with additional grid layers required for very steep dips above 65$^{\circ}$.
+The primary conclusion of this paper is that, given appropriate discretization and cell connectivity, the XT3D multi-point flux approximation implemented in MODFLOW 6 can be combined with vertically offset grids to efficiently and accurately model flows through steeply dipping sedimentary structures.
+Prior to this work, the capability to efficiently and accurately model flow through steeply dipping sedimentary structures was generally restricted to finite element simulators.
 
@@
 
-Points:
-
-1) Furthermore, we point out the nuances in specific discharge calculations and the ambiguity introduced when assessing model behaviour.  MODPATH and MT3D
-
-3) (Random thought) The use of full-connectivity grids improved the accuracy of the transect solutions; however, the results of \citep{bardot2022} using unstructured meshes that conformed to the aquifer geometry preformed better in isotropic grids.
-If you force the normal flux to be the same at the interface of the domain and aquifer, the analytical solution would suggest a uniform gradient either side, do we introduce an error if the faces don't align with the aquifer correctly?
-Would figure \ref{fig:fig2} show a correction for this if the anisotropic scenario was tested?
-{\color{red} (AMP: Is the underlying question whether using an unstructured mesh that conforms to the aquifer geometry [which MF6 can't do in transect] would improve the accuracy more than a full-connectivity vertically offset grid, assuming xt3d is used in both cases?
-Let's discuss on our next call.)}
-
- {\color{red} CDL: Maybe worthwhile to implement full connectivity approaches for DIS and DISV grids in mf6.}
-
 \section{Acknowledgments}
-We would like to acknowledge.... {\color{red} that this simple paper has been in the works far too long. Kindly address complaints to the first author.}
+We would like to acknowledge.... {\color{red} that this paper is almost ready to GO!! Yeeeha!}
 

### 2024-06-10 latex/article.tex

@@
 
-\title{Efficient and Accurate Modeling of Flow through Dipping Aquifers with MODFLOW 6}
+%\title{Improving Accuracy of Simulated Flows in Steeply Dipping Layers Using Full-Connectivity Grids with the XT3D Multi-Point Flux Approximation in MODFLOW 6}
+
+% suggested title revision
+\title{Efficient and Accurate Modeling of Flow through Sedimentary Structures with MODFLOW 6}
+
+%\title{Efficient and Accurate Modeling of Flow through Dipping Aquifers with MODFLOW 6}
+
+\title{The Importance of Cell Connectivity for Efficient and Accurate Modeling of Flow through Dipping Aquifers with MODFLOW 6}
 
@@
 
-\textbf{Article impact statement: Groundwater flow through dipping aquifers can be accurately simulated using a vertically offset grid combined with an unstructured grid approach and the XT3D multi-point flow formulation in MODFLOW 6.}
+%\textbf{Article impact statement: Groundwater flow through dipping aquifers can be accurately simulated using a vertically offset grid combined with an unstructured grid approach and the XT3D multi-point flow formulation in MODFLOW 6.}
+
+\textbf{Article impact statement:} Accurate simulation of groundwater flow through dipping aquifers using MODFLOW 6 can require enhanced cell connectivity and use of a multi-point formulation for flow between cells.
+This can be achieved using an unstructured grid and the XT3D flow formulation.
 
 \begin{abstract}
-Groundwater flow through dipping hydrogeologic layers is often modeled with MODFLOW by specifying top and bottom elevations of model layers to coincide with top and bottom elevations of hydrogeologic layers.
-With this approach, adjacent cells within a model layer are vertically offset from one another.
-While the standard formulation in MODFLOW does not rigorously account for these offsets in the underlying flow calculations, results are generally expected to be accurate provided dip angles do not exceed $10^{\circ}$  {\color{red} (KB: I would remove this sentence as even 10 degrees if anisotropic say 10:1 is still  a 50\% error)}.
-To improve the accuracy of flow simulations on unstructured grids, and to allow accurate simulation of arbitrarily oriented two- or three-dimensional anisotropy, the optional XT3D multi-point flow formulation was introduced in MODFLOW 6.
-The XT3D flow formulation is designed to account for geometric irregularities in the grid, including vertical offsets between laterally adjacent cells, and to provide accurate results for both isotropic and anisotropic groundwater flow.
-\cite{bardot2022} showed, however, that neither the standard formulation nor the XT3D formulation produced the correct flow direction and magnitude in simple tests involving flow through a steeply dipping, permeable hydrogeologic layer.
-In this paper, we explain that the inability of XT3D to give accurate results in the steeply dipping test problem of \cite{bardot2022} was caused by the limited cell connectivity that is inherent in the most commonly used discretization packages in MODFLOW 6.
-The DIS and DISV discretization packages are based on layered connectivity in which cells have lateral connections only with cells in the same grid layer.
-We demonstrate that layered lateral connectivity is insufficient to allow accurate simulation of flow through a steeply dipping hydrogeologic layer on a vertically offset grid.
-When lateral connections between cells in adjacent grid layers are included using the fully unstructured (DISU) grid type, XT3D is able to calculate the flow magnitude and direction accurately for both isotropic and anisotropic conductivity oriented along the permeable hydrogeologic layer.
-To realize this improved accuracy, the permeable hydrogeologic layer must be discretized vertically using at least two model layers.
-{\color{red} (AMP: Given the idealized nature of the test problem, how general can we be with our conclusions?
-I pulled it back here relative to what you had originally, CDL, but left a slightly stronger statement in the Conclusions, at least for now.})  Our findings suggest that, given appropriate discretization and cell connectivity, the combination of vertically offset grids with the XT3D flow formulation in MODFLOW 6 can provide a computationally efficient and accurate method for simulating flow through isotropic and anisotropic steeply dipping layers, a capability that was previously reserved to finite-element-based simulators.
-{\color{red} (AMP: Any Pal and Edwards-type MPFA stuff that's relevant to that final phrase?
-See also Conclusions)}
 
-{\color{red} Here's another option for a simpler abstract! }
-
-	Dipping aquifers are often modeled using a deformed grid whereby each hydrogeological layer is represented using a single grid layer.
-However, until the release of MODFLOW 6, accurate modeling of groundwater flow through dipping anisotropic layers has not been possible due to limitations of its finite difference conductance-based solution.
-The introduction of the XT3D multi-point flow formulation has allowed accurate simulation of the full conductivity tensor, as well as handling the flow solution for irregular cell geometries for unstructured grids.
-Through the use of a benchmark, \cite{bardot2022} showcased MODFLOW 6's capability to model flow through complex sedimentary structures, previously restricted to finite element simulators.
-However, the benchmark study also brought to light the inadequacy of the deformed vertically offset grid for accurate modeling of dipping layers, with or without XT3D.
-This paper explains that the limitation of the deformed grid is a result of poor cell connectivity between vertically adjacent model layers.
-Nonetheless, we also demonstrate that dipping layers may be efficiently and accurately modeled using a deformed vertically offset grid with full-connectivity, which can be achieved by combining MODFLOW 6's DISU discretization package with the XT3D flow formulation.
-A repeat benchmark study shows that this method produces accurate head and flux results in isotropic and anisotropic dipping layers, providing a minimum of two grid layers per hydrogeological layer are deployed which allows for sufficient vertical connectivity.
+Groundwater flow through a dipping aquifer is often modeled using a single layer of model cells to represent the aquifer.
+The layer of cells is ``deformed'' to follow the top and bottom elevations of the aquifer.
+When this approach is used in MODFLOW, adjacent cells within the model layer are vertically offset from one another, and the standard conductance-based (two-point) formulation for flow between cells does not rigorously account for these offsets.
+The XT3D multi-point flow formulation introduced in MODFLOW 6 is designed to account for geometric irregularities in the grid, including vertical offsets, and to provide accurate results for both isotropic and anisotropic groundwater flow.
+A recent study evaluated the performance of the standard formulation and XT3D using a simple, synthetic benchmark model of a steeply dipping aquifer.
+Although XT3D generally improved the accuracy of flow simulations relative to the standard formulation as expected, neither formulation produced accurate flows in cases that involved large vertical offsets.
+In this paper, we explain that the inability of XT3D to produce accurate flows in the steeply dipping aquifer benchmark was not due to an inherent limitation of the flow formulation, but to the limited cell connectivity inherent in the most commonly used discretization packages in MODFLOW 6.
+Furthermore, we demonstrate that XT3D is able to produce the expected accuracy when adequate cell connectivity is introduced using MODFLOW's unstructured grid type and the aquifer is discretized vertically using at least two model layers.
 
@@
 
-MODFLOW is commonly used to simulate groundwater flow in a wide variety of geological settings including sedimentary, alluvial and fractured rock aquifers \cite{modflow2005,Kumar2019}.
-Accurate simulation of the hydrostratigraphic sequences encountered in these geological settings requires that the model grid discretisation and solution technique can represent the scale and effects of important geologial features \cite{Reilly2004,bardot2022}, where permeability tensors are aligned along key structures \cite{hoaglund2003}.
-MODFLOW versions up to and including MODFLOW-2005 employ a structured grid based on a fixed number of layers, rows and columns \citep{modflow2005,modflow84,modflow88}.
-Further, through assuming that the hydraulic conductivity tensor aligns with the grid coordinates, the standard finite difference approximation can be applied accurately and efficiently to simulate groundwater flow \citep{modflow84}.
-The main trade-off in this approximation is that important hydrological and geological features are required to be mapped to the structured grids and orinetations need to be aligned to Cartesian coordinates.
-These requirements can result in pixelated features in plan view and, more significantly, create problems where geological layers are discontinuous at unconformities and faults, and where they are dipping and misaligned with the model grid \citep{bardot2022}.
+Numerical simulation of layered aquifer systems can involve a tradeoff between efficiency and accuracy.
+A traditional approach to discretizing layered aquifer systems, called the ``grid overlay'' method by \cite{hoaglund2003}, maps the heterogeneous hydrologic properties onto a rectilinear grid \citep{modflow84}.
+While this method preserves the benefits of the rectilinear grid structure, it can require a large number of model grid layers to adequately discretize hydrogeogic layers of variable dip and thickness \citep{Zyvoloski2006}.
 
-In meeting the requirements for the standard finite difference simulation, two methods have traditionally been used to represent dipping geological features in MODFLOW in versions up to and including MODFLOW-2005.
-One approach is to map the geological features to a rectilinear grid \citep{modflow84} in a method referred to as ‘grid overlay’  \citep{hoaglund2003}.
-Grid overlay preserves the control volume finite difference approximation but may require a high level of vertical discretization to accurately represent variable layering \citep{Zyvoloski2006}.
-An alternative approach is to set each hydrostratigraphic unit as its own model layer, resulting in 'deformed layers' \cite{anderson2015applied,Reilly2004} in a 'vertically offset' or 'conforming finite difference grid' \citep{Zyvoloski2006}.
-When the layers in the offset grid are not strictly horizontal and of a constant thickness, the standard finite difference approximation is violated \citep{Zheng1994,anderson2015applied,Reilly2004}.
-To maintain a reasonable level of accuracy, \cite{anderson2015applied} recommend a $10^{\circ}$ dip as a practical upper limit for using vertical offsets.
+An alternative approach, called the ``boundary-matching" method by \cite{hoaglund2003}, allows model grid layers to "deform" vertically to follow the undulations of hydrogeologic layers.
+This method ``minimize[s] the number of model [grid] layers required to simulate an aquifer system'' \citep{modflow84}.
+In fact, it is common practice to represent each hydrogeologic layer using a single, boundary-matching model grid layer.
+However, the non-orthogonality of the resulting grid can be a source of significant model error \citep{hoaglund2003}.
 
-The capability to represent complex ‘un-layered’ geology using unstructured grids was developed in versions of MODFLOW after MODFLOW 2005, namely MODFLOW-USG \citep{modflowusg} and MODFLOW 6 \citep{modflow6software}.
-Unstructured versions offered greater flexibility in representing the shape of geological features in plan view.
-To overcome the violations of the control volume finite difference approximations bought about by unsturctured grids, Ghost Node corrections were implemented in MODFLOW-USG \citep{modflowusg}, with MODFLOW 6 incorporating the XT3D option, which solves the full tensorial version of Darcy’s law \citep{modflow6xt3d}.
-Two of the discretisation options in MODFLOW 6 (known as the DIS and DISV options) preserve the layered connectivity approach from earlier MODFLOW versions \citep{modflow6framework}; however, both MODFLOW-USG, and the DISU option within MODLFOW 6 offer the capability to hydraulically connect multiple cells horizontally to the same face using the vertically staggered option with full connectivity, hence revolutionising the traditional layered-approach.
-We note that this type of connection is not implicit within the unstructured versions of the code, but needs to be defined by the user.
+In models based on the Control Volume Finite Difference method {(\color{red} ref for this?)}, which is the method used in the MODFLOW family of hydrologic simulation codes \citep{modflow6gwf}{\color{red} (ref to our new MODFLOW overview paper!)}, the loss of accuracy associated with the boundary-matching approach results from the use of a "two-point" formulation to calculate flows between adjacent grid cells.
+In the two-point formulation, the flow between two cells is proportional to the difference between the heads calculated in those two cells.
+For simulations of flow through isotropic porous media, the two-point flow formulation is most accurate when the grid satisfies the so-called ``CVFD requirement" \citep{modflow6gwf} that the straight-line connection between two cell centers must intersect the midpoint of the cell interface at a right angle \citep{narasimhan1976integrated}.
+However, boundary-matching grids violate the CVFD requirement at vertical interfaces between cells in the same model layer because the corresponding cell connections are not strictly horizontal.
 
-Recently Bardot et al.
-(2022) revisited the capacity of MODFLOW 6 to represent non-orthogonal geologic structures.
-A simple test problem was created using a highly conductive channel misaligned with a Cartesian grid in both plan and transect to test the code's capability to solve flow using flexible gridding in plan, and dipping anisotropic conductivity in transect.
-Through comparison with an analytical solution, the authors demonstrated that the application of XT3D produced excellent results in plan view with flexible gridding, highlighting the improved capabilities of MODFLOW 6.
-However, when attempting to represent an analytically identical problem in transect view, the authors experienced large errors in flow magnitudes and directions when using a vertically offset grid both with and without the use of XT3D.
-They recommended the use of the grid overlay method in conjunction with XT3D for simulating steeply dipping layers.
-A preliminary analysis by two authors of the present work (Provost and Langevin) hypothesized that the limited cell connectivity offered by the vertically offset grid was inadequate for allowing an accurate flow solution as it does not consider full connectivity between the offset layers.
+In MODFLOW, the top and bottom surfaces of model cells are horizontal \citep{modflow6gwf}.
+Thus, when boundary matching is implemented in a two-dimensional (2D) cross-sectional or three-dimensional (3D) MODFLOW model, the tops of laterally adjacent cells are vertically offset from each other (as are the cell bottoms) and can follow the undulations of the aquifer boundaries only on average.
+Following \cite{bardot2022}, we call this type of grid ``vertically offset." In a vertically offset grid, subhorizontal cell connections across vertical cell interfaces violate the CVFD requirement, which reduces the accuracy of two-point flow formulation.
+For an isotropic aquifer, \cite{anderson2015applied} recommend a $10^{\circ}$ dip as a practical upper limit when using the two-point flow formulation.
+{\color{red} (AMP: Was this offered in the context of MODFLOW, or more generally???
+I don't have that book.)}
 
-Although the grid overlay method accurately simulates groundwater flow in dipping layers, the high vertical resolution required to represent geological structure becomes increasingly significant as dip increases, making the solution computationally prohibitive in some cases due to the large number of cells.
-This paper investigates the limitations of vertically offset grids and develops a solution for accurate and efficient grid representation in the vertical direction.
-We present a theoretical background which explains in detail the connectivity issues associated with vertically distorted grids, and propose a method to upgrade its accuracy using full connectivity grids together with XT3D option.
-The effectiveness of the proposed approach is validated through the benchmark problem of \cite{bardot2022}.
+The introduction of unstructured grids into MODFLOW \citep{modflowusg, modflow6gwf} presented new ways for a grid to violate the CVFD requirements.
+To help overcome the limitations of the two-point flow formulation on geometrically irregular grids, MODFLOW-USG and MODFLOW 6 offer a Ghost-Node Correction (GNC) Package \citep{modflowusg, modflow6gwf}{\color{red} (ref to Dickenson and others 2007 ?)} However, configuration of optimal ghost node locations and interpolation weights is highly problem-dependent, and we are not aware of guidelines for handling general unstructured grids or arbitrarily oriented 3D anisotropy.
+
+The optional XT3D flow formulation \citep{modflow6xt3d} introduced in MODFLOW 6 offers a more ``automatic" alternative to ghost nodes.
+XT3D is a multi-point flow formulation {\color{red} (refs for multi-point)} based on gradient reconstruction {\color{red} (refs for gradient reconstruction)}.
+By performing flow calculations using Darcy's Law in its tensorial form using a multi-point approximation of the head gradient vector, XT3D accounts for both arbitrarily oriented anisotropy and geometric irregularity of the grid.
+In theory, XT3D should be able to compensate for the geometric irregularity introduced by vertical offsets between adjacent cells in the same grid layer.
+
+\cite{bardot2022} tested the ability of MODFLOW 6 to accurately simulate flow through sedimentary structures, both with and without XT3D.
+The simplest 2D cross-sectional (``transect") version of their benchmark model simulated flow through a steeply dipping, homogeneous, isotropic aquifer of uniform thickness embedded between two confining units.
+In test runs that used the boundary-matching method on a vertically offset grid, both the two-point flow formulation and XT3D produced flows that differed markedly from the expected magnitude and direction throughout the aquifer, regardless of how finely the aquifer was discretized.
+They noted, however, that a preliminary analysis by two authors of the present work (Provost and Langevin) suggested that the limited cell connectivity offered by the layered, vertically offset grids was inadequate to allow an accurate flow solution.
+
+To our knowledge, the implications of limited cell connectivity in layered, vertically offset grids for the accuracy of simulated flows, and the potential to improve accuracy using enhanced connectivity between model layers, have not been previously recognized or investigated.
+In this paper, we present theoretical explanations for why ``layered" connectivity in a vertically offset grid is inadequate for accurate simulation of flows in steeply dipping hydrogeologic layers, and why using what we call ``full" cell connectivity together with XT3D should improve accuracy.
+Then, in a set of test problems similar to those of \cite{bardot2022}, we demonstrate the effectiveness of using full connectivity with XT3D.
 Finally, we discuss the implications of our findings for practical groundwater problems involving steeply dipping hydrogeologic layers.
 
-\section{Theoretical Background}
+\section{Theory}
+
+Versions of MODFLOW prior to MODFLOW-USG \citep{modflowusg} supported only structured grids composed of rows, columns, and layers of cells, which could be vertically offset.
+This grid type, called ``DIS," continues to be supported in MODFLOW-USG and MODFLOW 6, which also offer an unstructured grid type called ``DISU" \citep{modflowusg, modflow6gwf}.
+In addition, MODFLOW 6 offers an  unstructured grid type called ``DISV," which is based on specification of cell vertices \citep{modflow6gwf}.
+The DIS and DISV grid types are based on model layers, with the same horizontal grid structure within each layer.
+The DISU grid type is more flexible and does not depend on the concept of model layers.
 
@@
 
-Figure \ref{fig:schem-conn-area-flux}a shows a group of four cells in a two-grid-layer, cross-sectional MODFLOW 6 model of fully saturated flow through a permeable hydrogeologic layer (aquifer) with impermeable top and bottom boundaries.
-The cells have horizontal tops and bottoms and can therefore follow the dip of the hydrogeologic layer only on average.
-If the grid is represented using a layered grid type (DIS or DISV) in MODFLOW 6, a cell is hydraulically connected to each adjacent cell in the same grid layer, regardless of whether or not the cells overlap, and with each overlying and underlying cell, as indicated by the blue lines that connect cell centers in Figure \ref{fig:schem-conn-area-flux}a.
-In this work we call this a vertically offset grid with ``layered connectivity.''  Note that cells that have overlapping vertical faces but are in different grid layers (e.g., cells 1A and 2B) do not have a direct hydraulic connection in a grid with layered connectivity.
-The addition of such connections between cells in different grid layers, as indicated by the red lines that connect cell centers in Figure \ref{fig:schem-conn-area-flux}a, results in a grid with ``full connectivity.''  In MODFLOW 6, only layered connectivity can be represented in structured (DIS) and vertex-based (DISV) grids.
-Fully unstructured (DISU) grids allow representation of any kind of cell connectivity, including layered or full connectivity.
+Figure \ref{fig:schem-conn-area-flux}a illustrates what we refer to as "layered" and "full" cell connectivity.
+The focus is on a group of four cells in a two-grid-layer, vertically offset MODFLOW 6 model of fully saturated flow through a permeable hydrogeologic layer (aquifer) with impermeable top and bottom boundaries.
+If the grid is represented using a layered grid type (DIS or DISV), the grid has what we call ``layered connectivity": a cell is hydraulically connected to each laterally adjacent cell in the same grid layer (regardless of whether or not the cell faces actually overlap), and with each overlying and underlying cell, as indicated by the blue lines that connect cell centers in Figure \ref{fig:schem-conn-area-flux}a.
+In layered connectivity, two cells that have overlapping vertical faces but are in different grid layers (e.g., cells 1A and 2B) notably do not have a direct hydraulic connection.
+The addition of such connections between cells in different grid layers, as indicated by the red lines that connect cell centers in Figure \ref{fig:schem-conn-area-flux}a, results in a grid with what we call ``full connectivity.''  Implementation of full connectivity in MODFLOW 6 requires that the vertically offset grid be represented using the fully unstructured (DISU) grid type.
 
-The default method for representing the flow between two hydraulically connected model cells in MODFLOW 6 is the ``conductance-based flow formulation'' \citep{modflow6gwf}.
-In this formulation, which is based on the commonly used two-point flux approximation which arises from the starndard finite difference formulation \citep{modflow84}, the flow is proportional to the difference in the hydraulic heads computed at the two cell centers and a hydraulic conductance based on an effective hydraulic conductivity for the connection between the cells.
-On strictly rectilinear grids, such as structured MODFLOW grids with no vertical offset between cells in the same grid layer, and hydraulic conductivity that is isotropic or has anisotropy aligned with the three mutually perpendicular grid directions, the conductance-based formulation is second-order accurate \citep{dehotin2010modeling, modflow6gwf}.
-This is because strictly rectilinear grids satisfy the ``control-volume finite-difference (CVFD) requirement'' that the straight-line connection between two cell centers must intersect the midpoint of the cell-cell interface at a right angle \citep{narasimhan1976integrated}.
-However, vertically offset grids violate the CFVD requirement because nominally ``horizontal'' connections between cell centers in the same grid layer are not strictly horizontal and therefore not perpendicular to the vertical cell-cell interfaces they intersect.
-For example, in Figure \ref{fig:schem-conn-area-flux}a, the connection between cells 1A and 1B is not perpendicular to the cell-cell interface, which compromises the accuracy of the conductance-based formulation.
-
-The conductance-based formulation was the only flow formulation available in versions of MODFLOW prior to MODFLOW 6.
-As mentioned earlier, the limitations of the conductance-based formulation on vertically offset grids were well recognized.
-However, the ability to use unstructured grids in MODFLOW-USG and MODFLOW 6 introduced new ways to violate the CVFD requirement, and thereby render the conductance-based formulation less accurate, even in the absence of vertical offsets.
-The concept of a ``ghost node'' was included in MODFLOW-USG \citep{modflowusg}, and subsequently in MODFLOW 6, as a simple and optional way to improve the accuracy for grids that violated the CVFD requirement.
-However, implementation of the ghost-node correction is highly problem dependent, and it is not obvious how to specify optimal ghost-node locations and weights on complex, unstructured grids.
-As a more ``automatic'' alternative to the ghost-node method for improving accuracy, and as a way to allow accurate simulation of arbitrarily oriented two- or three-dimensional anisotropy, the optional XT3D flow formulation \citep{modflow6xt3d} was introduced in MODFLOW 6.
-XT3D formulates the flow between two cells by interpolating head values from the two cells and their neighboring cells to construct an estimate of the full, two- or three-dimensional head-gradient vector on each side of the cell-cell interface; applying the cell conductivity tensors to obtain an estimate of the groundwater flux  normal to the interface on each side of the interface; and reconciling the two flux estimates to ensure continuity of flow across the interface.
-By performing flow calculations using Darcy's Law in its tensorial form, based on a ``multi-point'' approximation of the head gradient, XT3D accounts for both arbitrarily oriented anisotropy and geometric irregularity of the grid.
-
-The dipping-hydrogeologic-layer benchmark problem of \cite{bardot2022} used a vertically offset layered grid, which is the type of grid used most often in MODFLOW models.
-Test simulations using the conductance-based flow formulation for a steeply dipping hydrogeologic layer (which we will call the ``aquifer'') embedded in a near-impermable domain yielded significant errors in the simulated flows.
-The simulated groundwater flux in the middle of the aquifer was not along the $30^{\circ}$ incline of the aquifer as expected, but nearly horizontal ($0.03^{\circ}$ incline), and its magnitude was overestimated by 15\%.
-Significant errors in the simulated flows were expected in this case, given that the conductance-based flow formulation does not rigorously account for violations of the CVFD requirement that the cell-cell connection be perpendicular to the cell-cell interface, and that it takes the connection length to be the horizontal distance between cell centers.
-Given the ability of XT3D to account rigorously for cell connections that are not perpendicular to cell-cell interfaces, and for the increase in connection length due to the slope of the connection, rerunning the simulation with XT3D activated could have been expected to improve the flow solution substantially.
-However, \cite{bardot2022} observed similarly poor results with XT3D: the flux was still nearly horizontal ($0.01^{\circ}$ incline), and its magnitude was overestimated by 12\%.
-Based on a preliminary analysis by two authors of the present work (Provost and Langevin), \cite{bardot2022} hypothesized that the unexpectedly large error in the flow solution with XT3D was related to inadequate cell connectivity.
-
-The role of cell connectivity in enabling accurate simulation of flow along a dipping hydrogeologic layer can be understood by considering Figure \ref{fig:schem-conn-area-flux}b.
+The role of full cell connectivity in enabling accurate simulation of flow along a dipping hydrogeologic layer can be understood by considering Figure \ref{fig:schem-conn-area-flux}b.
 Components of the groundwater flux (specific discharge) vector are displayed on each cell-cell interface and are representative of steady, uniform flow along the hydrogeologic layer.
@@
 Although this concept has been illustrated using a two-grid-layer model for simplicity, the same reasoning applies given any number of grid layers.
+
+Note how the vertically offset grid violates the CFVD requirement because nominally ``horizontal'' connections between cell centers in the same grid layer are not strictly horizontal and therefore not perpendicular to the vertical cell-cell interfaces they intersect.
+For example, in Figure \ref{fig:schem-conn-area-flux}a, the connection between cells 1A and 1B is not perpendicular to the cell-cell interface, which reduces the accuracy of the conductance-based formulation.
 
@@
 \label{eqn:head_analyt_along}
-h = - x \cos \theta - z \sin \theta.
+h = - x \cos \theta - z \sin \theta,
 \end{equation}
@@
 
-Two methods of discretization are used to simulate the test problem: (1) vertically offset grid of the MODFLOW 6 DIS (structured) type, and (2) vertically offset grid of MODFLOW 6 DISU (unstructured) type.
-The cell geometry in these two discretization methods is identical, but the cell connectivity is different.
-The DIS grid uses layered connectivity, and the DISU grid includes the additional connections between cells in adjacent grid layers needed to realize full connectivity.
-The DISU grid is constructed by starting with a DIS grid, converting it to an equivalent DISU grid, and then implementing full connectivity.
+The test simulations use grids of MODFLOW 6 DISU (unstructured) type, which offers the flexibility to implement either layered or full connectivity.
+In all cases, a grid of MODFLOW 6 DIS (structured) type is first created and then converted a DISU grid.
+In simulations that use layered connectivity, the DISU grid is formulated to have the same cell geometry and connectivity as the original DIS grid.
+In simulations that use full connectivity, additional connections between adjacent layers are added during the conversion process.
 The base case uses grids consisting of 11 columns and 9 layers of cells, with 3 layers within the aquifer, 3 layers in the domain above the aquifer, and 3 layers in the domain below the aquifer.
-The model width was set to 11 m such that each column was 1m wide.
-Cells within the aquifer were chosen to stay square (i.e.
-1 x 1 m) for the purpose of the test problem to eliminate distortion effects.
+The model is 11 m wide, and each column is 1 m wide.
+Cells within the aquifer are square (1 m x 1 m).
 Corner cells in the lower left and upper right corner were also set to be the same dimension as the aquifer cells.
@@
 
-Jupyter notebooks created for the test problem are available in the Supporting Information that accompanies this paper.
+Jupyter notebooks created for the test problem are available in the Supporting Information that accompanies this paper {\color{red} (best place???)}.
 The notebooks leverage the capabilities of FloPy \citep{bakker2016scripting, hughes2023flopy} to assist in setting up input for and processing results from the MODFLOW 6 simulations.
@@
 
-To demonstrate the effect of cell connectivity on the accuracy of simulated flow through a steeply dipping hydrogeologic layer (aquifer), results for the base case described in the previous section are compared for simulations that use either layered or full connectivity, and either the standard or the XT3D flow formulation.
+To demonstrate the effect of cell connectivity on the accuracy of simulated flow through a steeply dipping aquifer, results for the base case described in the previous section are compared for simulations that use either layered or full connectivity, and either the two-point flow formulation (also known as the "standard conductance-based formulation" in MODFLOW 6) or the XT3D flow formulation.
 The base case is then modified to further investigate the effects of grid resolution, aquifer dip angle, conductivity contrast between the aquifer and the surrounding domain, and anisotropic hydraulic conductivity.
@@
 Plots of specific discharge (groundwater flux) at cell centers and flows across cell-cell interfaces (face flows) calculated by MODFLOW 6 are used to illustrate the accuracy of the simulated flows.
-The magnitude and direction of specific discharge at the cell in the center of the aquifer, at which boundary effects are presumably minimized, and the total volumetric flow rate entering the aquifer through the left-hand boundary, which consists of constant-head (CHD) cells, are also reported as ``representative'' values of flow.
+The magnitude and direction of specific discharge at the cell in the center of the aquifer, at which boundary effects are presumably minimized, and the total volumetric flow rate entering the aquifer through the left-hand boundary, which consists of constant-head (CHD) cells, are also reported as representative values of flow.
 Maximum head errors are also reported for the base case.
@@
 
-The numerical solution for the layered connectivity grid confirms the findings of \cite{bardot2022} in that fluxes are predominantly horizontal (top row, Figure \ref{fig:fig2}).
+The numerical solution for the layered-connectivity grid confirms the findings of \cite{bardot2022} in that fluxes are predominantly horizontal (Figure \ref{fig:fig2}a).
 We see that despite an enforced hydraulic gradient of $30^{\circ}$, the vertical component of the flux cannot propagate through the aquifer given that cells along the aquifer top are hydraulically disconnected from the low permeability domain, hence forcing flow horizontally (red arrows on right panel).
-The numerical solution for this grid type is a problem of cell connectivity, and thus the XT3D formulation does not resolve this issue (second row, Figure \ref{fig:fig2}).
-Results for the full-connectivity grid with the standard conductance-based formulation shows an improvement in the flow solution (third row, Figure \ref{fig:fig2}), with the XT3D formulation successfully reproducing the analytical solution (bottom row, Figure \ref{fig:fig2}).
+The numerical solution for this grid type is a problem of cell connectivity, and thus the XT3D formulation does not resolve this issue (Figure \ref{fig:fig2}b).
+Results for the full-connectivity grid with the standard conductance-based formulation shows an improvement in the flow solution (Figure \ref{fig:fig2}c), with the XT3D formulation successfully reproducing the analytical solution (Figure \ref{fig:fig2}d).
 We see that the grid with full connectivity allows incorporation of the vertical flow component, thus permitting cross-connection between grid layers.
 
-The results shown in Figure \ref{fig:fig2}a and b , which are based on a vertically offset grid with layered connectivity, are consistent with the findings of \cite{bardot2022}.
+The results shown in Figures \ref{fig:fig2}a and b , which are based on a vertically offset grid with layered connectivity, are consistent with the findings of \cite{bardot2022}.
 Despite boundary heads designed to induce specific discharge of 1 m/d along the $30^{\circ}$ incline of the aquifer, the simulated flow within the aquifer is predominantly horizontal, and the magnitude of the specific discharge at the center of the aquifer and the total flow entering the aquifer are both overestimated.
@@
 
-The vertically offset grid with full connectivity (Figure \ref{fig:fig2}c and d) show significantly improved results, even for the standard flow formulation (Figure \ref{fig:fig2}c).
+The vertically offset grid with full connectivity (Figures \ref{fig:fig2}c and d) shows significantly improved results, even for the standard flow formulation (Figure \ref{fig:fig2}c).
 Using the XT3D flow formulation provides the most accurate results, as expected (Figure \ref{fig:fig2}d).
@@
 	\includegraphics[scale=0.8]{../figures/fig2_paper.png}
-	\caption{Numerical results for the test problem using base-case settings for layered-connectivity and full-connectivity grids, with the standard conductance based formulation as well as the XT3D formulation.
-The left panel of each scenario shows the calculated specific discharge at each cell centre within the aquifer (black arrows) and domain (grey arrows).
-The middle panel shows the face flows (red arrows).
+	\caption{Numerical results for the test problem using base-case settings for layered-connectivity and full-connectivity grids, with the standard conductance-based formulation as well as the XT3D formulation.
+The left panel of each scenario shows the calculated specific discharge at each cell center (arrows).
+The middle panel shows the face flows (arrows).
 The right panel shows head contours for the analytical solution (black) as well as the numerical solution (blue).}
@@
 	\includegraphics[scale=0.9]{../figures/fig3paper.png}
-	\caption{Graphs indicating how increasing the number of grid layers within a single hydrogeologic layer (in this case, the aquifer) affects flux magnitude of middle cell (top), flux direction of middle cell (middle) and volumetric flow through the hydrogeologic layer (bottom).
-We immediately notice the errors layered-connectivity grids incur (left panel), and how using a full-connectivity grid (right panel), particularly in conjunction with XT3D (orange line) reproduces the analytical solution from two grid layers used to discretize the hydrogeologic layer.}
+	\caption{Graphs indicating how the number of grid layers within the aquifer affects the flux magnitude in the center cell (top), the flux direction in center cell (middle) and the volumetric flow through the aquifer (bottom).
+Errors are minimized by using full connectivity (right column), the XT3D flow formulation (orange lines) and at least two grids layers in the aquifer.}
 	\label{fig:fig3}
@@
 
-The base case arbitrarily uses 3 grid layers to represent the aquifer, but it is important to address the question: \emph{How many grid layers are required to adequately represent a hydrogeological layer?}.
-Layered-connectivity grids, which are commonly used by groundwater modelers \citep{Reilly2004}, are almost always configured as one grid layer per hydrogeological unit when simulating groundwater flow.
-{\color{red} (AMP: Could maybe mention that it's generally well known you want multiple layers for accurate transport?
-JM: Added "when simulating groundwater flow" to just specify flow explicitly and keep things simple)} However, although computationally efficient, we investigate the effect of altering resolution within the aquifer by using the base case and varying number of cells per aquifer width from one to nine (Figure \ref{fig:fig3}).
-When increasing resolution, cells inside the aquifer were maintained as a square shape to allow fair comparison between models of varying refinement, uncomplicated by cell distortion.
-Therefore, the height of the model extended as each additional grid layer was added to represent the aquifer.
-We compare modeled flux magnitude and direction in the center cell, as well as volumetric flow against the analytical solution.
+Layered-connectivity grids, which are commonly used in groundwater models \citep{Reilly2004}, are often configured using one model grid layer per hydrogeologic unit when simulating groundwater flow.
+The simulations in this subsection investigate the effect of vertically discretizing the aquifer using multiple grid layers (Figure \ref{fig:fig3}).
+The base case uses three grid layers to represent the aquifer.
+When the number of grid layers is altered, the horizontal discretization is also adjusted to maintain square cells inside the aquifer, thus, maintaining a constant cell aspect ratio.
+We compare modeled flux magnitude and direction in cell at the center of the aquifer, as well as volumetric flow through the aquifer, against the analytical solution.
 
-Flux magnitude error for layered-connectivity grids peaked at around 10\% but improved and tended to zero with refinement.
-Flux direction for layered-connectivity grids is clearly a major issue, being reported as horizontal for one grid layer per aquifer width, and only improved to 12$^{\circ}$ for 9 grid layers per aquifer width.
-Flow through the aquifer peaked at 13\% error but decreased to only a 4\% error for 9 grid layers per aquifer width.
+Like the results in Figure \ref{fig:fig2}, the results in Figure \ref{fig:fig3} show that the use of full connectivity and the XT3D flow formulation reduces the errors in calculated flows, provided at least two grid layers are used to vertically discretize the aquifer.
+When a single grid layer is used to represent the aquifer, the enhanced connectivity needed to achieve an accurate solution is not possible, and the XT3D flow formulation is unable to compensate for the inadequate connectivity.
 
-The results for the full-connectivity grid clearly show that a minimum of two grid layers per aquifer width is required to match the analytical result for flux magnitude, direction and volumetric flow.
-This is an important finding as it demonstrates that computational efficiency can be made using a full-connectivity grid over a rectilinear grid overlay approach, but it is important to use a minimum of two grid layers per hydrogeological unit to allow vertical fluxes to be incorporated into the flow solution.
+{\color{red} Saved just in case: This is an important finding as it demonstrates that computational efficiency can be made using a full-connectivity grid over a rectilinear grid overlay approach, but it is important to use a minimum of two grid layers per hydrogeological unit to allow vertical fluxes to be incorporated into the flow solution.}
 
@@
 
-The base case uses a dip angle of 30$^{\circ}$ and an extreme conductivity contrast between aquifer and domain of 1:$10^{-6}$.
-Here, we investigate the behavior of the flow solution for a wide range of aquifer inclines by systematically changing the dip angle by 2.5$^{\circ}$ from 0 to 80$^{\circ}$.
-Similarly, we examine the solution for multiple conductivity contrasts by changing the domain conductivity to 2, 5, 10 and 100 times less than the aquifer.
-Simulated flux magnitude and direction of the center cell for all combinations are plotted against the analytical solution in Figure 4.
+The base case uses an aquifer dip angle of 30$^{\circ}$ and a conductivity contrast between domain and aquifer of $10^{-6}$:1.
+In this subsection, we investigate the behavior of the flow solution for aquifer dip angles ranging from 0 to 80$^{\circ}$.
+We also examine the behavior for a range of conductivity contrasts by setting the domain conductivity to 1/2, 1/5, 1/10 and 1/100 that of the aquifer.
+Simulated flux magnitude and direction at the center cell for all combinations are plotted against the analytical solution in Figure \ref{fig:fig4} {\color{red} THERE'S NO GENERIC REF FOR FIG 4}.
 
-Traditional layered-connectivity grids without XT3D shows rapid diversion of flux magnitude and direction from the analytical solution as dip increases (Figure 4a).
-Use of the XT3D option keeps the flux magnitude somewhat on-track until about 30$^{\circ}$ when the solution starts peeling away from the analytical (Figure 4b).
-Flux direction is improved for the homogeneous scenario, but not for the heterogeneous scenarios.
-The full-connectivity grid on the other hand, even without XT3D, roughly follows the analytical solution for flux magnitude and direction until around 65$^{\circ}$ when it starts to ``fall-apart'' (Figure 4c).
-There are also spikes at certain dip angles which are related to transitions between cell connections due to changing cell face overlaps.
-The full-connectivity grid with XT3D clearly boasts the best results (Figure 4d) with the flux magnitude and direction matching the analytical until 65$^{\circ}$, with the exception of a slight decrease in flux direction after 45$^{\circ}$.
-The errors reflect the discretization choice of the model.
-For the 1 m by 1 m cells, the aquifer center cell is fully connected to other aquifer cells up to an angle of 45$^{\circ}$.
-After 45$^{\circ}$, some aquifer cell faces share greater overlap with domain cell faces than the adjacent aquifer cell face.
-After 65$^{\circ}$, the aquifer center cell loses all lateral connections to adjacent aquifer cells such that only vertical aquifer connections remain, forcing flow vertically (90$^{\circ}$).
-To rectify this issue when encountering dips greater than 45$^{\circ}$, and especially after 65$^{\circ}$, finer discretization in the vertical and horizontal directions should be made to maximize inter-aquifer cell connections.
+With layered connectivity and the standard conductance-based flow formulation (Figure \ref{fig:fig4a}), the magnitude of the error in the flux magnitude (left panel) is similar for all of the conductivity ratios simulated and increases with increasing aquifer dip angle.
+The magnitude of the error in the flux direction (right panel) increases with decreasing conductivity ratio and reaches a maximum at a dip angle of approximately {\color{red}X-Y}\%.
 
-These results confirm that full-connectivity grids with XT3D not only outperform the alternatives when modelling dipping hydrogeological units, but come extremely close to the analytical solution.
-We note that these results relate to the chosen discretization.
-The key element is to ensure that the discretization maintains connectivity at the most extreme angles.
-{\color{red} (JLM: is the advice here that the ratio of vertical to horizontal dicretisation needs to be approximately $\arctan(\theta)$ for three sub-layers.
-Could be a general formula for $\theta$ and number of sub cells)}
+With layered connectivity and the XT3D flow formulation (Figure \ref{fig:fig4b}), the magnitude of the error in the flux magnitude (left panel) increases with decreasing conductivity ratio and increasing aquifer dip angle.
+The magnitude of the error in the flux direction (right panel) increases monotonically with dip angle when there is no conductivity contrast between the domain and aquifer (conductivity ratio = 1.0).
+For the lower conductivity ratios simulated, the magnitude of the error in the flux direction increases with decreasing conductivity ratio and reaches a maximum at a dip angle of approximately {\color{red}X-Y}\%.
+
+With full connectivity and the standard conductance-based flow formulation (Figure \ref{fig:fig4c}), the magnitude of the error in the flux magnitude is no greater than {\color{red}X}\% of the analytical solution of 1 m/d, and the magnitude of the error in the flux direction is no greater than {\color{red}X}$^{\circ}$, at dip angles less than 40$^{\circ}$ for all of the conductivity ratios simulated.
+Given that the cells within the aquifer are square, cells that are nominally in the same grid layer within the aquifer no longer share any interfacial area, i.e., the become hydraulically disconnected, at a dip angle of 45$^{\circ}$.
+A second transition in connectivity occurs at a dip angle of approximately 63$^{\circ}$, and at a dip angle of approximately 72$^{\circ}$ the vertical columns of cells within the aquifer are completely disconnected from their neighbors to either side.
+When the discretization used to represent the aquifer becomes substantially disconnected, it cannot be expected to support accurate results.
+Note that for coarser horizontal discretization, transitions in connectivity within the aquifer would occur at smaller dip angles.
+
+With full connectivity and the XT3D flow formulation (Figure \ref{fig:fig4d}), the magnitude of the error in the flux magnitude is no greater than {\color{red}X}\% of the analytical solution of 1 m/d, and the magnitude of the error in the flux direction is no greater than {\color{red}X}$^{\circ}$, at dip angles less than 40$^{\circ}$ for all of the conductivity ratios simulated.
+The effects of hydraulic disconnection of cells with the aquifer are evident at approximately 63$^{\circ}$ for the magnitude of the error in flow magnitude and at 45$^{\circ}$  for the magnitude of the error in flow direction.
+Overall, use of full connectivity with the XT3D flow formulation reduces the error in calculated flows.
 
@@
 	\includegraphics[width=\textwidth]{../figures/fig4_0_paper.png}
+%	\caption{layered-connectivity grid, standard conductance-based formulation}
+	\label{fig:fig4a}
 \end{subfigure}
@@
 	\includegraphics[width=\textwidth]{../figures/fig4_1_paper.png}
+%	\caption{layered-connectivity grid, XT3D}
+	\label{fig:fig4b}
 \end{subfigure}
@@
 	\includegraphics[width=\textwidth]{../figures/fig4_2_paper.png}
+%	\caption{full-connectivity grid, standard conductance-based formulation}
+	\label{fig:fig4c}
 \end{subfigure}
@@
 	\includegraphics[width=\textwidth]{../figures/fig4_3_paper.png}
+%	\caption{full-connectivity grid, XT3D}
+	\label{fig:fig4d}
 \end{subfigure}
 
-\caption{Flux magnitude and direction for centre cells for varying contrasts in log conductivity (blue, orange, green, red lines) for varying dip angles (x-axis).
-The analytical value is shown in black dashed line.}
+\caption{Flux magnitude and direction at the cell in the center of the aquifer as a function of aquifer dip angle for various ratios of domain to aquifer conductivity (blue, orange, green, red lines).
+The analytical solution is shown as a black dashed line.
+{\color{red} CHANGE RATIO NUMBERS TO THEIR RECIPROCALS}}
 \label{fig:figures}
@@
 
-The base case and subsequent scenarios have assumed isotropic conductivity.
-However, sedimentary layers often exhibit anisotropic conductivity, and therefore we examine the gridding strategies using anisotropic and dipping conductivity tensors within the aquifer, with reduced conductivity perpendicular to the aquifer by 10, 100, 1000 and 10,000.
-We also include the isotropic scenario (ratio of 1).
+The base case and variations up to this point have used isotropic hydraulic conductivity.
+However, sedimentary layers often exhibit anisotropic conductivity.
+Therefore, we examine the effect of anisotropic and dipping conductivity tensors within the aquifer, with conductivity perpendicular to the aquifer reduced by factors of 10, 100, 1000 and 10,000.
+We also include the isotropic scenario (anisotropy ratio of 1).
 Results are presented in Figure \ref{fig:fig5}.
-Layered-connectivity grids produce flux magnitude and flow results that significantly deviate from the analytical solution both with and without the XT3D formulation (Figure \ref{fig:fig5}a), whilst full-connectivity grids with XT3D reproduce the analytical solution (Figure \ref{fig:fig5}b, orange line).
-These results verify, once again, our proposed method of using full-connectivity grids to model dipping anisotropic hydrogeologic layers.
+Layered-connectivity grids produce results that deviate significantly from the analytical solution both with and without the XT3D formulation (Figure \ref{fig:fig5}a), while full-connectivity grids with XT3D reproduce the analytical solution (Figure \ref{fig:fig5}b, orange lines).
 
@@
 	\includegraphics[scale=0.9]{../figures/fig5paper.png}
-	\caption{Varying anisotropy ratios of conductivity withing the aquifer and the resulting flux magnitude of middle cell (top panel), flux direction of middle cell (middle panel) and volumetric flow through the aquifer (bottom panel).
-Plots include results for the standard conductance-based formulation (blue) as well as the XT3D formulation (orange).}
+	\caption{Graphs indicating how the anisotropy ratio within the aquifer affects the flux magnitude in the center cell (top), the flux direction in center cell (middle) and the volumetric flow through the aquifer (bottom).
+Errors are minimized by using full connectivity (right column) and the XT3D flow formulation (orange lines).}
 	\label{fig:fig5}
@@
 They concluded that the XT3D multi-point flux approximation in MODFLOW 6 significantly improves the accuracy of simulated flows when used with a grid-overlay approach in which locations of sedimentary structures, which can be steeply dipping, are mapped onto a relatively fine, rectilinear model grid.
-They also concluded that the XT3D multi-point flux approximation does not perform as well as anticipated when simulating flow through a steeply dipping hydrogeologic layer using deformed vertically offset grids, which are advantageous for significantly reducing the number of grid layers required for many problems.
+They also concluded that the XT3D multi-point flux approximation does not perform as well as anticipated when simulating flow through a steeply dipping hydrogeologic layer using deformed (vertically offset) grids, which are advantageous for significantly reducing the number of grid layers required for many problems.
 They hypothesized that the inherently limited lateral connectivity between cells in DIS and DISV layered grids in MODFLOW 6 was responsible for the inability to obtain accurate flow results in their steeply dipping test problems using vertically offsets grids, with or without XT3D.
@@
 
+{\color{red} AMP: mention that our method reestablishes boundary matching as an efficient alterternative to grid overlay on vertically offset grids}
+
+{\color{red} AMP: mention crossflow}
+
 \section{Acknowledgments}

### 2024-06-10 latex/article.tex
edits by AMP and KB
edits by AMP and KB

@@
 
-%\title{Improving Accuracy of Simulated Flows in Steeply Dipping Layers Using Full-Connectivity Grids with the XT3D Multi-Point Flux Approximation in MODFLOW 6}
-
-% suggested title revision
-\title{Efficient and Accurate Modeling of Flow through Sedimentary Structures with MODFLOW 6}
-
-%\title{Efficient and Accurate Modeling of Flow through Dipping Aquifers with MODFLOW 6}
-
 \title{The Importance of Cell Connectivity for Efficient and Accurate Modeling of Flow through Dipping Aquifers with MODFLOW 6}
@@
 
-\textbf{Key words:} groundwater flow simulation, discretization, dipping layers, cell connectivity, multi-point flux approximation, sedimentary structures
-
-%\textbf{Article impact statement: Groundwater flow through dipping aquifers can be accurately simulated using a vertically offset grid combined with an unstructured grid approach and the XT3D multi-point flow formulation in MODFLOW 6.}
+\textbf{Key words:} groundwater flow simulation, discretization, vertically offset grid, dipping layers, cell connectivity, multi-point flux approximation, sedimentary structures
 
@@
 
-In models based on the Control Volume Finite Difference method {(\color{red} ref for this?)}, which is the method used in the MODFLOW family of hydrologic simulation codes \citep{modflow6gwf}{\color{red} (ref to our new MODFLOW overview paper!)}, the loss of accuracy associated with the boundary-matching approach results from the use of a "two-point" formulation to calculate flows between adjacent grid cells.
+In models based on the Control Volume Finite Difference method, which is the method used in the MODFLOW family of hydrologic simulation codes \citep{modflow6gwf, langevin2024}, the loss of accuracy associated with the boundary-matching approach results from the use of a "two-point" formulation to calculate flows between adjacent grid cells.
 In the two-point formulation, the flow between two cells is proportional to the difference between the heads calculated in those two cells.
@@
 Following \cite{bardot2022}, we call this type of grid ``vertically offset." In a vertically offset grid, subhorizontal cell connections across vertical cell interfaces violate the CVFD requirement, which reduces the accuracy of two-point flow formulation.
-For an isotropic aquifer, \cite{anderson2015applied} recommend a $10^{\circ}$ dip as a practical upper limit when using the two-point flow formulation.
-{\color{red} (AMP: Was this offered in the context of MODFLOW, or more generally???
-I don't have that book.)}
+For an isotropic aquifer, \cite{hoaglund2003} note that ``the error range is smaller for smaller dips, generally $<20$\% within $10^{\circ}$ of dip."
 
 The introduction of unstructured grids into MODFLOW \citep{modflowusg, modflow6gwf} presented new ways for a grid to violate the CVFD requirements.
-To help overcome the limitations of the two-point flow formulation on geometrically irregular grids, MODFLOW-USG and MODFLOW 6 offer a Ghost-Node Correction (GNC) Package \citep{modflowusg, modflow6gwf}{\color{red} (ref to Dickenson and others 2007 ?)} However, configuration of optimal ghost node locations and interpolation weights is highly problem-dependent, and we are not aware of guidelines for handling general unstructured grids or arbitrarily oriented 3D anisotropy.
+To help overcome the limitations of the two-point flow formulation on geometrically irregular grids, MODFLOW-USG and MODFLOW 6 offer a Ghost-Node Correction (GNC) Package \citep{modflowusg, modflow6gwf}.
+However, configuration of optimal ghost node locations and interpolation weights is highly problem-dependent, and we are not aware of guidelines for handling general unstructured grids or arbitrarily oriented 3D anisotropy.
 
 The optional XT3D flow formulation \citep{modflow6xt3d} introduced in MODFLOW 6 offers a more ``automatic" alternative to ghost nodes.
-XT3D is a multi-point flow formulation {\color{red} (refs for multi-point)} based on gradient reconstruction {\color{red} (refs for gradient reconstruction)}.
+XT3D is a multi-point flow formulation \citep{edwards1998, aavatsmark2002} based on gradient reconstruction \citep{mavriplis2003leastsq, diskin2008accuracy}.
 By performing flow calculations using Darcy's Law in its tensorial form using a multi-point approximation of the head gradient vector, XT3D accounts for both arbitrarily oriented anisotropy and geometric irregularity of the grid.
@@
 
-{\color{red} Saved just in case: This is an important finding as it demonstrates that computational efficiency can be made using a full-connectivity grid over a rectilinear grid overlay approach, but it is important to use a minimum of two grid layers per hydrogeological unit to allow vertical fluxes to be incorporated into the flow solution.}
-
 \subsection{Effects of Dip Angle and Conductivity Contrast}
@@
 We also examine the behavior for a range of conductivity contrasts by setting the domain conductivity to 1/2, 1/5, 1/10 and 1/100 that of the aquifer.
-Simulated flux magnitude and direction at the center cell for all combinations are plotted against the analytical solution in Figure \ref{fig:fig4} {\color{red} THERE'S NO GENERIC REF FOR FIG 4}.
+Simulated flux magnitude and direction at the center cell for all combinations are plotted against the analytical solution in Figure \ref{fig:fig4}.
 
-With layered connectivity and the standard conductance-based flow formulation (Figure \ref{fig:fig4a}), the magnitude of the error in the flux magnitude (left panel) is similar for all of the conductivity ratios simulated and increases with increasing aquifer dip angle.
-The magnitude of the error in the flux direction (right panel) increases with decreasing conductivity ratio and reaches a maximum at a dip angle of approximately {\color{red}X-Y}\%.
+With layered connectivity and the standard conductance-based flow formulation (Figure \ref{fig:fig4}a), the magnitude of the error in the flux magnitude (left panel) is similar for all of the conductivity ratios simulated and increases with increasing aquifer dip angle.
+The magnitude of the error in the flux direction (right panel) increases with decreasing conductivity ratio.
+As the surrounding domain becomes much less conductive than the aquifer, the flow direction tends toward horizontal.
 
-With layered connectivity and the XT3D flow formulation (Figure \ref{fig:fig4b}), the magnitude of the error in the flux magnitude (left panel) increases with decreasing conductivity ratio and increasing aquifer dip angle.
-The magnitude of the error in the flux direction (right panel) increases monotonically with dip angle when there is no conductivity contrast between the domain and aquifer (conductivity ratio = 1.0).
-For the lower conductivity ratios simulated, the magnitude of the error in the flux direction increases with decreasing conductivity ratio and reaches a maximum at a dip angle of approximately {\color{red}X-Y}\%.
+With layered connectivity and the XT3D flow formulation (Figure \ref{fig:fig4}b), the magnitude of the error in the flux magnitude (left panel) increases with decreasing conductivity ratio and increasing aquifer dip angle.
+The magnitude of the error in the flux direction (right panel) increases with decreasing conductivity ratio.
+As the surrounding domain becomes much less conductive than the aquifer, the flow direction tends toward horizontal, as with the standard flow formulation.
 
-With full connectivity and the standard conductance-based flow formulation (Figure \ref{fig:fig4c}), the magnitude of the error in the flux magnitude is no greater than {\color{red}X}\% of the analytical solution of 1 m/d, and the magnitude of the error in the flux direction is no greater than {\color{red}X}$^{\circ}$, at dip angles less than 40$^{\circ}$ for all of the conductivity ratios simulated.
+With full connectivity and the standard conductance-based flow formulation (Figure \ref{fig:fig4}c), the magnitude of the error in the flux magnitude is less than 17\% of the analytical solution of 1 m/d, and the magnitude of the error in the flux direction is less than 10$^{\circ}$, at dip angles less than 40$^{\circ}$ for all of the conductivity ratios simulated.
 Given that the cells within the aquifer are square, cells that are nominally in the same grid layer within the aquifer no longer share any interfacial area, i.e., the become hydraulically disconnected, at a dip angle of 45$^{\circ}$.
@@
 
-With full connectivity and the XT3D flow formulation (Figure \ref{fig:fig4d}), the magnitude of the error in the flux magnitude is no greater than {\color{red}X}\% of the analytical solution of 1 m/d, and the magnitude of the error in the flux direction is no greater than {\color{red}X}$^{\circ}$, at dip angles less than 40$^{\circ}$ for all of the conductivity ratios simulated.
-The effects of hydraulic disconnection of cells with the aquifer are evident at approximately 63$^{\circ}$ for the magnitude of the error in flow magnitude and at 45$^{\circ}$  for the magnitude of the error in flow direction.
-Overall, use of full connectivity with the XT3D flow formulation reduces the error in calculated flows.
+With full connectivity and the XT3D flow formulation (Figure \ref{fig:fig4}d), the magnitude of the error in the flux magnitude is less than 3\% of the analytical solution of 1 m/d, and the magnitude of the error in the flux direction is less than 1$^{\circ}$, at dip angles less than 40$^{\circ}$ for all of the conductivity ratios simulated.
+The effects of hydraulic disconnection of cells within the aquifer are evident at approximately 63$^{\circ}$ for the magnitude of the error in flow magnitude and at 45$^{\circ}$  for the magnitude of the error in flow direction.
+Overall, use of full connectivity with the XT3D flow formulation substantially reduces the error in calculated flows.
 
@@
 %	\caption{layered-connectivity grid, standard conductance-based formulation}
-	\label{fig:fig4a}
+%	\label{fig:fig4a}
 \end{subfigure}
@@
 %	\caption{layered-connectivity grid, XT3D}
-	\label{fig:fig4b}
+%	\label{fig:fig4b}
 \end{subfigure}
@@
 %	\caption{full-connectivity grid, standard conductance-based formulation}
-	\label{fig:fig4c}
+%	\label{fig:fig4c}
 \end{subfigure}
@@
 %	\caption{full-connectivity grid, XT3D}
-	\label{fig:fig4d}
+%	\label{fig:fig4d}
 \end{subfigure}
@@
 \caption{Flux magnitude and direction at the cell in the center of the aquifer as a function of aquifer dip angle for various ratios of domain to aquifer conductivity (blue, orange, green, red lines).
-The analytical solution is shown as a black dashed line.
-{\color{red} CHANGE RATIO NUMBERS TO THEIR RECIPROCALS}}
-\label{fig:figures}
+The analytical solution is shown as a black dashed line.}
+\label{fig:fig4}
 \end{figure}
@@
 
-\cite{bardot2022} revisited the capability of MODFLOW to accurately simulate groundwater flow through sedimentary structures.
-They concluded that the XT3D multi-point flux approximation in MODFLOW 6 significantly improves the accuracy of simulated flows when used with a grid-overlay approach in which locations of sedimentary structures, which can be steeply dipping, are mapped onto a relatively fine, rectilinear model grid.
-They also concluded that the XT3D multi-point flux approximation does not perform as well as anticipated when simulating flow through a steeply dipping hydrogeologic layer using deformed (vertically offset) grids, which are advantageous for significantly reducing the number of grid layers required for many problems.
-They hypothesized that the inherently limited lateral connectivity between cells in DIS and DISV layered grids in MODFLOW 6 was responsible for the inability to obtain accurate flow results in their steeply dipping test problems using vertically offsets grids, with or without XT3D.
+\cite{bardot2022} revisited the capability of MODFLOW 6 to accurately simulate groundwater flow through sedimentary structures.
+They observed that calculated flows involved significant error in a benchmark problem that simulated flow through a steeply dipping hydrogeologic layer using vertically offset grids, whether or not the XT3D multi-point flow formulation was used.
+They hypothesized that inadequate lateral connectivity between model cells was responsible for the inability to obtain accurate flow results in their benchmark problem.
 
-In this paper, we have confirmed that the layered lateral connectivity implemented in DIS and DISV model grids is inadequate for simulating groundwater flow through steeply dipping sedimentary structures, even when the XT3D multi-point flux approximation is used.
-However, when connections between laterally adjacent cells in different grid layers are added to comprise a grid with full connectivity, there is a substantial increase in the accuracy of simulated flows.
-The additional connectivity is implemented using the fully unstructured (DISU) grid type.
-Importantly, this accuracy improvement requires that at least two grid layers be used to represent the dipping aquifers, with additional grid layers required for very steep dips above 65$^{\circ}$.
-The primary conclusion of this paper is that, given appropriate discretization and cell connectivity, the XT3D multi-point flux approximation implemented in MODFLOW 6 can be combined with vertically offset grids to efficiently and accurately model flows through steeply dipping sedimentary structures.
-Prior to this work, the capability to efficiently and accurately model flow through steeply dipping sedimentary structures was generally restricted to finite element simulators.
+In this paper, we have confirmed that the layered lateral connectivity offered by DIS and DISV model grids can be inadequate for simulating groundwater flow through steeply dipping sedimentary structures, even when the XT3D multi-point flow formulation is used.
+However, when connections between laterally adjacent cells in different grid layers are added to comprise a grid with full connectivity, there can be a substantial increase in the accuracy of simulated flows.
+The additional connectivity is implemented using the fully unstructured (DISU) grid type in MODFLOW 6.
+This accuracy improvement requires that at least two grid layers be used to represent the dipping aquifers, with additional grid layers required for very steep dips.
 
-The importance of adequate connectivity has been explained and evaluated in this work specifically in terms of hydrogeologic layers that correspond to sedimentary structures.
-However, the impact of cell connectivity on the accuracy of simulated flows should be considered in any MODFLOW application that models flow through a heterogeneous hydraulic conductivity field using a grid with large vertical offsets between laterally adjacent cells, whether or not the grid has a layered structure.
-The full connectivity needed to improve accuracy can be implemented using the DISU grid types in both MODFLOW 6 and MODFLOW-USG.
+Our benchmark results suggest that, given appropriate discretization and cell connectivity, the XT3D multi-point flow formulation implemented in MODFLOW 6 can be combined with vertically offset grids to efficiently and accurately model flows through steeply dipping sedimentary structures.
+Future work may investigate the accuracy of calculated flows under different hydraulic gradients, as well as the implications of grid connectivity for more complex and realistic flow systems.
 
-{\color{red} AMP: mention that our method reestablishes boundary matching as an efficient alterternative to grid overlay on vertically offset grids}
-
-{\color{red} AMP: mention crossflow}
+The importance of adequate grid connectivity has been explained and evaluated in this work specifically in terms of hydrogeologic layers that correspond to sedimentary structures.
+However, the impact of cell connectivity on the accuracy of simulated flows could be important in any MODFLOW application that models flow through a heterogeneous hydraulic conductivity field using a grid with large vertical offsets between laterally adjacent cells, whether or not the grid has an explicitly layered structure.
 
 \section{Acknowledgments}
-We would like to acknowledge.... {\color{red} that this paper is almost ready to GO!! Yeeeha!}
+{\color{red} (Will acknowledge reviewers)}
 
 \section{Supporting Information}
-
-\section{Appendix}
+{\color{red} (To be determined)}
 

### 2024-06-10 latex/article.tex
Add mfusg citation to section on CVFD requirements; add modflow 6.5.0 citation as "current version" in case we add full connectivity as an option to DIS/DISV

@@
 In the two-point formulation, the flow between two cells is proportional to the difference between the heads calculated in those two cells.
-For simulations of flow through isotropic porous media, the two-point flow formulation is most accurate when the grid satisfies the so-called ``CVFD requirement" \citep{modflow6gwf} that the straight-line connection between two cell centers must intersect the midpoint of the cell interface at a right angle \citep{narasimhan1976integrated}.
+For simulations of flow through isotropic porous media, the two-point flow formulation is most accurate when the grid satisfies the so-called ``CVFD requirement" \citep{modflowusg, modflow6gwf} that the straight-line connection between two cell centers must intersect the midpoint of the cell interface at a right angle \citep{narasimhan1976integrated}.
 However, boundary-matching grids violate the CVFD requirement at vertical interfaces between cells in the same model layer because the corresponding cell connections are not strictly horizontal.
@@
 In layered connectivity, two cells that have overlapping vertical faces but are in different grid layers (e.g., cells 1A and 2B) notably do not have a direct hydraulic connection.
-The addition of such connections between cells in different grid layers, as indicated by the red lines that connect cell centers in Figure \ref{fig:schem-conn-area-flux}a, results in a grid with what we call ``full connectivity.''  Implementation of full connectivity in MODFLOW 6 requires that the vertically offset grid be represented using the fully unstructured (DISU) grid type.
+The addition of such connections between cells in different grid layers, as indicated by the red lines that connect cell centers in Figure \ref{fig:schem-conn-area-flux}a, results in a grid with what we call ``full connectivity.''  Implementation of full connectivity in the current version of MODFLOW 6 \citep{modflow650software} requires that the vertically offset grid be represented using the fully unstructured (DISU) grid type.
 

### 2024-06-13 latex/article.tex
Preparations for submission
* clean out old figures
* move figures to the end (endfloat)
* update/complete references for bardot2023 and hughes2024flopy
* make edits related to Supporting Information
* edit text and fig 5 caption to reflect the fact that anisotropy is also in the domain, not just the aquifer
* add support acknowledgment
* clean up some commented-out lines and some stuff that was being printed in notebooks
* clean out older notebooks

@@
 \bibliographystyle{apalike}
+%\bibliographystyle{agsm}
 \usepackage{lineno}
@@
 \captionsetup[figure]{font={stretch=2}}
+\usepackage[nomarkers,figuresonly]{endfloat}
+\usepackage[nottoc]{tocbibind}
 
@@
 Thus, when boundary matching is implemented in a two-dimensional (2D) cross-sectional or three-dimensional (3D) MODFLOW model, the tops of laterally adjacent cells are vertically offset from each other (as are the cell bottoms) and can follow the undulations of the aquifer boundaries only on average.
-Following \cite{bardot2022}, we call this type of grid ``vertically offset." In a vertically offset grid, subhorizontal cell connections across vertical cell interfaces violate the CVFD requirement, which reduces the accuracy of two-point flow formulation.
+Following \cite{bardot2023}, we call this type of grid ``vertically offset." In a vertically offset grid, subhorizontal cell connections across vertical cell interfaces violate the CVFD requirement, which reduces the accuracy of two-point flow formulation.
 For an isotropic aquifer, \cite{hoaglund2003} note that ``the error range is smaller for smaller dips, generally $<20$\% within $10^{\circ}$ of dip."
@@
 
-\cite{bardot2022} tested the ability of MODFLOW 6 to accurately simulate flow through sedimentary structures, both with and without XT3D.
+\cite{bardot2023} tested the ability of MODFLOW 6 to accurately simulate flow through sedimentary structures, both with and without XT3D.
 The simplest 2D cross-sectional (``transect") version of their benchmark model simulated flow through a steeply dipping, homogeneous, isotropic aquifer of uniform thickness embedded between two confining units.
@@
 In this paper, we present theoretical explanations for why ``layered" connectivity in a vertically offset grid is inadequate for accurate simulation of flows in steeply dipping hydrogeologic layers, and why using what we call ``full" cell connectivity together with XT3D should improve accuracy.
-Then, in a set of test problems similar to those of \cite{bardot2022}, we demonstrate the effectiveness of using full connectivity with XT3D.
+Then, in a set of test problems similar to those of \cite{bardot2023}, we demonstrate the effectiveness of using full connectivity with XT3D.
 Finally, we discuss the implications of our findings for practical groundwater problems involving steeply dipping hydrogeologic layers.
@@
 However, XT3D alone cannot compensate for cell connectivity that does not provide adequate pathways for flow.
-This explains why using XT3D did not improve the simulation results substantially on the layered-connectivity grid in the benchmark tests of \cite{bardot2022}.
+This explains why using XT3D did not improve the simulation results substantially on the layered-connectivity grid in the benchmark tests of \cite{bardot2023}.
 In the middle section of the hydrogeologic layer, away from the end boundary conditions, the flow solution approached steady, uniform flow the only way that it could given the limited connectivity: by suppressing vertical flows between grid layers, thereby rendering the overall flow approximately horizontal.
@@
 
-The test problem is patterned after the cross-sectional (transect) benchmark of \cite{bardot2022}.
+The test problem is patterned after the cross-sectional (transect) benchmark of \cite{bardot2023}.
 It attempts to simulate uniform flow in an inclined permeable hydrogeologic layer (``aquifer'') embedded in less permeable material (the ``domain'').
@@
 
-Jupyter notebooks created for the test problem are available in the Supporting Information that accompanies this paper {\color{red} (best place???)}.
-The notebooks leverage the capabilities of FloPy \citep{bakker2016scripting, hughes2023flopy} to assist in setting up input for and processing results from the MODFLOW 6 simulations.
+A link to the Jupyter notebooks and MODFLOW version 6.5.0 executable code used to create and run the test problem is available in the Supporting Information section.
+The notebooks leverage the capabilities of FloPy \citep{bakker2016scripting, hughes2024flopy} to assist in setting up input for and processing results from the MODFLOW 6 simulations.
 A Python script for converting structured, DIS grids to unstructured, DISU grids with full connectivity is included.
@@
 
-The numerical solution for the layered-connectivity grid confirms the findings of \cite{bardot2022} in that fluxes are predominantly horizontal (Figure \ref{fig:fig2}a).
+The numerical solution for the layered-connectivity grid confirms the findings of \cite{bardot2023} in that fluxes are predominantly horizontal (Figure \ref{fig:fig2}a).
 We see that despite an enforced hydraulic gradient of $30^{\circ}$, the vertical component of the flux cannot propagate through the aquifer given that cells along the aquifer top are hydraulically disconnected from the low permeability domain, hence forcing flow horizontally (red arrows on right panel).
@@
 
-The results shown in Figures \ref{fig:fig2}a and b , which are based on a vertically offset grid with layered connectivity, are consistent with the findings of \cite{bardot2022}.
+The results shown in Figures \ref{fig:fig2}a and b, which are based on a vertically offset grid with layered connectivity, are consistent with the findings of \cite{bardot2023}.
 Despite boundary heads designed to induce specific discharge of 1 m/d along the $30^{\circ}$ incline of the aquifer, the simulated flow within the aquifer is predominantly horizontal, and the magnitude of the specific discharge at the center of the aquifer and the total flow entering the aquifer are both overestimated.
@@
 However, sedimentary layers often exhibit anisotropic conductivity.
-Therefore, we examine the effect of anisotropic and dipping conductivity tensors within the aquifer, with conductivity perpendicular to the aquifer reduced by factors of 10, 100, 1000 and 10,000.
+Therefore, we examine the effect of anisotropic and dipping conductivity tensors within the aquifer and domain, with conductivity perpendicular to the aquifer reduced by factors of 10, 100, 1000 and 10,000.
 We also include the isotropic scenario (anisotropy ratio of 1).
@@
 	\includegraphics[scale=0.9]{../figures/fig5paper.png}
-	\caption{Graphs indicating how the anisotropy ratio within the aquifer affects the flux magnitude in the center cell (top), the flux direction in center cell (middle) and the volumetric flow through the aquifer (bottom).
+	\caption{Graphs indicating how the anisotropy ratio within the aquifer and domain affects the flux magnitude in the center cell (top), the flux direction in center cell (middle) and the volumetric flow through the aquifer (bottom).
 Errors are minimized by using full connectivity (right column) and the XT3D flow formulation (orange lines).}
@@
 
-\cite{bardot2022} revisited the capability of MODFLOW 6 to accurately simulate groundwater flow through sedimentary structures.
+\cite{bardot2023} revisited the capability of MODFLOW 6 to accurately simulate groundwater flow through sedimentary structures.
 They observed that calculated flows involved significant error in a benchmark problem that simulated flow through a steeply dipping hydrogeologic layer using vertically offset grids, whether or not the XT3D multi-point flow formulation was used.
@@
 \section{Acknowledgments}
-{\color{red} (Will acknowledge reviewers)}
+Kerry Bardot is supported by The Australian Research Council, The Western Australian Department of Water and Environmental Regulation and The Water Corporation of Western Australia through grant number LP18101153.
+The authors offer their sincere thanks to the reviewers of this manuscript.
 
 \section{Supporting Information}
-{\color{red} (To be determined)}
+During review, the files that support the creation and running of the test problem are available at {\color{red} XXXXXXXXXX}.
+Once these files are finalized prior to publication, they will be made available at a permanent URL that will be noted here.
 

### 2024-11-19 latex/article_clean_FINAL.tex
intermediate review drafts & final draft in progress
intermediate review drafts & final draft in progress

--- 2024-06-13 latex/article.tex
@@
 \usepackage{natbib}
+%\setcitestyle{aysep={}}
 \usepackage[labelfont=bf]{caption}
@@
 \usepackage{float}
-\bibliographystyle{apalike}
+%\bibliographystyle{apalike}
 %\bibliographystyle{agsm}
+\bibliographystyle{groundwater}
 \usepackage{lineno}
@@
 \usepackage{xcolor}
+\usepackage{ulem}
 
 \newcommand{\matr}[1]{\mathbf{#1}}
+\renewcommand{\bibname}{References}
 
-\title{The Importance of Cell Connectivity for Efficient and Accurate Modeling of Flow through Dipping Aquifers with MODFLOW 6}
+\title{Accurate Simulation of Flow through Dipping Aquifers with MODFLOW 6 Using Enhanced Cell Connectivity}
 
 \author{
-	Alden M. Provost, U.S. Geological Survey, Integrated Modeling and Prediction Division, U.S. Geological Survey, 12201 Sunrise Valley Dr, Reston, VA, USA  \\
+	Alden M. Provost (corresponding author), U.S. Geological Survey, Integrated Modeling and Prediction Division, U.S. Geological Survey, 12201 Sunrise Valley Dr, Reston, VA, USA; aprovost@usgs.gov \\
 	\and 
-	Kerry Bardot, School of Earth Sciences, University of Western Australia, Perth, Australia \\
+	Kerry Bardot, School of Earth Sciences, University of Western Australia, Perth, Australia; kerry.bardot@research.uwa.edu.au \\
 	\and 
-	Christian D. Langevin, U.S. Geological Survey, Integrated Modeling and Prediction Division, 2280 Woodale Drive, Mounds View, MN, USA \\
+	Christian D. Langevin, U.S. Geological Survey, Integrated Modeling and Prediction Division, 2280 Woodale Drive, Mounds View, MN, USA; langevin@usgs.gov \\
 	\and 
-	James L. McCallum, School of Earth Sciences, University of Western Australia, Perth, Australia \\
+	James L. McCallum, School of Earth Sciences, University of Western Australia, Perth, Australia; james.mccallum@uwa.edu.au \\
 	}
@@
 
-\maketitle
+\noindent Methods Brief/
+
+%\maketitle
+{\let\newpage\relax\maketitle}
 
@@
 
-Groundwater flow through a dipping aquifer is often modeled using a single layer of model cells to represent the aquifer.
-The layer of cells is ``deformed'' to follow the top and bottom elevations of the aquifer.
-When this approach is used in MODFLOW, adjacent cells within the model layer are vertically offset from one another, and the standard conductance-based (two-point) formulation for flow between cells does not rigorously account for these offsets.
-The XT3D multi-point flow formulation introduced in MODFLOW 6 is designed to account for geometric irregularities in the grid, including vertical offsets, and to provide accurate results for both isotropic and anisotropic groundwater flow.
+In simulations of groundwater flow through dipping aquifers, layers of model cells are often ``deformed'' to follow the top and bottom elevations of the aquifers.
+When this approach is used in MODFLOW, adjacent cells within the same model layer are vertically offset from one another, and the standard conductance-based (two-point) formulation for flow between cells does not rigorously account for these offsets.
+The XT3D multi-point flow formulation in MODFLOW 6 is designed to account for geometric irregularities in the grid, including vertical offsets, and to provide accurate results for both isotropic and anisotropic groundwater flow.
 A recent study evaluated the performance of the standard formulation and XT3D using a simple, synthetic benchmark model of a steeply dipping aquifer.
 Although XT3D generally improved the accuracy of flow simulations relative to the standard formulation as expected, neither formulation produced accurate flows in cases that involved large vertical offsets.
-In this paper, we explain that the inability of XT3D to produce accurate flows in the steeply dipping aquifer benchmark was not due to an inherent limitation of the flow formulation, but to the limited cell connectivity inherent in the most commonly used discretization packages in MODFLOW 6.
+In this paper, we explain that the inability of XT3D to produce accurate flows in the steeply dipping aquifer benchmark was not due to an inherent limitation of the flow formulation, but rather to the limited cell connectivity inherent in the most commonly used discretization packages in MODFLOW 6.
 Furthermore, we demonstrate that XT3D is able to produce the expected accuracy when adequate cell connectivity is introduced using MODFLOW's unstructured grid type and the aquifer is discretized vertically using at least two model layers.
@@
 
-\section{Introduction}
+\section*{Introduction}
 
@@
 A traditional approach to discretizing layered aquifer systems, called the ``grid overlay'' method by \cite{hoaglund2003}, maps the heterogeneous hydrologic properties onto a rectilinear grid \citep{modflow84}.
-While this method preserves the benefits of the rectilinear grid structure, it can require a large number of model grid layers to adequately discretize hydrogeogic layers of variable dip and thickness \citep{Zyvoloski2006}.
+Although this method preserves the benefits of the rectilinear grid structure, it can require a large number of model grid layers to adequately discretize hydrogeologic layers of variable dip and thickness \citep{Zyvoloski2006}.
 
 An alternative approach, called the ``boundary-matching" method by \cite{hoaglund2003}, allows model grid layers to "deform" vertically to follow the undulations of hydrogeologic layers.
-This method ``minimize[s] the number of model [grid] layers required to simulate an aquifer system'' \citep{modflow84}.
-In fact, it is common practice to represent each hydrogeologic layer using a single, boundary-matching model grid layer.
-However, the non-orthogonality of the resulting grid can be a source of significant model error \citep{hoaglund2003}.
+This commonly used method ``minimize[s] the number of model [grid] layers required to simulate an aquifer system'' \citep{modflow84}.
+However, the non-orthogonality of the resulting grid can be a source of substantial model error \citep{hoaglund2003}.
 
-In models based on the Control Volume Finite Difference method, which is the method used in the MODFLOW family of hydrologic simulation codes \citep{modflow6gwf, langevin2024}, the loss of accuracy associated with the boundary-matching approach results from the use of a "two-point" formulation to calculate flows between adjacent grid cells.
+In models based on the Control Volume Finite Difference (CVFD) method, which is the method used in the MODFLOW family of hydrologic simulation codes \citep{modflow6gwf, langevin2024}, the loss of accuracy associated with the boundary-matching approach results from the use of a "two-point" formulation to calculate flows between adjacent grid cells.
 In the two-point formulation, the flow between two cells is proportional to the difference between the heads calculated in those two cells.
@@
 To help overcome the limitations of the two-point flow formulation on geometrically irregular grids, MODFLOW-USG and MODFLOW 6 offer a Ghost-Node Correction (GNC) Package \citep{modflowusg, modflow6gwf}.
-However, configuration of optimal ghost node locations and interpolation weights is highly problem-dependent, and we are not aware of guidelines for handling general unstructured grids or arbitrarily oriented 3D anisotropy.
+However, configuration of optimal ghost node locations and interpolation weights is problem-dependent, and we are not aware of guidelines for handling general unstructured grids or arbitrarily oriented 3D anisotropy.
 
-The optional XT3D flow formulation \citep{modflow6xt3d} introduced in MODFLOW 6 offers a more ``automatic" alternative to ghost nodes.
+The optional XT3D flow formulation \citep{modflow6xt3d} in MODFLOW 6 offers a more ``automatic" alternative to ghost nodes.
 XT3D is a multi-point flow formulation \citep{edwards1998, aavatsmark2002} based on gradient reconstruction \citep{mavriplis2003leastsq, diskin2008accuracy}.
@@
 The simplest 2D cross-sectional (``transect") version of their benchmark model simulated flow through a steeply dipping, homogeneous, isotropic aquifer of uniform thickness embedded between two confining units.
-In test runs that used the boundary-matching method on a vertically offset grid, both the two-point flow formulation and XT3D produced flows that differed markedly from the expected magnitude and direction throughout the aquifer, regardless of how finely the aquifer was discretized.
+In test runs that used the boundary-matching method on a vertically offset grid, both the two-point flow formulation and XT3D produced flows that differed markedly from the expected magnitude and direction throughout the aquifer, regardless of how finely the aquifer was discretized both vertically and horizontally.
 They noted, however, that a preliminary analysis by two authors of the present work (Provost and Langevin) suggested that the limited cell connectivity offered by the layered, vertically offset grids was inadequate to allow an accurate flow solution.
@@
 
-\section{Theory}
+\section*{Theory}
 
@@
 The DIS and DISV grid types are based on model layers, with the same horizontal grid structure within each layer.
-The DISU grid type is more flexible and does not depend on the concept of model layers.
+In a DIS or DISV grid, a connection between two cells in different model layers is created automatically if and only if one cell is directly below the other and neither cell has been optionally excluded from the model domain by the user.
+The DISU grid type is more flexible and, in MODFLOW 6, does not require the grid to have a layered structure.
+(The DISU grid type in MODFLOW-USG is similarly flexible but retains the concept of layers.) In a DISU grid, the user explicitly specifies all connections between grid cells.
+Thus, when used to represent a grid with a layered structure, as in this work, the DISU grid type allows connections to exist between any two cells, whether they are in the same layer or different layers.
+Grid geometry and connectivity information supplied by the discretization package (DIS, DISV, or DISU) is used by the MODFLOW 6 Node Property Flow (NPF) Package to formulate expressions for groundwater flow between cells in terms of heads calculated at cell centers.
+By default, the NPF Package uses a two-point flow formulation, which we call the ``conductance-based'' formulation.
+The optional, multi-point XT3D flow formulation can be activated simply by including the corresponding keyword in the NPF Package input.
 
@@
 
-The role of full cell connectivity in enabling accurate simulation of flow along a dipping hydrogeologic layer can be understood by considering Figure \ref{fig:schem-conn-area-flux}b.
+The need for full cell connectivity to accurately simulate flow along the dipping hydrogeologic layer is illustrated in Figure \ref{fig:schem-conn-area-flux}b.
 Components of the groundwater flux (specific discharge) vector are displayed on each cell-cell interface and are representative of steady, uniform flow along the hydrogeologic layer.
@@
 The blue vectors in Figure \ref{fig:schem-conn-area-flux}b show that to simulate steady, uniform flow along the dipping hydrogeologic layer, cells must exchange water not only ``horizontally'' with neighboring cells in the same grid layer, but also vertically with cells in the adjacent grid layer.
-Specifically, there must be flow from cells in the bottom grid layer to cells in the top grid layer, which corresponds to the vertical component of the groundwater flux.
+Specifically, there must be flow from cells in the lower grid layer to cells in the upper grid layer, which corresponds to the vertical component of the groundwater flux.
 Note, however, that the blue vectors alone are incompatible with steady, uniform flow along the hydrogeologic layer because the exchange of water between the two grid layers is in one direction only: upward.
-Such a steady-state flow configuration would imply continuous depletion of flow within the bottom grid layer and accumulation of flow within the top grid layer as one moves along the hydrogeologic layer in the direction of flow.
-Thus, steady, uniform flow along the hydrogeologic layer cannot be simulated accurately without some mechanism for returning flow from the top grid layer to the bottom grid layer.
+Such a steady-state flow configuration would imply continuous depletion of flow within the lower grid layer and accumulation of flow within the upper grid layer as one moves along the hydrogeologic layer in the direction of flow.
+Thus, the steady, uniform flow represented by the gray vectors cannot be simulated accurately without some mechanism for returning flow from the upper grid layer to the lower grid layer.
 The full-connectivity grid provides such a mechanism by offering additional connections between cells in adjacent grid layers.
-Vertical flows from the bottom grid layer to the top grid layer, represented by vertical blue vectors in Figure \ref{fig:schem-conn-area-flux}b, can be returned to the bottom grid layer by nominally ``horizontal'' flows, represented by red vectors in Figure \ref{fig:schem-conn-area-flux}b.
+Vertical flows from the lower grid layer to the upper grid layer, represented by vertical blue vectors in Figure \ref{fig:schem-conn-area-flux}b, can be returned to the lower grid layer by nominally ``horizontal'' flows, represented by red vectors in Figure \ref{fig:schem-conn-area-flux}b.
 Although this concept has been illustrated using a two-grid-layer model for simplicity, the same reasoning applies given any number of grid layers.
 
-Note how the vertically offset grid violates the CFVD requirement because nominally ``horizontal'' connections between cell centers in the same grid layer are not strictly horizontal and therefore not perpendicular to the vertical cell-cell interfaces they intersect.
+Vertically offset grids violate the CFVD requirement because nominally ``horizontal'' connections between cell centers in the same grid layer are not strictly horizontal and therefore not perpendicular to the vertical cell-cell interfaces they intersect.
 For example, in Figure \ref{fig:schem-conn-area-flux}a, the connection between cells 1A and 1B is not perpendicular to the cell-cell interface, which reduces the accuracy of the conductance-based formulation.
@@
 
-\section{Description of Test Problem}
+\section*{Description of Test Problem}
 
@@
 The test simulations use grids of MODFLOW 6 DISU (unstructured) type, which offers the flexibility to implement either layered or full connectivity.
-In all cases, a grid of MODFLOW 6 DIS (structured) type is first created and then converted a DISU grid.
+In all cases, a grid of MODFLOW 6 DIS (structured) type is first created and then converted to a DISU grid.
 In simulations that use layered connectivity, the DISU grid is formulated to have the same cell geometry and connectivity as the original DIS grid.
@@
 
-\section{Results and Discussion}
+\section*{Results and Discussion}
 
-To demonstrate the effect of cell connectivity on the accuracy of simulated flow through a steeply dipping aquifer, results for the base case described in the previous section are compared for simulations that use either layered or full connectivity, and either the two-point flow formulation (also known as the "standard conductance-based formulation" in MODFLOW 6) or the XT3D flow formulation.
-The base case is then modified to further investigate the effects of grid resolution, aquifer dip angle, conductivity contrast between the aquifer and the surrounding domain, and anisotropic hydraulic conductivity.
+To demonstrate the effect of cell connectivity on the accuracy of simulated flow through a steeply dipping aquifer, results for the base case described in the previous section are compared for simulations that use either layered or full connectivity, and either the standard conductance-based (two-point) flow formulation or the XT3D (multi-point) flow formulation.
+The base case is then modified to investigate the effects of grid resolution, aquifer dip angle, conductivity contrast between the aquifer and the surrounding domain, and anisotropic hydraulic conductivity.
 
@@
 Likewise, specific discharge in cells within the surrounding domain is calculated using only face flows between cells within the domain.
-Specific discharge values reported in the binary budget output file by MODFLOW 6 are different than the values reported here for cells that are adjacent to the boundary between the aquifer and the domain.
-This is because MODFLOW interpolates specific discharge in a cell using all of the face flows for that cell, including flows across cell faces on the aquifer-domain boundary.
-Those flows represent an ``average'' of the different flows on either side of the aquifer-domain boundary.
+These specific discharge values are different from those reported in the binary budget output file by MODFLOW 6 because MODFLOW interpolates specific discharge in a cell using all of the face flows for that cell, including flows across cell faces on the aquifer-domain boundary.
+The flows reported by MODFLOW represent an ``average'' of the different flows on either side of the aquifer-domain boundary.
 Therefore, specific discharge calculated in a cell adjacent to the aquifer-domain boundary is affected by the different flow rate on the other side of the boundary.
-While this is an unavoidable consequence of vertically offset discretization in practical applications, and not necessarily a problem, users should be aware of these effects when interpreting the results of MODFLOW 6 dispersive-transport and MODPATH particle-tracking simulations.
+Although this is an unavoidable consequence of vertically offset discretization in practical applications, and not necessarily a problem, users should be aware of these effects when interpreting the results of MODFLOW 6 dispersive-transport and MODPATH particle-tracking simulations.
 
-\subsection{Effects of Cell Connectivity and Flow Formulation}
-
-The numerical solution for the layered-connectivity grid confirms the findings of \cite{bardot2023} in that fluxes are predominantly horizontal (Figure \ref{fig:fig2}a).
-We see that despite an enforced hydraulic gradient of $30^{\circ}$, the vertical component of the flux cannot propagate through the aquifer given that cells along the aquifer top are hydraulically disconnected from the low permeability domain, hence forcing flow horizontally (red arrows on right panel).
-The numerical solution for this grid type is a problem of cell connectivity, and thus the XT3D formulation does not resolve this issue (Figure \ref{fig:fig2}b).
-Results for the full-connectivity grid with the standard conductance-based formulation shows an improvement in the flow solution (Figure \ref{fig:fig2}c), with the XT3D formulation successfully reproducing the analytical solution (Figure \ref{fig:fig2}d).
-We see that the grid with full connectivity allows incorporation of the vertical flow component, thus permitting cross-connection between grid layers.
+\subsection*{Effects of Cell Connectivity and Flow Formulation}
 
 The results shown in Figures \ref{fig:fig2}a and b, which are based on a vertically offset grid with layered connectivity, are consistent with the findings of \cite{bardot2023}.
-Despite boundary heads designed to induce specific discharge of 1 m/d along the $30^{\circ}$ incline of the aquifer, the simulated flow within the aquifer is predominantly horizontal, and the magnitude of the specific discharge at the center of the aquifer and the total flow entering the aquifer are both overestimated.
+Despite boundary heads designed to induce specific discharge of 1 m/d along the $30^{\circ}$ incline of the aquifer, the simulated flow within the aquifer is predominantly horizontal, and the magnitude of the specific discharge at the center of the aquifer (``flux magnitude'') and the total flow entering the aquifer are overestimated by approximately 15\% and 18\%, respectively.
 The XT3D flow formulation is unable to compensate for the inadequate cell connectivity (Figure \ref{fig:fig2}b), which does not allow flow between laterally adjacent cells in different grid layers.
 
-The vertically offset grid with full connectivity (Figures \ref{fig:fig2}c and d) shows significantly improved results, even for the standard flow formulation (Figure \ref{fig:fig2}c).
-Using the XT3D flow formulation provides the most accurate results, as expected (Figure \ref{fig:fig2}d).
+The vertically offset grid with full connectivity (Figures \ref{fig:fig2}c and d) shows substantially improved results, even for the standard flow formulation (Figure \ref{fig:fig2}c), which overestimates the flux magnitude and total flow by approximately 3\% and 9\%, respectively.
+Using the XT3D flow formulation provides the most accurate results, as expected (Figure \ref{fig:fig2}d), reducing the errors in flux magnitude and total flow to approximately $-2 \times 10^{-4}$\% and $2 \times 10^{-4}$\%, respectively.
 
@@
 	\begin{center}
-	\includegraphics[scale=0.8]{../figures/fig2_paper.png}
+	\includegraphics[scale=0.75]{../figures/fig2_paper.png}
 	\caption{Numerical results for the test problem using base-case settings for layered-connectivity and full-connectivity grids, with the standard conductance-based formulation as well as the XT3D formulation.
@@
 The middle panel shows the face flows (arrows).
-The right panel shows head contours for the analytical solution (black) as well as the numerical solution (blue).}
+The right panel shows head contours for the analytical solution (black) as well as the numerical solution (blue).
+Flux magnitudes, flux directions (angles), and head errors are reported to two decimal places.
+Percent errors in flow are reported to one decimal place.}
 	\label{fig:fig2}
@@
 
-\subsection{Effects of Grid Resolution}
+\subsection*{Effects of Grid Resolution}
 
@@
 
-Layered-connectivity grids, which are commonly used in groundwater models \citep{Reilly2004}, are often configured using one model grid layer per hydrogeologic unit when simulating groundwater flow.
+Groundwater models often use layered-connectivity grids \citep{Reilly2004}.
+Although the practice is less common today than it once was, models are sometimes configured using one grid layer per hydrogeologic unit.
 The simulations in this subsection investigate the effect of vertically discretizing the aquifer using multiple grid layers (Figure \ref{fig:fig3}).
@@
 When the number of grid layers is altered, the horizontal discretization is also adjusted to maintain square cells inside the aquifer, thus, maintaining a constant cell aspect ratio.
-We compare modeled flux magnitude and direction in cell at the center of the aquifer, as well as volumetric flow through the aquifer, against the analytical solution.
+We compare the simulated flux magnitude and direction in the cell at the center of the aquifer, as well as volumetric flow through the aquifer, against the analytical solution.
 
-Like the results in Figure \ref{fig:fig2}, the results in Figure \ref{fig:fig3} show that the use of full connectivity and the XT3D flow formulation reduces the errors in calculated flows, provided at least two grid layers are used to vertically discretize the aquifer.
+The results in Figure \ref{fig:fig3} confirm that the use of full connectivity and the XT3D flow formulation reduces the errors in calculated flows, provided at least two grid layers are used to vertically discretize the aquifer.
+For example, when the aquifer is represented using two layers with layered connectivity, the flow error is approximately 18\% and 5\% for the standard and XT3D flow formulations, respectively.
+Use of full connectivity reduces these errors to approximately 11\% and $3 \times 10^{-4}$\%, respectively.
 When a single grid layer is used to represent the aquifer, the enhanced connectivity needed to achieve an accurate solution is not possible, and the XT3D flow formulation is unable to compensate for the inadequate connectivity.
+With full connectivity and the XT3D flow formulation, using more than two layers to discretize the aquifer provides no substantial benefit in our benchmark problem (Figure \ref{fig:fig3}b).
 
-\subsection{Effects of Dip Angle and Conductivity Contrast}
+\subsection*{Effects of Dip Angle and Conductivity Contrast}
 
@@
 With layered connectivity and the standard conductance-based flow formulation (Figure \ref{fig:fig4}a), the magnitude of the error in the flux magnitude (left panel) is similar for all of the conductivity ratios simulated and increases with increasing aquifer dip angle.
-The magnitude of the error in the flux direction (right panel) increases with decreasing conductivity ratio.
-As the surrounding domain becomes much less conductive than the aquifer, the flow direction tends toward horizontal.
+The magnitude of the error in the flux direction (right panel) increases with decreasing ratio of domain conductivity to aquifer conductivity (increasing conductivity contrast).
+As the surrounding domain becomes much less conductive than the aquifer, the flow direction tends toward horizontal, as shown by the ``flattening'' of the curve toward a flux direction of zero degrees as the conductivity ratio decreases from 1.0 to 0.01 in Figure \ref{fig:fig4}a.
 
@@
 The magnitude of the error in the flux direction (right panel) increases with decreasing conductivity ratio.
-As the surrounding domain becomes much less conductive than the aquifer, the flow direction tends toward horizontal, as with the standard flow formulation.
+As the surrounding domain becomes much less conductive than the aquifer, the flow direction tends toward horizontal in Figure \ref{fig:fig4}b, as occurs with the standard flow formulation in Figure \ref{fig:fig4}a.
 
-With full connectivity and the standard conductance-based flow formulation (Figure \ref{fig:fig4}c), the magnitude of the error in the flux magnitude is less than 17\% of the analytical solution of 1 m/d, and the magnitude of the error in the flux direction is less than 10$^{\circ}$, at dip angles less than 40$^{\circ}$ for all of the conductivity ratios simulated.
-Given that the cells within the aquifer are square, cells that are nominally in the same grid layer within the aquifer no longer share any interfacial area, i.e., the become hydraulically disconnected, at a dip angle of 45$^{\circ}$.
+With full connectivity and the standard conductance-based flow formulation (Figure \ref{fig:fig4}c), the maximum error in the flux magnitude is less than 17\% of the analytical solution of 1 m/d and the maximum error in the flux direction is less than 10$^{\circ}$ at dip angles less than 40$^{\circ}$ for all of the conductivity ratios simulated.
+At a dip angle of  45$^{\circ}$, cells that are nominally in the same grid layer within the aquifer no longer share any interfacial area, thus becoming hydraulically disconnected from one another.
 A second transition in connectivity occurs at a dip angle of approximately 63$^{\circ}$, and at a dip angle of approximately 72$^{\circ}$ the vertical columns of cells within the aquifer are completely disconnected from their neighbors to either side.
 When the discretization used to represent the aquifer becomes substantially disconnected, it cannot be expected to support accurate results.
-Note that for coarser horizontal discretization, transitions in connectivity within the aquifer would occur at smaller dip angles.
+Note that in our benchmark problem the cells within the aquifer are square.
+For the same vertical discretization but coarser horizontal discretization, transitions in connectivity within the aquifer would occur at smaller dip angles due to the smaller vertical-to-horizontal aspect ratio of the cells.
 
-With full connectivity and the XT3D flow formulation (Figure \ref{fig:fig4}d), the magnitude of the error in the flux magnitude is less than 3\% of the analytical solution of 1 m/d, and the magnitude of the error in the flux direction is less than 1$^{\circ}$, at dip angles less than 40$^{\circ}$ for all of the conductivity ratios simulated.
-The effects of hydraulic disconnection of cells within the aquifer are evident at approximately 63$^{\circ}$ for the magnitude of the error in flow magnitude and at 45$^{\circ}$  for the magnitude of the error in flow direction.
+With full connectivity and the XT3D flow formulation (Figure \ref{fig:fig4}d), the maximum error in the flux magnitude is less than 3\% of the analytical solution of 1 m/d and the maximum error in the flux direction is less than 1$^{\circ}$ at dip angles less than 40$^{\circ}$ for all of the conductivity ratios simulated.
+The effects of hydraulic disconnection of cells on errors in flow magnitude and flow direction are evident at approximately 63$^{\circ}$ and 45$^{\circ}$, respectively.
 Overall, use of full connectivity with the XT3D flow formulation substantially reduces the error in calculated flows.
@@
 
-\subsection{Effect of Anisotropy}
+\subsection*{Effect of Anisotropy}
 
-The base case and variations up to this point have used isotropic hydraulic conductivity.
+The base case and variations on the base case up to this point have used isotropic hydraulic conductivity.
 However, sedimentary layers often exhibit anisotropic conductivity.
-Therefore, we examine the effect of anisotropic and dipping conductivity tensors within the aquifer and domain, with conductivity perpendicular to the aquifer reduced by factors of 10, 100, 1000 and 10,000.
-We also include the isotropic scenario (anisotropy ratio of 1).
-Results are presented in Figure \ref{fig:fig5}.
-Layered-connectivity grids produce results that deviate significantly from the analytical solution both with and without the XT3D formulation (Figure \ref{fig:fig5}a), while full-connectivity grids with XT3D reproduce the analytical solution (Figure \ref{fig:fig5}b, orange lines).
+Therefore, we examine the effect of anisotropy within the aquifer and domain by dividing the original conductivity perpendicular to the dip of the aquifer by factors of 1, 10, 100, 1,000 and 10,000 to give anisotropy ratios ranging from 1.0 (isotropic) to 0.0001.
+Orientation of the principal directions of anisotropy parallel and perpendicular to the dip of the aquifer is accomplished by setting three angles in the NPF Package input.
+The first angle is the rotation within the horizontal plane of the first principal direction (typically the direction of maximum hydraulic conductivity) with respect to the positive $x$ axis.
+The second angle is the rotation of the first principal direction upward from the horizontal plane.
+The third angle is a rotation of the second and third principal directions about the axis of the first principal direction.
+For this two-dimensional, cross-sectional problem, the first and third angles are set to zero because they represent three-dimensional effects, and the second angle is set equal to the dip angle ($30^{\circ}$).
+
+Layered-connectivity grids produce results that deviate substantially from the analytical solution both with and without the XT3D formulation (Figure \ref{fig:fig5}a), whereas full-connectivity grids with XT3D reproduce the analytical solution (Figure \ref{fig:fig5}b, orange lines).
+For example, at an anisotropy ratio of 0.01, use of layered connnectivity produces flow errors of approximately -96\% for both the standard and XT3D flow formulations.
+Use of full connectivity produces errors of approximately 30\% and -3\% for the standard and XT3D flow formulations, respectively.
+These results suggest that use of full grid connectivity and the XT3D flow formulation in MODFLOW 6 simulations can be particularly important in the presence of anisotropy.
 
@@
 
-\section{Conclusions}
+\section*{Conclusions}
 
 \cite{bardot2023} revisited the capability of MODFLOW 6 to accurately simulate groundwater flow through sedimentary structures.
-They observed that calculated flows involved significant error in a benchmark problem that simulated flow through a steeply dipping hydrogeologic layer using vertically offset grids, whether or not the XT3D multi-point flow formulation was used.
+They observed that calculated flows involved substantial error in a benchmark problem that simulated flow through a steeply dipping hydrogeologic layer using vertically offset grids, whether or not the XT3D multi-point flow formulation was used.
 They hypothesized that inadequate lateral connectivity between model cells was responsible for the inability to obtain accurate flow results in their benchmark problem.
@@
 In this paper, we have confirmed that the layered lateral connectivity offered by DIS and DISV model grids can be inadequate for simulating groundwater flow through steeply dipping sedimentary structures, even when the XT3D multi-point flow formulation is used.
-However, when connections between laterally adjacent cells in different grid layers are added to comprise a grid with full connectivity, there can be a substantial increase in the accuracy of simulated flows.
-The additional connectivity is implemented using the fully unstructured (DISU) grid type in MODFLOW 6.
+However, when connections between laterally adjacent cells in different grid layers are added using the fully unstructured (DISU) grid type to comprise a grid with full connectivity, the accuracy of simulated flows can increase substantially.
 This accuracy improvement requires that at least two grid layers be used to represent the dipping aquifers, with additional grid layers required for very steep dips.
 
-Our benchmark results suggest that, given appropriate discretization and cell connectivity, the XT3D multi-point flow formulation implemented in MODFLOW 6 can be combined with vertically offset grids to efficiently and accurately model flows through steeply dipping sedimentary structures.
-Future work may investigate the accuracy of calculated flows under different hydraulic gradients, as well as the implications of grid connectivity for more complex and realistic flow systems.
+Note that vertical offsets could be avoided when modeling a 2D cross-section using MODFLOW 6 by formulating the model in plan view, such that rows of neighboring cells play the role of model layers.
+This would allow a DISV grid to follow the top and bottom surfaces of a sloping aquifer in a piecewise-linear fashion without vertical offsets.
+For a 3D model, an alternative could be to use a simulator based on the finite element method \citep[see, e.g.,][]{yager2009anisotropy}.
+In evaluating any alternative approach, the ability to accurately simulate flows on the resulting grid, possibly in the presence of anisotropy that is not aligned with the grid, remains one of the considerations.
 
-The importance of adequate grid connectivity has been explained and evaluated in this work specifically in terms of hydrogeologic layers that correspond to sedimentary structures.
-However, the impact of cell connectivity on the accuracy of simulated flows could be important in any MODFLOW application that models flow through a heterogeneous hydraulic conductivity field using a grid with large vertical offsets between laterally adjacent cells, whether or not the grid has an explicitly layered structure.
+Our benchmark results suggest that, given appropriate discretization and cell connectivity, the XT3D multi-point flow formulation implemented in MODFLOW 6 can be combined with vertically offset grids to efficiently and accurately simulate flows through steeply dipping sedimentary structures.
+Our benchmark problem, which is patterned after that of \cite{bardot2023}, features a two-dimensional cross-section with flow along a dipping aquifer of uniform thickness that is discretized using square model cells.
+Additional investigation is needed to assess the accuracy of calculated flows under different hydraulic gradients and more typical cell aspect ratios, and to better understand the implications of grid connectivity for more complex and realistic flow systems.
 
-\section{Acknowledgments}
+The importance of adequate grid connectivity has been explained and demonstrated in this work specifically in terms of hydrogeologic layers that correspond to sedimentary structures.
+However, the effect of cell connectivity on the accuracy of simulated flows could be important in any MODFLOW application that simulates flow through a heterogeneous hydraulic conductivity field using a grid with large vertical offsets between laterally adjacent cells, whether or not the grid has an explicitly layered structure.
+
+Our idealized benchmark problem serves as a reminder of the assumptions and limitations inherent in the use of vertically offset grids and points to conditions under which full cell connectivity might be of benefit in practical modeling studies.
+When such conditions exist in a MODFLOW 6 model, investigation of the effect of cell connectivity may be warranted.
+In a complex, real-world model, uncertainty in the structure and properties of hydrogeologic units, stresses on the system, and other model parameters could lead to compensation for the effects of inadequate cell connectivity during model calibration.
+However, when unrealistic flows or parameter values are observed in a calibrated model, inadequate cell connectivity can be considered among the possible sources of structural error, and full connectivity can be considered as a remedy.
+
+\section*{Acknowledgments}
 Kerry Bardot is supported by The Australian Research Council, The Western Australian Department of Water and Environmental Regulation and The Water Corporation of Western Australia through grant number LP18101153.
-The authors offer their sincere thanks to the reviewers of this manuscript.
+The authors sincerely thank James Rumbaugh, John Masterson, Kalle Jahn, Jacob LaFontaine, John Brakebill, Heather Welch, Randall Hunt, and Richard Yager for their review comments, which improved the clarity of the manuscript and refined our conclusions regarding the practical implications of this study.
+Any use of trade, firm, or product names is for descriptive purposes only and does not imply endorsement by the U.S.
+Government.
 
-\section{Supporting Information}
-During review, the files that support the creation and running of the test problem are available at {\color{red} XXXXXXXXXX}.
-Once these files are finalized prior to publication, they will be made available at a permanent URL that will be noted here.
+\section*{Supporting Information}
+The files that support the creation and running of the test problem are available at https://doi.org/10.5066/P13BNARA.
 

### 2024-11-19 latex/article_tracked_changes_FINAL.tex
intermediate review drafts & final draft in progress
intermediate review drafts & final draft in progress

--- 2024-06-13 latex/article.tex
@@
 \usepackage{xcolor}
+\usepackage{ulem}
 
 \newcommand{\matr}[1]{\mathbf{#1}}
+\renewcommand{\bibname}{References}
 
-\title{The Importance of Cell Connectivity for Efficient and Accurate Modeling of Flow through Dipping Aquifers with MODFLOW 6}
+\title{{\color{red} \sout{The Importance of Cell Connectivity for Efficient and }}Accurate {\color{red} \sout{Modeling} Simulation} of Flow through Dipping Aquifers with MODFLOW 6 {\color{red} Using Enhanced Cell Connectivity}}
 
 \author{
-	Alden M. Provost, U.S. Geological Survey, Integrated Modeling and Prediction Division, U.S. Geological Survey, 12201 Sunrise Valley Dr, Reston, VA, USA  \\
+	Alden M.
+Provost {\color{red} (corresponding author)}, U.S.
+Geological Survey, Integrated Modeling and Prediction Division, U.S.
+Geological Survey, 12201 Sunrise Valley Dr, Reston, VA, USA{\color{red}; aprovost@usgs.gov} \\
 	\and 
-	Kerry Bardot, School of Earth Sciences, University of Western Australia, Perth, Australia \\
+	Kerry Bardot, School of Earth Sciences, University of Western Australia, Perth, Australia{\color{red}; kerry.bardot@research.uwa.edu.au} \\
 	\and 
-	Christian D. Langevin, U.S. Geological Survey, Integrated Modeling and Prediction Division, 2280 Woodale Drive, Mounds View, MN, USA \\
+	Christian D. Langevin, U.S. Geological Survey, Integrated Modeling and Prediction Division, 2280 Woodale Drive, Mounds View, MN, USA{\color{red}; langevin@usgs.gov} \\
 	\and 
-	James L. McCallum, School of Earth Sciences, University of Western Australia, Perth, Australia \\
+	James L. McCallum, School of Earth Sciences, University of Western Australia, Perth, Australia{\color{red}; james.mccallum@uwa.edu.au} \\
 	}
@@
 
-\maketitle
+\noindent {\color{red} Methods Brief/}
+
+%\maketitle
+{\let\newpage\relax\maketitle}
 
@@
 
-Groundwater flow through a dipping aquifer is often modeled using a single layer of model cells to represent the aquifer.
-The layer of cells is ``deformed'' to follow the top and bottom elevations of the aquifer.
-When this approach is used in MODFLOW, adjacent cells within the model layer are vertically offset from one another, and the standard conductance-based (two-point) formulation for flow between cells does not rigorously account for these offsets.
+{\color{red} In simulations of groundwater flow through dipping aquifers, layers of model cells are often\sout{Groundwater flow through a dipping aquifer is often modeled using a single layer of model cells to represent the aquifer.
+The layer of cells is }}``deformed'' to follow the top and bottom elevations of the aquifer{\color{red}s}.
+When this approach is used in MODFLOW, adjacent cells within the {\color{red} same} model layer are vertically offset from one another, and the standard conductance-based (two-point) formulation for flow between cells does not rigorously account for these offsets.
 The XT3D multi-point flow formulation introduced in MODFLOW 6 is designed to account for geometric irregularities in the grid, including vertical offsets, and to provide accurate results for both isotropic and anisotropic groundwater flow.
@@
 Although XT3D generally improved the accuracy of flow simulations relative to the standard formulation as expected, neither formulation produced accurate flows in cases that involved large vertical offsets.
-In this paper, we explain that the inability of XT3D to produce accurate flows in the steeply dipping aquifer benchmark was not due to an inherent limitation of the flow formulation, but to the limited cell connectivity inherent in the most commonly used discretization packages in MODFLOW 6.
+In this paper, we explain that the inability of XT3D to produce accurate flows in the steeply dipping aquifer benchmark was not due to an inherent limitation of the flow formulation, but {\color{red} rather} to the limited cell connectivity inherent in the most commonly used discretization packages in MODFLOW 6.
 Furthermore, we demonstrate that XT3D is able to produce the expected accuracy when adequate cell connectivity is introduced using MODFLOW's unstructured grid type and the aquifer is discretized vertically using at least two model layers.
@@
 
-\section{Introduction}
+\noindent{\color{red} \textit{\break Note to editor: removed all section numbers}}
+
+\section*{Introduction}
 
@@
 An alternative approach, called the ``boundary-matching" method by \cite{hoaglund2003}, allows model grid layers to "deform" vertically to follow the undulations of hydrogeologic layers.
-This method ``minimize[s] the number of model [grid] layers required to simulate an aquifer system'' \citep{modflow84}.
-In fact, it is common practice to represent each hydrogeologic layer using a single, boundary-matching model grid layer.
-However, the non-orthogonality of the resulting grid can be a source of significant model error \citep{hoaglund2003}.
+This {\color{red} commonly used} method ``minimize[s] the number of model [grid] layers required to simulate an aquifer system'' \citep{modflow84}.
+{\color{red} \sout{In fact, it is common practice to represent each hydrogeologic layer using a single, boundary-matching model grid layer.
+}}However, the non-orthogonality of the resulting grid can be a source of {\color{red} \sout{significant}substantial} model error \citep{hoaglund2003}.
 
-In models based on the Control Volume Finite Difference method, which is the method used in the MODFLOW family of hydrologic simulation codes \citep{modflow6gwf, langevin2024}, the loss of accuracy associated with the boundary-matching approach results from the use of a "two-point" formulation to calculate flows between adjacent grid cells.
+In models based on the Control Volume Finite Difference {\color{red} (CVFD)} method, which is the method used in the MODFLOW family of hydrologic simulation codes \citep{modflow6gwf, langevin2024}, the loss of accuracy associated with the boundary-matching approach results from the use of a "two-point" formulation to calculate flows between adjacent grid cells.
 In the two-point formulation, the flow between two cells is proportional to the difference between the heads calculated in those two cells.
@@
 The simplest 2D cross-sectional (``transect") version of their benchmark model simulated flow through a steeply dipping, homogeneous, isotropic aquifer of uniform thickness embedded between two confining units.
-In test runs that used the boundary-matching method on a vertically offset grid, both the two-point flow formulation and XT3D produced flows that differed markedly from the expected magnitude and direction throughout the aquifer, regardless of how finely the aquifer was discretized.
+In test runs that used the boundary-matching method on a vertically offset grid, both the two-point flow formulation and XT3D produced flows that differed markedly from the expected magnitude and direction throughout the aquifer, regardless of how finely the aquifer was discretized {\color{red} both vertically and horizontally}.
 They noted, however, that a preliminary analysis by two authors of the present work (Provost and Langevin) suggested that the limited cell connectivity offered by the layered, vertically offset grids was inadequate to allow an accurate flow solution.
@@
 The DIS and DISV grid types are based on model layers, with the same horizontal grid structure within each layer.
-The DISU grid type is more flexible and does not depend on the concept of model layers.
+{\color{red} In a DIS or DISV grid, a connection between two cells in different model layers is created automatically if and only if one cell is directly below the other and neither cell has been optionally excluded from the model domain by the user.} The DISU grid type is more flexible and{\color{red}, in MODFLOW 6,} does not {\color{red} \sout{depend on the concept of model layers.}}{\color{red} require the grid to have a layered structure.
+(The DISU grid type in MODFLOW-USG is similarly flexible but retains the concept of layers.) In a DISU grid, the user explicitly specifies all connections between grid cells.
+Thus, when used to represent a grid with a layered structure, as in this work, the DISU grid type allows connections to exist between any two cells, whether they are in the same layer or different layers.
+Grid geometry and connectivity information supplied by the discretization package (DIS, DISV, or DISU) is used by the MODFLOW 6 Node Property Flow (NPF) Package to formulate expressions for groundwater flow between cells in terms of heads calculated at cells centers.
+By default, the NPF Package uses a two-point flow formulation, which we call the ``conductance-based'' formulation.
+The optional, multi-point XT3D flow formulation can be activated simply by including the corresponding keyword in the NPF Package input.}
 
@@
 
-The role of full cell connectivity in enabling accurate simulation of flow along a dipping hydrogeologic layer can be understood by considering Figure \ref{fig:schem-conn-area-flux}b.
+{\color{red} \sout{The role of full cell connectivity in enabling accurate simulation of flow along a dipping hydrogeologic layer can be understood by considering}The need for full cell connectivity to accurately simulate flow along the dipping hydrogeologic layer is illustrated in} Figure \ref{fig:schem-conn-area-flux}b.
 Components of the groundwater flux (specific discharge) vector are displayed on each cell-cell interface and are representative of steady, uniform flow along the hydrogeologic layer.
@@
 The blue vectors in Figure \ref{fig:schem-conn-area-flux}b show that to simulate steady, uniform flow along the dipping hydrogeologic layer, cells must exchange water not only ``horizontally'' with neighboring cells in the same grid layer, but also vertically with cells in the adjacent grid layer.
-Specifically, there must be flow from cells in the bottom grid layer to cells in the top grid layer, which corresponds to the vertical component of the groundwater flux.
+Specifically, there must be flow from cells in the {\color{red} \sout{bottom}lower} grid layer to cells in the {\color{red} \sout{top}upper} grid layer, which corresponds to the vertical component of the groundwater flux.
 Note, however, that the blue vectors alone are incompatible with steady, uniform flow along the hydrogeologic layer because the exchange of water between the two grid layers is in one direction only: upward.
-Such a steady-state flow configuration would imply continuous depletion of flow within the bottom grid layer and accumulation of flow within the top grid layer as one moves along the hydrogeologic layer in the direction of flow.
-Thus, steady, uniform flow along the hydrogeologic layer cannot be simulated accurately without some mechanism for returning flow from the top grid layer to the bottom grid layer.
+Such a steady-state flow configuration would imply continuous depletion of flow within the {\color{red} \sout{bottom}lower} grid layer and accumulation of flow within the {\color{red} \sout{top}upper} grid layer as one moves along the hydrogeologic layer in the direction of flow.
+Thus, {\color{red} the} steady, uniform flow {\color{red} \sout{along the hydrogeologic layer}represented by the gray vectors} cannot be simulated accurately without some mechanism for returning flow from the {\color{red} \sout{top}upper} grid layer to the {\color{red} \sout{bottom}lower} grid layer.
 The full-connectivity grid provides such a mechanism by offering additional connections between cells in adjacent grid layers.
-Vertical flows from the bottom grid layer to the top grid layer, represented by vertical blue vectors in Figure \ref{fig:schem-conn-area-flux}b, can be returned to the bottom grid layer by nominally ``horizontal'' flows, represented by red vectors in Figure \ref{fig:schem-conn-area-flux}b.
+Vertical flows from the {\color{red} \sout{bottom}lower} grid layer to the {\color{red} \sout{top}upper} grid layer, represented by vertical blue vectors in Figure \ref{fig:schem-conn-area-flux}b, can be returned to the {\color{red} \sout{bottom}lower} grid layer by nominally ``horizontal'' flows, represented by red vectors in Figure \ref{fig:schem-conn-area-flux}b.
 Although this concept has been illustrated using a two-grid-layer model for simplicity, the same reasoning applies given any number of grid layers.
@@
 
-\section{Results and Discussion}
+\section*{Results and Discussion}
 
-To demonstrate the effect of cell connectivity on the accuracy of simulated flow through a steeply dipping aquifer, results for the base case described in the previous section are compared for simulations that use either layered or full connectivity, and either the two-point flow formulation (also known as the "standard conductance-based formulation" in MODFLOW 6) or the XT3D flow formulation.
-The base case is then modified to further investigate the effects of grid resolution, aquifer dip angle, conductivity contrast between the aquifer and the surrounding domain, and anisotropic hydraulic conductivity.
+To demonstrate the effect of cell connectivity on the accuracy of simulated flow through a steeply dipping aquifer, results for the base case described in the previous section are compared for simulations that use either layered or full connectivity, and either the {\color{red} \sout{two-point flow formulation (also known as the "standard conductance-based formulation" in MODFLOW 6)}standard conductance-based (two-point) flow formulation} or the XT3D {\color{red}(multi-point)} flow formulation.
+The base case is then modified to {\color{red} \sout{further}} investigate the effects of grid resolution, aquifer dip angle, conductivity contrast between the aquifer and the surrounding domain, and anisotropic hydraulic conductivity.
 
@@
 Likewise, specific discharge in cells within the surrounding domain is calculated using only face flows between cells within the domain.
-Specific discharge values reported in the binary budget output file by MODFLOW 6 are different than the values reported here for cells that are adjacent to the boundary between the aquifer and the domain.
-This is because MODFLOW interpolates specific discharge in a cell using all of the face flows for that cell, including flows across cell faces on the aquifer-domain boundary.
-Those flows represent an ``average'' of the different flows on either side of the aquifer-domain boundary.
+{\color{red} These s\sout{S}}pecific discharge values {\color{red} are different from those} reported in the binary budget output file by MODFLOW 6 {\color{red} \sout{are different than the values reported here for cells that are adjacent to the boundary between the aquifer and the domain.
+This is }}because MODFLOW interpolates specific discharge in a cell using all of the face flows for that cell, including flows across cell faces on the aquifer-domain boundary.
+{\color{red} \sout{Those flows}The flows reported by MODFLOW} represent an ``average'' of the different flows on either side of the aquifer-domain boundary.
 Therefore, specific discharge calculated in a cell adjacent to the aquifer-domain boundary is affected by the different flow rate on the other side of the boundary.
@@
 
-\subsection{Effects of Cell Connectivity and Flow Formulation}
+\subsection*{Effects of Cell Connectivity and Flow Formulation}
 
-The numerical solution for the layered-connectivity grid confirms the findings of \cite{bardot2023} in that fluxes are predominantly horizontal (Figure \ref{fig:fig2}a).
+% Bardot et al. (2023) changed from live ref to text to allow \sout
+{\color{red} \sout{The numerical solution for the layered-connectivity grid confirms the findings of Bardot et al.
+(2023) in that fluxes are predominantly horizontal (Figure \ref{fig:fig2}a).
 We see that despite an enforced hydraulic gradient of $30^{\circ}$, the vertical component of the flux cannot propagate through the aquifer given that cells along the aquifer top are hydraulically disconnected from the low permeability domain, hence forcing flow horizontally (red arrows on right panel).
@@
 Results for the full-connectivity grid with the standard conductance-based formulation shows an improvement in the flow solution (Figure \ref{fig:fig2}c), with the XT3D formulation successfully reproducing the analytical solution (Figure \ref{fig:fig2}d).
-We see that the grid with full connectivity allows incorporation of the vertical flow component, thus permitting cross-connection between grid layers.
+We see that the grid with full connectivity allows incorporation of the vertical flow component, thus permitting cross-connection between grid layers.}}
 
 The results shown in Figures \ref{fig:fig2}a and b, which are based on a vertically offset grid with layered connectivity, are consistent with the findings of \cite{bardot2023}.
-Despite boundary heads designed to induce specific discharge of 1 m/d along the $30^{\circ}$ incline of the aquifer, the simulated flow within the aquifer is predominantly horizontal, and the magnitude of the specific discharge at the center of the aquifer and the total flow entering the aquifer are both overestimated.
+Despite boundary heads designed to induce specific discharge of 1 m/d along the $30^{\circ}$ incline of the aquifer, the simulated flow within the aquifer is predominantly horizontal, and the magnitude of the specific discharge at the center of the aquifer {\color{red} (``flux magnitude'')} and the total flow entering the aquifer are {\color{red} \sout{both}} overestimated {\color{red}by approximately 15\% and 18\%, respectively}.
 The XT3D flow formulation is unable to compensate for the inadequate cell connectivity (Figure \ref{fig:fig2}b), which does not allow flow between laterally adjacent cells in different grid layers.
 
-The vertically offset grid with full connectivity (Figures \ref{fig:fig2}c and d) shows significantly improved results, even for the standard flow formulation (Figure \ref{fig:fig2}c).
-Using the XT3D flow formulation provides the most accurate results, as expected (Figure \ref{fig:fig2}d).
+The vertically offset grid with full connectivity (Figures \ref{fig:fig2}c and d) shows {\color{red} \sout{significantly}substantially} improved results, even for the standard flow formulation (Figure \ref{fig:fig2}c){\color{red}, which overestimates the flux magnitude and total flow by approximately 3\% and 9\%, respectively.} Using the XT3D flow formulation provides the most accurate results, as expected (Figure \ref{fig:fig2}d){\color{red} \sout{.}, reducing the errors in flux magnitude and total flow to approximately $-2 \times 10^{-4}$\% and $2 \times 10^{-4}$\%, respectively.}
 
@@
 The middle panel shows the face flows (arrows).
-The right panel shows head contours for the analytical solution (black) as well as the numerical solution (blue).}
+The right panel shows head contours for the analytical solution (black) as well as the numerical solution (blue).
+{\color{red} Flux magnitudes, flux directions (angles), and head errors are reported to two decimal places.
+Percent errors in flow are reported to one decimal place.}}
 	\label{fig:fig2}
@@
 	\caption{Graphs indicating how the number of grid layers within the aquifer affects the flux magnitude in the center cell (top), the flux direction in center cell (middle) and the volumetric flow through the aquifer (bottom).
-Errors are minimized by using full connectivity (right column), the XT3D flow formulation (orange lines) and at least two grids layers in the aquifer.}
+Errors are minimized by using full connectivity (right column), the XT3D flow formulation (orange lines) and at least two grids layers in the aquifer.
+{\color{red} \textit{Note to editor and reviewers: "channel" replaced with "aquifer" in axis labels.}}}
 	\label{fig:fig3}
@@
 
-Layered-connectivity grids, which are commonly used in groundwater models \citep{Reilly2004}, are often configured using one model grid layer per hydrogeologic unit when simulating groundwater flow.
-The simulations in this subsection investigate the effect of vertically discretizing the aquifer using multiple grid layers (Figure \ref{fig:fig3}).
+% Reilly and Harbaugh, 2004 changed from live link to text to enable \sout
+{\color{red} \sout{Layered-connectivity grids, which are commonly used in groundwater models (Reilly and Harbaugh, 2004), are often configured using one model grid layer per hydrogeologic unit when simulating groundwater flow.}
+Groundwater models often use layered-connectivity grids \citep{Reilly2004}.
+Although the practice is less common today than it once was, models are sometimes configured using one grid layer per hydrogeologic unit.} The simulations in this subsection investigate the effect of vertically discretizing the aquifer using multiple grid layers (Figure \ref{fig:fig3}).
 The base case uses three grid layers to represent the aquifer.
 When the number of grid layers is altered, the horizontal discretization is also adjusted to maintain square cells inside the aquifer, thus, maintaining a constant cell aspect ratio.
-We compare modeled flux magnitude and direction in cell at the center of the aquifer, as well as volumetric flow through the aquifer, against the analytical solution.
+We compare {\color{red} \sout{modeled}the simulated} flux magnitude and direction in {\color{red} the} cell at the center of the aquifer, as well as volumetric flow through the aquifer, against the analytical solution.
 
-Like the results in Figure \ref{fig:fig2}, the results in Figure \ref{fig:fig3} show that the use of full connectivity and the XT3D flow formulation reduces the errors in calculated flows, provided at least two grid layers are used to vertically discretize the aquifer.
-When a single grid layer is used to represent the aquifer, the enhanced connectivity needed to achieve an accurate solution is not possible, and the XT3D flow formulation is unable to compensate for the inadequate connectivity.
+{\color{red} \sout{Like the results in Figure \ref{fig:fig2}, t}T}he results in Figure \ref{fig:fig3} {\color{red} \sout{show}confirm} that the use of full connectivity and the XT3D flow formulation reduces the errors in calculated flows, provided at least two grid layers are used to vertically discretize the aquifer.
+{\color{red}For example, when the aquifer is represented using two layers with layered connectivity, the flow error is approximately 18\% and 5\% for the standard and XT3D flow formulations, respectively.
+Use of full connectivity reduces these errors to approximately 11\% and $3 \times 10^{-4}$\%, respectively.} When a single grid layer is used to represent the aquifer, the enhanced connectivity needed to achieve an accurate solution is not possible, and the XT3D flow formulation is unable to compensate for the inadequate connectivity.
+{\color{red} With full connectivity and the XT3D flow formulation, using more than two layers to discretize the aquifer provides no substantial benefit in our benchmark problem (Figure \ref{fig:fig3}b).}
 
-\subsection{Effects of Dip Angle and Conductivity Contrast}
+\subsection*{Effects of Dip Angle and Conductivity Contrast}
 
@@
 With layered connectivity and the standard conductance-based flow formulation (Figure \ref{fig:fig4}a), the magnitude of the error in the flux magnitude (left panel) is similar for all of the conductivity ratios simulated and increases with increasing aquifer dip angle.
-The magnitude of the error in the flux direction (right panel) increases with decreasing conductivity ratio.
-As the surrounding domain becomes much less conductive than the aquifer, the flow direction tends toward horizontal.
+The magnitude of the error in the flux direction (right panel) increases with {\color{red} \sout{decreasing conductivity ratio}decreasing ratio of domain conductivity to aquifer conductivity (increasing conductivity contrast)}.
+As the surrounding domain becomes much less conductive than the aquifer, the flow direction tends toward horizontal{\color{red}, as shown by the ``flattening'' of the curve toward a flux direction of zero degrees as the conductivity ratio decreases from 1.0 to 0.01 in Figure \ref{fig:fig4}a}.
 
@@
 The magnitude of the error in the flux direction (right panel) increases with decreasing conductivity ratio.
-As the surrounding domain becomes much less conductive than the aquifer, the flow direction tends toward horizontal, as with the standard flow formulation.
+As the surrounding domain becomes much less conductive than the aquifer, the flow direction tends toward horizontal {\color{red} in Figure \ref{fig:fig4}b}, as {\color{red} occurs} with the standard flow formulation {\color{red} in Figure \ref{fig:fig4}a}.
 
-With full connectivity and the standard conductance-based flow formulation (Figure \ref{fig:fig4}c), the magnitude of the error in the flux magnitude is less than 17\% of the analytical solution of 1 m/d, and the magnitude of the error in the flux direction is less than 10$^{\circ}$, at dip angles less than 40$^{\circ}$ for all of the conductivity ratios simulated.
-Given that the cells within the aquifer are square, cells that are nominally in the same grid layer within the aquifer no longer share any interfacial area, i.e., the become hydraulically disconnected, at a dip angle of 45$^{\circ}$.
+With full connectivity and the standard conductance-based flow formulation (Figure \ref{fig:fig4}c), the {\color{red}\sout{magnitude of the}maximum} error in the flux magnitude is less than 17\% of the analytical solution of 1 m/d{\color{red} \sout{,}} and the {\color{red}\sout{magnitude of the}maximum} error in the flux direction is less than 10$^{\circ}${\color{red} \sout{,}} at dip angles less than 40$^{\circ}$ for all of the conductivity ratios simulated.
+{\color{red} \sout{Given that the cells within the aquifer are square}At a dip angle of  45$^{\circ}$}, cells that are nominally in the same grid layer within the aquifer no longer share any interfacial area, {\color{red} \sout{i.e., they become}thus becoming} hydraulically disconnected from one another{\color{red} \sout{, at a dip angle of 45$^{\circ}$}}.
 A second transition in connectivity occurs at a dip angle of approximately 63$^{\circ}$, and at a dip angle of approximately 72$^{\circ}$ the vertical columns of cells within the aquifer are completely disconnected from their neighbors to either side.
 When the discretization used to represent the aquifer becomes substantially disconnected, it cannot be expected to support accurate results.
-Note that for coarser horizontal discretization, transitions in connectivity within the aquifer would occur at smaller dip angles.
+Note that {\color{red}in our benchmark problem the cells within the aquifer are square.
+\sout{for coarser horizontal discretization}For the same vertical discretization but coarser horizontal discretization}, transitions in connectivity within the aquifer would occur at smaller dip angles {\color{red} due to the smaller vertical-to-horizontal aspect ratio of the cells}.
 
-With full connectivity and the XT3D flow formulation (Figure \ref{fig:fig4}d), the magnitude of the error in the flux magnitude is less than 3\% of the analytical solution of 1 m/d, and the magnitude of the error in the flux direction is less than 1$^{\circ}$, at dip angles less than 40$^{\circ}$ for all of the conductivity ratios simulated.
-The effects of hydraulic disconnection of cells within the aquifer are evident at approximately 63$^{\circ}$ for the magnitude of the error in flow magnitude and at 45$^{\circ}$  for the magnitude of the error in flow direction.
+With full connectivity and the XT3D flow formulation (Figure \ref{fig:fig4}d), the {\color{red}\sout{magnitude of the}maximum} error in the flux magnitude is less than 3\% of the analytical solution of 1 m/d{\color{red} \sout{,}} and the {\color{red}\sout{magnitude of the}maximum} error in the flux direction is less than 1$^{\circ}${\color{red} \sout{,}} at dip angles less than 40$^{\circ}$ for all of the conductivity ratios simulated.
+The effects of hydraulic disconnection of cells {\color{red} \sout{within the aquifer}on errors in flow magnitude and flow direction} are evident at approximately 63$^{\circ}$ {\color{red} \sout{for the magnitude of the error in flow magnitude and at}and} 45$^{\circ}${\color{red} \sout{ for the magnitude of the error in flow direction}, respectively}.
 Overall, use of full connectivity with the XT3D flow formulation substantially reduces the error in calculated flows.
@@
 
-\subsection{Effect of Anisotropy}
+\subsection*{Effect of Anisotropy}
 
-The base case and variations up to this point have used isotropic hydraulic conductivity.
+The base case and variations {\color{red} on the base case} up to this point have used isotropic hydraulic conductivity.
 However, sedimentary layers often exhibit anisotropic conductivity.
-Therefore, we examine the effect of anisotropic and dipping conductivity tensors within the aquifer and domain, with conductivity perpendicular to the aquifer reduced by factors of 10, 100, 1000 and 10,000.
-We also include the isotropic scenario (anisotropy ratio of 1).
+{\color{red} \sout{Therefore, we examine the effect of anisotropic and dipping conductivity tensors within the aquifer and domain, with conductivity perpendicular to the aquifer reduced by factors of 10, 100, 1000 and 10,000.
+We also include the isotropic scenario (anisotropy ratio of 1).}Therefore, we examine the effect of anisotropy within the aquifer and domain by dividing the original conductivity perpendicular to the dip of the aquifer by factors of 1, 10, 100, 1000 and 10,000 to give anisotropy ratios ranging from 1.0 (isotropic) to 0.0001.} {\color{red} Orientation of the principal directions of anisotropy parallel and perpendicular to the dip of the aquifer is accomplished by setting three angles in the NPF Package input.
+The first angle is the rotation within the horizontal plane of the first principal direction (typically the direction of maximum hydraulic conductivity) with respect to the positive $x$ axis.
+The second angle is the rotation of the first principal direction upward from the horizontal plane.
+The third angle is a rotation of the second and third principal directions about the axis of the first principal direction.
+For this two-dimensional, cross-sectional problem, the first and third angles are set to zero because they represent three-dimensional effects, and the second angle is set equal to the dip angle ($30^{\circ}$).}
+
 Results are presented in Figure \ref{fig:fig5}.
-Layered-connectivity grids produce results that deviate significantly from the analytical solution both with and without the XT3D formulation (Figure \ref{fig:fig5}a), while full-connectivity grids with XT3D reproduce the analytical solution (Figure \ref{fig:fig5}b, orange lines).
+Layered-connectivity grids produce results that deviate {\color{red} \sout{significantly}substantially} from the analytical solution both with and without the XT3D formulation (Figure \ref{fig:fig5}a), while full-connectivity grids with XT3D reproduce the analytical solution (Figure \ref{fig:fig5}b, orange lines).
+{\color{red} For example, at an anisotropy ratio of 0.01, use of layered connnectivity produces flow errors of approximately -96\% for both the standard and XT3D flow formulations.
+Use of full connectivity produces errors of approximately 30\% and -3\% for the standard and XT3D flow formulations, respectively.} {\color{red} These results suggest that use of full grid connectivity and the XT3D flow formulation in MODFLOW 6 simulations can be particularly important in the presence of anisotropy.}
 
@@
 
-\section{Conclusions}
+\section*{Conclusions}
 
 \cite{bardot2023} revisited the capability of MODFLOW 6 to accurately simulate groundwater flow through sedimentary structures.
-They observed that calculated flows involved significant error in a benchmark problem that simulated flow through a steeply dipping hydrogeologic layer using vertically offset grids, whether or not the XT3D multi-point flow formulation was used.
+They observed that calculated flows involved {\color{red} \sout{significant}substantial} error in a benchmark problem that simulated flow through a steeply dipping hydrogeologic layer using vertically offset grids, whether or not the XT3D multi-point flow formulation was used.
 They hypothesized that inadequate lateral connectivity between model cells was responsible for the inability to obtain accurate flow results in their benchmark problem.
@@
 In this paper, we have confirmed that the layered lateral connectivity offered by DIS and DISV model grids can be inadequate for simulating groundwater flow through steeply dipping sedimentary structures, even when the XT3D multi-point flow formulation is used.
-However, when connections between laterally adjacent cells in different grid layers are added to comprise a grid with full connectivity, there can be a substantial increase in the accuracy of simulated flows.
-The additional connectivity is implemented using the fully unstructured (DISU) grid type in MODFLOW 6.
+However, when connections between laterally adjacent cells in different grid layers are added {\color{red}using the fully unstructured (DISU) grid type} to comprise a grid with full connectivity, {\color{red} \sout{there can be a substantial increase in }}the accuracy of simulated flows {\color{red}can increase substantially.\sout{  The additional connectivity is implemented using the fully unstructured (DISU) grid type in MODFLOW 6}}.
 This accuracy improvement requires that at least two grid layers be used to represent the dipping aquifers, with additional grid layers required for very steep dips.
 
-Our benchmark results suggest that, given appropriate discretization and cell connectivity, the XT3D multi-point flow formulation implemented in MODFLOW 6 can be combined with vertically offset grids to efficiently and accurately model flows through steeply dipping sedimentary structures.
-Future work may investigate the accuracy of calculated flows under different hydraulic gradients, as well as the implications of grid connectivity for more complex and realistic flow systems.
+{\color{red} Note that vertical offsets could be avoided when modeling a 2D cross-section using MODFLOW 6 by formulating the model in plan view, such that rows of neighboring cells play the role of model layers.
+This would allow a DISV grid to follow the top and bottom surfaces of a sloping aquifer in a piecewise-linear fashion without vertical offsets.
+For a 3D model, an alternative could be to use a simulator based on the finite element method \citep[see, e.g.,][]{yager2009anisotropy}.
+In evaluating any alternative approach, the ability to accurately simulate flows on the resulting grid, possibly in the presence of anisotropy that is not aligned with the grid, remains one of the considerations.}
 
-The importance of adequate grid connectivity has been explained and evaluated in this work specifically in terms of hydrogeologic layers that correspond to sedimentary structures.
-However, the impact of cell connectivity on the accuracy of simulated flows could be important in any MODFLOW application that models flow through a heterogeneous hydraulic conductivity field using a grid with large vertical offsets between laterally adjacent cells, whether or not the grid has an explicitly layered structure.
+Our benchmark results suggest that, given appropriate discretization and cell connectivity, the XT3D multi-point flow formulation implemented in MODFLOW 6 can be combined with vertically offset grids to efficiently and accurately {\color{red} \sout{model}simulate} flows through steeply dipping sedimentary structures.
+{\color{red} Our benchmark problem, which is patterned after that of \cite{bardot2023}, features a two-dimensional cross-section with flow along a dipping aquifer of uniform thickness that is discretized using square model cells.} {\color{red} \sout{Future work may investigate}Additional investigation is needed to assess} the accuracy of calculated flows under different hydraulic gradients {\color{red} and more typical cell aspect ratios}, {\color{red} \sout{as well as}and to better understand} the implications of grid connectivity for more complex and realistic flow systems.
 
-\section{Acknowledgments}
+The importance of adequate grid connectivity has been explained and {\color{red} \sout{evaluated}demonstrated} in this work specifically in terms of hydrogeologic layers that correspond to sedimentary structures.
+However, the {\color{red} \sout{impact}effect} of cell connectivity on the accuracy of simulated flows could be important in any MODFLOW application that {\color{red} \sout{models}simulates} flow through a heterogeneous hydraulic conductivity field using a grid with large vertical offsets between laterally adjacent cells, whether or not the grid has an explicitly layered structure.
+
+{\color{red} Our idealized benchmark problem serves as a reminder of the assumptions and limitations inherent in the use of vertically offset grids and points to conditions under which full cell connectivity might be of benefit in practical modeling studies.
+When such conditions exist in a MODFLOW 6 model, investigation of the effect of cell connectivity may be warranted.
+In a complex, real-world model, uncertainty in the structure and properties of hydrogeologic units, stresses on the system, and other model parameters could lead to compensation for the effects of inadequate cell connectivity during model calibration.
+However, when unrealistic flows or parameters values are observed in a calibrated model, inadequate cell connectivity can be considered among the possible sources of structural error, and full connectivity can be considered as remedy.}
+
+\section*{Acknowledgments}
 Kerry Bardot is supported by The Australian Research Council, The Western Australian Department of Water and Environmental Regulation and The Water Corporation of Western Australia through grant number LP18101153.
-The authors offer their sincere thanks to the reviewers of this manuscript.
+{\color{red} \sout{The authors offer their sincere thanks to the reviewers of this manuscript.
+}The authors sincerely thank James Rumbaugh, John Masterson, Kalle Jahn, Jacob LaFontaine, Randall Hunt, and Richard Yager for their review comments, which improved the clarity of the manuscript and refined our conclusions regarding the practical implications of this study.}
 
-\section{Supporting Information}
-During review, the files that support the creation and running of the test problem are available at {\color{red} XXXXXXXXXX}.
-Once these files are finalized prior to publication, they will be made available at a permanent URL that will be noted here.
+\section*{Supporting Information}
+{\color{red} \sout{During review, the files that support the creation and running of the test problem are available at https://github.com/langevin-usgs/xt3d-vertically-staggered/releases/download/v1.0/suppinfo.zip.
+Once these files are finalized prior to publication, they will be made available at a permanent URL that will be noted here.}}{\color{red} The files that support the creation and running of the test problem are available at https://doi.org/10.5066/P13BNARA.}
+
+\noindent{\color{red} \textit{\break Note to editor: changed heading "Bibliography" to "References".}}
+
+\noindent{\color{red} \textit{Note to editor: implemented Harvard style for references.}}
 

### 2024-11-19 latex/saved drafts/article.tex
intermediate review drafts & final draft in progress
intermediate review drafts & final draft in progress

--- 2024-06-13 latex/article.tex
@@
 With full connectivity and the standard conductance-based flow formulation (Figure \ref{fig:fig4}c), the magnitude of the error in the flux magnitude is less than 17\% of the analytical solution of 1 m/d, and the magnitude of the error in the flux direction is less than 10$^{\circ}$, at dip angles less than 40$^{\circ}$ for all of the conductivity ratios simulated.
-Given that the cells within the aquifer are square, cells that are nominally in the same grid layer within the aquifer no longer share any interfacial area, i.e., the become hydraulically disconnected, at a dip angle of 45$^{\circ}$.
+Given that the cells within the aquifer are square, cells that are nominally in the same grid layer within the aquifer no longer share any interfacial area, i.e., they become hydraulically disconnected, at a dip angle of 45$^{\circ}$.
 A second transition in connectivity occurs at a dip angle of approximately 63$^{\circ}$, and at a dip angle of approximately 72$^{\circ}$ the vertical columns of cells within the aquifer are completely disconnected from their neighbors to either side.
@@
 \section{Supporting Information}
-During review, the files that support the creation and running of the test problem are available at {\color{red} XXXXXXXXXX}.
+During review, the files that support the creation and running of the test problem are available at https://github.com/langevin-usgs/xt3d-vertically-staggered/releases/download/v1.0/suppinfo.zip.
 Once these files are finalized prior to publication, they will be made available at a permanent URL that will be noted here.

### 2024-11-19 latex/saved drafts/article_clean.tex
intermediate review drafts & final draft in progress
intermediate review drafts & final draft in progress

--- 2024-11-19 latex/article_clean_FINAL.tex
@@
 When this approach is used in MODFLOW, adjacent cells within the same model layer are vertically offset from one another, and the standard conductance-based (two-point) formulation for flow between cells does not rigorously account for these offsets.
-The XT3D multi-point flow formulation in MODFLOW 6 is designed to account for geometric irregularities in the grid, including vertical offsets, and to provide accurate results for both isotropic and anisotropic groundwater flow.
+The XT3D multi-point flow formulation introduced in MODFLOW 6 is designed to account for geometric irregularities in the grid, including vertical offsets, and to provide accurate results for both isotropic and anisotropic groundwater flow.
 A recent study evaluated the performance of the standard formulation and XT3D using a simple, synthetic benchmark model of a steeply dipping aquifer.
@@
 A traditional approach to discretizing layered aquifer systems, called the ``grid overlay'' method by \cite{hoaglund2003}, maps the heterogeneous hydrologic properties onto a rectilinear grid \citep{modflow84}.
-Although this method preserves the benefits of the rectilinear grid structure, it can require a large number of model grid layers to adequately discretize hydrogeologic layers of variable dip and thickness \citep{Zyvoloski2006}.
+While this method preserves the benefits of the rectilinear grid structure, it can require a large number of model grid layers to adequately discretize hydrogeogic layers of variable dip and thickness \citep{Zyvoloski2006}.
 
@@
 This commonly used method ``minimize[s] the number of model [grid] layers required to simulate an aquifer system'' \citep{modflow84}.
-However, the non-orthogonality of the resulting grid can be a source of substantial model error \citep{hoaglund2003}.
+However, the non-orthogonality of the resulting grid can be a source of significant model error \citep{hoaglund2003}.
 
@@
 To help overcome the limitations of the two-point flow formulation on geometrically irregular grids, MODFLOW-USG and MODFLOW 6 offer a Ghost-Node Correction (GNC) Package \citep{modflowusg, modflow6gwf}.
-However, configuration of optimal ghost node locations and interpolation weights is problem-dependent, and we are not aware of guidelines for handling general unstructured grids or arbitrarily oriented 3D anisotropy.
+However, configuration of optimal ghost node locations and interpolation weights is highly problem-dependent, and we are not aware of guidelines for handling general unstructured grids or arbitrarily oriented 3D anisotropy.
 
-The optional XT3D flow formulation \citep{modflow6xt3d} in MODFLOW 6 offers a more ``automatic" alternative to ghost nodes.
+The optional XT3D flow formulation \citep{modflow6xt3d} introduced in MODFLOW 6 offers a more ``automatic" alternative to ghost nodes.
 XT3D is a multi-point flow formulation \citep{edwards1998, aavatsmark2002} based on gradient reconstruction \citep{mavriplis2003leastsq, diskin2008accuracy}.
@@
 Thus, when used to represent a grid with a layered structure, as in this work, the DISU grid type allows connections to exist between any two cells, whether they are in the same layer or different layers.
-Grid geometry and connectivity information supplied by the discretization package (DIS, DISV, or DISU) is used by the MODFLOW 6 Node Property Flow (NPF) Package to formulate expressions for groundwater flow between cells in terms of heads calculated at cell centers.
+Grid geometry and connectivity information supplied by the discretization package (DIS, DISV, or DISU) is used by the MODFLOW 6 Node Property Flow (NPF) Package to formulate expressions for groundwater flow between cells in terms of heads calculated at cells centers.
 By default, the NPF Package uses a two-point flow formulation, which we call the ``conductance-based'' formulation.
@@
 
-Vertically offset grids violate the CFVD requirement because nominally ``horizontal'' connections between cell centers in the same grid layer are not strictly horizontal and therefore not perpendicular to the vertical cell-cell interfaces they intersect.
+Note how the vertically offset grid violates the CFVD requirement because nominally ``horizontal'' connections between cell centers in the same grid layer are not strictly horizontal and therefore not perpendicular to the vertical cell-cell interfaces they intersect.
 For example, in Figure \ref{fig:schem-conn-area-flux}a, the connection between cells 1A and 1B is not perpendicular to the cell-cell interface, which reduces the accuracy of the conductance-based formulation.
@@
 The test simulations use grids of MODFLOW 6 DISU (unstructured) type, which offers the flexibility to implement either layered or full connectivity.
-In all cases, a grid of MODFLOW 6 DIS (structured) type is first created and then converted to a DISU grid.
+In all cases, a grid of MODFLOW 6 DIS (structured) type is first created and then converted a DISU grid.
 In simulations that use layered connectivity, the DISU grid is formulated to have the same cell geometry and connectivity as the original DIS grid.
@@
 Therefore, specific discharge calculated in a cell adjacent to the aquifer-domain boundary is affected by the different flow rate on the other side of the boundary.
-Although this is an unavoidable consequence of vertically offset discretization in practical applications, and not necessarily a problem, users should be aware of these effects when interpreting the results of MODFLOW 6 dispersive-transport and MODPATH particle-tracking simulations.
+While this is an unavoidable consequence of vertically offset discretization in practical applications, and not necessarily a problem, users should be aware of these effects when interpreting the results of MODFLOW 6 dispersive-transport and MODPATH particle-tracking simulations.
 
@@
 
-The vertically offset grid with full connectivity (Figures \ref{fig:fig2}c and d) shows substantially improved results, even for the standard flow formulation (Figure \ref{fig:fig2}c), which overestimates the flux magnitude and total flow by approximately 3\% and 9\%, respectively.
+The vertically offset grid with full connectivity (Figures \ref{fig:fig2}c and d) shows significantly improved results, even for the standard flow formulation (Figure \ref{fig:fig2}c), which overestimates the flux magnitude and total flow by approximately 3\% and 9\%, respectively.
 Using the XT3D flow formulation provides the most accurate results, as expected (Figure \ref{fig:fig2}d), reducing the errors in flux magnitude and total flow to approximately $-2 \times 10^{-4}$\% and $2 \times 10^{-4}$\%, respectively.
@@
 When a single grid layer is used to represent the aquifer, the enhanced connectivity needed to achieve an accurate solution is not possible, and the XT3D flow formulation is unable to compensate for the inadequate connectivity.
-With full connectivity and the XT3D flow formulation, using more than two layers to discretize the aquifer provides no substantial benefit in our benchmark problem (Figure \ref{fig:fig3}b).
+With full connectivity and the XT3D flow formulation, using more than two layers to discretize the aquifer provides no significant benefit in our benchmark problem (Figure \ref{fig:fig3}b).
 
@@
 However, sedimentary layers often exhibit anisotropic conductivity.
-Therefore, we examine the effect of anisotropy within the aquifer and domain by dividing the original conductivity perpendicular to the dip of the aquifer by factors of 1, 10, 100, 1,000 and 10,000 to give anisotropy ratios ranging from 1.0 (isotropic) to 0.0001.
+Therefore, we examine the effect of anisotropy within the aquifer and domain by dividing the original conductivity perpendicular to the dip of the aquifer by factors of 1, 10, 100, 1000 and 10,000 to give anisotropy ratios ranging from 1.0 (isotropic) to 0.0001.
 Orientation of the principal directions of anisotropy parallel and perpendicular to the dip of the aquifer is accomplished by setting three angles in the NPF Package input.
@@
 
-Layered-connectivity grids produce results that deviate substantially from the analytical solution both with and without the XT3D formulation (Figure \ref{fig:fig5}a), whereas full-connectivity grids with XT3D reproduce the analytical solution (Figure \ref{fig:fig5}b, orange lines).
+Results are presented in Figure \ref{fig:fig5}.
+Layered-connectivity grids produce results that deviate significantly from the analytical solution both with and without the XT3D formulation (Figure \ref{fig:fig5}a), while full-connectivity grids with XT3D reproduce the analytical solution (Figure \ref{fig:fig5}b, orange lines).
 For example, at an anisotropy ratio of 0.01, use of layered connnectivity produces flow errors of approximately -96\% for both the standard and XT3D flow formulations.
@@
 \cite{bardot2023} revisited the capability of MODFLOW 6 to accurately simulate groundwater flow through sedimentary structures.
-They observed that calculated flows involved substantial error in a benchmark problem that simulated flow through a steeply dipping hydrogeologic layer using vertically offset grids, whether or not the XT3D multi-point flow formulation was used.
+They observed that calculated flows involved significant error in a benchmark problem that simulated flow through a steeply dipping hydrogeologic layer using vertically offset grids, whether or not the XT3D multi-point flow formulation was used.
 They hypothesized that inadequate lateral connectivity between model cells was responsible for the inability to obtain accurate flow results in their benchmark problem.
@@
 In a complex, real-world model, uncertainty in the structure and properties of hydrogeologic units, stresses on the system, and other model parameters could lead to compensation for the effects of inadequate cell connectivity during model calibration.
-However, when unrealistic flows or parameter values are observed in a calibrated model, inadequate cell connectivity can be considered among the possible sources of structural error, and full connectivity can be considered as a remedy.
+However, when unrealistic flows or parameters values are observed in a calibrated model, inadequate cell connectivity can be considered among the possible sources of structural error, and full connectivity can be considered as remedy.
 
@@
 Kerry Bardot is supported by The Australian Research Council, The Western Australian Department of Water and Environmental Regulation and The Water Corporation of Western Australia through grant number LP18101153.
-The authors sincerely thank James Rumbaugh, John Masterson, Kalle Jahn, Jacob LaFontaine, John Brakebill, Heather Welch, Randall Hunt, and Richard Yager for their review comments, which improved the clarity of the manuscript and refined our conclusions regarding the practical implications of this study.
-Any use of trade, firm, or product names is for descriptive purposes only and does not imply endorsement by the U.S.
-Government.
+The authors sincerely thank James Rumbaugh, John Masterson, Kalle Jahn, Jacob LaFontaine, Randall Hunt, and Richard Yager for their review comments, which improved the clarity of the manuscript and refined our conclusions regarding the practical implications of this study.
 

### 2024-11-19 latex/saved drafts/article_clean_v2.tex
intermediate review drafts & final draft in progress
intermediate review drafts & final draft in progress

--- 2024-11-19 latex/saved drafts/article_clean.tex
@@
 This commonly used method ``minimize[s] the number of model [grid] layers required to simulate an aquifer system'' \citep{modflow84}.
-However, the non-orthogonality of the resulting grid can be a source of significant model error \citep{hoaglund2003}.
+However, the non-orthogonality of the resulting grid can be a source of substantial model error \citep{hoaglund2003}.
 
@@
 
-The vertically offset grid with full connectivity (Figures \ref{fig:fig2}c and d) shows significantly improved results, even for the standard flow formulation (Figure \ref{fig:fig2}c), which overestimates the flux magnitude and total flow by approximately 3\% and 9\%, respectively.
+The vertically offset grid with full connectivity (Figures \ref{fig:fig2}c and d) shows substantially improved results, even for the standard flow formulation (Figure \ref{fig:fig2}c), which overestimates the flux magnitude and total flow by approximately 3\% and 9\%, respectively.
 Using the XT3D flow formulation provides the most accurate results, as expected (Figure \ref{fig:fig2}d), reducing the errors in flux magnitude and total flow to approximately $-2 \times 10^{-4}$\% and $2 \times 10^{-4}$\%, respectively.
@@
 When a single grid layer is used to represent the aquifer, the enhanced connectivity needed to achieve an accurate solution is not possible, and the XT3D flow formulation is unable to compensate for the inadequate connectivity.
-With full connectivity and the XT3D flow formulation, using more than two layers to discretize the aquifer provides no significant benefit in our benchmark problem (Figure \ref{fig:fig3}b).
+With full connectivity and the XT3D flow formulation, using more than two layers to discretize the aquifer provides no substantial benefit in our benchmark problem (Figure \ref{fig:fig3}b).
 
@@
 Results are presented in Figure \ref{fig:fig5}.
-Layered-connectivity grids produce results that deviate significantly from the analytical solution both with and without the XT3D formulation (Figure \ref{fig:fig5}a), while full-connectivity grids with XT3D reproduce the analytical solution (Figure \ref{fig:fig5}b, orange lines).
+Layered-connectivity grids produce results that deviate substantially from the analytical solution both with and without the XT3D formulation (Figure \ref{fig:fig5}a), while full-connectivity grids with XT3D reproduce the analytical solution (Figure \ref{fig:fig5}b, orange lines).
 For example, at an anisotropy ratio of 0.01, use of layered connnectivity produces flow errors of approximately -96\% for both the standard and XT3D flow formulations.
@@
 \cite{bardot2023} revisited the capability of MODFLOW 6 to accurately simulate groundwater flow through sedimentary structures.
-They observed that calculated flows involved significant error in a benchmark problem that simulated flow through a steeply dipping hydrogeologic layer using vertically offset grids, whether or not the XT3D multi-point flow formulation was used.
+They observed that calculated flows involved substantial error in a benchmark problem that simulated flow through a steeply dipping hydrogeologic layer using vertically offset grids, whether or not the XT3D multi-point flow formulation was used.
 They hypothesized that inadequate lateral connectivity between model cells was responsible for the inability to obtain accurate flow results in their benchmark problem.

### 2024-11-19 latex/saved drafts/article_tracked_changes.tex
intermediate review drafts & final draft in progress
intermediate review drafts & final draft in progress

--- 2024-11-19 latex/article_tracked_changes_FINAL.tex
@@
 {\color{red} \sout{In fact, it is common practice to represent each hydrogeologic layer using a single, boundary-matching model grid layer.
-}}However, the non-orthogonality of the resulting grid can be a source of {\color{red} \sout{significant}substantial} model error \citep{hoaglund2003}.
+}}However, the non-orthogonality of the resulting grid can be a source of significant model error \citep{hoaglund2003}.
 
@@
 
-The vertically offset grid with full connectivity (Figures \ref{fig:fig2}c and d) shows {\color{red} \sout{significantly}substantially} improved results, even for the standard flow formulation (Figure \ref{fig:fig2}c){\color{red}, which overestimates the flux magnitude and total flow by approximately 3\% and 9\%, respectively.} Using the XT3D flow formulation provides the most accurate results, as expected (Figure \ref{fig:fig2}d){\color{red} \sout{.}, reducing the errors in flux magnitude and total flow to approximately $-2 \times 10^{-4}$\% and $2 \times 10^{-4}$\%, respectively.}
+The vertically offset grid with full connectivity (Figures \ref{fig:fig2}c and d) shows significantly improved results, even for the standard flow formulation (Figure \ref{fig:fig2}c){\color{red}, which overestimates the flux magnitude and total flow by approximately 3\% and 9\%, respectively.} Using the XT3D flow formulation provides the most accurate results, as expected (Figure \ref{fig:fig2}d){\color{red} \sout{.}, reducing the errors in flux magnitude and total flow to approximately $-2 \times 10^{-4}$\% and $2 \times 10^{-4}$\%, respectively.}
 
@@
 Use of full connectivity reduces these errors to approximately 11\% and $3 \times 10^{-4}$\%, respectively.} When a single grid layer is used to represent the aquifer, the enhanced connectivity needed to achieve an accurate solution is not possible, and the XT3D flow formulation is unable to compensate for the inadequate connectivity.
-{\color{red} With full connectivity and the XT3D flow formulation, using more than two layers to discretize the aquifer provides no substantial benefit in our benchmark problem (Figure \ref{fig:fig3}b).}
+{\color{red} With full connectivity and the XT3D flow formulation, using more than two layers to discretize the aquifer provides no significant benefit in our benchmark problem (Figure \ref{fig:fig3}b).}
 
@@
 Results are presented in Figure \ref{fig:fig5}.
-Layered-connectivity grids produce results that deviate {\color{red} \sout{significantly}substantially} from the analytical solution both with and without the XT3D formulation (Figure \ref{fig:fig5}a), while full-connectivity grids with XT3D reproduce the analytical solution (Figure \ref{fig:fig5}b, orange lines).
+Layered-connectivity grids produce results that deviate significantly from the analytical solution both with and without the XT3D formulation (Figure \ref{fig:fig5}a), while full-connectivity grids with XT3D reproduce the analytical solution (Figure \ref{fig:fig5}b, orange lines).
 {\color{red} For example, at an anisotropy ratio of 0.01, use of layered connnectivity produces flow errors of approximately -96\% for both the standard and XT3D flow formulations.
@@
 \cite{bardot2023} revisited the capability of MODFLOW 6 to accurately simulate groundwater flow through sedimentary structures.
-They observed that calculated flows involved {\color{red} \sout{significant}substantial} error in a benchmark problem that simulated flow through a steeply dipping hydrogeologic layer using vertically offset grids, whether or not the XT3D multi-point flow formulation was used.
+They observed that calculated flows involved significant error in a benchmark problem that simulated flow through a steeply dipping hydrogeologic layer using vertically offset grids, whether or not the XT3D multi-point flow formulation was used.
 They hypothesized that inadequate lateral connectivity between model cells was responsible for the inability to obtain accurate flow results in their benchmark problem.

### 2024-11-22 latex/article_clean_FINAL.tex
make figures more accessible
* change colors and symbols used in plots to make them more accessible for colorblind readers
* update manuscript accordingly

@@
 	\begin{center}
-	\includegraphics[scale=0.6]{../figures/schem_conn_area_flux.png}
+	\includegraphics[scale=0.5]{../figures/fig1_paper.png}
 	\caption{Schematics showing (a) cell connectivity and (b) cell-cell interface fluxes in a two-grid-layer model of a dipping hydrogeologic layer.
@@
 In (a), black circles represent cell centers.
-Blue lines represent hydraulic connections between cells in a layered-connectivity grid.
-Red lines represent additional connections that must be made to achieve full connectivity.
-In (b), gray arrows represent the uniform groundwater flux the model is attempting to simulate.
-Blue arrows represent components of the uniform groundwater flux normal to the horizontal and vertical cell-cell interfaces in the layered-connectivity grid.
-Red arrows represent components of the uniform groundwater flux normal to the additional vertical cell-cell interfaces introduced to represent the full connectivity.}
+Solid blue lines represent hydraulic connections between cells in a layered-connectivity grid.
+Dashed red lines represent additional connections that must be made to achieve full connectivity.
+In (b), solid gray arrows represent the uniform groundwater flux the model is attempting to simulate.
+Solid blue arrows represent components of the uniform groundwater flux normal to the horizontal and vertical cell-cell interfaces in the layered-connectivity grid.
+Dashed red arrows represent components of the uniform groundwater flux normal to the additional vertical cell-cell interfaces introduced to represent the full connectivity.}
 	\label{fig:schem-conn-area-flux}
@@
 The focus is on a group of four cells in a two-grid-layer, vertically offset MODFLOW 6 model of fully saturated flow through a permeable hydrogeologic layer (aquifer) with impermeable top and bottom boundaries.
-If the grid is represented using a layered grid type (DIS or DISV), the grid has what we call ``layered connectivity": a cell is hydraulically connected to each laterally adjacent cell in the same grid layer (regardless of whether or not the cell faces actually overlap), and with each overlying and underlying cell, as indicated by the blue lines that connect cell centers in Figure \ref{fig:schem-conn-area-flux}a.
+If the grid is represented using a layered grid type (DIS or DISV), the grid has what we call ``layered connectivity": a cell is hydraulically connected to each laterally adjacent cell in the same grid layer (regardless of whether or not the cell faces actually overlap), and with each overlying and underlying cell, as indicated by the solid blue lines that connect cell centers in Figure \ref{fig:schem-conn-area-flux}a.
 In layered connectivity, two cells that have overlapping vertical faces but are in different grid layers (e.g., cells 1A and 2B) notably do not have a direct hydraulic connection.
-The addition of such connections between cells in different grid layers, as indicated by the red lines that connect cell centers in Figure \ref{fig:schem-conn-area-flux}a, results in a grid with what we call ``full connectivity.''  Implementation of full connectivity in the current version of MODFLOW 6 \citep{modflow650software} requires that the vertically offset grid be represented using the fully unstructured (DISU) grid type.
+The addition of such connections between cells in different grid layers, as indicated by the dashed red lines that connect cell centers in Figure \ref{fig:schem-conn-area-flux}a, results in a grid with what we call ``full connectivity.''  Implementation of full connectivity in the current version of MODFLOW 6 \citep{modflow650software} requires that the vertically offset grid be represented using the fully unstructured (DISU) grid type.
 
@@
 Gray vectors represent the uniform flux oriented along the hydrogeologic layer, which the model is attempting to simulate.
-Blue vectors represent the flux components normal to cell-cell interfaces that correspond to blue connections in Figure \ref{fig:schem-conn-area-flux}a, which are common to both the layered-connectivity and full-connectivity grids.
-Red vectors represent the flux components normal to cell-cell interfaces that correspond to red connections in Figure \ref{fig:schem-conn-area-flux}a, which are lacking in the layered-connectivity grid.
+Solid blue vectors represent the flux components normal to cell-cell interfaces that correspond to solid blue connections in Figure \ref{fig:schem-conn-area-flux}a, which are common to both the layered-connectivity and full-connectivity grids.
+Dashed red vectors represent the flux components normal to cell-cell interfaces that correspond to dashed red connections in Figure \ref{fig:schem-conn-area-flux}a, which are lacking in the layered-connectivity grid.
 
-The blue vectors in Figure \ref{fig:schem-conn-area-flux}b show that to simulate steady, uniform flow along the dipping hydrogeologic layer, cells must exchange water not only ``horizontally'' with neighboring cells in the same grid layer, but also vertically with cells in the adjacent grid layer.
+The solid blue vectors in Figure \ref{fig:schem-conn-area-flux}b show that to simulate steady, uniform flow along the dipping hydrogeologic layer, cells must exchange water not only ``horizontally'' with neighboring cells in the same grid layer, but also vertically with cells in the adjacent grid layer.
 Specifically, there must be flow from cells in the lower grid layer to cells in the upper grid layer, which corresponds to the vertical component of the groundwater flux.
-Note, however, that the blue vectors alone are incompatible with steady, uniform flow along the hydrogeologic layer because the exchange of water between the two grid layers is in one direction only: upward.
+Note, however, that the solid blue vectors alone are incompatible with steady, uniform flow along the hydrogeologic layer because the exchange of water between the two grid layers is in one direction only: upward.
 Such a steady-state flow configuration would imply continuous depletion of flow within the lower grid layer and accumulation of flow within the upper grid layer as one moves along the hydrogeologic layer in the direction of flow.
@@
 The full-connectivity grid provides such a mechanism by offering additional connections between cells in adjacent grid layers.
-Vertical flows from the lower grid layer to the upper grid layer, represented by vertical blue vectors in Figure \ref{fig:schem-conn-area-flux}b, can be returned to the lower grid layer by nominally ``horizontal'' flows, represented by red vectors in Figure \ref{fig:schem-conn-area-flux}b.
+Vertical flows from the lower grid layer to the upper grid layer, represented by vertical solid blue vectors in Figure \ref{fig:schem-conn-area-flux}b, can be returned to the lower grid layer by nominally ``horizontal'' flows, represented by dashed red vectors in Figure \ref{fig:schem-conn-area-flux}b.
 Although this concept has been illustrated using a two-grid-layer model for simplicity, the same reasoning applies given any number of grid layers.
@@
 The middle panel shows the face flows (arrows).
-The right panel shows head contours for the analytical solution (black) as well as the numerical solution (blue).
+The right panel shows head contours for the analytical solution (dashed black) as well as the numerical solution (solid blue).
 Flux magnitudes, flux directions (angles), and head errors are reported to two decimal places.
@@
 	\begin{center}
-	\includegraphics[scale=0.9]{../figures/fig3paper.png}
+	\includegraphics[scale=0.9]{../figures/fig3_paper.png}
 	\caption{Graphs indicating how the number of grid layers within the aquifer affects the flux magnitude in the center cell (top), the flux direction in center cell (middle) and the volumetric flow through the aquifer (bottom).
-Errors are minimized by using full connectivity (right column), the XT3D flow formulation (orange lines) and at least two grids layers in the aquifer.}
+Errors are minimized by using full connectivity (right column), the XT3D flow formulation (orange lines with solid circles) and at least two grids layers in the aquifer.}
 	\label{fig:fig3}
@@
 In this subsection, we investigate the behavior of the flow solution for aquifer dip angles ranging from 0 to 80$^{\circ}$.
-We also examine the behavior for a range of conductivity contrasts by setting the domain conductivity to 1/2, 1/5, 1/10 and 1/100 that of the aquifer.
+We also examine the behavior for a range of conductivity contrasts by setting the domain conductivity equal to and to 1/2, 1/5, 1/10 and 1/100 that of the aquifer.
 Simulated flux magnitude and direction at the center cell for all combinations are plotted against the analytical solution in Figure \ref{fig:fig4}.
@@
 
-\caption{Flux magnitude and direction at the cell in the center of the aquifer as a function of aquifer dip angle for various ratios of domain to aquifer conductivity (blue, orange, green, red lines).
-The analytical solution is shown as a black dashed line.}
+\caption{Flux magnitude and direction at the cell in the center of the aquifer as a function of aquifer dip angle for various ratios of domain to aquifer conductivity (solid and dash-dotted lines of varying color).
+The analytical solution is shown as a dashed black line.}
 \label{fig:fig4}
@@
 
-Layered-connectivity grids produce results that deviate substantially from the analytical solution both with and without the XT3D formulation (Figure \ref{fig:fig5}a), whereas full-connectivity grids with XT3D reproduce the analytical solution (Figure \ref{fig:fig5}b, orange lines).
+Layered-connectivity grids produce results that deviate substantially from the analytical solution both with and without the XT3D formulation (Figure \ref{fig:fig5}a), whereas full-connectivity grids with XT3D reproduce the analytical solution (Figure \ref{fig:fig5}b).
 For example, at an anisotropy ratio of 0.01, use of layered connnectivity produces flow errors of approximately -96\% for both the standard and XT3D flow formulations.
@@
 	\begin{center}
-	\includegraphics[scale=0.9]{../figures/fig5paper.png}
+	\includegraphics[scale=0.9]{../figures/fig5_paper.png}
 	\caption{Graphs indicating how the anisotropy ratio within the aquifer and domain affects the flux magnitude in the center cell (top), the flux direction in center cell (middle) and the volumetric flow through the aquifer (bottom).
-Errors are minimized by using full connectivity (right column) and the XT3D flow formulation (orange lines).}
+Errors are minimized by using full connectivity (right column) and the XT3D flow formulation (orange lines with solid circles).}
 	\label{fig:fig5}

### 2024-12-05 latex/article_clean_FINAL - Copy.tex
last-minute tweaks made before resubmission
last-minute tweaks made before resubmission

--- 2024-11-22 latex/article_clean_FINAL.tex
@@
 	\begin{center}
-	\includegraphics[scale=0.5]{../figures/fig1_paper.png}
+	\includegraphics[scale=1.0]{../figures/fig1_paper.png}
 	\caption{Schematics showing (a) cell connectivity and (b) cell-cell interface fluxes in a two-grid-layer model of a dipping hydrogeologic layer.
@@
 Such a steady-state flow configuration would imply continuous depletion of flow within the lower grid layer and accumulation of flow within the upper grid layer as one moves along the hydrogeologic layer in the direction of flow.
-Thus, the steady, uniform flow represented by the gray vectors cannot be simulated accurately without some mechanism for returning flow from the upper grid layer to the lower grid layer.
+Thus, the steady, uniform flow represented by the solid gray vectors cannot be simulated accurately without some mechanism for returning flow from the upper grid layer to the lower grid layer.
 The full-connectivity grid provides such a mechanism by offering additional connections between cells in adjacent grid layers.
@@
 
-A link to the Jupyter notebooks and MODFLOW version 6.5.0 executable code used to create and run the test problem is available in the Supporting Information section.
+A link to the Jupyter notebooks \citep{jupyterproject} and MODFLOW version 6.5.0 executable code used to create and run the test problem is available in the Supporting Information section.
 The notebooks leverage the capabilities of FloPy \citep{bakker2016scripting, hughes2024flopy} to assist in setting up input for and processing results from the MODFLOW 6 simulations.
-A Python script for converting structured, DIS grids to unstructured, DISU grids with full connectivity is included.
+A Python \citep{pythonsoftware} script for converting structured, DIS grids to unstructured, DISU grids with full connectivity is included.
 
@@
 Kerry Bardot is supported by The Australian Research Council, The Western Australian Department of Water and Environmental Regulation and The Water Corporation of Western Australia through grant number LP18101153.
-The authors sincerely thank James Rumbaugh, John Masterson, Kalle Jahn, Jacob LaFontaine, John Brakebill, Heather Welch, Randall Hunt, and Richard Yager for their review comments, which improved the clarity of the manuscript and refined our conclusions regarding the practical implications of this study.
+The authors sincerely thank James Rumbaugh, John Masterson, Kalle Jahn, Jacob LaFontaine, Randall Hunt, and Richard Yager for their review comments, which improved the clarity of the manuscript and refined our conclusions regarding the practical implications of this study.
 Any use of trade, firm, or product names is for descriptive purposes only and does not imply endorsement by the U.S.
@@
 \section*{Supporting Information}
-The files that support the creation and running of the test problem are available at https://doi.org/10.5066/P13BNARA.
+The files that support the creation and running of the test problem are available at https://doi.org/10.5066/P13BNARA \citep{suppinfodatarelease}.
 

### 2024-12-05 latex/article_tracked_changes_FINAL - Copy.tex
last-minute tweaks made before resubmission
last-minute tweaks made before resubmission

--- 2024-11-19 latex/article_tracked_changes_FINAL.tex
@@
 When this approach is used in MODFLOW, adjacent cells within the {\color{red} same} model layer are vertically offset from one another, and the standard conductance-based (two-point) formulation for flow between cells does not rigorously account for these offsets.
-The XT3D multi-point flow formulation introduced in MODFLOW 6 is designed to account for geometric irregularities in the grid, including vertical offsets, and to provide accurate results for both isotropic and anisotropic groundwater flow.
+The XT3D multi-point flow formulation {\color{red} \sout{introduced }}in MODFLOW 6 is designed to account for geometric irregularities in the grid, including vertical offsets, and to provide accurate results for both isotropic and anisotropic groundwater flow.
 A recent study evaluated the performance of the standard formulation and XT3D using a simple, synthetic benchmark model of a steeply dipping aquifer.
@@
 A traditional approach to discretizing layered aquifer systems, called the ``grid overlay'' method by \cite{hoaglund2003}, maps the heterogeneous hydrologic properties onto a rectilinear grid \citep{modflow84}.
-While this method preserves the benefits of the rectilinear grid structure, it can require a large number of model grid layers to adequately discretize hydrogeogic layers of variable dip and thickness \citep{Zyvoloski2006}.
+{\color{red} \sout{While}Although} this method preserves the benefits of the rectilinear grid structure, it can require a large number of model grid layers to adequately discretize hydrogeogic layers of variable dip and thickness \citep{Zyvoloski2006}.
 
@@
 To help overcome the limitations of the two-point flow formulation on geometrically irregular grids, MODFLOW-USG and MODFLOW 6 offer a Ghost-Node Correction (GNC) Package \citep{modflowusg, modflow6gwf}.
-However, configuration of optimal ghost node locations and interpolation weights is highly problem-dependent, and we are not aware of guidelines for handling general unstructured grids or arbitrarily oriented 3D anisotropy.
+However, configuration of optimal ghost node locations and interpolation weights is {\color{red} \sout{highly }}problem-dependent, and we are not aware of guidelines for handling general unstructured grids or arbitrarily oriented 3D anisotropy.
 
-The optional XT3D flow formulation \citep{modflow6xt3d} introduced in MODFLOW 6 offers a more ``automatic" alternative to ghost nodes.
+The optional XT3D flow formulation \citep{modflow6xt3d} {\color{red} \sout{introduced }}in MODFLOW 6 offers a more ``automatic" alternative to ghost nodes.
 XT3D is a multi-point flow formulation \citep{edwards1998, aavatsmark2002} based on gradient reconstruction \citep{mavriplis2003leastsq, diskin2008accuracy}.
@@
 Thus, when used to represent a grid with a layered structure, as in this work, the DISU grid type allows connections to exist between any two cells, whether they are in the same layer or different layers.
-Grid geometry and connectivity information supplied by the discretization package (DIS, DISV, or DISU) is used by the MODFLOW 6 Node Property Flow (NPF) Package to formulate expressions for groundwater flow between cells in terms of heads calculated at cells centers.
+Grid geometry and connectivity information supplied by the discretization package (DIS, DISV, or DISU) is used by the MODFLOW 6 Node Property Flow (NPF) Package to formulate expressions for groundwater flow between cells in terms of heads calculated at cell centers.
 By default, the NPF Package uses a two-point flow formulation, which we call the ``conductance-based'' formulation.
@@
 	\begin{center}
-	\includegraphics[scale=0.6]{../figures/schem_conn_area_flux.png}
+	\includegraphics[scale=1.0]{../figures/fig1_paper.png}
 	\caption{Schematics showing (a) cell connectivity and (b) cell-cell interface fluxes in a two-grid-layer model of a dipping hydrogeologic layer.
@@
 In (a), black circles represent cell centers.
-Blue lines represent hydraulic connections between cells in a layered-connectivity grid.
-Red lines represent additional connections that must be made to achieve full connectivity.
-In (b), gray arrows represent the uniform groundwater flux the model is attempting to simulate.
-Blue arrows represent components of the uniform groundwater flux normal to the horizontal and vertical cell-cell interfaces in the layered-connectivity grid.
-Red arrows represent components of the uniform groundwater flux normal to the additional vertical cell-cell interfaces introduced to represent the full connectivity.}
+{\color{red} \sout{Blue}Solid Blue} lines represent hydraulic connections between cells in a layered-connectivity grid.
+{\color{red} \sout{Red}Dashed red} lines represent additional connections that must be made to achieve full connectivity.
+In (b), {\color{red}solid} gray arrows represent the uniform groundwater flux the model is attempting to simulate.
+{\color{red} \sout{Blue}Solid Blue} arrows represent components of the uniform groundwater flux normal to the horizontal and vertical cell-cell interfaces in the layered-connectivity grid.
+{\color{red} \sout{Red}Dashed red} arrows represent components of the uniform groundwater flux normal to the additional vertical cell-cell interfaces introduced to represent the full connectivity.}
 	\label{fig:schem-conn-area-flux}
@@
 The focus is on a group of four cells in a two-grid-layer, vertically offset MODFLOW 6 model of fully saturated flow through a permeable hydrogeologic layer (aquifer) with impermeable top and bottom boundaries.
-If the grid is represented using a layered grid type (DIS or DISV), the grid has what we call ``layered connectivity": a cell is hydraulically connected to each laterally adjacent cell in the same grid layer (regardless of whether or not the cell faces actually overlap), and with each overlying and underlying cell, as indicated by the blue lines that connect cell centers in Figure \ref{fig:schem-conn-area-flux}a.
+If the grid is represented using a layered grid type (DIS or DISV), the grid has what we call ``layered connectivity": a cell is hydraulically connected to each laterally adjacent cell in the same grid layer (regardless of whether or not the cell faces actually overlap), and with each overlying and underlying cell, as indicated by the {\color{red}solid} blue lines that connect cell centers in Figure \ref{fig:schem-conn-area-flux}a.
 In layered connectivity, two cells that have overlapping vertical faces but are in different grid layers (e.g., cells 1A and 2B) notably do not have a direct hydraulic connection.
-The addition of such connections between cells in different grid layers, as indicated by the red lines that connect cell centers in Figure \ref{fig:schem-conn-area-flux}a, results in a grid with what we call ``full connectivity.''  Implementation of full connectivity in the current version of MODFLOW 6 \citep{modflow650software} requires that the vertically offset grid be represented using the fully unstructured (DISU) grid type.
+The addition of such connections between cells in different grid layers, as indicated by the {\color{red}dashed} red lines that connect cell centers in Figure \ref{fig:schem-conn-area-flux}a, results in a grid with what we call ``full connectivity.''  Implementation of full connectivity in the current version of MODFLOW 6 \citep{modflow650software} requires that the vertically offset grid be represented using the fully unstructured (DISU) grid type.
 
@@
 Gray vectors represent the uniform flux oriented along the hydrogeologic layer, which the model is attempting to simulate.
-Blue vectors represent the flux components normal to cell-cell interfaces that correspond to blue connections in Figure \ref{fig:schem-conn-area-flux}a, which are common to both the layered-connectivity and full-connectivity grids.
-Red vectors represent the flux components normal to cell-cell interfaces that correspond to red connections in Figure \ref{fig:schem-conn-area-flux}a, which are lacking in the layered-connectivity grid.
+{\color{red} \sout{Blue}Solid blue} vectors represent the flux components normal to cell-cell interfaces that correspond to {\color{red}solid} blue connections in Figure \ref{fig:schem-conn-area-flux}a, which are common to both the layered-connectivity and full-connectivity grids.
+{\color{red} \sout{Red}Dashed red} vectors represent the flux components normal to cell-cell interfaces that correspond to {\color{red}dashed} red connections in Figure \ref{fig:schem-conn-area-flux}a, which are lacking in the layered-connectivity grid.
 
-The blue vectors in Figure \ref{fig:schem-conn-area-flux}b show that to simulate steady, uniform flow along the dipping hydrogeologic layer, cells must exchange water not only ``horizontally'' with neighboring cells in the same grid layer, but also vertically with cells in the adjacent grid layer.
+The {\color{red}solid} blue vectors in Figure \ref{fig:schem-conn-area-flux}b show that to simulate steady, uniform flow along the dipping hydrogeologic layer, cells must exchange water not only ``horizontally'' with neighboring cells in the same grid layer, but also vertically with cells in the adjacent grid layer.
 Specifically, there must be flow from cells in the {\color{red} \sout{bottom}lower} grid layer to cells in the {\color{red} \sout{top}upper} grid layer, which corresponds to the vertical component of the groundwater flux.
-Note, however, that the blue vectors alone are incompatible with steady, uniform flow along the hydrogeologic layer because the exchange of water between the two grid layers is in one direction only: upward.
+Note, however, that the {\color{red}solid} blue vectors alone are incompatible with steady, uniform flow along the hydrogeologic layer because the exchange of water between the two grid layers is in one direction only: upward.
 Such a steady-state flow configuration would imply continuous depletion of flow within the {\color{red} \sout{bottom}lower} grid layer and accumulation of flow within the {\color{red} \sout{top}upper} grid layer as one moves along the hydrogeologic layer in the direction of flow.
-Thus, {\color{red} the} steady, uniform flow {\color{red} \sout{along the hydrogeologic layer}represented by the gray vectors} cannot be simulated accurately without some mechanism for returning flow from the {\color{red} \sout{top}upper} grid layer to the {\color{red} \sout{bottom}lower} grid layer.
+Thus, {\color{red} the} steady, uniform flow {\color{red} \sout{along the hydrogeologic layer}represented by the {\color{red}solid} gray vectors} cannot be simulated accurately without some mechanism for returning flow from the {\color{red} \sout{top}upper} grid layer to the {\color{red} \sout{bottom}lower} grid layer.
 The full-connectivity grid provides such a mechanism by offering additional connections between cells in adjacent grid layers.
-Vertical flows from the {\color{red} \sout{bottom}lower} grid layer to the {\color{red} \sout{top}upper} grid layer, represented by vertical blue vectors in Figure \ref{fig:schem-conn-area-flux}b, can be returned to the {\color{red} \sout{bottom}lower} grid layer by nominally ``horizontal'' flows, represented by red vectors in Figure \ref{fig:schem-conn-area-flux}b.
+Vertical flows from the {\color{red} \sout{bottom}lower} grid layer to the {\color{red} \sout{top}upper} grid layer, represented by vertical {\color{red}solid} blue vectors in Figure \ref{fig:schem-conn-area-flux}b, can be returned to the {\color{red} \sout{bottom}lower} grid layer by nominally ``horizontal'' flows, represented by {\color{red}dashed} red vectors in Figure \ref{fig:schem-conn-area-flux}b.
 Although this concept has been illustrated using a two-grid-layer model for simplicity, the same reasoning applies given any number of grid layers.
 
-Note how the vertically offset grid violates the CFVD requirement because nominally ``horizontal'' connections between cell centers in the same grid layer are not strictly horizontal and therefore not perpendicular to the vertical cell-cell interfaces they intersect.
+{\color{red} \sout{Note how the vertically offset grid violates}Vertically offset grids violate} the CFVD requirement because nominally ``horizontal'' connections between cell centers in the same grid layer are not strictly horizontal and therefore not perpendicular to the vertical cell-cell interfaces they intersect.
 For example, in Figure \ref{fig:schem-conn-area-flux}a, the connection between cells 1A and 1B is not perpendicular to the cell-cell interface, which reduces the accuracy of the conductance-based formulation.
@@
 The test simulations use grids of MODFLOW 6 DISU (unstructured) type, which offers the flexibility to implement either layered or full connectivity.
-In all cases, a grid of MODFLOW 6 DIS (structured) type is first created and then converted a DISU grid.
+In all cases, a grid of MODFLOW 6 DIS (structured) type is first created and then converted {\color{red}to} a DISU grid.
 In simulations that use layered connectivity, the DISU grid is formulated to have the same cell geometry and connectivity as the original DIS grid.
@@
 
-A link to the Jupyter notebooks and MODFLOW version 6.5.0 executable code used to create and run the test problem is available in the Supporting Information section.
+A link to the Jupyter notebooks {\color{red} \citep{jupyterproject}} and MODFLOW version 6.5.0 executable code used to create and run the test problem is available in the Supporting Information section.
 The notebooks leverage the capabilities of FloPy \citep{bakker2016scripting, hughes2024flopy} to assist in setting up input for and processing results from the MODFLOW 6 simulations.
-A Python script for converting structured, DIS grids to unstructured, DISU grids with full connectivity is included.
+A Python {\color{red} \citep{pythonsoftware}} script for converting structured, DIS grids to unstructured, DISU grids with full connectivity is included.
 
@@
 Therefore, specific discharge calculated in a cell adjacent to the aquifer-domain boundary is affected by the different flow rate on the other side of the boundary.
-While this is an unavoidable consequence of vertically offset discretization in practical applications, and not necessarily a problem, users should be aware of these effects when interpreting the results of MODFLOW 6 dispersive-transport and MODPATH particle-tracking simulations.
+{\color{red} \sout{While}Although} this is an unavoidable consequence of vertically offset discretization in practical applications, and not necessarily a problem, users should be aware of these effects when interpreting the results of MODFLOW 6 dispersive-transport and MODPATH particle-tracking simulations.
 
@@
 The middle panel shows the face flows (arrows).
-The right panel shows head contours for the analytical solution (black) as well as the numerical solution (blue).
+The right panel shows head contours for the analytical solution ({\color{red}dashed} black) as well as the numerical solution ({\color{red}solid} blue).
 {\color{red} Flux magnitudes, flux directions (angles), and head errors are reported to two decimal places.
@@
 	\begin{center}
-	\includegraphics[scale=0.9]{../figures/fig3paper.png}
+	\includegraphics[scale=0.9]{../figures/fig3_paper.png}
 	\caption{Graphs indicating how the number of grid layers within the aquifer affects the flux magnitude in the center cell (top), the flux direction in center cell (middle) and the volumetric flow through the aquifer (bottom).
-Errors are minimized by using full connectivity (right column), the XT3D flow formulation (orange lines) and at least two grids layers in the aquifer.
+Errors are minimized by using full connectivity (right column), the XT3D flow formulation (orange lines {\color{red}with solid circles}) and at least two grids layers in the aquifer.
 {\color{red} \textit{Note to editor and reviewers: "channel" replaced with "aquifer" in axis labels.}}}
@@
 In this subsection, we investigate the behavior of the flow solution for aquifer dip angles ranging from 0 to 80$^{\circ}$.
-We also examine the behavior for a range of conductivity contrasts by setting the domain conductivity to 1/2, 1/5, 1/10 and 1/100 that of the aquifer.
+We also examine the behavior for a range of conductivity contrasts by setting the domain conductivity {\color{red}equal to and} to 1/2, 1/5, 1/10 and 1/100 that of the aquifer.
 Simulated flux magnitude and direction at the center cell for all combinations are plotted against the analytical solution in Figure \ref{fig:fig4}.
@@
 
-\caption{Flux magnitude and direction at the cell in the center of the aquifer as a function of aquifer dip angle for various ratios of domain to aquifer conductivity (blue, orange, green, red lines).
-The analytical solution is shown as a black dashed line.}
+\caption{Flux magnitude and direction at the cell in the center of the aquifer as a function of aquifer dip angle for various ratios of domain to aquifer conductivity ({\color{red} \sout{blue, orange, green, red lines}solid and dash-dotted lines of varying color}).
+The analytical solution is shown as a {\color{red} \sout{black dashed}dashed black} line.}
 \label{fig:fig4}
@@
 {\color{red} \sout{Therefore, we examine the effect of anisotropic and dipping conductivity tensors within the aquifer and domain, with conductivity perpendicular to the aquifer reduced by factors of 10, 100, 1000 and 10,000.
-We also include the isotropic scenario (anisotropy ratio of 1).}Therefore, we examine the effect of anisotropy within the aquifer and domain by dividing the original conductivity perpendicular to the dip of the aquifer by factors of 1, 10, 100, 1000 and 10,000 to give anisotropy ratios ranging from 1.0 (isotropic) to 0.0001.} {\color{red} Orientation of the principal directions of anisotropy parallel and perpendicular to the dip of the aquifer is accomplished by setting three angles in the NPF Package input.
+We also include the isotropic scenario (anisotropy ratio of 1).}Therefore, we examine the effect of anisotropy within the aquifer and domain by dividing the original conductivity perpendicular to the dip of the aquifer by factors of 1, 10, 100, 1,000 and 10,000 to give anisotropy ratios ranging from 1.0 (isotropic) to 0.0001.} {\color{red} Orientation of the principal directions of anisotropy parallel and perpendicular to the dip of the aquifer is accomplished by setting three angles in the NPF Package input.
 The first angle is the rotation within the horizontal plane of the first principal direction (typically the direction of maximum hydraulic conductivity) with respect to the positive $x$ axis.
@@
 
-Results are presented in Figure \ref{fig:fig5}.
-Layered-connectivity grids produce results that deviate {\color{red} \sout{significantly}substantially} from the analytical solution both with and without the XT3D formulation (Figure \ref{fig:fig5}a), while full-connectivity grids with XT3D reproduce the analytical solution (Figure \ref{fig:fig5}b, orange lines).
+{\color{red} \sout{Results are presented in Figure \ref{fig:fig5}.
+}}Layered-connectivity grids produce results that deviate {\color{red} \sout{significantly}substantially} from the analytical solution both with and without the XT3D formulation (Figure \ref{fig:fig5}a), {\color{red} \sout{while}whereas} full-connectivity grids with XT3D reproduce the analytical solution (Figure \ref{fig:fig5}b{\color{red} \sout{, orange lines}}).
 {\color{red} For example, at an anisotropy ratio of 0.01, use of layered connnectivity produces flow errors of approximately -96\% for both the standard and XT3D flow formulations.
@@
 	\begin{center}
-	\includegraphics[scale=0.9]{../figures/fig5paper.png}
+	\includegraphics[scale=0.9]{../figures/fig5_paper.png}
 	\caption{Graphs indicating how the anisotropy ratio within the aquifer and domain affects the flux magnitude in the center cell (top), the flux direction in center cell (middle) and the volumetric flow through the aquifer (bottom).
-Errors are minimized by using full connectivity (right column) and the XT3D flow formulation (orange lines).}
+Errors are minimized by using full connectivity (right column) and the XT3D flow formulation (orange lines {\color{red}with solid circles}).}
 	\label{fig:fig5}
@@
 In a complex, real-world model, uncertainty in the structure and properties of hydrogeologic units, stresses on the system, and other model parameters could lead to compensation for the effects of inadequate cell connectivity during model calibration.
-However, when unrealistic flows or parameters values are observed in a calibrated model, inadequate cell connectivity can be considered among the possible sources of structural error, and full connectivity can be considered as remedy.}
+However, when unrealistic flows or parameter values are observed in a calibrated model, inadequate cell connectivity can be considered among the possible sources of structural error, and full connectivity can be considered as a remedy.}
 
@@
 {\color{red} \sout{The authors offer their sincere thanks to the reviewers of this manuscript.
-}The authors sincerely thank James Rumbaugh, John Masterson, Kalle Jahn, Jacob LaFontaine, Randall Hunt, and Richard Yager for their review comments, which improved the clarity of the manuscript and refined our conclusions regarding the practical implications of this study.}
+}The authors sincerely thank James Rumbaugh, John Masterson, Kalle Jahn, Jacob LaFontaine, Randall Hunt, and Richard Yager for their review comments, which improved the clarity of the manuscript and refined our conclusions regarding the practical implications of this study.
+Any use of trade, firm, or product names is for descriptive purposes only and does not imply endorsement by the U.S.
+Government.}
 
@@
 {\color{red} \sout{During review, the files that support the creation and running of the test problem are available at https://github.com/langevin-usgs/xt3d-vertically-staggered/releases/download/v1.0/suppinfo.zip.
-Once these files are finalized prior to publication, they will be made available at a permanent URL that will be noted here.}}{\color{red} The files that support the creation and running of the test problem are available at https://doi.org/10.5066/P13BNARA.}
+Once these files are finalized prior to publication, they will be made available at a permanent URL that will be noted here.}}{\color{red} The files that support the creation and running of the test problem are available at https://doi.org/10.5066/P13BNARA \citep{suppinfodatarelease}.}
 
@@
 
+\noindent{\color{red} \textit{Note to editor: colors and symbols in the figures were modified to improve accessibility for colorblind readers.}}
+
 \bibliography{references.bib}
